### 2013-07-22 thesis/src/chapters/chapter-1.tex
Initial commit of the written-part of the thesis
Introduces the directory structure, makefile for compiling the sources
and layout of the frontpage.

\chapter{Draft chapter}
\section{Draft section}

section contents

### 2013-07-22 auto-scaling/cloud-provider/src/opennebula/appstage_client.rb
[#18] getting ip address of a container

@@ -9,6 +9,11 @@ module AutoScaling
 
     @@logger = Logger.new(STDOUT)
 
+    CLIENT = {
+        :retries => 3,
+        :sleep => 2
+    }
+
     def initialize(options)
       @options = options
     end
@@ -19,10 +24,24 @@ module AutoScaling
         # output is in a format: "VM ID: 168"
         id = output.split(' ')[2]
 
-        @@logger.debug("Instantiated vm: #{id} with appstage_id #{appstage_id} and tempalte #{template_id}")
+        @@logger.debug("Instantiated vm: #{id} with appstage_id #{appstage_id} and template #{template_id}")
 
-        # get ip address here
+        # get ip address
+        retry_count = 0
         ip = ''
+        begin
+          if ip == nil
+            @@logger.debug "Waiting for scheduling vm to get an ip address"
+            sleep CLIENT[:sleep]
+          end
+
+          xml = ssh.exec!("onevm show #{id} --xml")
+          ip = extract_ip xml
+          retry_count += 1
+        end while retry_count <= CLIENT[:retries] and ip != nil
+
+        raise RuntimeError, "Can't get container ip" if ip == nil
+        @@logger.debug "VM #{id} has an address: #{ip}"
 
         {:id => id.to_i, :ip => ip}
       end

### 2013-07-22 auto-scaling/cloud-provider/src/opennebula/opennebula_client.rb

@@ -39,6 +39,10 @@ module AutoScaling
       @appstage.instantiate_container appstage_id, template_id
     end
 
+    def delete_container(container_id)
+      @appstage.delete_container container_id
+    end
+
     def render(service, bindings)
       ServiceRenderer::render(service, bindings)
     end

### 2013-07-22 auto-scaling/cloud-provider/test/integration/opennebula/opennebula_client_test.rb

@@ -63,10 +63,12 @@ eos
       appstage_id = 25
       template_id = 7
 
-      instance_id = @opennebula_client.instantiate_container(appstage_id, template_id)
-      assert_equal true, instance_id >= 0
+      container_info = @opennebula_client.instantiate_container(appstage_id, template_id)
+      assert_equal true, container_info[:id] >= 0
+      assert_equal true, container_info[:ip] != ''
+
       # cleanup
-      @opennebula_client.appstage.delete_container instance_id
+      @opennebula_client.delete_container container_info[:id]
     end
 
   end

### 2013-07-22 auto-scaling/service-controller/src/executor/service_executor.rb
[#18] configuring master after node addition

@@ -59,7 +59,7 @@ module AutoScaling
 
       container_info = @cloud_provider.instantiate_container(appstage_id, template_id)
 
-      # TODO add ip
+      # persist data
       container = Container.create(
           :id => container_info[:id],
           :ip => container_info[:ip]
@@ -67,28 +67,16 @@ module AutoScaling
       stack.containers << container
 
       stack.save
+
+      # reconfigure master
+      configure(Container.master(stack))
     end
 
     def converge(service, container_id)
       update(service) if(service.status != :converged)
       container = ::AutoScaling::Container.get(container_id)
 
-      if(container.type == :master)
-        chef = ChefRenderer.render container.stack
-        url = "http://#{container.ip}:4567/chef"
-
-        @@logger.debug "Sending configuration #{chef} to #{url}"
-
-        RestClient.post(url, :node_object_data => Base64::encode64(chef)) { |response, request, result, &block|
-          case response.code
-            when 200
-              @@logger.debug "Successfully sent configuration #{response}"
-              response
-            else
-              raise RuntimeError, "An error occurred during sending configuration, status: #{response.code}"
-          end
-        }
-      end
+      configure(container) if container.type == :master
     end
 
     private
@@ -122,6 +110,23 @@ module AutoScaling
       # TODO add synchronization
       service.status = :converged
       service.save
+    end
+
+    def configure(container)
+      chef = ChefRenderer.render container.stack
+      url = "http://#{container.ip}:4567/chef"
+
+      @@logger.debug "Sending configuration #{chef} to #{url}"
+
+      RestClient.post(url, :node_object_data => Base64::encode64(chef)) { |response, request, result, &block|
+        case response.code
+          when 200
+            @@logger.debug "Successfully sent configuration #{response}"
+            response
+          else
+            raise RuntimeError, "An error occurred during sending configuration, status: #{response.code}"
+        end
+      }
     end
 
   end

### 2013-07-22 auto-scaling/service-controller/src/models/container.rb

@@ -9,7 +9,7 @@ module AutoScaling
     property :id, Integer, :key => true
     belongs_to :stack
 
-    property :ip, IPAddress
+    property :ip, IPAddress, :required => true
     property :type, Enum[ :master, :slave ], :default => :slave, :required => true
 
     def self.master(stack)

### 2013-07-22 auto-scaling/service-controller/test/unit/executor/service_executor_test.rb

@@ -138,6 +138,7 @@ module AutoScaling
 
     def test_shall_deploy_new_container
       instance_id = 69
+      master_ip = '192.168.122.200'
 
       containers = [
           Container.create(
@@ -146,7 +147,7 @@ module AutoScaling
           ),
           Container.create(
               :id => 0,
-              :ip => '192.168.122.200',
+              :ip => master_ip,
               :type => :master
           )
       ]
@@ -164,7 +165,13 @@ module AutoScaling
       )
 
       container_id = 800
-      @cloud_provider.expects(:instantiate_container).with(25, 7).returns(container_id)
+      container_ip = '192.168.122.69'
+      response = 'Od przedszkola do Opola, kupuj szybko dzis bejzbola'
+      @cloud_provider.expects(:instantiate_container).with(25, 7).returns({:id => container_id, :ip=> container_ip})
+      FakeWeb.register_uri(:post,
+                           "http://#{master_ip}:4567/chef",
+                           :body => response,
+                           :status => ["200", "OK"])
 
       id = @executor.deploy_container stack, MAPPINGS
       assert_equal 2, Container.slaves(stack).size

### 2013-07-22 auto-scaling/service-controller/src/executor/service_executor.rb
[#18] deleting container from stack

@@ -72,6 +72,18 @@ module AutoScaling
       configure(Container.master(stack))
     end
 
+    def delete_container(stack, mappings = {})
+      slaves = Container.slaves(stack)
+      raise RuntimeError, "Can't delete slave from stack: #{stack.id}, there aren't any left" if slaves.size == 0
+
+      slave = slaves.pop
+      @cloud_provider.delete_container slave.id
+      stack.save
+
+      # reconfigure master
+      configure(Container.master(stack))
+    end
+
     def converge(service, container_id)
       update(service) if(service.status != :converged)
       container = ::AutoScaling::Container.get(container_id)

### 2013-07-22 auto-scaling/service-controller/test/unit/executor/service_executor_test.rb

@@ -179,6 +179,64 @@ module AutoScaling
       assert Container.slaves(stack).include?(Container.get(container_id))
     end
 
+    def test_shall_delete_container
+      instance_id = 69
+      master_ip = '192.168.122.200'
+
+      containers = [
+          Container.create(
+              :id => 10,
+              :ip => '192.168.122.1'
+          ),
+          Container.create(
+              :id => 0,
+              :ip => master_ip,
+              :type => :master
+          )
+      ]
+
+      stack = Stack.create(
+          :type => 'java',
+          :containers => containers
+      )
+
+      Service.create(
+          :id => instance_id,
+          :name => 'service-name',
+          :stacks => [stack],
+          :status => :converged
+      )
+
+      response = 'Hulaj dusza, baki ida'
+      FakeWeb.register_uri(:post,
+                           "http://#{master_ip}:4567/chef",
+                           :body => response,
+                           :status => ["200", "OK"])
+      @cloud_provider.expects(:delete_container).with(10)
+
+      @executor.delete_container(Stack.get(stack.id))
+      assert_equal 1, Container.slaves(stack).size
+    end
+
+    def test_shall_throw_exception_when_there_is_no_slaves_left
+      instance_id = 69
+      stack = Stack.create(
+          :type => 'java',
+          :containers => []
+      )
+
+      Service.create(
+          :id => instance_id,
+          :name => 'service-name',
+          :stacks => [stack],
+          :status => :converged
+      )
+
+      assert_raise RuntimeError do
+        @executor.delete_container(Stack.get(stack.id))
+      end
+    end
+
   end
 
 end

### 2013-07-22 auto-scaling/bin/auto_scaling_server.rb
[#18] added scale up / down interfaces

@@ -81,3 +81,31 @@ post '/service/:service_id/container/:container_id' do |service_id, container_id
   end
 
 end
+
+put '/service/:service_id/stack/:stack_id' do |service_id, stack_id|
+  stack = ::AutoScaling::Stack.get(stack_id)
+  logger.debug("Scaling up stack #{stack.id} (service: #{service_id})")
+
+  begin
+    service_executor.deploy_container stack
+    status 200
+  rescue RuntimeError => e
+    logger.error e
+    status 503
+  end
+
+end
+
+delete '/service/:service_id/stack/:stack_id' do |service_id, stack_id|
+  stack = ::AutoScaling::Stack.get(stack_id)
+  logger.debug("Scaling down stack #{stack.id} (service: #{service_id})")
+
+  begin
+    service_executor.delete_container stack
+    status 200
+  rescue RuntimeError => e
+    logger.error e
+    status 503
+  end
+
+end

### 2013-07-22 auto-scaling/cloud-provider/src/opennebula/appstage_client.rb
[#18] fixed scaling up - passing SERVICE_ID during appstage invocation

@@ -18,9 +18,9 @@ module AutoScaling
       @options = options
     end
 
-    def instantiate_container(appstage_id, template_id)
+    def instantiate_container(appstage_id, template_id, service_id)
       Net::SSH.start( host, @options[:username], :password => @options[:password] ) do|ssh|
-        output = ssh.exec!("appstage instantiate -t #{template_id} #{appstage_id}")
+        output = ssh.exec!("appstage instantiate -t #{template_id} #{appstage_id} -d SERVICE_ID=#{service_id},VM_ID=$VMID")
         # output is in a format: "VM ID: 168"
         id = output.split(' ')[2]
 

### 2013-07-22 auto-scaling/cloud-provider/src/opennebula/opennebula_client.rb

@@ -35,8 +35,8 @@ module AutoScaling
       @appflow.configuration service_id
     end
 
-    def instantiate_container(appstage_id, template_id)
-      @appstage.instantiate_container appstage_id, template_id
+    def instantiate_container(appstage_id, template_id, service_id)
+      @appstage.instantiate_container appstage_id, template_id, service_id
     end
 
     def delete_container(container_id)

### 2013-07-22 auto-scaling/cloud-provider/test/integration/opennebula/opennebula_client_test.rb

@@ -62,8 +62,9 @@ eos
     def test_shall_instantiate_container
       appstage_id = 25
       template_id = 7
+      service_id = 1
 
-      container_info = @opennebula_client.instantiate_container(appstage_id, template_id)
+      container_info = @opennebula_client.instantiate_container(appstage_id, template_id, service_id)
       assert_equal true, container_info[:id] >= 0
       assert_equal true, container_info[:ip] != ''
 

### 2013-07-22 auto-scaling/service-controller/src/executor/service_executor.rb

@@ -57,7 +57,7 @@ module AutoScaling
       appstage_id = mappings[:appstage][stack.type.to_sym][:slave]
       template_id = mappings[:onetemplate_id]
 
-      container_info = @cloud_provider.instantiate_container(appstage_id, template_id)
+      container_info = @cloud_provider.instantiate_container(appstage_id, template_id, stack.service.id)
 
       # persist data
       container = Container.create(
@@ -78,6 +78,7 @@ module AutoScaling
 
       slave = slaves.pop
       @cloud_provider.delete_container slave.id
+      slave.destroy
       stack.save
 
       # reconfigure master

### 2013-07-22 auto-scaling/service-controller/test/unit/executor/service_executor_test.rb

@@ -167,7 +167,7 @@ module AutoScaling
       container_id = 800
       container_ip = '192.168.122.69'
       response = 'Od przedszkola do Opola, kupuj szybko dzis bejzbola'
-      @cloud_provider.expects(:instantiate_container).with(25, 7).returns({:id => container_id, :ip=> container_ip})
+      @cloud_provider.expects(:instantiate_container).with(25, 7, instance_id).returns({:id => container_id, :ip=> container_ip})
       FakeWeb.register_uri(:post,
                            "http://#{master_ip}:4567/chef",
                            :body => response,
@@ -215,7 +215,7 @@ module AutoScaling
       @cloud_provider.expects(:delete_container).with(10)
 
       @executor.delete_container(Stack.get(stack.id))
-      assert_equal 1, Container.slaves(stack).size
+      assert_equal 0, Container.slaves(stack).size
     end
 
     def test_shall_throw_exception_when_there_is_no_slaves_left

### 2013-07-23 thesis/src/abstract.tex
Add draft of the abstract

\vspace*{0.5cm}

\begin{center}
  \LARGE Abstract
\end{center}

Cloud computing has become an attractive model for provisioning on demand
computing resources as services to end-users. It is based on the assumption
that almost anything can be viewed as a service, starting from applications
delivered over Internet, through hardware in the data centers and ending on
computing power. The model appears to be so attractive as from the user point
of view the offered resources are infinite, transparent, robust and ready to
consume at any time. What is more, at present the ease and elasticity at
which the services are available is on an unprecedented scale, allowing both
individuals or organizations and companies, regardless of their needs, set up
and deploy the most sophisticated environments and infrastructures in a
matter of minutes or hours. Additionally, most of the times end-users do not
know in advance what the demand for the service is. This creates a
requirement in which their systems are auto-scalable, i.e. they support
sudden spikes in demand followed by underutilization at other times.

In order to ensure these growing and demanding aspects of Quality of Service
it is a common case when cloud providers cooperate with one-another and
mutually share their resources.

Therefore it is absolutely vital an architecture that enables seamless
cooperation among cloud providers and takes into account various QoS
requirements of end-users be developed.

The InterCloud architecture is one of the first attempts that had been made
in this direction. However, until now there are no known implementations of
it and all the tests were executed on a simulator.

The main goal of this thesis is to propose and implement our own architecture
which is based of the aforementioned work. Additionally, we want to use quite
new virtualization technology, linux containers, as a foundation of
provisioning resources of data centers.

The thesis is organized as follows \ldots

### 2013-07-27 thesis/src/abstract.tex
Abstract v0.2

@@
 
-Cloud computing has become an attractive model for provisioning on demand
-computing resources as services to end-users. It is based on the assumption
-that almost anything can be viewed as a service, starting from applications
-delivered over Internet, through hardware in the data centers and ending on
-computing power. The model appears to be so attractive as from the user point
-of view the offered resources are infinite, transparent, robust and ready to
-consume at any time. What is more, at present the ease and elasticity at
-which the services are available is on an unprecedented scale, allowing both
-individuals or organizations and companies, regardless of their needs, set up
-and deploy the most sophisticated environments and infrastructures in a
-matter of minutes or hours. Additionally, most of the times end-users do not
-know in advance what the demand for the service is. This creates a
-requirement in which their systems are auto-scalable, i.e. they support
-sudden spikes in demand followed by underutilization at other times.
+Cloud computing has become an attractive model for provisioning on demand computing resources as services to end-users.
+It is based on the assumption that almost anything can be viewed as a service, starting from applications delivered over Internet, through hardware in the data centers and ending on computing power.
+The model appears to be so attractive as from the user point of view the offered resources are infinite, transparent, robust and ready to consume at any time.
+The model appears to be so attractive as from the user point of view the offered resources are infinite, transparent, robust and ready to consume at any time.
 
-In order to ensure these growing and demanding aspects of Quality of Service
-it is a common case when cloud providers cooperate with one-another and
-mutually share their resources.
+Most of the times end-users do not know in advance what the demand for the service is.
+This creates a requirement in which their systems are auto-scalable, i.e.
+they support sudden spikes in demand followed by underutilization at other times (wtf?).
+An architecture of a cloud computing system, that meets these growing and demanding requirements of Quality of Service, has a characteristic of a multi-hierarchical autonomous system, where different layers corresponds to different levels where cloud operates: starting from application layer, through the hypervisor, service to end on cloud instance.
 
-Therefore it is absolutely vital an architecture that enables seamless
-cooperation among cloud providers and takes into account various QoS
-requirements of end-users be developed.
+Problem complexity raises challenges in a variety of aspects, especially in terms of providing cooperation and mutual sharing of resources, that may belong to different kind of cloud providers.
+Therefore, an architecture that enables seamless cooperation among cloud providers and takes into account various QoS requirements of end-users be developed is absolutely vital.
+The InterCloud architecture is one of the first attempts that had been made in this direction.
 
-The InterCloud architecture is one of the first attempts that had been made
-in this direction. However, until now there are no known implementations of
-it and all the tests were executed on a simulator.
-
-The main goal of this thesis is to propose and implement our own architecture
-which is based of the aforementioned work. Additionally, we want to use quite
-new virtualization technology, linux containers, as a foundation of
-provisioning resources of data centers.
-
-The thesis is organized as follows \ldots
+The main contributions of this thesis are as follows: a) proposition of an architecture that enables aforementioned scenarios, b) implementation of that architecture c) simulation and laboratory tests.

### 2013-07-27 auto-scaling/service-controller/src/monitor/service_monitor.rb

@@ -1,0 +1,3 @@
+class ServiceMonitor
+  # To change this template use File | Settings | File Templates.
+end

### 2013-07-28 intercloud/cloud-broker/lib/cloud_broker.rb
[#26] Initial commit of a cloud-broker

@@ -1,0 +1,18 @@
+require 'rubygems'
+require 'bundler/setup'
+
+module Intercloud
+  class CloudBroker
+
+    def initialize(settings)
+
+    end
+
+    def deploy(service_spec, client_endpoint)
+    end
+
+    def valid?(service_spec)
+      true
+    end
+  end
+end

### 2013-07-28 intercloud/cloud-broker/spec/cloud_broker_client_endpoint_spec.rb

@@ -1,0 +1,50 @@
+require 'app'
+require 'rack/test'
+require 'json'
+
+module Intercloud
+
+  describe "CloudBroker-ClientEndpoint" do
+    include Rack::Test::Methods
+
+      class CloudBrokerDouble
+        def initialize(database)
+
+        end
+        def valid?(request_body)
+          true
+        end
+        def deploy(request_body, client_endpoint)
+
+        end
+      end
+
+      def app
+        if @app
+          @app
+        else
+          CloudBrokerClientEndpoint.set(:cloud_broker, CloudBrokerDouble.new(nil))
+          CloudBrokerClientEndpoint.set(:environment, :test)
+          @app = CloudBrokerClientEndpoint
+        end
+      end
+
+      it "should accept requests with client endpoint and json payload" do
+        body = JSON.generate({ :service => 'name' })
+        post '/service', body , {'HTTP_IC_RETURN_ENDPOINT' => '192.168.21.153:8712', 'HTTP_ACCEPT' => "application/json"}
+        last_response.should be_ok
+      end
+
+      it "should reject requests without client endpoint or json payload" do
+        post '/service'
+        last_response.should_not be_ok
+
+        post '/environment', {}, {'HTTP_IC_RETURN_ENDPOINT' => '192.168.21.153:8712'}
+        last_response.should_not be_ok
+
+        post '/service', {}, {'HTTP_ACCEPT' => "application/json"}
+        last_response.should_not be_ok
+      end
+
+    end
+end

### 2013-07-28 intercloud/client/lib/command_line_parser.rb
[#26] Initial commit of a cloud-client

@@ -1,0 +1,30 @@
+require 'rubygems'
+require 'bundler/setup'
+
+require 'trollop'
+
+COMMANDS = %w(deploy)
+
+module Intercloud
+  class CommandLineParser
+    def self.parse
+      Trollop::options do
+        banner "Deploy an environment using `deploy`"
+        stop_on COMMANDS
+      end
+
+      action = ARGV.shift
+      arguments = nil
+      case action
+        when "deploy"
+          o = Trollop::options do
+            opt :environment, "Specification of an environment", :type => :string, :short => '-e'
+          end
+          arguments = o[:environment]
+        else
+          Trollop::die "Unknown command #{action.inspect}"
+      end
+      return  {:command => action.inspect.gsub('"', ''), :arguments => arguments }
+    end
+  end
+end

### 2013-07-28 intercloud/client/lib/http_sender.rb

@@ -1,0 +1,21 @@
+require 'ostruct'
+require 'net/http'
+
+module Intercloud
+  class HttpSender
+    def send(msg)
+      http_handler = Net::HTTP.new(msg.to['host'], msg.to['port'])
+      response = http_handler.request_post('/service', msg.body, msg.headers)
+    end
+
+    def prepare_deploy_message(data_source, body)
+      msg = OpenStruct.new
+      msg.to = data_source['cloud_broker']
+      msg.headers = {}
+      msg.headers['Accept'] = 'application/json'
+      msg.headers['IC_RETURN_ENDPOINT'] = data_source['endpoint']['host'] + ':' + data_source['endpoint']['port'].to_s
+      msg.body = body
+      msg
+    end
+  end
+end

### 2013-07-28 intercloud/client/spec/client_spec.rb

@@ -1,0 +1,25 @@
+require 'rspec'
+require 'client'
+
+module Intercloud
+
+  describe 'Intercloud client' do
+    before(:each) do
+      @db = double('storage', :set => 1, :get => 1)
+      @sender = double('sender', :send => 1, :prepare_deploy_message => 1)
+      @client = Client.new(@sender, @db)
+    end
+
+    it 'saves id of the deployed env' do
+      @db.should_receive(:set)
+      @client.deploy(nil)
+    end
+
+    it 'sends a deploy messsage' do
+      @sender.should_receive(:send)
+      @sender.should_receive(:prepare_deploy_message)
+      @client.deploy(nil)
+    end
+  end
+
+end

### 2013-07-28 auto-scaling/service-controller/test/unit/monitor/service_monitor_test.rb
Creted thesis structure

@@ -25,8 +25,8 @@ module AutoScaling
       set_2 = [["1374678040", "524288"], ["1374678083", "524288"], ["1374678113", "524288"], ["1374678155", "524288"], ["1374678198", "524288"], ["1374678241", "524288"], ["1374678284", "524288"], ["1374678327", "524288"], ["1374678370", "524288"], ["1374678413", "524288"]]
 
 
-      assert_equal set_1, @client.send(:last, set_1)
-      assert_equal set_2 - set_1, @client.send(:last, set_2)
+      assert_equal set_1, @client.send(:monitor_container, set_1)
+      assert_equal set_2, @client.send(:monitor_container, set_2)
     end
 
   end

### 2013-08-01 intercloud/client/app.rb
[#26] Add client endpoint

@@ -1,3 +1,5 @@
+#!/usr/bin/env ruby
+
 $LOAD_PATH << File.expand_path(File.dirname(__FILE__)) + '/lib'
 
 require 'rubygems'

### 2013-08-01 intercloud/client/bin/client_endpoint.rb

@@ -1,0 +1,30 @@
+#!/usr/bin/env ruby
+
+$LOAD_PATH << File.expand_path(File.dirname(__FILE__)) + '/..'
+
+require 'rubygems'
+require 'bundler/setup'
+require 'redis'
+require 'yaml'
+require 'pp'
+require 'sinatra/base'
+
+
+module Intercloud
+  class ClientEndpoint < Sinatra::Base
+
+    set :config, YAML.load_file('config/client_endpoint.yaml')
+    set :port, config['port']
+    set :environment, ENV['INTERCLOUD_ENV'] || ENV['RACK_ENV'] || :development
+    set :db, Redis.new(:host => config['redis']['host'], :port => config['redis']['port'])
+
+    post '/service_status' do
+      service_id = params[:service_id]
+      status     = params[:status]
+      db.set(service_id, status)
+      200
+    end
+
+    run! if app_file == $0
+  end
+end

### 2013-08-01 intercloud/client/start_endpoint.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+
+LOG_DIR=log/client_endpoint.log
+
+bin/client_endpoint.rb 2> $LOG_DIR &
+
+if [ $? -eq 0 ]; then
+  echo "Client rest service started or failed. Check logs that are located at $LOG_DIR"
+fi

### 2013-08-01 auto-scaling/bin/auto_scaling_server.rb
[#22] mapping conlusions to executor actions

@@ -58,7 +58,7 @@ post '/service' do
   logger.info("Got service creation request: #{service}")
 
   begin
-    service_planner.plan service
+    service_planner.plan_deployment service
   rescue RuntimeError => e
     logger.error e
     error 400

### 2013-08-01 auto-scaling/service-controller/src/analyzer/service_analyzer.rb

@@ -50,6 +50,7 @@ module AutoScaling
       end
 
       # filter out healthy conclusions if there were some issues
+      # TODO analyze here global problems
       conclusions.each do |stack, issues|
         issues.delete(:healthy) if issues.size > 1
       end

### 2013-08-01 auto-scaling/service-controller/src/analyzer/threshold_model.rb

@@ -9,8 +9,12 @@ module AutoScaling
       @max = max
     end
 
-    # list of values in form: [[timestamp, value], [timestamp, value], [timestamp, value]], ex:
-    # data = [["10", "5" , "8", "60"], ["10", "50" , "80"], ["11", "15" , "50"]]
+    # Analyzes data using threshold model
+    #
+    # * *Args* :
+    # - +data+ -> list of data, ex: ["10", "243", "432"]
+    #
+    # Return decision whether supplied data fits / is lesser / is greater than range [@min, @max]
     def analyze(data)
       # wrap data if it is not an array (ie single value, aggregated beforehand)
       data = [data] unless data.kind_of?(Array)

### 2013-08-01 auto-scaling/service-controller/src/planner/service_planner.rb

@@ -6,11 +6,13 @@ require 'executor/service_executor'
 module AutoScaling
   class ServicePlanner
 
+    @@logger = Logger.new(STDOUT)
+
     def initialize(executor)
       @executor = executor
     end
 
-    def plan(service)
+    def plan_deployment(service)
       # are there enough resources?
 
       # reserve resources
@@ -18,5 +20,44 @@ module AutoScaling
       # deploy
       @executor.deploy_service service
     end
+
+    # Analyzes data using supplied model
+    #
+    # * *Args* :
+    # - +data+ -> hashmap that contains analysis conclusions
+    # {
+    #   stack => [:insufficient_slaves, :overloaded_master],
+    #   stack => [:healthy]
+    # }
+    #
+    def plan(data)
+      data.each do |stack, conclusions|
+        conclusions.each do |conclusion|
+          self.send(conclusion, stack)
+        end
+      end
+    end
+
+    private
+    def insufficient_slaves(stack)
+      # TODO reserve resources
+
+      @executor.deploy_container(stack)
+    end
+
+    def redundant(stack)
+      # TODO free resources
+
+      @executor.delete_container(stack)
+    end
+
+    def overloaded_master(stack)
+      @@logger.warn "Stack #{stack} has overloaded master. Currently, nothing can be done"
+    end
+
+    def healthy(stack)
+      @@logger.info "Stack #{stack} is healthy. Ignoring"
+    end
+
   end
 end

### 2013-08-01 auto-scaling/service-controller/test/unit/planner/service_planner_test.rb

@@ -19,7 +19,7 @@ module AutoScaling
 
     end
 
-    def test_plan
+    def test_shall_plan_service_deployment
       service = {
           'stack' => 'java',
           'instances' => 2,
@@ -27,8 +27,25 @@ module AutoScaling
       }
       @executor.expects(:deploy_service).with(service)
 
-      @planner.plan service
+      @planner.plan_deployment service
     end
+
+    def test_shall_plan_appropriate_executor_actions
+      stack_1 = mock()
+      stack_2 = mock()
+
+      data = {
+        stack_1 => [:insufficient_slaves, :overloaded_master],
+        mock() => [:healthy],
+        stack_2 => [:redundant]
+      }
+
+      @executor.expects(:deploy_container).with(stack_1)
+      @executor.expects(:delete_container).with(stack_2)
+
+      @planner.plan(data)
+    end
+
   end
 
 end

### 2013-08-01 auto-scaling/service-controller/service_controller.rb
[#22] scheduling job execution

@@ -2,8 +2,49 @@ $: << File.join(File.dirname(File.expand_path(__FILE__)), 'src')
 $: << File.join(File.dirname(File.expand_path(__FILE__)), 'lib')
 $: << File.join(File.dirname(File.expand_path(__FILE__)), '..')
 
+require 'rubygems'
+require 'logger'
+require 'rufus-scheduler'
+
 require 'planner/service_planner'
 require 'executor/service_executor'
 require 'monitor/service_monitor'
 require 'analyzer/service_analyzer'
 require 'models/models'
+
+module AutoScaling
+
+  class ServiceJob
+    @@logger = Logger.new(STDOUT)
+
+    def initialize(service, controller)
+      @service = service
+      @controller = controller
+    end
+
+    def call(job)
+      @@logger.debug "Executing job: #{job}"
+
+      monitoring_data = @controller.monitor.monitor(@service)
+      conclusions = @controller.analyzer.analyze(monitoring_data)
+      @controller.planner.plan(conclusions)
+    end
+  end
+
+  class ServiceController
+
+    attr_reader :monitor, :analyzer, :planner
+
+    def initialize(monitor, analyzer, planner, scheduler = nil)
+      @monitor, @analyzer, @planner = monitor, analyzer, planner
+      #scheduler ||= Rufus::Scheduler.new
+
+      @scheduler = scheduler
+    end
+
+    def schedule(job, interval)
+      @scheduler.every interval, job
+    end
+
+  end
+end

### 2013-08-01 auto-scaling/service-controller/test/unit/service_controller_test.rb

@@ -1,0 +1,37 @@
+require 'rubygems'
+require 'test/unit'
+require 'mocha/setup'
+
+require 'utils'
+require 'service_controller'
+
+module AutoScaling
+
+  class ServiceControllerTest < Test::Unit::TestCase
+
+    def setup
+      @monitor, @analyzer, @planner = mock(), mock(), mock()
+      @scheduler = Rufus::Scheduler.new
+
+      @controller = ServiceController.new @monitor, @analyzer, @planner, @scheduler
+    end
+
+    def test_shall_perform_full_cycle
+      service = mock()
+      job = ServiceJob.new service, @controller
+
+      monitoring_data, conclusions = mock(), mock()
+
+      @monitor.expects(:monitor).with(service).returns(monitoring_data)
+      @analyzer.expects(:analyze).with(monitoring_data).returns(conclusions)
+      @planner.expects(:plan).with(conclusions)
+
+      scheduled_job = @controller.schedule job, '1s'
+
+      # give some time to finish job and then pause scheduler
+      sleep 2; scheduled_job.pause
+    end
+
+  end
+
+end

### 2013-08-01 thesis/src/abstract.tex
[#29] abstract v0.3

@@
 This creates a requirement in which their systems are auto-scalable, i.e.
-they support sudden spikes in demand followed by underutilization at other times (wtf?).
-An architecture of a cloud computing system, that meets these growing and demanding requirements of Quality of Service, has a characteristic of a multi-hierarchical autonomous system, where different layers corresponds to different levels where cloud operates: starting from application layer, through the hypervisor, service to end on cloud instance.
+they support sudden spikes in demand followed by underutilization at other times.
+An architecture of a cloud computing system, that meets these requirements, has a characteristic of a multi-hierarchical autonomous system, where different layers corresponds to different levels where cloud operates: starting from an application layer, through the hypervisor, service to end on cloud instance.
 
@@
 The InterCloud architecture is one of the first attempts that had been made in this direction.
+Having characteristic of an application platform in mind, we propose a variation of that architecture that supports cooperation among these application platform and fulfills need decentralized environment.
 

### 2013-08-27 thesis/src/abstract.tex
Typo

@@
 
-Problem complexity raises challenges in a variety of aspects, especially in terms of providing cooperation and mutual sharing of resources, that may belong to different kind of cloud providers.
+Problem complexity raises challenges in a variety of aspects, especially in terms of providing cooperation and mutual sharing of resources that may belong to different kind of cloud providers.
 Therefore, an architecture that enables seamless cooperation among cloud providers and takes into account various QoS requirements of end-users be developed is absolutely vital.
 The InterCloud architecture is one of the first attempts that had been made in this direction.
-Having characteristic of an application platform in mind, we propose a variation of that architecture that supports cooperation among these application platform and fulfills need decentralized environment.
+Having characteristic of an application platform in mind, we propose a variation of that architecture that supports cooperation among these application platform and fulfils needs of a decentralized environment.
 

### 2013-08-27 thesis/src/chapters/chapter-1.tex
Creted thesis structure
structure
Add the first version of 'main goals' & 'impact'

@@
-\chapter{Draft chapter}
-\section{Draft section}
+\chapter{Introduction}
 
-section contents
+\section{Motivation}
+
+\section{Contributions}
+The main contributions of this dissertation are as follows:
+\begin{itemize}
+  \item A proposal of an architecture of federated cloud computing environment, which is based on and can be viewed as a simplified version of \emph{InterCloud}
+  \item The notion of considering each service model as an autonomic system
+  \item The implementation of the proposed architecture using OpenNebula technology stack
+\end{itemize}
+
+\section{Impact}
+We hope that the ideas contained in this work be beneficial to the OpenNebula ecosystem as they provide insights into the ways the following features can be implemented:
+\begin{itemize}
+  \item autoscaling capabilities
+  \item virtual machines provisioning
+  \item managing of a group of virtual machines which forms a logical entity
+\end{itemize}
+What is more, we believe that our successful attempt to implement a simplified variant of an InterCloud architecture will cause its gain in interest and popularity.
+And finally, the idea of representing each level of an autoscaling subsystem as an autonomic one can be thought-provoking for cloud computing scientists.
+\section{Thesis structure}

### 2013-08-27 thesis/src/chapters/chapter-1.tex
Motivation draft

@@
 \section{Motivation}
+One of the keys factors that has driven transformation of computing industry in the last years is the perception of computing utilities as a ordinary property, which can be easily accessed and adjusted to a specific needs.
+That point of view resulted in profusion of different services, often collectively referred as a cloud computing.
+Similarly to services common to other markets, customers expect them to be accessible on demand and in easy manner, while paying only for the consumed goods.
+Furthermore, customers are interested in a particular service only when it is eligible to provide appropriate quality of service.
+
+Having that requirements in mind, it is crucial that system which provides services to user is able to auto-adapt itself to meet customer requirements.
+Due to the complexity of system under consideration, there are different levels where such adaptation is possible:
+\begin{itemize}
+	\item infrastructure
+	\item application platform
+	\item user application
+\end{itemize} 
+However, the fact that single service provider is constrained by his finite amount of resources poses a risk that it may not be able to serve customer all the time.
+Consequently, it is expected that adaptation at a service provider level is also possible, i.e.
+provider can offload some traffic to a different provider, as long as it satisfies a customer.
+
+While architectures that enables auto-adaptivity have a long history, none of them has been directly applied to a multi-layered problem that exists in a cloud computing environment.
+Especially, the research area at the last level of adaptivity, that sizes across different service providers, is new.
+Although, architecture known as InterCloud investigates problem of cooperation and negotiation at cloud level, it neither has been implemented nor presented in context of layered adaptive system.
 

### 2013-08-27 thesis/src/chapters/chapter-1.tex
Motiviation draft2

--- 2013-08-27 thesis/src/chapters/chapter-1.tex
@@
 \section{Motivation}
+One of the keys factors that has driven transformation of computing industry in the last years is the perception of computing utilities as a ordinary property, which can be easily accessed and adjusted to a specific needs.
+That point of view resulted in profusion of different services, often collectively referred as a cloud computing.
+Similarly to services common to traditional markets, customers expect them to be accessible on demand and in easy manner, while paying only for the consumed goods.
+Furthermore, customers are interested in a given service provider only when it is eligible to guarantee appropriate quality of service.
+
+The particular service providers that are addressed by this paper are the ones that supply users with an application execution platform, what is widely known as providing Platform-as-a-Service.
+In that case, a customer is an entity that has developed application and is eager to deploy it on an application platform that is able to fulfil his specific requirements, both in terms of quality and cost.
+
+Having customer requirements in mind, it is crucial that service provider is able to adapt itself to meet them.
+For example, such adaptation can be triggered by a sudden spike in resource demand and may result in provisioning additional application platforms.
+However, due to the complexity of a system under consideration, there are different levels where adaptation is possible:
+\begin{itemize}
+	\item user application
+	\item application platform
+	\item infrastructure
+\end{itemize} 
+What is more, the fact that single service provider is constrained by his finite amount of resources poses a risk that it may not be able to serve customer all the time.
+Consequently, it is expected that adaptation at a service provider level is also possible, i.e.
+provider can offload some traffic to a different provider, as long as it satisfies a customer.
+
+While autonomic computing has a long history, it has not been directly applied to a multi-layered problem that exists in a cloud computing environment.
+Especially, the research area at the last layer, which sizes across different service providers, is new.
+Although, architecture known as InterCloud investigates problem of cooperation and negotiation at cloud level, it neither has been implemented nor presented in context of autonomic system.
+
+\textit{business potential - dobry hajs, wymagania do QoS, lista produktow i ze nie maja pierdolniecia}
 

### 2013-08-27 thesis/src/chapters/chapter-1.tex
Thesis impact v.0.2

@@
 \section{Motivation}
-One of the keys factors that has driven transformation of computing industry in the last years is the perception of computing utilities as a ordinary property, which can be easily accessed and adjusted to a specific needs.
+One of the keys factors that has driven transformation of computing industry in the last years is the perception of computing utilities as an ordinary property, which can be easily accessed and adjusted to a specific needs.
 That point of view resulted in profusion of different services, often collectively referred as a cloud computing.
@@
 \section{Impact}
-We hope that the ideas contained in this work be beneficial to the OpenNebula ecosystem as they provide insights into the ways the following features can be implemented:
+We hope that the concept of representing each level of an autoscaling subsystem as an autonomic one can be thought-provoking for cloud computing scientists.
+What is more, we believe that our successful attempt to implement a simplified variant of an InterCloud architecture will cause its gain in interest and popularity.
+Finally, we consider the ideas contained in this work be beneficial to the OpenNebula ecosystem as they provide insights into the ways Quality of Service can be ensured:
 \begin{itemize}
-  \item autoscaling capabilities
-  \item virtual machines provisioning
-  \item managing of a group of virtual machines which forms a logical entity
+  \item implementing autoscaling capabilities
+  \item designing \emph{cloud infrastructure} in accordance with InterCloud architecture
 \end{itemize}
@@
 And finally, the idea of representing each level of an autoscaling subsystem as an autonomic one can be thought-provoking for cloud computing scientists.
+
 \section{Thesis structure}

### 2013-08-27 thesis/src/abstract.tex
Few fixes in an abstract

@@
 It is based on the assumption that almost anything can be viewed as a service, starting from applications delivered over Internet, through hardware in the data centers and ending on computing power.
-The model appears to be so attractive as from the user point of view the offered resources are infinite, transparent, robust and ready to consume at any time.
-The model appears to be so attractive as from the user point of view the offered resources are infinite, transparent, robust and ready to consume at any time.
+That model appears to be so attractive as from the user point of view the offered resources are infinite, transparent, robust and ready to consume at any time.
 
@@
 they support sudden spikes in demand followed by underutilization at other times.
-An architecture of a cloud computing system, that meets these requirements, has a characteristic of a multi-hierarchical autonomous system, where different layers corresponds to different levels where cloud operates: starting from an application layer, through the hypervisor, service to end on cloud instance.
+An architecture of a cloud computing system, that meets these requirements, has a characteristic of a multi-hierarchical autonomic system, where different layers corresponds to different levels where cloud operates: starting from an application layer, through the application platform, infrastructure to end on a cloud instance.
 

### 2013-08-27 thesis/src/chapters/chapter-2.tex
Few fixes in an abstract

\chapter{Scaling applications}

\section{Introduction}

### 2013-08-27 thesis/assets/figures-data/generate-pdf-figure.sh
Add a script that helps generating pdf figures

@@ -1,0 +1,44 @@
+#!/bin/bash
+
+if [ $# -ne 1 ]; then
+  echo "Wrong number of arguments! The only argument should be  " \
+    "the name of the gnuplot input file  (with .gnu extension). " \
+    "Output must have the same name as the input file."
+  exit 1
+fi
+
+DIRECTORY=`echo "$1" | sed 's/\(.*\)\/\(.*\)$/\1/'` 
+BARE_FILE=`echo "$1" | sed 's/\(.*\)\/\(.*\)$/\2/'` 
+CURRENT_DIR=`pwd`
+GNUPLOT_FILE=$BARE_FILE.gnu
+EPS_FILE=$BARE_FILE-inc.eps
+TEX_FILE=$BARE_FILE.tex
+
+function cleanup {
+  rm -rf *.log *.aux *.eps *-inc.pdf *.tex
+}
+
+cd $DIRECTORY
+
+if [ ! -f $GNUPLOT_FILE ]; then 
+  echo "File $GNUPLOT_FILE not found!" 
+  cd $CURRENT_DIR
+  exit 1
+fi
+
+# this command should create a 'eps' file
+gnuplot $GNUPLOT_FILE
+
+if [ ! -f $EPS_FILE ]; then 
+  echo "File $EPS_FILE not found!"
+  cleanup
+  cd $CURRENT_DIR
+  exit 1
+fi
+
+epstopdf $EPS_FILE
+pdflatex $TEX_FILE
+
+cleanup
+cd $CURRENT_DIR
+

### 2013-08-27 thesis/src/chapters/chapter-1.tex
Add business subsection to the motivation

--- 2013-08-27 thesis/src/chapters/chapter-1.tex
@@
 
+\subsection{Business}
+The rapid growth of interest in cloud computing in recent years resulted in huge sums of money being invested in the field.
+Figure \ref{ch1:public-cloud-services-market-size} shows the size of the public cloud services market in 2012 and the forecast of its nearly two times growth in 2016.
+This data suggests that the subject is attractive for IT industry from the economic point of view.
+However, higher amounts of money spent on cloud services involves higher expectation of theirs quality.
+
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics{chapter-1/public-cloud-services-market-size}
+  \end{center}
+  \caption{Public Cloud Services Market Size, 2010-2016 (forecast)}
+  \label{ch1:public-cloud-services-market-size}
+\end{figure}
+
 \textit{business potential - dobry hajs, wymagania do QoS, lista produktow i ze nie maja pierdolniecia}

### 2013-08-27 thesis/assets/figures-data/generate-pdf-figure.sh

@@ -9,6 +9,7 @@ fi
 
 DIRECTORY=`echo "$1" | sed 's/\(.*\)\/\(.*\)$/\1/'` 
 BARE_FILE=`echo "$1" | sed 's/\(.*\)\/\(.*\)$/\2/'` 
+DESTINATION_DIR=../figures/$DIRECTORY
 CURRENT_DIR=`pwd`
 GNUPLOT_FILE=$BARE_FILE.gnu
 EPS_FILE=$BARE_FILE-inc.eps
@@ -41,4 +42,6 @@ pdflatex $TEX_FILE
 
 cleanup
 cd $CURRENT_DIR
+
+mv $1.pdf $DESTINATION_DIR
 

### 2013-08-28 thesis/src/chapters/chapter-1.tex
Remove duplicated txt and add figure source

@@
   \end{center}
-  \caption{Public Cloud Services Market Size, 2010-2016 (forecast)}
+  \caption{Public Cloud Services Market Size, 2010-2016 (forecast). Source: \textit{Gartner, 08/2012}}
   \label{ch1:public-cloud-services-market-size}
@@
 \end{itemize}
-What is more, we believe that our successful attempt to implement a simplified variant of an InterCloud architecture will cause its gain in interest and popularity.
-And finally, the idea of representing each level of an autoscaling subsystem as an autonomic one can be thought-provoking for cloud computing scientists.
 

### 2013-08-28 thesis/src/chapters/chapter-1.tex
Add description of 3 another PaaS providers

@@
 This data suggests that the subject is attractive for IT industry from the economic point of view.
-However, higher amounts of money spent on cloud services involves higher expectation of theirs quality.
+However, higher amounts of money spent on cloud services involve higher expectations of theirs quality from customers.
+Although the most significant players in cloud computing have been in the field for quite a long time, it is still possible to outline some deficiencies their products have.
+Additionally, lack of common standards hinder cooperation among different cloud providers.
+For example it is nearly impossible to create an autoscaling cloud federation with Amazon Web Services (current leader in providing cloud services\footnote{Gartner: Magic Quadrant for Cloud Infrastructure as a Service (Aug 2013)}) and another provider.
+What is more, Amazon AWS users cannot use more lightweight virtualization methods, such as linux containers.
+Nevertheless, when compared to other companies especially in terms of autoscaling capabilities, Amazon really shines.
+OpenShift, RedHat PaaS solution, ensures application scaling but with very limited possibilities of customisation of the process -- the user can only choose if their application should scale and the whole algorithm is solely based on the number of concurrent requests to the application.
+Users of Heroku, another PaaS solution, have no automation tool that would control the number of instances (\emph{dynos} in Heroku nomenclature) their application is running on -- they can change it manually.
+
+The proposed solution in this dissertation tries to deal with the aforementioned providers problems by outlining an example architecture that enables seamless cooperation among cloud providers and provides auto-scaling capabilities.
 
@@
 \end{figure}
-
-\textit{business potential - dobry hajs, wymagania do QoS, lista produktow i ze nie maja pierdolniecia}
 

### 2013-08-28 thesis/src/chapters/chapter-1.tex
Bibtex fix
Change footnote for a bib reference

@@
 One of the keys factors that has driven transformation of computing industry in the last years is the perception of computing utilities as an ordinary property, which can be easily accessed and adjusted to a specific needs.
-That point of view resulted in profusion of different services, often collectively referred as a cloud computing.
+That point of view resulted in profusion of different services, often collectively referred as a cloud computing \cite{MeGr11}.
 Similarly to services common to traditional markets, customers expect them to be accessible on demand and in easy manner, while paying only for the consumed goods.
@@
 Additionally, lack of common standards hinder cooperation among different cloud providers.
-For example it is nearly impossible to create an autoscaling cloud federation with Amazon Web Services (current leader in providing cloud services\footnote{Gartner: Magic Quadrant for Cloud Infrastructure as a Service (Aug 2013)}) and another provider.
+For example it is nearly impossible to create an autoscaling cloud federation with Amazon Web Services (current leader in providing cloud services\cite{GartnerMagicQuadrantSep2013}) and another provider.
 What is more, Amazon AWS users cannot use more lightweight virtualization methods, such as linux containers.

### 2013-08-28 thesis/src/chapters/chapter-2.tex
Bibtex fix
Added new chapter

@@
 \section{Introduction}
+
+The reason why this dissertation treats about scaling application is the fact that it is widely agreed measure for improving application performance, consequently increasing offered Quality-of-Service.
+Enriching system with capability to scale entails avoiding additional costs that are related to coping with excessive traffic.
+In some cases, these costs may be caused by not handling extra traffic at all or may involve aspects such as: increased response time, processing overhead, space, memory, or money \cite{Bo00}.
+
+While scalability is a widely used term, it still lacks a clear and concise definition.
+Over the time, there were a few attempts to define it, yet not all of them were claimed as successful \cite{Hi90} \cite{DuRoWi06}.
+Hence, it is necessary to clarify this term before going into further discussion.
+Instinctively, scalability is perceived as ability of a system to accommodate an increasing number of elements or objects to process.
+In particular, we can define different aspects of scalability \cite{Bo00}:
+\begin{itemize}
+	\item Load scalability,
+	\item Space scalability,
+	\item Space-time scalability,
+	\item Structural scalability,
+\end{itemize}
+
+With all that said, there is no silver bullet - not matter what underlying mechanism provider decides to use, the application developer is still responsible for creating an application with scaling in-mind.
+The idea behind this statement has been already proven in 1967 by Amdahl law, which in short states that sequential component of parallel algorithms impacts efficiency for a sufficiently large number processors \cite{Am67}.

### 2013-08-29 thesis/src/chapters/chapter-2.tex
Added new chapter
Added introduction to chapter-2

--- 2013-08-27 thesis/src/chapters/chapter-2.tex
@@
 \section{Introduction}
+
+The reason why scaling application lies in our area of interest is the fact that it is widely accepted measure for improving application performance, consequently increasing offered Quality-of-Service.
+Enriching system with capability to scale entails avoiding additional costs that are related to coping with excessive traffic.
+In some cases, these costs may be caused by not handling extra traffic at all or may involve aspects such as: increased response time, processing overhead, space, memory, or money \cite{Bo00}.
+
+While scalability is a widely used term, it still lacks a clear and concise definition.
+Over the time, there were a few attempts to define it, yet not all of them were claimed as successful \cite{Hi90} \cite{DuRoWi06}.
+Hence, it is necessary to clarify this term before going into further discussion.
+Instinctively, scalability is perceived as ability of a system to accommodate an increasing number of elements or objects to process.
+In particular, we can point out different types of scalability that are affected by increased number of requests: \cite{Bo00}:
+\begin{itemize}
+	\item \textit{load scalability} - ability to work without delays and unproductive resource consumption at light, moderate, or heavy loads while making good use of available resources.
+Factors that may hinder load scalability include aspects such as: scheduling shared resource, self-expansion, inadequate exploitation of parallelism
+	\item \textit{space scalability} - memory requirements do not grow to intolerable levels as the number of items it supports increases
+	\item \textit{space-time scalability} - system continues to function gracefully as the number of objects it encompasses increases by orders of magnitude
+	\item \textit{structural scalability} - implementation or standards do not impede the growth of the number of objects it encompasses
+\end{itemize}
+Although, all of the aforementioned aspects are crucial for functioning of any healthy application, our work focuses only on the first type of scalability.
+The reasoning behind this statement is that while all of these scalability types lies in direct responsibility of an application developer, the load scalability can be additionally improved by adding additional resources to a system.
+This brings us to a question what kind of resources are used by an application or more appropriately to context of this dissertation: \textit{what kind of resources can we add to improve application performance?} Required resources varies from an application to an application.
+However, among the most common ones we can distinguish:
+\begin{itemize}
+	\item CPU
+	\item memory
+	\item storage
+	\item network bandwidth
+\end{itemize}
+
+It is commonly agreed that there are two main possible ways the resource can be added to a server:
+\begin{itemize}
+	\item \textit{horizontal scaling (scaling out)} - adding more nodes to a system, such as servers in context of distributed application
+	\item \textit{vertical scaling (scaling up)} - increasing capacity of a single node in a system, i.e. adding additional memory, CPU, storage, etc.
+\end{itemize}
+
+Having that said, common sense dictates that adding resources is only a part of the success - it should be accompanied by tuning application platform configuration.
+For example, adding supplementary CPUs without increasing thread pools that handle request make a little sense.
+Similarly, we have to increase heap size, in context of a Java application, to make a good use of extra memory.
+While importance of application tuning cannot be underestimated, its detailed analysis lies outside of the scope of this dissertation.
+
+With all that said, there is no silver bullet - not matter what underlying mechanism platform provider decides to use, the application developer is still responsible for creating an application with scaling in-mind.
+This statement has been already proven in 1967 by Amdahl law, which in short states that sequential component of parallel algorithms impacts efficiency for a sufficiently large number processors \cite{Am67}.
+In other words, adding supplementary resources to a poorly written application (i.e.
+having a lot of sequential or synchronized components) can be beneficial only to a certain degree.
+
+The rest of this chapter elaborates in details about horizontal, vertical scaling and application tuning taking into account mechanisms used in Platform-as-a-Service solutions that are available on the market.
+
+\section{Horizontal scaling}
+
+\section{Vertical scaling}

### 2013-08-29 thesis/src/chapters/chapter-2.tex
Added chapter-5 draft

@@
 
-It is commonly agreed that there are two main possible ways the resource can be added to a server:
+It is commonly agreed that there are two main possible ways the resource can be added:
 \begin{itemize}
@@
 Having that said, common sense dictates that adding resources is only a part of the success - it should be accompanied by tuning application platform configuration.
-For example, adding supplementary CPUs without increasing thread pools that handle request make a little sense.
+For example, adding supplementary CPUs without increasing thread pool size that handle requests make a little sense.
 Similarly, we have to increase heap size, in context of a Java application, to make a good use of extra memory.
 While importance of application tuning cannot be underestimated, its detailed analysis lies outside of the scope of this dissertation.
+
+What makes scaling application particularly interesting are the benefits offered by a cloud computing, especially the illusion of a virtually infinite computing infrastructure \cite{VaRoBu11}.
+Making use of virtualization technologies, which often underpins cloud computing platform, allows for resource manipulation in a dynamic, on-demand manner.
 
@@
 
-The rest of this chapter elaborates in details about horizontal, vertical scaling and application tuning taking into account mechanisms used in Platform-as-a-Service solutions that are available on the market.
+The rest of this chapter elaborates in details about horizontal and vertical scaling taking into account mechanisms used in Platform-as-a-Service solutions that are available on the market.
 
@@
 
+As it was outlined in a previous section, vital element of horizontal scaling is adding additional resources.
+In context of an application execution environment, node can be represented by an application server, database or other resources that affect application performance such as network interface \cite{VaRoBu11}.
+Dynamic capabilities of a cloud computing platform offer a couple of ways this can be done:
+\begin{itemize}
+	\item virtual machine replication
+	\item network slicing
+	\item database replication
+\end{itemize}
+
+jebnac mapowanie na warstwy clouda - infr, platform, app
+
+\subsection{Virtual machine replication}
+
+\subsection{Network slicing}
+
+\subsection{Database replication}
+
 \section{Vertical scaling}

### 2013-08-29 thesis/src/chapters/chapter-5.tex
Added chapter-5 draft

\chapter{Solution}

\section{Introduction}

Taking requirements into consideration, desired platform clearly has features common to a autonomic system, especially the need for automatic environment reconfiguration. One of the first architectures that satisfied that requirement was proposed by \cite{IBM06}. Beside this, system under consideration has a character of a multi-layered one (i.e. it operates on a application / container / service / cloud level). Optimally, each level should be characterized by different actions, what can be meet by implementing multi-hierarchical autonomic system, as proposed in \cite{LiWoZh05}.

Considering that observation, auto-scaling subsystem spans across different layers:

\begin{figure}[!ht]
  \begin{center}
    \includegraphics[width=\textwidth]{chapter-5/hierarchical-autonomic-system}
  \end{center}
  \caption{Hierarchical autonomic system}
  \label{ch5:hierarchical-autonomic-system}
\end{figure}

Remaining of this section breaks apart each functional module of an autonomic component, whereas further sections cover in details each autonomic level.

\subsection{Autonomic component}

\begin{figure}[!ht]
  \begin{center}
    \includegraphics[width=0.8\textwidth]{chapter-5/autnomic-component}
  \end{center}
  \caption{Autonomic component}
  \label{ch5:autonomic-component}
\end{figure}

\subsubsection{Monitor}

Monitor module is responsible for gathering monitoring data from \textit{Sensors} and store them in a database, correlated with a specific autonomous component (ie. application / container / service / cloud). Problem of a multi-layered, hierarchical self-adaptive monitoring system is described in \cite{KaKoGoKyMeVa12}.

\subsubsection{Analyse}

Analyse module performs analysis of a gathered data. There is a variety of models that helps to achieve this goal. For example, the \cite{LiWoZh05} proposes:

\begin{itemize}
	\item Queue-based Performance Models
	\item Dynamic models
	\item Monotonic static models
	\item Policy based models
\end{itemize}

Details about which level leverage which model are presented in detailed architecture description. Important part in a data analysis plays a predictive approach \cite{JiPeLiCh11} and tracking changes in a system \cite{ZhYaWo05}.

\subsubsection{Plan}

This module is responsible for scheduling actions that have to be taken to solve problems reported by an earlier analysis. This may involve:
\begin{itemize}
	\item reserving some resources (for example to deploy a new virtual machine)
	\item handling situation where multiple entities compete about the resources (for example by a prioritization)
\end{itemize}

It is possible, however, that resolving issues is not possible at current level. It that case, the upper-level is responsible for handling it.

\subsubsection{Execute}

Execute module is directly responsible for managing resource, by performing various operations on resource's effector. Each layer is characterized by different actions that can be taken, for example: application server reconfiguration, vertical or horizontal scaling.

\subsubsection{Knowledge base}

Knowledge base is represented as a set of rules, policies shared across autonomous components.

\subsection{Managed resource}

\subsubsection{Effector}

Enables management of a resources by exposing some API compatible with an Executor.

\subsubsection{Sensor}

Aim of a sensor is probing a resource and sending back results to a Monitor. There are different kind of metrics that should be considered such as: CPU load, memory usage or response delay.

\section{Application controller}

\section{Container controller}

\section{Service controller}

\section{Cloud controller}

### 2013-08-29 thesis/src/chapters/chapter-4.tex
Add the first part of 'hybrid cloud' section

\chapter{Interoperability of clouds}

\chapterintro{This chapter introduces the notion of a hybrid cloud and explains its role in cloud computing industry. On top of this deployment model, the concept of InterCloud is presented and elaborated with the emphasis on its application in ensuring scalability of users' services.}

\section{Introduction}
From the perspective of a user of PaaS services it is vital that they are able to deploy seamlessly their applications using libraries, tools and services supported by the cloud provider\cite{MeGr11}. Judging by such factors as the popularity of Heroku -- currently one of the most popular PaaS providers which does not offer more advanced features which would enable management of the infrastructure underpinning the deployment platform, the fact that Microsoft added auto-scaling to its Azure platform as late as in June 2013, it is perfectly possible most PaaS users are satisfied with the current offers of their providers and do need another, more sophisticated functionalities. However, there are more complex applications and systems whose requirements regarding technology stack, availability and scalability are considerably more demanding. For such services there ought to be designed slightly specialized features that would require cooperation among different cloud providers.

\section{Hybrid cloud}
One can imagine scenarios in which customers of cloud services know their applications are vulnerable to sudden variations in demand and their responsiveness must be kept at the same level all the time. In such cases, they want them to scale dynamically according to current load or other predefined or manually specified metrics. What is more, in order to ensure high availability of their services, customers do not want to confine themselves to only one provider -- in the best scenario they want their applications (or their logical parts, such as persistence layer) to be spanned across different providers and be able to cooperate with one another at the same time. This leads to the concept of a \emph{hybrid cloud}\cite{MeGr11} -- the case in which the cloud infrastructure is a composition of two or more distinct infrastructures which are unique entities, but there are technological means that make it possible to port data and applications among them.

\subsection{Usage in industry}

\section{Federation of clouds -- InterCloud}

\subsection{Usage in industry}

### 2013-08-29 thesis/src/chapters/chapter-2.tex
Added horizontal scaling stub

@@
 Making use of virtualization technologies, which often underpins cloud computing platform, allows for resource manipulation in a dynamic, on-demand manner.
+Although, cloud computing offers additional scaling capabilities, it increases solution complexity since they operate in different layers: server, platform, network as stated in \cite{VaRoBu11}.
 
@@
 
-The rest of this chapter elaborates in details about horizontal and vertical scaling taking into account mechanisms used in Platform-as-a-Service solutions that are available on the market.
+The rest of this chapter elaborates in detail about server, platform and network scaling taking into account mechanisms used in Platform-as-a-Service solutions that are available on the market.
 
-\section{Horizontal scaling}
+\section{Server level}
 
-As it was outlined in a previous section, vital element of horizontal scaling is adding additional resources.
-In context of an application execution environment, node can be represented by an application server, database or other resources that affect application performance such as network interface \cite{VaRoBu11}.
-Dynamic capabilities of a cloud computing platform offer a couple of ways this can be done:
-\begin{itemize}
-	\item virtual machine replication
-	\item network slicing
-	\item database replication
-\end{itemize}
+\subsection{Horizontal scaling}
+As outlined in previous section, horizontal scaling is all about adding supplementary nodes to a system.
+As it is common to cloud computing, nodes are represented as virtual machines and this assumption is used in further discussion.
+As a consequence, adding server comes down to cloning new virtual machine from a template and possibly installing additional software or reconfiguring it later.
+While mechanism of creating new virtual machine from a template is offered literally in every IaaS platform currently available (OpenStack, OpenNebula, OpenStack, CloudStack or Eucalyptus to name a few) and is similar in manner, the underlying virtualization mechanism influences how fast provisioning is done.
+Table 1 describes in details different capabilities of commonly used virtualization technologies.
 
-jebnac mapowanie na warstwy clouda - infr, platform, app
+Provisioning new server is only a first step in scaling an application, it is required to configure some load balancing mechanism to make user of this additional node.
 
-\subsection{Virtual machine replication}
+\subsection{Vertical scaling}
 
-\subsection{Network slicing}
+\section{Platform level}
 
-\subsection{Database replication}
-
-\section{Vertical scaling}
+\section{Network level}

### 2013-08-30 thesis/src/chapters/chapter-2.tex
Added vertical scaling stub

@@
 \chapter{Scaling applications}
+
+\chapterintro{This chapter is devoted to the concept of scaling users' application from the perspective of a cloud platform provider.
+To achieve that, it presents attainments of research groups working in that area as well as it considers mechanisms used in products currently available on the market.}
 
@@
 As a consequence, adding server comes down to cloning new virtual machine from a template and possibly installing additional software or reconfiguring it later.
-While mechanism of creating new virtual machine from a template is offered literally in every IaaS platform currently available (OpenStack, OpenNebula, OpenStack, CloudStack or Eucalyptus to name a few) and is similar in manner, the underlying virtualization mechanism influences how fast provisioning is done.
-Table 1 describes in details different capabilities of commonly used virtualization technologies.
+While mechanism of creating new virtual machine from a template is offered literally in every IaaS platform currently available (OpenStack \cite{OpenStack}, OpenNebula \cite{OpenNebula}, CloudStack \cite{CloudStack} or Eucalyptus \cite{Eucalyptus} to name a few) and is similar in manner, the underlying virtualization mechanism influences how fast provisioning is done.
+Table \ref{tab:hypervisors-provisioning} describes in details different capabilities of commonly used virtualization technologies.
 
-Provisioning new server is only a first step in scaling an application, it is required to configure some load balancing mechanism to make user of this additional node.
+\begin{table}[!htbp]
+\begin{tabularx}{\textwidth}{l | X | X}
+ &  & \\
+\hline 
+KVM &  & \\
+\hline
+Xen &  & \\
+\hline
+VMware &  & \\
+\hline
+OpenVZ &  & \\
+\hline
+LXC &  &  \\
+\end{tabularx}
+\caption{Comparison of hypervisors provisioning capabilities}
+\label{tab:hypervisors-provisioning}
+\end{table}
+
+Provisioning new server is only a first step in scaling an application, it is required to configure load balancing mechanism to make use of additional node.
+Two important aspects that has to be consider are: load-balancing algorithms and scalability.
+
+\subsubsection*{Load-balancing algorithms}
+Generally, there are two types of load-balancers: hardware and software based.
+Due to the dynamic nature of system under consideration, we focus only on the latter as it offers a greater deal of flexibility.
+Among the most common algorithms we can distinguish:
+\begin{itemize}
+ \item \textit{round-robin scheduling} -
+ \item \textit{first in first out} -
+ \item \textit{shortest remaining time} -
+\end{itemize}
+
+Situation gets further complicated when considering real-world web application that sends user information using cookies, what imposes requirement on load-balancer for session stickiness \cite{StBaMa11}.
+Table \ref{tab:load-balancers} summaries available load-balancing mechanism including their utilization in IaaS products.
+
+\begin{table}[!htbp]
+\begin{tabularx}{\textwidth}{l | X | X}
+ &  & \\
+\hline 
+HAProxy &  & \\
+\hline
+Apache modproxy &  & \\
+\hline
+... &  & \\
+\end{tabularx}
+\caption{Comparision of load balancers}
+\label{tab:load-balancers}
+\end{table}
+
+\subsubsection*{Load-balancing scalability}
+Although, it may seem that balancing workloads eliminates problem of a single point of failure (SPOF) among different servers, it is in fact shifted to load-balancing layer.
+In other words, load-balancer becomes a new SPOF.
+Therefore, in cases where high availability is required, multi-tiered load balancing architecture should be considered.
+This, however, seems not to be a case among IaaS or PaaS providers - none of them unequivocally specifies whether their provide redundancy at load-balancer level \textbf{TODO check :)}
 
 \subsection{Vertical scaling}
+
+Essentially, vertical scaling is concentrated upon increasing capacity of single node.
+Again, when considering technical advancements that comes with cloud computing and virtualization, we can differ two categories of scaling: virtual machine resizing and virtual machines replacement.
+This distinction is dictated by limitation hypervisors - not all of them are able to resize virtual machine without shutting it down, in that case provisioning new one is a necessity.
+
+\subsubsection*{Virtual machine resizing}
+
+\begin{table}[!htbp]
+\begin{tabularx}{\textwidth}{l | X | X}
+ &  & \\
+\hline 
+KVM &  & \\
+\hline
+Xen &  & \\
+\hline
+VMware &  & \\
+\hline
+OpenVZ &  & \\
+\hline
+LXC &  &  \\
+\end{tabularx}
+\caption{Comparision of hypervisors resizing capabilities}
+\label{tab:hypervisors-resizing}
+\end{table}
+
+\subsubsection*{Virtual machine replacement}
 

### 2013-09-06 thesis/src/chapters/chapter-4.tex
Add the 'deployment' section

@@
 
-\chapterintro{This chapter introduces the notion of a hybrid cloud and explains its role in cloud computing industry.
+\chapterintro{This chapter introduces the notion of a hybrid cloud and explains its role in IT industry.
 On top of this deployment model, the concept of InterCloud is presented and elaborated with the emphasis on its application in ensuring scalability of users' services.}
@@
 What is more, in order to ensure high availability of their services, customers do not want to confine themselves to only one provider -- in the best scenario they want their applications (or their logical parts, such as persistence layer) to be spanned across different providers and be able to cooperate with one another at the same time.
-This leads to the concept of a \emph{hybrid cloud}\cite{MeGr11} -- the case in which the cloud infrastructure is a composition of two or more distinct infrastructures which are unique entities, but there are technological means that make it possible to port data and applications among them.
+This leads to the concept of a \emph{hybrid cloud}\cite{MeGr11} -- the case in which the cloud is a composition of two or more distinct infrastructures which are unique entities, but there are technological means that make it possible to port data and applications among them.
 
-\subsection{Usage in industry}
+\subsection{Deployment models}
+The informal introduction to the concept of a \emph{hybrid cloud} in the previous section requires a strict definion, but it is virtually impossible without introduction of other deployment models:
+\begin{itemize}
+  \item Private Cloud -- The provisioned cloud infrastructure is used exclusively by a single organization (that may consist of many business units) and may be owned, managed and operated by the organization or a third party.
+  \item Public Cloud -- The provisioned cloud infrastructre is used by general public and may be owned, managed and operated by a business, academic or government organization or some combination of them.
+It exists on the premises of the cloud provider.
+  \item Community Cloud -- The cloud infrastructure is provisioned for exclusive use by a specific community of consumers from organizations that have shared concerns (e.g., mission, security requirements, policy , and compliance considerations).
+It may be owned, managed , and operated by one or more of the organizations in the community, a third party, or some combination of them, and it may exist on or off premises.
+\end{itemize}
+Having defined those models, we can see that \emph{hybrid cloud} can be placed among them and be defined as a model in which the provisioned infrastructure is a composition of two or more other infrastructures - \emph{private},\emph{community} or \emph{public}.
+
+\subsection{Current usage and trends}
 

### 2013-09-09 thesis/src/chapters/chapter-2.tex
Vertical scaling

@@
 As a consequence, adding server comes down to cloning new virtual machine from a template and possibly installing additional software or reconfiguring it later.
-While mechanism of creating new virtual machine from a template is offered literally in every IaaS platform currently available (OpenStack \cite{OpenStack}, OpenNebula \cite{OpenNebula}, CloudStack \cite{CloudStack} or Eucalyptus \cite{Eucalyptus} to name a few) and is similar in manner, the underlying virtualization mechanism influences how fast provisioning is done.
+While mechanism of creating new virtual machine from a template is offered literally in every IaaS platform currently available (OpenStack \cite{OpenStack}, OpenNebula \cite{OpenNebula}, CloudStack \cite{CloudStack} or Eucalyptus \cite{Eucalyptus} to name a few) and is similar in manner, the underlying virtualization mechanism determines how fast provisioning is done.
 Table \ref{tab:hypervisors-provisioning} describes in details different capabilities of commonly used virtualization technologies.
@@
 
-\subsubsection*{Load-balancing algorithms}
+\subsubsection{Load-balancing algorithms}
 Generally, there are two types of load-balancers: hardware and software based.
@@
  \item \textit{round-robin scheduling} - 
- \item \textit{first in first out} -
- \item \textit{shortest remaining time} -
+ \item \textit{least connection} - 
+ \item \textit{source routing} - 
+ \item \textit{URI hashing} - 
 \end{itemize}
@@
 Situation gets further complicated when considering real-world web application that sends user information using cookies, what imposes requirement on load-balancer for session stickiness \cite{StBaMa11}.
-Table \ref{tab:load-balancers} summaries available load-balancing mechanism including their utilization in IaaS products.
+Below are presented one of the most common load-balancers, including their key performance features.
 
-\begin{table}[!htbp]
-\begin{tabularx}{\textwidth}{l | X | X}
- &  & \\
-\hline 
-HAProxy &  & \\
-\hline
-Apache modproxy &  & \\
-\hline
-... &  & \\
-\end{tabularx}
-\caption{Comparision of load balancers}
-\label{tab:load-balancers}
-\end{table}
+\subsubsection*{HAProxy}
 
-\subsubsection*{Load-balancing scalability}
+HAProxy \cite{HAProxy} is a load-balancer developed by. Noticeably, it's used by OpenShift \cite{OpenShift} to distribute load among gears (CITE NEEDED). Its key features include:
+\begin{itemize}
+	\item a single-process, event-driven model reduces the cost of context switch and the memory usage
+	\item O(1) event checker
+	\item Single-buffering without copying data between reads and writes
+	\item Zero-copy forwarding
+	\item Optimized HTTP header analysis : headers are parsed an interpreted on the fly
+\end{itemize}
+
+Requests are dispatched accordingly to chosen algorithm:
+\begin{itemize}
+ \item \textit{round-robin scheduling}
+ \item \textit{least connection}
+ \item \textit{source routing}
+ \item \textit{URI hashing}
+\end{itemize}
+
+\subsubsection*{Apache HTTP with modproxy}
+
+\subsubsection*{Apache HTTP with mod_jk}
+
+\subsubsection{Load-balancing scalability}
 Although, it may seem that balancing workloads eliminates problem of a single point of failure (SPOF) among different servers, it is in fact shifted to load-balancing layer.

### 2013-09-09 thesis/src/chapters/chapter-2.tex
Load balancing algorithms

@@
 Due to the dynamic nature of system under consideration, we focus only on the latter as it offers a greater deal of flexibility.
-Among the most common algorithms we can distinguish:
+Among the most common algorithms we can distinguish \cite{HaProxyDoc}:
 \begin{itemize}
- \item \textit{round-robin scheduling} - 
- \item \textit{least connection} - 
- \item \textit{source routing} - 
- \item \textit{URI hashing} - 
+ \item \textit{round-robin scheduling} - request are sent to sucessive nodes, according to their weights.
+This algorithm is fairest when the server's processing time remains equally distributed \cite{HaProxyDoc}
+ \item \textit{least connection} - the server with the lowest number of connections receives the connection
+ \item \textit{source routing} - source IP address is hashed, the same client IP address always reaches the same server
+ \item \textit{URI hashing} - URI that designates resource is hashed and divided by the total weight of the running servers.
+Such hash designates which server that receives the request.
+In practice, this algorithm is commonly used with proxy caches and anti-virus proxies in order to maximize the cache hit rate.
 \end{itemize}
@@
 Situation gets further complicated when considering real-world web application that sends user information using cookies, what imposes requirement on load-balancer for session stickiness \cite{StBaMa11}.
+
+\subsubsection{Load-balancing scalability}
+Although, it may seem that balancing workloads eliminates problem of a single point of failure (SPOF) among different servers, it is in fact shifted to load-balancing layer.
+In other words, load-balancer becomes a new SPOF.
+Therefore, in cases where high availability is required, multi-tiered load balancing architecture should be considered.
+This, however, seems not to be a case among IaaS or PaaS providers - none of them unequivocally specifies whether their provide redundancy at load-balancer level.
+
 Below are presented one of the most common load-balancers, including their key performance features.
@@
 
-HAProxy \cite{HAProxy} is a load-balancer developed by. Noticeably, it's used by OpenShift \cite{OpenShift} to distribute load among gears (CITE NEEDED). Its key features include:
+HAProxy \cite{HAProxy} is a load-balancer developed by. Noticeably, it's used by OpenShift \cite{OpenShift} to distribute load among gears \cite{OpenShiftScaling}. Its key features include:
 \begin{itemize}
@@
 
-\subsubsection*{Apache HTTP with modproxy}
+\subsubsection*{BIG-IP Local Traffic Manager (LTM)}
+BIG-IP Local Traffic Manager is a solution offered by F5 \cite{F5}.
+Although LTM is a hardware solution, omitted in this section, it also has its virtualized counterpart.
+Key features offered by LTM are \cite{LTM}:
+\begin{itemize}
+  \item managing at application services level rather than at individual devices and objects
+  \item scripting language that allows administrator to intercept, inspect, transform, and direct application traffic
+  \item built-in firewall protection, application security, and access control
+  \item real-time protocol and traffic management decisions
+\end{itemize}
 
-\subsubsection*{Apache HTTP with mod_jk}
+\subsubsection*{Apache HTTPD}
+Apache HTTPD \cite{ApacheHTTPD} is a popular HTTP server.
+When enhanced with additional modules, it can behave as a proxy or load-balancer.
+Over the time, there were serveral attempts to develop such modules: mod\_jk \cite{ApacheModJk}, mod\_proxy\_balancer \cite{ApacheModProxyBalancer}, to name a few.
+While the former is AJP13 oriented, the latter supports different protocols: HTTP, FTP and AJP13, hence, it is described here.
 
-\subsubsection{Load-balancing scalability}
-Although, it may seem that balancing workloads eliminates problem of a single point of failure (SPOF) among different servers, it is in fact shifted to load-balancing layer.
-In other words, load-balancer becomes a new SPOF.
-Therefore, in cases where high availability is required, multi-tiered load balancing architecture should be considered.
-This, however, seems not to be a case among IaaS or PaaS providers - none of them unequivocally specifies whether their provide redundancy at load-balancer level \textbf{TODO check :)}
+Among the mod\_proxy\_balancer features we may distinguish:
+\begin{itemize}
+ \item support for different load-balancing algorithms:
+    \begin{itemize}
+      \item Request Counting Algorithm - load is distributed the requests among the various workers, ensuring that each gets their configured share of the number of requests
+      \item Weighted Traffic Counting Algorithm - variation of above-mentioned algorithm with a difference that it is focused on bytes rather than number of request
+      \item Pending Request Counting Algorithm - scheduler keeps track of how many requests each worker is assigned at present.
+A new request is automatically assigned to the worker with the lowest number of active requests
+    \end{itemize}
+
+ \item support for session stickiness by using cookies and URL encoding.
+This approach \cite{ApacheModProxyBalancer} avoids unequal load distribution if clients are hidden behind proxies and stickyness errors when a client uses a dynamic IP address that changes during a session
+\end{itemize}
 
 \subsection{Vertical scaling}
-
 Essentially, vertical scaling is concentrated upon increasing capacity of single node.

### 2013-09-09 thesis/src/chapters/chapter-4.tex
Add clients' view on cloud computing section

@@
 \subsection{Deployment models}
-The informal introduction to the concept of a \emph{hybrid cloud} in the previous section requires a strict definion, but it is virtually impossible without introduction of other deployment models:
+The informal introduction to the concept of a \emph{hybrid cloud} in the previous section requires a strict definition, but it is virtually impossible without defining other deployment models:
 \begin{itemize}
   \item Private Cloud -- The provisioned cloud infrastructure is used exclusively by a single organization (that may consist of many business units) and may be owned, managed and operated by the organization or a third party.
-  \item Public Cloud -- The provisioned cloud infrastructre is used by general public and may be owned, managed and operated by a business, academic or government organization or some combination of them.
+  \item Public Cloud -- The provisioned cloud infrastructure is used by general public and may be owned, managed and operated by a business, academic or government organization or some combination of them.
 It exists on the premises of the cloud provider.
@@
 \subsection{Current usage and trends}
+\subsubsection*{Cloud -- clients' view}
+Before digging into the details of current usage and popularity of the hybrid model, it is worth discussing the general attitude of clients towards cloud computing.
+As the recent survey \cite{RackspaceSurvey13} shows, the major factor that prevents companies from adopting cloud solutions is their concern over security -- in 2012 as much as 52\% responders considered it as a main concern with the regard to cloud in general.
+However, the tendency is that more and more enterprises do not find it a major issue as in 2012 the number declined to 46\%.
+Complexity related to the managament of cloud components, Vendor lock-in, interoperability and reliability were among the most frequent obsctacles to adoption in 2013 for they constitued 46\%, 35\%, 27\% and 22.3\% of responders' votes respectively.
+Total results are shown in the figure \ref{ch4:cloud-computing-adoption-obstacles-2013}.
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics{chapter-4/cloud-computing-adoption-obstacles-2013}
+  \end{center}
+  \caption{Major obstacles to cloud adoption in 2013 according to \cite{RackspaceSurvey13}}
+  \label{ch4:cloud-computing-adoption-obstacles-2013}
+\end{figure}
+
+\subsubsection*{View on hybrid cloud}
 

### 2013-09-09 thesis/src/chapters/chapter-4.tex
Add clients' view on cloud computing section - continued

@@
 Before digging into the details of current usage and popularity of the hybrid model, it is worth discussing the general attitude of clients towards cloud computing.
-As the recent survey \cite{RackspaceSurvey13} shows, the major factor that prevents companies from adopting cloud solutions is their concern over security -- in 2012 as much as 52\% responders considered it as a main concern with the regard to cloud in general.
+As the recent survey \cite{NBSurvey13} shows, the major factor that prevents companies from adopting cloud solutions is their concern over security -- in 2012 as much as 52\% responders considered it as a main concern with the regard to cloud in general.
 However, the tendency is that more and more enterprises do not find it a major issue as in 2012 the number declined to 46\%.
-Complexity related to the managament of cloud components, Vendor lock-in, interoperability and reliability were among the most frequent obsctacles to adoption in 2013 for they constitued 46\%, 35\%, 27\% and 22.3\% of responders' votes respectively.
+Complexity related to the management of cloud components, Vendor lock-in, interoperability and reliability were among the most frequent obstacles to adoption in 2013 for they constituted 46\%, 35\%, 27\% and 22.3\% of responders' votes respectively.
 Total results are shown in the figure \ref{ch4:cloud-computing-adoption-obstacles-2013}.
@@
   \end{center}
-  \caption{Major obstacles to cloud adoption in 2013 according to \cite{RackspaceSurvey13}}
+  \caption{Major obstacles to cloud adoption in 2013 according to \cite{NBSurvey13}}
   \label{ch4:cloud-computing-adoption-obstacles-2013}
 \end{figure}
+
+The same survey shows that the cloud adoption growth rate is high -- 75 percent of responders stated usage of some sort of a cloud platform.
+This means 8 percent growth when compared to the results obtained in 2012.
+The expectations for the total worldwide addressable market for cloud computing are to reach \$158.8B by 2014 -- an increase of 126.5 percent from 2011.
 

### 2013-09-09 thesis/src/chapters/chapter-4.tex
Add view on a hybrid cloud

@@
 \subsubsection*{View on hybrid cloud}
+When it comes to the application of a hybrid model in industry, in most cases the definition introduced in the previous chapter now becomes a 'public-private' composition.
+And this is how the term should be understood when discussing the results of the surveys which aimed to provide insights onto the view on a hybrid cloud from the customers' perspective.
+The study \cite{NBSurvey13} forecasts 16 percentage growth in the hybrid cloud adoption in 5 years, from 27 to 43 percent.
+At the same time, the usage of a public model will decline from 39 to 32 percent.
+The other survey, conducted by Rackspace \cite{RackspaceSurvey13}, provides more detailed data about current usage and popularity of a hybrid model.
+The first interesting finding is that as much as 60\% responders, which included 1300 companies in the UK and US, have moved or are planning to move certain applications either partially (41\%) or completely (19\%) off the public cloud because of its limitations or the potential benefits of other models, e.g.
+the hybrid one.
+The second one is about the pros of adopting the hybrid cloud  -- potential users find more control (59\%) and better security (54\%) top benefits of using this deployment model.
+The other most responded benefits are shown in the figure \ref{ch4:hybrid-cloud-top-benefits}.
+
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics{chapter-4/hybrid-cloud-top-benefits}
+  \end{center}
+  \caption{Top benefits of using the hybrid model according to \cite{RackspaceSurvey13}}
+  \label{ch4:hybrid-cloud-top-benefits}
+\end{figure}
 

### 2013-09-09 thesis/src/chapters/chapter-2.tex
Finished application scaling

@@
 Enriching system with capability to scale entails avoiding additional costs that are related to coping with excessive traffic.
-In some cases, these costs may be caused by not handling extra traffic at all or may involve aspects such as: increased response time, processing overhead, space, memory, or money \cite{Bo00}.
+In some cases, these costs may be caused by not handling extra traffic at all and may involve aspects such as: increased response time, processing overhead, space, memory, or money \cite{Bo00}.
 
@@
 	\item \textit{load scalability} - ability to work without delays and unproductive resource consumption at light, moderate, or heavy loads while making good use of available resources.
-Factors that may hinder load scalability include aspects such as: scheduling shared resource, self-expansion, inadequate exploitation of parallelism
-	\item \textit{space scalability} - memory requirements do not grow to intolerable levels as the number of items it supports increases
+Factors that may hinder load scalability include: scheduling shared resource, self-expansion, inadequate exploitation of parallelism
+	\item \textit{space scalability} - memory requirements do not grow to intolerable levels as the number of items system supports increases
 	\item \textit{space-time scalability} - system continues to function gracefully as the number of objects it encompasses increases by orders of magnitude
-	\item \textit{structural scalability} - implementation or standards do not impede the growth of the number of objects it encompasses
+	\item \textit{structural scalability} - implementation or standards do not impede the growth of the number of objects system encompasses
 \end{itemize}
 Although, all of the aforementioned aspects are crucial for functioning of any healthy application, our work focuses only on the first type of scalability.
-The reasoning behind this statement is that while all of these scalability types lies in direct responsibility of an application developer, the load scalability can be additionally improved by adding additional resources to a system.
+The reasoning behind this statement is that, while all of these scalability types lies in direct responsibility of an application developer, the load scalability can be additionally improved by adding additional resources to a system.
 This brings us to a question what kind of resources are used by an application or more appropriately to context of this dissertation: \textit{what kind of resources can we add to improve application performance?} Required resources varies from an application to an application.
@@
 Having that said, common sense dictates that adding resources is only a part of the success - it should be accompanied by tuning application platform configuration.
-For example, adding supplementary CPUs without increasing thread pool size that handle requests make a little sense.
-Similarly, we have to increase heap size, in context of a Java application, to make a good use of extra memory.
+For example, adding supplementary CPUs without increasing thread pool size that handle requests makes a little sense.
+Similarly, in context of a Java application, we have to increase heap size, , to make a good use of extra memory.
 While importance of application tuning cannot be underestimated, its detailed analysis lies outside of the scope of this dissertation.
@@
 Although, cloud computing offers additional scaling capabilities, it increases solution complexity since they operate in different layers: server, platform, network as stated in \cite{VaRoBu11}.
+However, since platform containers are often represented either as virtual machines or another isolated environment (e.g.
+OpenShift leverages SELinux and cgroups) they are similar in nature to server scaling and supports both scaling up and out.
+
+<mapping layer => actions>
 
@@
 
-The rest of this chapter elaborates in detail about server, platform and network scaling taking into account mechanisms used in Platform-as-a-Service solutions that are available on the market.
+The rest of this chapter elaborates in detail about horizontal and vertical scaling taking into account mechanisms used in Platform-as-a-Service solutions that are available on the market.
 
-\section{Server level}
-
-\subsection{Horizontal scaling}
-As outlined in previous section, horizontal scaling is all about adding supplementary nodes to a system.
+\section{Horizontal scaling}
+As outlined in previous section, horizontal scaling is about adding supplementary nodes to a system.
 As it is common to cloud computing, nodes are represented as virtual machines and this assumption is used in further discussion.
-As a consequence, adding server comes down to cloning new virtual machine from a template and possibly installing additional software or reconfiguring it later.
-While mechanism of creating new virtual machine from a template is offered literally in every IaaS platform currently available (OpenStack \cite{OpenStack}, OpenNebula \cite{OpenNebula}, CloudStack \cite{CloudStack} or Eucalyptus \cite{Eucalyptus} to name a few) and is similar in manner, the underlying virtualization mechanism determines how fast provisioning is done.
-Table \ref{tab:hypervisors-provisioning} describes in details different capabilities of commonly used virtualization technologies.
-
-\begin{table}[!htbp]
-\begin{tabularx}{\textwidth}{l | X | X}
- &  & \\
-\hline 
-KVM &  & \\
-\hline
-Xen &  & \\
-\hline
-VMware &  & \\
-\hline
-OpenVZ &  & \\
-\hline
-LXC &  &  \\
-\end{tabularx}
-\caption{Comparison of hypervisors provisioning capabilities}
-\label{tab:hypervisors-provisioning}
-\end{table}
+Consequently, adding server comes down to cloning a new virtual machine from a template and possibly installing additional software and reconfiguring it later.
+While mechanism of creating new virtual machine from a template is offered literally in every IaaS platform currently available (OpenStack \cite{OpenStack}, OpenNebula \cite{OpenNebula}, CloudStack \cite{CloudStack} or Eucalyptus \cite{Eucalyptus} to name a few) and is similar in manner, the underlying hardware and virtualization mechanism determines how fast provisioning is done.
 
 Provisioning new server is only a first step in scaling an application, it is required to configure load balancing mechanism to make use of additional node.
-Two important aspects that has to be consider are: load-balancing algorithms and scalability.
+The two important aspects that have to be consider are: load-balancing algorithms and scalability.
 
-\subsubsection{Load-balancing algorithms}
+\subsection{Load-balancing algorithms}
 Generally, there are two types of load-balancers: hardware and software based.
@@
 
-\subsubsection{Load-balancing scalability}
-Although, it may seem that balancing workloads eliminates problem of a single point of failure (SPOF) among different servers, it is in fact shifted to load-balancing layer.
-In other words, load-balancer becomes a new SPOF.
-Therefore, in cases where high availability is required, multi-tiered load balancing architecture should be considered.
-This, however, seems not to be a case among IaaS or PaaS providers - none of them unequivocally specifies whether their provide redundancy at load-balancer level.
-
 Below are presented one of the most common load-balancers, including their key performance features.
@@
 
-HAProxy \cite{HAProxy} is a load-balancer developed by. Noticeably, it's used by OpenShift \cite{OpenShift} to distribute load among gears \cite{OpenShiftScaling}. Its key features include:
+HAProxy \cite{HAProxy} is a load-balancer initially written by Willy Tarreau.
+Noticeably, it's used by OpenShift \cite{OpenShift} to distribute load among gears \cite{OpenShiftScaling}.
+Its key features include:
 \begin{itemize}
@@
 
-\subsection{Vertical scaling}
+\subsection{Load-balancing scalability}
+Although, it may seem that balancing workloads eliminates problem of a single point of failure (SPOF) among different servers, it is in fact shifted to load-balancing layer.
+In other words, load-balancer becomes a new SPOF.
+Therefore, in cases where high availability is required, multi-tiered load balancing architecture should be considered.
+This, however, seems not to be a case among IaaS or PaaS providers - none of them unequivocally specifies whether their provide redundancy at load-balancer level.
+
+\section{Vertical scaling}
 Essentially, vertical scaling is concentrated upon increasing capacity of single node.
 Again, when considering technical advancements that comes with cloud computing and virtualization, we can differ two categories of scaling: virtual machine resizing and virtual machines replacement.
-This distinction is dictated by limitation hypervisors - not all of them are able to resize virtual machine without shutting it down, in that case provisioning new one is a necessity.
+This distinction is dictated by limitation hypervisors - not all of them are able to resize virtual machine without shutting it down.
 
-\subsubsection*{Virtual machine resizing}
+\subsection*{Virtual machine resizing}
 
 \begin{table}[!htbp]
-\begin{tabularx}{\textwidth}{l | X | X}
- &  & \\
+\begin{tabularx}{\textwidth}{| l | X | X | X |}
 \hline
-KVM &  & \\
+ & \textbf{Memory} & \textbf{CPU} & \textbf{Disk} \\
 \hline 
-Xen &  & \\
-\hline
-VMware &  & \\
-\hline
-OpenVZ &  & \\
-\hline
-LXC &  &  \\
+KVM &
+-  & 
+dynamic pinning CPU to a specific virtual machine (depending on underlying hardware)
+& 
+adding a disk to a LVM group
+
+\\ \hline
+Xen & 
+- changing the amount of host physical memory assigned to virtual machine without rebooting it
+
+- start additional virtual machines on a host whose physical memory is currently full, by automatically reducing the memory allocations of existing virtual machines in order to make space
+&
+dynamic pinning CPU to a specific virtual machine (depending on underlying hardware)
+&
+dynamic block attaching, adding a disk to a LVM group
+
+\\ \hline
+VMware ESX &
+hot-plugging memory, ex. using VMware vSphere
+&
+hot-plugging CPU, ex. using VMware vSphere
+&
+adding additional disks to existing virtual machine
+
+\\ \hline
+OpenVZ &
+configurable via user beancounters
+&
+configurable via user beancounters 
+& 
+configurable via user beancounters
+\\ \hline
 \end{tabularx}
@@
 
-\subsubsection*{Virtual machine replacement}
-
-\section{Platform level}
-
-\section{Network level}
+\subsection*{Virtual machine replacement}
+As it was highlighted in previous section, reasoning behind virtual machine replacement is that, in case when dynamic resizing is not possible, a new virtual machine with a desired configuration can be provisioned and replace the old one.
+Since this is a basic operation, all above-mentioned hypervisors supports this scenario as long as required resources are available.

### 2013-09-10 thesis/src/chapters/chapter-2.tex
Added amdahl chart, new chapter

@@
 
-<mapping layer => actions>
-
 With all that said, there is no silver bullet - not matter what underlying mechanism platform provider decides to use, the application developer is still responsible for creating an application with scaling in-mind.
-This statement has been already proven in 1967 by Amdahl law, which in short states that sequential component of parallel algorithms impacts efficiency for a sufficiently large number processors \cite{Am67}.
+This statement has been already proven in 1967 by Amdahl law, which in short states that sequential component of parallel algorithms impacts efficiency for a sufficiently large number processors \cite{Am67} as shown in \ref{ch2:amdahl-lawl}.
 In other words, adding supplementary resources to a poorly written application (i.e.
 having a lot of sequential or synchronized components) can be beneficial only to a certain degree.
+
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics{chapter-2/amdahl-law}
+  \end{center}
+  \caption{Amdahl's law}
+  \label{ch2:amdahl-law}
+\end{figure}
 
@@
     \begin{itemize}
-      \item Request Counting Algorithm - load is distributed the requests among the various workers, ensuring that each gets their configured share of the number of requests
-      \item Weighted Traffic Counting Algorithm - variation of above-mentioned algorithm with a difference that it is focused on bytes rather than number of request
-      \item Pending Request Counting Algorithm - scheduler keeps track of how many requests each worker is assigned at present.
+      \item \textit{request counting algorithm} - load is distributed the requests among the various workers, ensuring that each gets their configured share of the number of requests
+      \item \textit{weighted traffic counting algorithm} - variation of above-mentioned algorithm with a difference that it is focused on bytes rather than number of request
+      \item \textit{pending request counting algorithm} - scheduler keeps track of how many requests each worker is assigned at present.
 A new request is automatically assigned to the worker with the lowest number of active requests

### 2013-09-10 thesis/src/chapters/chapter-3.tex
Added amdahl chart, new chapter

\chapter{Platform adaptivity}

\chapterintro{This chapter introduces concepts and mechanisms that enhance a platform with adaptivity capabilities, which are often achieved by fusion of rules, policies and scaling techniques.}

\section{Introduction}
In short, platform adaptivity adds a auto-scaling features to a solution offered by Platform-as-a-Service provider. Key concept is to have a Elasticity Controller which gathers probes from virtual machines and uses that knowledge to execute appropriate action on Cloud instance, indirectly modifying consecutive probes \cite{VaRoBu11}. This concept illustrates diagram (IMAGE).

\section{Policies}
While auto-scaling is offered by a vast amount of cloud providers (e.g AWS, OpenShift, OpenNebula) it often lack a sophisticated mechanisms allowing for specific scaling policies, limited to only one predefined rule as it is in case of OpenShift for example.

In a environment consisting of hundred nodes, provisioned and shut down in a dynamic manner, automatization

### 2013-09-10 thesis/src/chapters/chapter-4.tex
Add InterCloud section

@@
 \section{Federation of clouds -- InterCloud}
+As stated in the previous sections, one of the major obstacles that prevents consumers from adopting cloud solutions is reliability.
+It appears that these statements are not only imaginary worries of entrepreneurs, but real problems -- there are cases, where some providers temporarily run short of capacity (e.g.
+because of provisioning too many virtual machines) in the face of hight demand \cite{Lavitt09}.
+What is more, the more demanding clients require specific QoS to be satisfied by their providers as negotiated in Service Level Agreements.
+In order to meet these challenges there is a need of a completely new approach to the problem of effective management of resources. The new solution should take into account such factors as:
+
+\begin{itemize}
+  \item users' requests priority
+  \item users' QoS requirements (e.g. the deadline by which some jobs have to be executed)
+  \item price that the clients pay for the usage of resources
+\end{itemize}
+
+Computer scientists in the field of cloud computing devised a model \cite{MarketOrientedCC08} in which resources are managed in a market-oriented fashion that enables dynamic regulation of the supply and demand of resources and promotes the mechanisms for their allocation that would take into account their priorities and levels of utilization.
+The extension of this model is a vision of creating the federated cloud computing environment, so called \emph{InterCloud}, that ``facilitates just-in-time, opportunistic, and scalable provisioning of application services, consistently achieving QoS targets under variable workload, resource and network conditions'' \cite{InterCloud10}.
+The elements of the proposed architecture are as follows:
+\begin{itemize}
+  \item Cloud Exchange -- acts as a market maker for bringing together both producers and consumers of services.
+It allows Cloud Brokers and Cloud Coordinators to match consumers with the fitting offers from providers.
+Such a market is a step forward towards creating a dynamic infrastructure for trading based on Service Level Agreements.
+  \item Cloud Coordinator -- manages the instance of a cloud and its membership in the overall federation; provides an environment (programming, deployment) for applications
+  \item Cloud Broker -- acts on behalf of the client; communicates with the Cloud Exchange to find the best cloud instances for the application
+\end{itemize}
 
 \subsection{Usage in industry}
+The depicted model has not yet been adopted in the industry, yet some simulations were carried out on a \emph{CloudSim} platform and the obtained results showed that this concept has ``immense potential'' \cite{InterCloud10}.

### 2013-09-11 thesis/src/appendices/listings.tex
Added policy description

\chapter{Code listings}

\section{Scaling policies}
\label{app:scaling-policies}

\lstinputlisting[caption=Auto Scaling policy\, AWS EC2]{scaling-policy.xml}

### 2013-09-11 thesis/src/chapters/chapter-2.tex
Added policy description

@@
 With all that said, there is no silver bullet - not matter what underlying mechanism platform provider decides to use, the application developer is still responsible for creating an application with scaling in-mind.
-This statement has been already proven in 1967 by Amdahl law, which in short states that sequential component of parallel algorithms impacts efficiency for a sufficiently large number processors \cite{Am67} as shown in \ref{ch2:amdahl-lawl}.
+This statement has been already proven in 1967 by Amdahl law, which in short states that sequential component of parallel algorithms impacts efficiency for a sufficiently large number processors \cite{Am67} as shown in Figure \ref{ch2:amdahl-law}.
 In other words, adding supplementary resources to a poorly written application (i.e.

### 2013-09-11 thesis/src/chapters/chapter-3.tex
Added policy description

@@
 
+The remaining of this chapter describes crucial elemenets that compose elasticity controler: policies, data analysis, triggered actions and presents a comparision of cloud providers.
+
 \section{Policies}
-While auto-scaling is offered by a vast amount of cloud providers (e.g AWS, OpenShift, OpenNebula) it often lack a sophisticated mechanisms allowing for specific scaling policies, limited to only one predefined rule as it is in case of OpenShift for example.
+While auto-scaling is offered by a vast amount of cloud providers (e.g AWS, OpenShift, OpenNebula) it often lacks a sophisticated mechanisms allowing for specific scaling policies, being limited to only one predefined rule as it is in case of OpenShift for example.
 
-In a environment consisting of hundred nodes, provisioned and shut down in a dynamic manner, automatization
+Policy denotes a condition which, when satisfied, triggers an action that is supposed to harness cloud instance in a way that future evalutions of condition will be unsuccessfull.
+Typically condition itself is accompanied by a minimal and maximal number of node instances, allowing for ensuring minimal QoS and controling maximal costs.
+Currently, industry leaders supports two main kind of policies \cite{AmazonAutoScaling}:
+\begin{itemize}
+ \item \textit{expression based} - allows to define how you to scale application in response to changing conditions, which can include factors such as memory, cpu usage, cost or some indirect, calulated metrics
+ \item \textit{scheduled} - allows you to scale your application in response to predictable load changes.
+For example, traffic increases during the weekends and decreases on working days.
+Hence, we may use that predictable traffic patterns to scale application based on current time.
+\end{itemize}
+
+Technically, policies are expressed in some human-readable format such as JSON, XML as it is in case of AWS EC2 or custom expression used for example by Carina.
+Appendinx \ref{app:scaling-policies} presents example configuration used by AWS E2 Auto-Scaling.
+
+\section{Data analysis}
+Having policies defined, their conditions are evaluated against data aquired from sensors.
+In a simplest case this evaluation can be based on a Threshold Model \cite{LiWoZh05}, which defines a valid range.
+In cases when given metric violates that condition (i.e.
+value is either smaller than minimal or bigger than maximal acceptable) corresponding resource is properly adjusted.
+While trivial in its form, cases of AWS, OpenShift, Carina, OneCloud proves it is useful in a real-world scenarios.
+Having that said, more sophisticated algorithms also exists:
+\begin{itemize}
+  \item \textit{integer programming} - autoscaling is reduced to server interger programming problems, which aims to minimize the cost or maximize the computing power with either computing power constraints or
+budget constraints \cite{MaLiHu10}
+  \item \textit{prediction error correction} - \cite{ShSuGuWi11}
+\end{itemize}
+
+\section{Triggered actions}
+Actions that are being triggered by a elasticity controller are focused on application scaling. Problem of scaling an applcication is described in detail in previous chapter.
+
+\section{Providers comparision}

### 2013-09-11 article/article.tex
Article -- initial commit

\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
\documentclass[twocolumn]{svjour3}          % twocolumn
%
\usepackage[utf8]{inputenc}
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
\journalname{Computer Science Journal -- AGH University of Science and Technology}
%
\begin{document}

\title{Automatic scaling of users' application deployed in a federation of cloud computing environments}

%\subtitle{Do you have a subtitle?\\ If so, write it here}

%\titlerunning{Short form of title}        % if too long for running head

\author{Dariusz Chrząścik          \and
        Radosław Morytko
}

%\authorrunning{Short form of author list} % if too long for running head

\institute{D. Chrząścik \at
  \email{dariusz@chrzascik.com}           %  \\
  \and
  R. Morytko \at
  \email{radoslaw@morytko.pl}
}

\date{Received: 11 Sep 2013 / Accepted: 15 Sep 2013}

\maketitle

\begin{abstract}
Insert your abstract here. Include keywords, PACS and mathematical
subject classification numbers as needed.
\keywords{cloud computing\and application scaling \and intercloud \and hybrid configuration}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

\section{Introduction}
\label{intro}
Your text comes here. Separate text sections with
\section{Section title}
\label{sec:1}
Text with citations \cite{RefB} and \cite{RefJ}.
\subsection{Subsection title}
\label{sec:2}
as required. Don't forget to give each section
and subsection a unique label (see Sect.~\ref{sec:1}).
\paragraph{Paragraph headings} Use paragraph headings as needed.
\begin{equation}
a^2+b^2=c^2
\end{equation}

% For one-column wide figures use
%\begin{figure}
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
%  \includegraphics{example.eps}
% figure caption is below the figure
%\caption{Please write your figure caption here}
%\label{fig:1}       % Give a unique label
%\end{figure}
%
% For two-column wide figures use
%\begin{figure*}
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
%  \includegraphics[width=0.75\textwidth]{example.eps}
% figure caption is below the figure
%\caption{Please write your figure caption here}
%\label{fig:2}       % Give a unique label
%\end{figure*}
%
% For tables use
\begin{table}
% table caption is above the table
\caption{Please write your table caption here}
\label{tab:1}       % Give a unique label
% For LaTeX tables use
\begin{tabular}{lll}
\hline\noalign{\smallskip}
first & second & third  \\
\noalign{\smallskip}\hline\noalign{\smallskip}
number & number & number \\
number & number & number \\
\noalign{\smallskip}\hline
\end{tabular}
\end{table}

%\begin{acknowledgements}
%If you'd like to thank anyone, place your comments here
%and remove the percent signs.
%\end{acknowledgements}

% BibTeX users please use one of
%\bibliographystyle{spbasic}      % basic style, author-year citations
%\bibliographystyle{spmpsci}      % mathematics and physical sciences
%\bibliographystyle{spphys}       % APS-like style for physics
%\bibliography{}   % name your BibTeX data base

% Non-BibTeX users please use
\begin{thebibliography}{}
%
% and use \bibitem to create references. Consult the Instructions
% for authors for reference list style.
%
\bibitem{RefJ}
% Format for Journal Reference
Author, Article title, Journal, Volume, page numbers (year)
% Format for books
\bibitem{RefB}
Author, Book title, page numbers. Publisher, place (year)
% etc
\end{thebibliography}

\end{document}

### 2013-09-12 thesis/src/chapters/chapter-3.tex
added 2 more data analysis modesl

@@
   \item \textit{prediction error correction} - \cite{ShSuGuWi11}
+  \item \textit{linearized dynamic control} - linearized correction model is based on control equations \cite{AbShBh02}:
+
+    \begin{equation}
+      x(k+1)=Ax(k)+Bu(k)
+    \end{equation}
+
+    \begin{equation}
+       y(k) = C x(k) + D u(k) + E z(k)
+    \end{equation}
+    where x denotes the state variable vector and coefficient matrices: A B C D E are fitted to historical data as a regression model - \cite{DiLuFrHePa03}
+
+  \item \textit{markov decision process model} - computes optimal reaction to state changes by using obsercation on the system with assumptions about the rate of changes expected in the future \cite{AbWo02}
 \end{itemize}

### 2013-09-12 thesis/src/chapters/chapter-2.tex
Added comparision of providers' scaling capabilites

@@
 Since this is a basic operation, all above-mentioned hypervisors supports this scenario as long as required resources are available.
+
+\section{Providers comparision}
+
+Table \ref{tab:cloud-providers-scaling} presents a summary of cloud provider scaling capabilites.
+Interestingly, all of them are focused solely on horizontal scaling, ignoring advantages offered by a fine-grained approach to scaling that leverage scaling up and application tuning.
+
+\begin{table}[!htbp]
+\begin{tabularx}{\textwidth}[]{ X  c  c  c }
+\specialrule{.1em}{.05em}{.05em}
+
+  & \textbf{Horizontal scaling} & \textbf{Vertical scaling} & \textbf{Application / resource tuning} \\
+\specialrule{.1em}{.05em}{.05em}
+
+\multicolumn{4}{ l }{\textbf{Infrastructure provider}} \\
+\specialrule{.1em}{.05em}{.05em}
+
+Carina & \checkmark & $\times$ & $\times$ \\ \hline
+
+OneFlow & \checkmark & $\times$ & $\times$ \\ \hline
+
+AWS EC2 & \checkmark & $\times$ & $\times$ \\ \hline
+
+\multicolumn{4}{ l }{\textbf{Platform provider}} \\
+\specialrule{.1em}{.05em}{.05em}
+
+CloudFoundry & $\times$ & $\times$ & $\times$ \\ \hline
+
+OpenShift & \checkmark & $\times$ & $\times$ \\ \hline
+
+AppEngine & \checkmark & $\times$ & $\times$ \\ \hline
+
+Azure & \checkmark & $\times$ & $\times$ \\ \hline
+
+Heroku & \checkmark & $\times$ & $\times$ \\ \hline
+\end{tabularx}
+
+\caption{Comparision of cloud providers scaling capabilites}
+\label{tab:cloud-providers-scaling}
+
+\end{table}

### 2013-09-12 article/article.tex
Add abstract to the article

@@
 
-\title{Automatic scaling of users' application deployed in a federation of cloud computing environments}
+\title{Self-adaptive platform for scaling users' application deployed in a federation of cloud computing environments}
 
@@
 \author{Dariusz Chrząścik          \and
-        Radosław Morytko
+        Radosław D. Morytko
 }
@@
 
-\institute{D. Chrząścik \at
+\institute{
+  D. Chrząścik \\
+  Chief Technology Officer at Software Mind SA \\
+  Enthusiast of gorgeous cars and fast women \\
+  Amateur poet, screenwriter and novelist\at
   \email{dariusz@chrzascik.com}           %  \\
   \and
-  R. Morytko \at
+  R.D. Morytko \\
+  Software Developer and Board Member at Dropsport \\
+  Devoted admirer of Gothic cathedrals \\
+  Amateur guitar player and zumba dancer\at
   \email{radoslaw@morytko.pl}
@@
 \begin{abstract}
-Insert your abstract here. Include keywords, PACS and mathematical
-subject classification numbers as needed.
+In order to retain the lead among its competitors, a software company needs to create a distributed system which is spanned across different geographical locations, vulnerable to sudden variations in demand and has remarkable QoS requirements.
+Following current trends it chooses Cloud as a deployment platform.
+Unfortunately, existing cloud providers do not have tools and mechanisms that would enable dynamic load distribution among different data centers to meet aforementioned requirements.
+In this article we want to outline the architecture of a self-adaptive platform which facilitates scalable provisioning users' applications and fulfills QoS needs under variable conditions.
+Our solution can be considered a multi-layered environment for users' services as it applies the notion of an autonomic system to its every layer on a which auto scaling can be executed -- application, container, service and cloud.
+To ensure meeting QoS requirements at the cloud level we use the recent concept of a utility-oriented federation of cloud environments (InterCloud).
+We present and compare current cloud solutions with the emphasis on their scaling capabilities. Finally, we demonstrate our preliminary results of conducted evaluation studies on the CloudSim toolkit.
+
 \keywords{cloud computing\and application scaling \and intercloud \and hybrid configuration}
@@
 \label{intro}
-Your text comes here. Separate text sections with
+Dizffko
 \section{Section title}
@@
 % For tables use
-\begin{table}
+%\begin{table}
 % table caption is above the table
-\caption{Please write your table caption here}
-\label{tab:1}       % Give a unique label
+%\caption{Please write your table caption here}
+%\label{tab:1}       % Give a unique label
 % For LaTeX tables use
-\begin{tabular}{lll}
-\hline\noalign{\smallskip}
-first & second & third  \\
-\noalign{\smallskip}\hline\noalign{\smallskip}
-number & number & number \\
-number & number & number \\
-\noalign{\smallskip}\hline
-\end{tabular}
-\end{table}
+%\begin{tabular}{lll}
+%\hline\noalign{\smallskip}
+%first & second & third  \\
+%\noalign{\smallskip}\hline\noalign{\smallskip}
+%number & number & number \\
+%number & number & number \\
+%\noalign{\smallskip}\hline
+%\end{tabular}
+%\end{table}
 

### 2013-09-13 article/article.tex
Add introduction and the shitty table

@@
 \usepackage{graphicx}
+\usepackage{amssymb}
+
 %
@@
 
-\title{Self-adaptive platform for scaling users' application deployed in a federation of cloud computing environments}
+\title{Cloud-SAP: Self-adaptive platform for scaling users' application deployed in a federation of cloud computing environments}
 
@@
 Unfortunately, existing cloud providers do not have tools and mechanisms that would enable dynamic load distribution among different data centers to meet aforementioned requirements.
-In this article we want to outline the architecture of a self-adaptive platform which facilitates scalable provisioning users' applications and fulfills QoS needs under variable conditions.
+In this article we want to outline the architecture of a self-adaptive platform (\emph{Cloud-SAP}) which facilitates scalable provisioning users' applications and fulfills QoS needs under variable conditions.
 Our solution can be considered a multi-layered environment for users' services as it applies the notion of an autonomic system to its every layer on a which auto scaling can be executed -- application, container, service and cloud.
@@
 \label{intro}
-Dizffko
+Cloud computing, a relatively new concept which denotes offering computing utilities on demand in a pay-as-you-go manner like any other services available in today's society, has already started the process of transforming IT industry, making an impact on the way the software/hardware is produced and perceived.
+Because of its widespread presence in the IT community, this model makes computing resources more and more attractive for broader audience, especially in diverse branches of business, e.g.
+pharmaceutical or construction.
+
+There are a few noticeable things which can be considered innovative in the way the services are provided.
+Firstly, clients of such services have no knowledge of the physical location of resources they use.
+Secondly, they are under illusions of infinite resources that are available.
+What is more, they pay providers only when they really use their services -- this means they do not need to invest heavily in building and maintaining hardware infrastructure.
+Last but not least, in order to ease dynamic resource management and manipulation, cloud providers use various virtualization technologies which operate on different levels, e.g.
+server, network or storage.
+
+\subsection{Service Models}
+Cloud computing classified the delivered services to three categories: Infrastructure as a Service (IaaS), Platform as a Service (PaaS) and Software as a Service (SaaS).
+They provide the consumer with infrastructure, deployment platform (application ecosystem) and software respectively.
+The just defined terms are commonly referred to in Cloud community as \emph{service models}.
+
+\subsection{Deployment models and cloud federation}
+
+\subsection{Application scaling}
+
+\section{Cloud providers -- comparison}
+
+\begin{table*}[!ht]
+  \renewcommand{\arraystretch}{2}
+\begin{tabular}{ l | c | c | c c }
+  \hline 
+  \multicolumn{1}{l}{} & \multicolumn{1}{l}{\textbf{Horizontal scaling}} & \multicolumn{1}{l}{\textbf{Vertical scaling}} & \textbf{Application / resource tuning} \\ \hline
+
+  \multicolumn{1}{l}{\textbf{Infrastructure provider}} & \multicolumn{4}{ l }{} \\ \hline
+
+Carina & \checkmark & $\times$ & $\times$ & \\ \hline
+
+OneFlow & \checkmark & $\times$ & $\times$ & \\ \hline
+
+AWS EC2 & \checkmark & $\times$ & $\times$ & \\ \hline
+
+\multicolumn{1}{l}{\textbf{Platform provider}} & \multicolumn{4}{ l }{} \\ \hline
+
+CloudFoundry & $\times$ & $\times$ & $\times$  &\\ \hline
+
+OpenShift & \checkmark & $\times$ & $\times$  &\\ \hline
+
+AppEngine & \checkmark & $\times$ & $\times$  &\\ \hline
+
+Azure & \checkmark & $\times$ & $\times$  &\\ \hline
+
+Heroku & \checkmark & $\times$ & $\times$ & \\ \hline
+\end{tabular}
+
+\caption{Comparision of cloud providers scaling capabilites}
+\label{tab:cloud-providers-scaling}
+
+\end{table*}
+
+\section{Architecture of SAP-Cloud}
+
+\section{Preliminary test results}
+
+\section{Summary and conclusion}
+
 \section{Section title}

### 2013-09-13 article/article.tex
Provide \titlerunning
Add intercloud section

@@
 %\titlerunning{Short form of title}        % if too long for running head
+\titlerunning{Cloud-SAP: Self-adaptive platform for scaling users' application in a federation of cloud computing environments}
 
@@
 They provide the consumer with infrastructure, deployment platform (application ecosystem) and software respectively.
-The just defined terms are commonly referred to in Cloud community as \emph{service models}.
+The just defined terms are commonly referred to in the Cloud community as \emph{service models}.
 
-\subsection{Deployment models and cloud federation}
+\subsection{Cloud federation -- InterCloud}
+However hard cloud providers would try to fulfill even the most demanding requirements from their clients, it is virtually impossible that resources of one of them would be satisfactory for all users -- firstly, because their resources are finite; secondly, there are providers which specialize in providing services to only one, specific types of clients and which raised the quality of their resources to such a level that other cloud computing companies cannot compete with.
+In particular, the last factor creates a need for a solution that would take into account different QoS expectations from the clients and map them onto the best resources offered by cloud providers.
+As it is impossible to predict the geographic distribution of consumers, the load coordination should be executed automatically and resources utilization should change in dependence of current load.
+
+Such a model has been devised and named \emph{InterCloud}.
+In this solution resources are managed in a market-oriented fashion that enables dynamic regulation of the supply and demand of resources and promotes the mechanisms for their allocation that would take into account their priorities and levels of utilization.
 
@@
 
-\section{Cloud providers -- comparison}
+\section{Who is Who in cloud provisioning}
 

### 2013-09-13 thesis/src/chapters/chapter-4.tex
Amendment in the chapter 4

@@
 What is more, in order to ensure high availability of their services, customers do not want to confine themselves to only one provider -- in the best scenario they want their applications (or their logical parts, such as persistence layer) to be spanned across different providers and be able to cooperate with one another at the same time.
-This leads to the concept of a \emph{hybrid cloud}\cite{MeGr11} -- the case in which the cloud is a composition of two or more distinct infrastructures which are unique entities, but there are technological means that make it possible to port data and applications among them.
+Additionally, due to privacy concerns of the sensible data, companies are reluctant to put it in the public cloud storage.
+All these factors lead to the concept of a \emph{hybrid cloud}\cite{MeGr11} -- the case in which the cloud is a composition of two or more distinct infrastructures which are unique entities, but there are technological means that make it possible to port data and applications among them.
 

### 2013-09-13 article/article.tex
Add a stub section of scaling in the article

@@
 \begin{abstract}
-In order to retain the lead among its competitors, a software company needs to create a distributed system which is spanned across different geographical locations, vulnerable to sudden variations in demand and has remarkable QoS requirements.
+In order to retain the lead among its competitors, a software company needs to create a distributed system which is spanned across different geographical locations, vulnerable to sudden variations in demand and has remarkable Quality-of-Service requirements.
 Following current trends it chooses Cloud as a deployment platform.
@@
 \subsection{Application scaling}
+Scalability is a widely accepted measure for improving application performance, consequently increasing offered Quality-of-Service.
+Despite the lack of a clear definition, in this article we consider this term an ability to work without delays and unproductive resource consumption at light, moderate, or heavy loads while making good use of available resources (in some works it is referred to as \emph{load scalability}).
+
+Having clarified the term we use, the case of scalability can be reduced to a question: \emph{what kind of resources can we add, remove or change to improve application's performance.}
+
+Resources can be managed in two, common agreed-upon ways: \emph{vertical} and \emph{horizontal}.
+By \emph{vertical} we understand changing the number of nodes a system is comprised, such as servers in the context of distributed application.
+\emph{Horizontal}, on the other hand, means changing the capacity of a single node in a system, e.g.
+adding additional memory, CPU, storage, etc.
+
+\paragraph{Horizontal scaling (scaling out)} 
+\paragraph{Vertical scaling (scaling up)}
 

### 2013-09-13 thesis/src/chapters/chapter-2.tex
refactored table

@@
 \end{itemize}
-Although, all of the aforementioned aspects are crucial for functioning of any healthy application, our work focuses only on the first type of scalability.
+Although, all of the aforementioned aspects are vital for any application, our work focuses solely on the first type of scalability.
 The reasoning behind this statement is that, while all of these scalability types lies in direct responsibility of an application developer, the load scalability can be additionally improved by adding additional resources to a system.
-This brings us to a question what kind of resources are used by an application or more appropriately to context of this dissertation: \textit{what kind of resources can we add to improve application performance?} Required resources varies from an application to an application.
+This brings us to a question what kind of resources are used by an application or more appropriately in context of this dissertation: \textit{what kind of resources can we add to improve application performance?} Required resources varies from an application to an application.
 However, among the most common ones we can distinguish:
@@
 With all that said, there is no silver bullet - not matter what underlying mechanism platform provider decides to use, the application developer is still responsible for creating an application with scaling in-mind.
-This statement has been already proven in 1967 by Amdahl law, which in short states that sequential component of parallel algorithms impacts efficiency for a sufficiently large number processors \cite{Am67} as shown in Figure \ref{ch2:amdahl-law}.
+This statement has been already proven in 1967 by Amdahl law, which in short states that sequential component of a parallel algorithm impacts efficiency for a sufficiently large number processors \cite{Am67} as shown in Figure \ref{ch2:amdahl-law}.
 In other words, adding supplementary resources to a poorly written application (i.e.
@@
 
+\newpage
 \section{Horizontal scaling}
@@
 In practice, this algorithm is commonly used with proxy caches and anti-virus proxies in order to maximize the cache hit rate.
-\end{itemize}
-
-Situation gets further complicated when considering real-world web application that sends user information using cookies, what imposes requirement on load-balancer for session stickiness \cite{StBaMa11}.
-
-Below are presented one of the most common load-balancers, including their key performance features.
-
-\subsubsection*{HAProxy}
-
-HAProxy \cite{HAProxy} is a load-balancer initially written by Willy Tarreau.
-Noticeably, it's used by OpenShift \cite{OpenShift} to distribute load among gears \cite{OpenShiftScaling}.
-Its key features include:
-\begin{itemize}
-	\item a single-process, event-driven model reduces the cost of context switch and the memory usage
-	\item O(1) event checker
-	\item Single-buffering without copying data between reads and writes
-	\item Zero-copy forwarding
-	\item Optimized HTTP header analysis : headers are parsed an interpreted on the fly
-\end{itemize}
-
-Requests are dispatched accordingly to chosen algorithm:
-\begin{itemize}
-  \item \textit{round-robin scheduling}
-  \item \textit{least connection}
-  \item \textit{source routing}
-  \item \textit{URI hashing}
-\end{itemize}
-
-\subsubsection*{BIG-IP Local Traffic Manager (LTM)}
-BIG-IP Local Traffic Manager is a solution offered by F5 \cite{F5}.
-Although LTM is a hardware solution, omitted in this section, it also has its virtualized counterpart.
-Key features offered by LTM are \cite{LTM}:
-\begin{itemize}
-  \item managing at application services level rather than at individual devices and objects
-  \item scripting language that allows administrator to intercept, inspect, transform, and direct application traffic
-  \item built-in firewall protection, application security, and access control
-  \item real-time protocol and traffic management decisions
-\end{itemize}
-
-\subsubsection*{Apache HTTPD}
-Apache HTTPD \cite{ApacheHTTPD} is a popular HTTP server.
-When enhanced with additional modules, it can behave as a proxy or load-balancer.
-Over the time, there were serveral attempts to develop such modules: mod\_jk \cite{ApacheModJk}, mod\_proxy\_balancer \cite{ApacheModProxyBalancer}, to name a few.
-While the former is AJP13 oriented, the latter supports different protocols: HTTP, FTP and AJP13, hence, it is described here.
-
-Among the mod\_proxy\_balancer features we may distinguish:
-\begin{itemize}
- \item support for different load-balancing algorithms:
-    \begin{itemize}
  \item \textit{request counting algorithm} - load is distributed the requests among the various workers, ensuring that each gets their configured share of the number of requests
@@
 
- \item support for session stickiness by using cookies and URL encoding.
-This approach \cite{ApacheModProxyBalancer} avoids unequal load distribution if clients are hidden behind proxies and stickyness errors when a client uses a dynamic IP address that changes during a session
-\end{itemize}
+Situation gets further complicated when considering real-world web application that sends user information using cookies, what imposes requirement on load-balancer for session stickiness \cite{StBaMa11}.
 
@@
 
+\subsection{Load-balancer comparision}
+
+While there are many load-balancers available on the market, following are credited to be most popular:
+\begin{itemize}
+ \item \textbf{HAProxy} \cite{HAProxy} - load-balancer initially written by Willy Tarreau. Noticeably, it's used by OpenShift \cite{OpenShift} to distribute load among gears \cite{OpenShiftScaling}
+ \item \textbf{BIG-IP Local Traffic Manager (LTM)} - solution offered by F5 \cite{F5}. Although LTM is a hardware solution, omitted in this section, it also has also its virtualized counterpart.
+ \item \textbf{Apache HTTPD} \cite{ApacheHTTPD} - popular HTTP server.
+When enhanced with additional modules, it can behave like a proxy or load-balancer.
+Over the time, there were serveral attempts to develop such modules: mod\_jk \cite{ApacheModJk}, mod\_proxy\_balancer \cite{ApacheModProxyBalancer}, to name a few.
+While the former is purely AJP13 oriented, the latter supports different protocols: HTTP, FTP and AJP13.
+As a consequnce, only mod\_proxy\_balancer was taken into account during comparision.
+\end{itemize}
+
+Table \ref{tab:load-balancer-comparision} presents they key performance features and algorithm used to schedule requests.
+
+\begin{table}[!htbp]
+\begin{tabularx}{\textwidth}[]{ X  X  X }
+\specialrule{.1em}{.05em}{.05em}
+
+  & \textbf{Performance features} & \textbf{Scheduling algorithms} \\
+\specialrule{.1em}{.05em}{.05em}
+
+HAProxy & 
+-- a single-process, event-driven model reduces the cost of context switch and the memory usage
+
+-- O(1) event checker
+
+-- single-buffering without copying data between reads and writes
+
+-- zero-copy forwarding
+
+-- optimized HTTP header analysis: headers are parsed an interpreted on the fly
+
+&
+-- round-robin scheduling
+
+-- least connection
+
+-- source routing
+
+-- URI hashing
+\\ \hline
+
+BIG-IP Local Traffic Manager &
+-- managing at application services level rather than at individual devices and objects
+
+-- scripting language that allows administrator to intercept, inspect, transform, and direct application traffic
+
+-- built-in firewall protection, application security, and access control
+
+-- real-time protocol and traffic management decisions
+&
+
+\\ \hline
+
+Apache HTTPD & 
+-- support for session stickiness by using cookies and URL encoding.
+This approach \cite{ApacheModProxyBalancer} avoids unequal load distribution if clients are hidden behind proxies and stickyness errors when a client uses a dynamic IP address that changes during a session
+& 
+-- request counting algorithm
+
+-- weighted traffic counting algorithm
+
+-- pending request counting algorithm
+\\ \hline
+
+\end{tabularx}
+
+\caption{Comparision of load balancers}
+\label{tab:load-balancer-comparision}
+
+\end{table}
+
+\newpage
 \section{Vertical scaling}
@@
 
-\subsection*{Virtual machine resizing}
+\subsection{Virtual machine resizing}
 
 \begin{table}[!htbp]
-\begin{tabularx}{\textwidth}{| l | X | X | X |}
-\hline
+\begin{tabularx}{\textwidth}{ l  X  X  X }
+\specialrule{.1em}{.05em}{.05em} 
  & \textbf{Memory} & \textbf{CPU} & \textbf{Disk} \\
-\hline 
+\specialrule{.1em}{.05em}{.05em}
+
 KVM &
--  & 
-dynamic pinning CPU to a specific virtual machine (depending on underlying hardware)
   & 
-adding a disk to a LVM group
+-- dynamic pinning CPU to a specific virtual machine (depending on underlying hardware)
+& 
+-- adding a disk to a LVM group
 
@@
 Xen & 
-- changing the amount of host physical memory assigned to virtual machine without rebooting it
+-- changing the amount of host physical memory assigned to virtual machine without rebooting it
 
-- start additional virtual machines on a host whose physical memory is currently full, by automatically reducing the memory allocations of existing virtual machines in order to make space
+-- start additional virtual machines on a host whose physical memory is currently full, by automatically reducing the memory allocations of existing virtual machines in order to make space
 &
-dynamic pinning CPU to a specific virtual machine (depending on underlying hardware)
+-- dynamic pinning CPU to a specific virtual machine (depending on underlying hardware)
 &
-dynamic block attaching, adding a disk to a LVM group
+-- dynamic block attaching, adding a disk to a LVM group
 
@@
 VMware ESX &
-hot-plugging memory, ex. using VMware vSphere
+-- hot-plugging memory, ex. using VMware vSphere
 &
-hot-plugging CPU, ex. using VMware vSphere
+-- hot-plugging CPU, ex. using VMware vSphere
 &
-adding additional disks to existing virtual machine
+-- adding additional disks to existing virtual machine
 
@@
 OpenVZ &
-configurable via user beancounters
+-- configurable via user beancounters
 &
-configurable via user beancounters 
+-- configurable via user beancounters 
 & 
-configurable via user beancounters
+-- configurable via user beancounters
 \\ \hline
@@
 
-\subsection*{Virtual machine replacement}
+\subsection{Virtual machine replacement}
 As it was highlighted in previous section, reasoning behind virtual machine replacement is that, in case when dynamic resizing is not possible, a new virtual machine with a desired configuration can be provisioned and replace the old one.
@@
 \begin{table}[!htbp]
-\begin{tabularx}{\textwidth}[]{ X  c  c  c }
+\begin{tabularx}{\textwidth}[]{ X  X  X  X}
 \specialrule{.1em}{.05em}{.05em}
 
-  & \textbf{Horizontal scaling} & \textbf{Vertical scaling} & \textbf{Application / resource tuning} \\
+  & \textbf{Horizontal scaling} & \textbf{Vertical scaling} & \textbf{Application tuning} \\
 \specialrule{.1em}{.05em}{.05em}

### 2013-09-13 thesis/src/chapters/chapter-3.tex
refactored table

@@
 \section{Providers comparision}
+
+Table \ref{tab:cloud-providers-adaptivity} summaries chapter with approaches to adaptivity taken by different cloud providers.
+Section 'triggered action' is omited for brevity - it was described extensively in previous chapter.
+
+\begin{table}[!htbp]
+\begin{tabularx}{\textwidth}[]{ X  X X }
+\specialrule{.1em}{.05em}{.05em}
+
+  & \textbf{Policies} & \textbf{Data analysis} \\
+\specialrule{.1em}{.05em}{.05em}
+
+\multicolumn{3}{ l }{\textbf{Infrastructure provider}} \\
+\specialrule{.1em}{.05em}{.05em}
+
+Carina & 
+- time frame based
+
+- expression based (only for CPU)
+&
+- threshold model
+
+\\ \hline
+
+OneFlow & 
+- time frame based
+
+- expression based
+&
+- threshold model
+\\ \hline
+
+AWS EC2 & 
+- time frame based
+
+- expression based, (CloudWatch alarm)
+&
+- threshold model
+\\ \hline
+
+\multicolumn{3}{ l }{\textbf{Platform provider}} \\
+\specialrule{.1em}{.05em}{.05em}
+
+CloudFoundry & $\times$ & $\times$ \\ \hline
+
+OpenShift & 
+$\times$ &
+single built-in threshold model 
+\\ \hline
+
+AppEngine & ? & ? \\ \hline
+
+Azure & ? & ? \\ \hline
+
+Heroku & ? & ? \\ \hline
+\end{tabularx}
+
+\caption{Comparision of cloud providers approach to adaptivity}
+\label{tab:cloud-providers-adaptivity}
+
+\end{table}

### 2013-09-13 thesis/src/chapters/chapter-2.tex
Spellings

@@
 \begin{itemize}
- \item \textit{round-robin scheduling} - request are sent to sucessive nodes, according to their weights.
+ \item \textit{round-robin scheduling} - request are sent to successive nodes, according to their weights.
 This algorithm is fairest when the server's processing time remains equally distributed \cite{HaProxyDoc}
@@
 When enhanced with additional modules, it can behave like a proxy or load-balancer.
-Over the time, there were serveral attempts to develop such modules: mod\_jk \cite{ApacheModJk}, mod\_proxy\_balancer \cite{ApacheModProxyBalancer}, to name a few.
+Over the time, there were several attempts to develop such modules: mod\_jk \cite{ApacheModJk}, mod\_proxy\_balancer \cite{ApacheModProxyBalancer}, to name a few.
 While the former is purely AJP13 oriented, the latter supports different protocols: HTTP, FTP and AJP13.
-As a consequnce, only mod\_proxy\_balancer was taken into account during comparision.
+As a consequence, only mod\_proxy\_balancer was taken into account during comparision.
 \end{itemize}
 
-Table \ref{tab:load-balancer-comparision} presents they key performance features and algorithm used to schedule requests.
+Table \ref{tab:load-balancer-comparison} presents they key performance features and algorithm used to schedule requests.
 
@@
 
-\caption{Comparision of load balancers}
-\label{tab:load-balancer-comparision}
+\caption{Comparison of load balancers}
+\label{tab:load-balancer-comparison}
 
@@
 \end{tabularx}
-\caption{Comparision of hypervisors resizing capabilities}
+\caption{Comparison of hypervisors resizing capabilities}
 \label{tab:hypervisors-resizing}
@@
 
-\section{Providers comparision}
+\section{Providers comparison}
 
-Table \ref{tab:cloud-providers-scaling} presents a summary of cloud provider scaling capabilites.
+Table \ref{tab:cloud-providers-scaling} presents a summary of cloud provider scaling capabilities.
 Interestingly, all of them are focused solely on horizontal scaling, ignoring advantages offered by a fine-grained approach to scaling that leverage scaling up and application tuning.
@@
 
-\caption{Comparision of cloud providers scaling capabilites}
+\caption{Comparison of cloud providers scaling capabilities}
 \label{tab:cloud-providers-scaling}

### 2013-09-13 thesis/src/chapters/chapter-3.tex
Spellings

@@
 
-The remaining of this chapter describes crucial elemenets that compose elasticity controler: policies, data analysis, triggered actions and presents a comparision of cloud providers.
+The remaining of this chapter describes crucial elements that compose elasticity controller: policies, data analysis, triggered actions and presents a comparison of cloud providers.
 
@@
 
-Policy denotes a condition which, when satisfied, triggers an action that is supposed to harness cloud instance in a way that future evalutions of condition will be unsuccessfull.
-Typically condition itself is accompanied by a minimal and maximal number of node instances, allowing for ensuring minimal QoS and controling maximal costs.
+Policy denotes a condition which, when satisfied, triggers an action that is supposed to harness cloud instance in a way that future evaluations of condition will be unsuccessful.
+Typically condition itself is accompanied by a minimal and maximal number of node instances, allowing for ensuring minimal QoS and controlling maximal costs.
 Currently, industry leaders supports two main kind of policies \cite{AmazonAutoScaling}:
@@
 \section{Data analysis}
-Having policies defined, their conditions are evaluated against data aquired from sensors.
+Having policies defined, their conditions are evaluated against data acquired from sensors.
 In a simplest case this evaluation can be based on a Threshold Model \cite{LiWoZh05}, which defines a valid range.
@@
 \begin{itemize}
-  \item \textit{integer programming} - autoscaling is reduced to server interger programming problems, which aims to minimize the cost or maximize the computing power with either computing power constraints or
+  \item \textit{integer programming} - auto scaling is reduced to server integer programming problems, which aims to minimize the cost or maximize the computing power with either computing power constraints or
 budget constraints \cite{MaLiHu10}
   \item \textit{prediction error correction} - \cite{ShSuGuWi11}
-  \item \textit{linearized dynamic control} - linearized correction model is based on control equations \cite{AbShBh02}:
+  \item \textit{linearised dynamic control} - linearised correction model is based on control equations \cite{AbShBh02}:
 
@@
 
-  \item \textit{markov decision process model} - computes optimal reaction to state changes by using obsercation on the system with assumptions about the rate of changes expected in the future \cite{AbWo02}
+  \item \textit{markov decision process model} - computes optimal reaction to state changes by using observation on the system with assumptions about the rate of changes expected in the future \cite{AbWo02}
 \end{itemize}
@@
 \section{Triggered actions}
-Actions that are being triggered by a elasticity controller are focused on application scaling. Problem of scaling an applcication is described in detail in previous chapter.
+Actions that are being triggered by a elasticity controller are focused on application scaling. Problem of scaling an application is described in detail in previous chapter.
 
-\section{Providers comparision}
+\section{Providers comparison}
 
 Table \ref{tab:cloud-providers-adaptivity} summaries chapter with approaches to adaptivity taken by different cloud providers.
-Section 'triggered action' is omited for brevity - it was described extensively in previous chapter.
+Section 'triggered action' is omitted for brevity - it was described extensively in previous chapter.
 
@@
 
-\caption{Comparision of cloud providers approach to adaptivity}
+\caption{Comparison of cloud providers approach to adaptivity}
 \label{tab:cloud-providers-adaptivity}

### 2013-09-13 thesis/src/chapters/chapter-3.tex
Spellings
Added two more models

@@
 budget constraints \cite{MaLiHu10}
-  \item \textit{prediction error correction} - \cite{ShSuGuWi11}
+  \item \textit{burst based padding} - employs a signal processing technique based on fast Fourier transform, burst pattern is extracted and used to calculate a padding value.
+Coefficients that represents the amplitude of each frequency component are used to calculate burst density.
+Depending of that value (i.e.
+is higher than 50\%) appropriate percentile of the burst values are used \cite{ShSuGuWi11}
+  \item \textit{remedial padding} - padding errors are being recorded and used in successive padding evaluations.
+In other words, let $e_1, e_2, ...
+e_k$ denote the recent prediction errors, next the weighted moving average is calculated.
+Actual applied padding is either padding itself or weighted average, depending which one is greater \cite{ShSuGuWi11}
   \item \textit{linearised dynamic control} - linearised correction model is based on control equations \cite{AbShBh02}:

### 2013-09-13 thesis/src/chapters/chapter-3.tex
Added description of oneflow, carina, aws

@@
 \begin{itemize}
- \item \textit{expression based} - allows to define how you to scale application in response to changing conditions, which can include factors such as memory, cpu usage, cost or some indirect, calulated metrics
- \item \textit{scheduled} - allows you to scale your application in response to predictable load changes.
+ \item \textit{expression based} - allows to define how you to scale application in response to changing conditions, which include factors such as memory, CPU usage, cost or some indirect, calculated metrics
+ \item \textit{scheduled} - allows to scale an application in response to predictable load changes.
 For example, traffic increases during the weekends and decreases on working days.
-Hence, we may use that predictable traffic patterns to scale application based on current time.
+Hence, that predictable traffic patterns is used to scale application based on current time.
 \end{itemize}
@@
 Technically, policies are expressed in some human-readable format such as JSON, XML as it is in case of AWS EC2 or custom expression used for example by Carina.
-Appendinx \ref{app:scaling-policies} presents example configuration used by AWS E2 Auto-Scaling.
+Appendix \ref{app:scaling-policies} presents example configuration used by AWS E2 Auto-Scaling.
 
@@
 \section{Triggered actions}
-Actions that are being triggered by a elasticity controller are focused on application scaling. Problem of scaling an application is described in detail in previous chapter.
+Actions that are being triggered by a elasticity controller are focused on application scaling. Previous chapter described that problem in detail.
 
@@
 Carina & 
-- time frame based
+-- time frame based
 
-- expression based (only for CPU)
+-- expression based (only for CPU)
 &
-- threshold model
+-- threshold model that takes into account minimal and maximal permitted instances of an application as well as application priority
 
@@
 
-OneFlow & 
-- time frame based
+OneFlow 4.2 & 
+-- time frame based with customizable padding
 
-- expression based
+-- expression based build on custom language, where all vm's metrics are supported
+
+-- customizable adjustment padding, cooldown time
 &
-- threshold model
+-- threshold model
+
 \\ \hline
@@
 AWS EC2 & 
-- time frame based
+-- time frame based
 
-- expression based, (CloudWatch alarm)
+-- expression based, where expressions corresponds to a AutoScalingGroup
+
+-- actions are triggered by a CloudWatch alarms
+
+-- customizable adjustments paddings, types, cooldown time
 &
-- threshold model
+-- threshold model, takes into account minimal and maximal permitted instances of an application as well as application priority
 \\ \hline
@@
 OpenShift & 
-$\times$ &
-single built-in threshold model 
+- single built-in policy
+
+ &
+- single built-in threshold model that scales an application when CPU load is greater than 50\% for a given period
 \\ \hline

### 2013-09-13 article/article.tex
Finish the scaling section in the article

@@
 \usepackage{amssymb}
+\usepackage{footnote}
+\usepackage{array}
+\newcolumntype{C}{>{\centering\arraybackslash}p{3.4cm}}
+\newcolumntype{L}{>{\arraybackslash}p{2.5cm}}
 
@@
 
-\author{Dariusz Chrząścik          \and
+\author{Dariusz M. Chrząścik          \and
         Radosław D. Morytko
@@
 \institute{
-  D. Chrząścik \\
+  D.M. Chrząścik \\
   Chief Technology Officer at Software Mind SA \\
@@
 
-Having clarified the term we use, the case of scalability can be reduced to a question: \emph{what kind of resources can we add, remove or change to improve application's performance.}
+Having clarified the term we use, the case of scalability can be reduced to a question: \emph{what kind of resources can we add, remove or change to improve application's performance.} However, it should be stressed that the first step towards effective performance of an application should be its proper tuning and configuration.
+Otherwise, no matter what kind of and how much resources is added it will not influence the overall application's utilisation.
 
 Resources can be managed in two, common agreed-upon ways: \emph{vertical} and \emph{horizontal}.
-By \emph{vertical} we understand changing the number of nodes a system is comprised, such as servers in the context of distributed application.
+By \emph{vertical} we understand changing the number of nodes a system is comprised, such as servers in the context of a distributed application.
 \emph{Horizontal}, on the other hand, means changing the capacity of a single node in a system, e.g.
@@
 
-\paragraph{Horizontal scaling (scaling out)} 
-\paragraph{Vertical scaling (scaling up)}
+\paragraph{Horizontal scaling (scaling out)} As it was stated in the previous paragraph, horizontal scaling refers to managing the number of instances the application runs on.
+Since provisioning a raw or subtracting an existing virtual machine is a common process which is performed by a hypervisor, the much more interesting one is the configuration of a node that would enable its cooperation among existing ones, e.g.
+the added instance should be used as a load balancer of the user's service.
+
+\paragraph{Vertical scaling (scaling up)} Vertical scaling concentrates on managing the capacity of a single node.
+Among the most common resources that can be changed in runtime of a machine are (a) CPU, (b) memory and (c) storage.
 
@@
 
-\begin{table*}[!ht]
+\begin{table*}[ht]
+  \renewcommand{\arraystretch}{2}
+  \begin{tabular}{p{2.5cm} | *{3}{C}}
+    \multicolumn{1}{c}{} & \textbf{Memory} & \textbf{CPU} & \textbf{Disk} \\ \hline
+
+   \textbf{KVM} & 
+   - &
+   dynamic pinning CPU to a specific VM \footnotemark[1] &
+   adding a disk to a LVM group 
+   \\ \hline
+
+   \textbf{Xen} &
+  - changing the amount of host physical memory assigned to virtual machine without rebooting it
+
+  - start additional virtual machines on a host whose physical memory is currently full, by automatically reducing the memory allocations of existing virtual machines in order to make space &
+  dynamic pinning CPU to a specific VM \footnotemark[1] &
+  dynamic block attaching, adding a disk to a LVM group
+  \\ \hline
+
+  \textbf{VMware ESX} &
+hot-plugging memory \footnotemark[2] & 
+hot-plugging CPU \footnotemark[2]  &
+adding additional disks to an existing VM
+\\ \hline
+
+\textbf{OpenVZ} & configurable via beancounters & configurable via beancounters & configurable via beancounters\\ \hline
+
+%OneFlow & \checkmark & $\times$ & $\times$ & \\ \hline
+
+%AWS EC2 & \checkmark & $\times$ & $\times$ & \\ \hline
+
+\end{tabular}
+
+\caption{Comparision of hypervisors' resizing capabilites}
+\label{tab:hypervisors-resizing}
+
+\end{table*}
+\footnotetext[1]{Depending on the underlying hardware}
+\footnotetext[2]{For example using VMware vSphere}
+
+\begin{table*}[ht]
   \renewcommand{\arraystretch}{2}

### 2013-09-13 thesis/src/chapters/chapter-3.tex
Added diagrams

@@
 In short, platform adaptivity adds a auto-scaling features to a solution offered by Platform-as-a-Service provider.
-Key concept is to have a Elasticity Controller which gathers probes from virtual machines and uses that knowledge to execute appropriate action on Cloud instance, indirectly modifying consecutive probes \cite{VaRoBu11}.
-This concept illustrates diagram (IMAGE).
+Key concept is to have a Elasticity Controller which gathers probes from virtual machines and uses that knowledge to execute appropriate action on cloud instance, indirectly modifying consecutive probes \cite{VaRoBu11}.
+This concept illustrates diagram \ref{ch3:elasticity-controller}.
+
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics{chapter-3/elasticity-controller}
+  \end{center}
+  \caption{Elasticity controller}
+  \label{ch3:elasticity-controller}
+\end{figure}
 
@@
 
-Heroku & ? & ? \\ \hline
+Heroku & $\times$ & $\times$ \\ \hline
 \end{tabularx}

### 2013-09-14 article/article.tex
Change table styles
Add the state-of-the-art in cc providers in the article

@@
 
-\section{Who is Who in cloud provisioning}
+\paragraph{Hypervisors} Being able to scale resources vertically requires hypervisors to have such possibilities.
+Table \ref{tab:hypervisors-resizing} presents resizing capabilities of common hypervisors.
 
@@
   \renewcommand{\arraystretch}{2}
-  \begin{tabular}{p{2.5cm} | *{3}{C}}
+  \begin{tabular}{p{4cm} | *{3}{C}}
     \multicolumn{1}{c}{} & \textbf{Memory} & \textbf{CPU} & \textbf{Disk} \\ \hline
@@
 
+\section{Who is Who in cloud provisioning}
+In the context of this article, the main point of interest are providers which offer features related to resource management, monitoring and utilization.
+In general, current key providers include Amazon \emph{Web Services}, \emph{Microsoft Azure}, Google \emph{AppEngine} and \emph{Heroku}.
+Despite not all of them operate on the same level of abstraction, it is possible to compare them as they expose functionalities which can be classified as PaaS.
+For example, on top of its infrastructure services, Amazon has such products as \emph{Elastic Load Balancer}, \emph{Cloud Watch} and \emph{Auto Scaling} which, to some extent, matches our needs.
+
+Table \ref{tab:cloud-providers-scaling} presents comparison of some Cloud providers with the emphasis on their scaling capabilities.
+Interestingly, all of them are focused solely on horizontal scaling, ignoring advantages offered by a fine-grained approach to scaling that leverage scaling up and application tuning.
+
 \begin{table*}[ht]
   \renewcommand{\arraystretch}{2}
-\begin{tabular}{ l | c | c | c c }
+  \begin{tabular}{ p{4cm} | C | C | C }
   \hline 
@@
 
-  \multicolumn{1}{l}{\textbf{Infrastructure provider}} & \multicolumn{4}{ l }{} \\ \hline
+  \multicolumn{1}{l}{\textbf{Infrastructure provider}} & \multicolumn{3}{ l }{} \\ \hline
 
-Carina & \checkmark & $\times$ & $\times$ & \\ \hline
+Carina & \checkmark & $\times$ & $\times$ \\ \hline
 
-OneFlow & \checkmark & $\times$ & $\times$ & \\ \hline
+OneFlow & \checkmark & $\times$ & $\times$ \\ \hline
 
-AWS EC2 & \checkmark & $\times$ & $\times$ & \\ \hline
+AWS EC2 & \checkmark & $\times$ & $\times$ \\ \hline
 
-\multicolumn{1}{l}{\textbf{Platform provider}} & \multicolumn{4}{ l }{} \\ \hline
+\multicolumn{1}{l}{\textbf{Platform provider}} & \multicolumn{3}{ l }{} \\ \hline
 
-CloudFoundry & $\times$ & $\times$ & $\times$  &\\ \hline
+CloudFoundry & $\times$ & $\times$ & $\times$  \\ \hline
 
-OpenShift & \checkmark & $\times$ & $\times$  &\\ \hline
+OpenShift & \checkmark & $\times$ & $\times$  \\ \hline
 
-AppEngine & \checkmark & $\times$ & $\times$  &\\ \hline
+AppEngine & \checkmark & $\times$ & $\times$  \\ \hline
 
-Azure & \checkmark & $\times$ & $\times$  &\\ \hline
+Azure & \checkmark & $\times$ & $\times$  \\ \hline
 
-Heroku & \checkmark & $\times$ & $\times$ & \\ \hline
+Heroku & \checkmark & $\times$ & $\times$ \\ \hline
 \end{tabular}
@@
 
+\section{Adaptivity}
 \section{Architecture of SAP-Cloud}

### 2013-09-14 thesis/src/chapters/chapter-2.tex
Added diagrams
Added threshold-model chart

@@
 \begin{itemize}
-	\item \textit{horizontal scaling (scaling out)} - adding more nodes to a system, such as servers in context of distributed application
+	\item \textit{horizontal scaling (scaling out)} - adding more nodes to a system, such as servers in a context of distributed application
 	\item \textit{vertical scaling (scaling up)} - increasing capacity of a single node in a system, i.e. adding additional memory, CPU, storage, etc.
 \end{itemize}
-
-Having that said, common sense dictates that adding resources is only a part of the success - it should be accompanied by tuning application platform configuration.
-For example, adding supplementary CPUs without increasing thread pool size that handle requests makes a little sense.
-Similarly, in context of a Java application, we have to increase heap size, , to make a good use of extra memory.
-While importance of application tuning cannot be underestimated, its detailed analysis lies outside of the scope of this dissertation.
 
@@
 OpenShift leverages SELinux and cgroups) they are similar in nature to server scaling and supports both scaling up and out.
+Therefore, the remaining of this chapter is focused solely on server scaling, omitting network scaling as it lays outside of scope of this dissertation.
+
+Having that said, common sense dictates that adding resources is only a part of the success - it should be accompanied by tuning application platform configuration.
+For example, adding supplementary CPUs without increasing thread pool size that handle requests makes a little sense.
+Similarly, in context of a Java application, we have to increase heap size, to make a good use of extra memory.
+While importance of application tuning cannot be underestimated, its detailed analysis lies outside of the scope of this dissertation.
+Figure \ref{ch2:scalability-layers} presents different scalability layers and actions that can be taken at each level to improve application performance.
+
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics{chapter-2/scalability-layers}
+  \end{center}
+  \caption{Scalability layers}
+  \label{ch2:scalability-layers}
+\end{figure}
 
@@
 
-\newpage
 \section{Horizontal scaling}
@@
 
-KVM &
+KVM 1.2.0 &
   & 
@@
 \\ \hline
-Xen & 
+Xen 4.3 & 
 -- changing the amount of host physical memory assigned to virtual machine without rebooting it
@@
 \\ \hline
-VMware ESX &
+VMware ESX 5.1 &
 -- hot-plugging memory, ex. using VMware vSphere
@@
 \\ \hline
-OpenVZ &
+OpenVZ (kernel: 042) &
 -- configurable via user beancounters

### 2013-09-14 thesis/src/chapters/chapter-3.tex
Added threshold-model chart
Added azure, appengine

@@
 In cases when given metric violates that condition (i.e.
-value is either smaller than minimal or bigger than maximal acceptable) corresponding resource is properly adjusted.
-While trivial in its form, cases of AWS, OpenShift, Carina, OneCloud proves it is useful in a real-world scenarios.
-Having that said, more sophisticated algorithms also exists:
+value is either smaller than minimal or bigger than maximal acceptable) corresponding resource is properly adjusted - figure \ref{ch3:threshold-model} illustrates that idea.
+
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics{chapter-3/threshold-model}
+  \end{center}
+  \caption{Threshold model}
+  \label{ch3:threshold-model}
+\end{figure}
+
+While trivial in its form, cases of AWS, OpenShift, Carina, OneCloud proves it is useful in a real-world scenarios. Having that said, more sophisticated algorithms also exists:
 \begin{itemize}
@@
 OpenShift & 
-- single built-in policy
+-- single built-in policy
 
  &
-- single built-in threshold model that scales an application when CPU load is greater than 50\% for a given period
+-- single built-in threshold model that scales an application when CPU load is greater than 50\% for a given period
 \\ \hline
 
-AppEngine & ? & ? \\ \hline
+AppEngine &
 
-Azure & ? & ? \\ \hline
+-- built-in policy based on request queue length
+
+-- adjustable minimal, maximal number of application instances, pending latency
+
+ &
+-- queue-based, new instance is provisioned if queue length got too long 
+  \\ \hline
+
+Azure & 
+-- time frame based
+
+-- expression based, where expression can involve either CPU usage or Queue length
+
+-- customizable adjustments paddings, types, cooldown time
+&
+-- threshold model, that takes into account minimal and maximal allowed instances
+ \\ \hline
 

### 2013-09-14 article/article.tex
Add the test section in the article

@@
 %\authorrunning{Short form of author list} % if too long for running head
-
 \institute{
@@
 
+As the proposed platform has not been implemented yet, we decided to evaluate the proposed model using the \emph{CloudSim} framework.
+
+\paragraph{First test case} The aim of the first test scenario is to prove that the proposed system delivers better performance when compared to current, non market-oriented and self-adaptive solutions.
+Preconditions are that there are 3 cloud providers and one consumer who deployed their application and as a result of shortage of resources of the current provider, the application must be migrated to another one.
+This involves (a) creating a clone of the virtual machine at the other provider (b) provisioning VM with the technology stack which needs to be pre-installed to ensure that the application will run correctly (c) performing a migration -- installing the application and configuring it so that any it cooperates with the other parts, if any.
+
+There are 20 instances of the application. Each one is deployed on a separate instance of the VM in two scenarios: the one involving Cloud-SAP and the other without it.
+
+The results of the test are shown in table \ref{tab:test1-results} -- turn-around time for each instance of the application and the overall makespan for the service.
+Moreover, we added the row containing information of the collected revenue to stress the business aspect of the model.
+
+The test case shows that using of the proposed platform reduced the average turn-around time by more than 30 percent and total makespan by over 20 percent.
+This means a considerable improvement to the user's application performance.
+
+\begin{table*}[ht]
+  \renewcommand{\arraystretch}{2}
+  \begin{tabular}{ | C | C | C | C | }
+  \hline 
+  \textbf{Performance metrics} & \textbf{With Cloud-SAP} & {\textbf{Without Cloud-SAP}} & \textbf{Improvement} \\ \hline
+
+  Average turn-around time (s) &
+  1671 &
+  2217 &
+  over 30\% \\ \hline
+
+  Makespan (s) &
+  5721 &
+  7038 &
+  over 20\% \\ \hline
+
+  Revenue from scaling capabilities (\$) &
+  (depends on the pricing) &
+  - &
+  proportional to the number of deployed nodes \\ \hline
+\end{tabular}
+\caption{Results of the first test case scenario}
+\label{tab:test1-results}
+\end{table*}
+
+\paragraph{Second test case}
+The second test case aims at drawing attention to the potential income earned by adopting the proposed platform.
+There are two scenarios: the first one involves the depicted solution, the other one not.
+In both of them, two clients want to use the same cloud resources of the one provider, but each of them has different QoS requirements and different budget -- client B can spend more money on the same service than the client A.
+Resources are limited -- the provider can provision as many as 18 virtual machines.
+
+The test scenario is as follows: in the beginning, client A and B have instantiated 10 and 3 virtual machines respectively.
+As the time goes by, there is a variation in load in customers' applications and to handle it additional resources need to be provided.
+In the scenario which does not assume usage of the Cloud-SAP, requests are not prioritized which results in generating lower income, what is shown in the figure \ref{fig:hor-scaling-no-prioritization}.
+In the second one is quite on the contrary -- as requests are not considered the same, the platform chooses the most profitable options for the provider (see figure \ref{fig:hor-scaling-with-prioritization}).
+
+This could raise a lot of question regarding handling the need of resources by the client A.
+It is worth noticing that because of the adoption of market-oriented federation, this request can be caught and processed by another provider.
+The main disadvantage in this case is the time burden.
+
+\begin{figure}
+  \includegraphics[width=0.47\textwidth]{figures/request-priority-revenue-1}
+\caption{Horizontal scaling without  requests' prioritization}
+\label{fig:hor-scaling-no-prioritization}       % Give a unique label
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=0.47\textwidth]{figures/request-priority-revenue-2}
+\caption{Horizontal scaling with requests' prioritization}
+\label{fig:hor-scaling-with-prioritization}       % Give a unique label
+\end{figure}
+
 \section{Summary and conclusion}

### 2013-09-15 article/article.tex
Add the 'conclusion' section in the article

@@
 \section{Summary and conclusion}
+Cloud computing emerged as a completely new paradigm for providing users with IT services in a pay-as-you-go manner.
+As the time goes by, the clients expectations and requirements in terms of providing services with stringent Quality-of-Service requirements have raised and forced Cloud computing scientists to invent and design new solutions capable of satisfying them.
+
+In this paper we gave a brief overview of current cloud providers capabilities with the stress on auto-scaling features.
+Then we outlined the architecture of Cloud-SAP platform which facilitates autonomic and scalable provisioning of users' applications.
+The results of the conducted tests indicate that this model has enormous business potential for both cloud providers and clients.
+
+We hope our work will be of a great benefit to computer scientists and engineers involved in designing and building cloud solutions and its outcome will make progress in enlightening new problems in the area of combining autonomous systems and market-oriented cloud federation.
+
+Our future work will concentrate on implementing the proposed solution and testing it in real-world hardware against business use cases.
 

### 2013-09-16 article/article.tex
Add some bibligraphy to the article

@@
 \usepackage{footnote}
+\usepackage{url}
 \usepackage{array}
@@
 Because of its widespread presence in the IT community, this model makes computing resources more and more attractive for broader audience, especially in diverse branches of business, e.g.
-pharmaceutical or construction.
+pharmaceutical or construction \cite{PharmaComputation}.
 
@@
 Firstly, clients of such services have no knowledge of the physical location of resources they use.
-Secondly, they are under illusions of infinite resources that are available.
+Secondly, they are under illusions of infinite resources that are available \cite{ScalingInDaCloud}.
 What is more, they pay providers only when they really use their services -- this means they do not need to invest heavily in building and maintaining hardware infrastructure.
@@
 They provide the consumer with infrastructure, deployment platform (application ecosystem) and software respectively.
-The just defined terms are commonly referred to in the Cloud community as \emph{service models}.
+The just defined terms are commonly referred to in the Cloud community as \emph{service models} \cite{NIST}.
 
@@
 
-Such a model has been devised and named \emph{InterCloud}.
-In this solution resources are managed in a market-oriented fashion that enables dynamic regulation of the supply and demand of resources and promotes the mechanisms for their allocation that would take into account their priorities and levels of utilization.
+Such a model has been devised and named \emph{InterCloud} \cite{InterCloud}.
+In this solution resources are managed in a market-oriented \cite{MarketOriented} fashion that enables dynamic regulation of the supply and demand of resources and promotes the mechanisms for their allocation that would take into account their priorities and levels of utilization.
 
@@
 Scalability is a widely accepted measure for improving application performance, consequently increasing offered Quality-of-Service.
-Despite the lack of a clear definition, in this article we consider this term an ability to work without delays and unproductive resource consumption at light, moderate, or heavy loads while making good use of available resources (in some works it is referred to as \emph{load scalability}).
+Despite the lack of a clear definition, in this article we consider this term an ability to work without delays and unproductive resource consumption at light, moderate, or heavy loads while making good use of available resources \cite{ScalabilityTerm} (in some works it is referred to as \emph{load scalability}).
 
@@
 In the context of this article, the main point of interest are providers which offer features related to resource management, monitoring and utilization.
-In general, current key providers include Amazon \emph{Web Services}, \emph{Microsoft Azure}, Google \emph{AppEngine} and \emph{Heroku}.
+In general, current key providers include Amazon \emph{Web Services} \cite{AWS}, \emph{Microsoft Azure} \cite{Azure}, Google \emph{AppEngine} \cite{GoogleAppEngine} and \emph{Heroku} \cite{Heroku}.
 Despite not all of them operate on the same level of abstraction, it is possible to compare them as they expose functionalities which can be classified as PaaS.
@@
 
-\section{Section title}
-\label{sec:1}
-Text with citations \cite{RefB} and \cite{RefJ}.
-\subsection{Subsection title}
-\label{sec:2}
-as required. Don't forget to give each section
-and subsection a unique label (see Sect.~\ref{sec:1}).
-\paragraph{Paragraph headings} Use paragraph headings as needed.
-\begin{equation}
-a^2+b^2=c^2
-\end{equation}
+%\section{Section title}
+%\label{sec:1}
+%Text with citations \cite{RefB} and \cite{RefJ}.
+%\subsection{Subsection title}
+%\label{sec:2}
+%as required. Don't forget to give each section
+%and subsection a unique label (see Sect.~\ref{sec:1}).
 
@@
 %
-\bibitem{RefJ}
+\bibitem{NIST}
+  {Mell, P. and Grance, T.}, {The NIST Definition of Cloud Computing}, 2011
+\bibitem{InterCloud}
+  {Buyya, R. Ranjan, R. and Calheiros, R.}, {InterCloud: Scaling of Applications across multiple Cloud Computing Environments}, 2010
+\bibitem{MarketOriented}
+  {Buyyaa, R., Yeoa, C., Venugopala, S., d Broberga, J.
+and Brandicc, I.}, {Cloud computing and emerging IT platforms: Vision, hype, and reality for delivering computing as the 5th utility}, {Future Generation Computer Systems}, vol.25, 2009
+\bibitem{ScalingInDaCloud}
+  {Vaquero, L., Rodero-Merino, L. and Buyya, R.}, {Dynamically Scaling Applications in the Cloud}, {ACM SIGCOMM Computer Communication Review}, vol.41, 2011
+\bibitem{ScalabilityTerm}
+  Bondi, A., {Characteristics of Scalability and Their Impact on Performance}, 2000
+\bibitem{PharmaComputation}
+  Jon Brodkin, \emph{\$1,279-per-hour, 30,000-core cluster built on Amazon EC2 cloud}, {\emph{Ars Technica}, \url{http://arstechnica.com/business/2011/09/30000-core-cluster-built-on-amazon-ec2-cloud/}}, last access: 16/09/2013
+\bibitem{AWS}
+  {Amazon Elastic Compute Cloud (EC2)}, {\url{http://aws.amazon.com/}}, last access: 16/09/2013 
+\bibitem{GoogleAppEngine}
+  {Google App Engine}, {\url{https://appengine.google.com/}}, last access: 16/09/2013 
+\bibitem{Heroku}
+  {Heroku}, {\url{https://www.heroku.com}}, last access: 16/09/2013 
+\bibitem{Azure}
+  {Windows Azure}, {\url{www.windowsazure.com/}}, last access: 16/09/2013 
 % Format for Journal Reference
-Author, Article title, Journal, Volume, page numbers (year)
+%Author, Article title, Journal, Volume, page numbers (year)
 % Format for books
-\bibitem{RefB}
-Author, Book title, page numbers. Publisher, place (year)
+%Author, Book title, page numbers. Publisher, place (year)
 % etc

### 2013-09-16 thesis/src/chapters/chapter-5.tex
Added arch of csap

@@
-\chapter{Solution}
+\chapter{Architecture of Cloud-SAP}
+
+\chapterintro{This chapter introduces architecture of Cloud-SAP, Platform-as-a-Service model that aims to satisfy requirements introduced in first chapter.}
 
@@
 
-Taking requirements into consideration, desired platform clearly has features common to a autonomic system, especially the need for automatic environment reconfiguration.
-One of the first architectures that satisfied that requirement was proposed by \cite{IBM06}.
-Beside this, system under consideration has a character of a multi-layered one (i.e.
-it operates on a application / container / service / cloud level).
-Optimally, each level should be characterized by different actions, what can be meet by implementing multi-hierarchical autonomic system, as proposed in \cite{LiWoZh05}.
-
-Considering that observation, auto-scaling subsystem spans across different layers:
-
-\begin{figure}[!ht]
-  \begin{center}
-    \includegraphics[width=\textwidth]{chapter-5/hierarchical-autonomic-system}
-  \end{center}
-  \caption{Hierarchical autonomic system}
-  \label{ch5:hierarchical-autonomic-system}
-\end{figure}
-
-Remaining of this section breaks apart each functional module of an autonomic component, whereas further sections cover in details each autonomic level.
-
-\subsection{Autonomic component}
-
-\begin{figure}[!ht]
-  \begin{center}
-    \includegraphics[width=0.8\textwidth]{chapter-5/autnomic-component}
-  \end{center}
-  \caption{Autonomic component}
-  \label{ch5:autonomic-component}
-\end{figure}
-
-\subsubsection{Monitor}
-
-Monitor module is responsible for gathering monitoring data from \textit{Sensors} and store them in a database, correlated with a specific autonomous component (ie.
-application / container / service / cloud).
-Problem of a multi-layered, hierarchical self-adaptive monitoring system is described in \cite{KaKoGoKyMeVa12}.
-
-\subsubsection{Analyse}
-
-Analyse module performs analysis of a gathered data. There is a variety of models that helps to achieve this goal. For example, the \cite{LiWoZh05} proposes:
-
+One can notice that elements that yields a solution for a problem stated in first chapter, which is ensuring that users' application provide appropriate Quality-of-Service for its customers, were introduced in previous chapters:
 \begin{itemize}
-	\item Queue-based Performance Models
-	\item Dynamic models
-	\item Monotonic static models
-	\item Policy based models
+	\item scalability - ability to improve application performance by enriching 
+	\item adaptivity - ability to adapt (i.e. scale) appropriately to current usage pattern
+	\item inter-cloud awareness - ability to cooperate with different cloud provider to supply application with extra resources
 \end{itemize}
 
-Details about which level leverage which model are presented in detailed architecture description.
-Important part in a data analysis plays a predictive approach \cite{JiPeLiCh11} and tracking changes in a system \cite{ZhYaWo05}.
+To begin with, different scalability models were introduced. Noticeably, they operated at different layers: 
+\begin{itemize}
+	\item application platform tuning
+	\item single-server level (vertical scaling)
+	\item multi-server level (horizontal scaling) 
+	\item inter-cloud level (scaling out across different cloud-providers, concept introduced in previous chapter)
+\end{itemize}
+Consequently, it is expected that proposed solution's architecture reflects that multi-layered model, which, as Table \ref{tab:cloud-providers-scaling} illustrates, hasn't been yet introduced to the marked.
+Additionally, each layer has to be characterized by an ability to adapt to a given application usage.
+As previous chapter stated, this can be achieved by adding a elasticity controller to each layer.
+This observation is a foundation of proposed architecture.
 
-\subsubsection{Plan}
+One of the first models that ensured system adaptivity is an autonomic component, concept based on a feed-back loop, initially proposed by IBM \cite{IBM06}. Figure () depicts that architecture.
 
-This module is responsible for scheduling actions that have to be taken to solve problems reported by an earlier analysis. This may involve:
+Since designed platform operates on multiple layers, we can extend that model by using concept of a multi-hierarchical autonomic system \cite{LiWoZh05}.
+Figure () illustrates that hierarchy, were each level represents a different perspective at application scaling.
+Using terminology characteristic to a autonomic system, first three hierarchy levels (application tuning, single-server, multi-server) are centralized and controlled by single elasticity controller at each level, while the last inter-cloud level is a decentralized one - each cloud instance is fully independent.
+
+As \cite{IBM06} states, each autonomic component has modules that are responsible for:
 \begin{itemize}
-	\item reserving some resources (for example to deploy a new virtual machine)
-	\item handling situation where multiple entities compete about the resources (for example by a prioritization)
+	\item monitoring
+	\item analysis
+	\item planning
+	\item action execution
 \end{itemize}
 
-It is possible, however, that resolving issues is not possible at current level. It that case, the upper-level is responsible for handling it.
+While the managed component has:
+\begin{itemize}
+	\item sensors
+	\item effectors
+\end{itemize}
 
-\subsubsection{Execute}
-
-Execute module is directly responsible for managing resource, by performing various operations on resource's effector.
-Each layer is characterized by different actions that can be taken, for example: application server reconfiguration, vertical or horizontal scaling.
-
-\subsubsection{Knowledge base}
-
-Knowledge base is represented as a set of rules, policies shared across autonomous components.
-
-\subsection{Managed resource}
-
-\subsubsection{Effector}
-
-Enables management of a resources by exposing some API compatible with an Executor.
-
-\subsubsection{Sensor}
-
-Aim of a sensor is probing a resource and sending back results to a Monitor. There are different kind of metrics that should be considered such as: CPU load, memory usage or response delay.
-
-\section{Application controller}
-
-\section{Container controller}
-
-\section{Service controller}
-
-\section{Cloud controller}
+Due to the hierarchy of our architecture, each level manages an underlying autonomic component, while being managed by an upper layer at the same time.
+This lead us to observation that they are autonomic components and managed components at the same time.
+Remaining of this chapter discusses in detail each layer taking into account that observation.

### 2013-09-16 article/article.tex
Add the adaptability section to the article

@@
 \section{Adaptivity}
+Platform adaptivity adds auto-scaling features to a solution offered by a Platform-as-a-Service provider.
+The key concept is to have a Elasticity Controller which gathers probes from virtual machines and uses that knowledge to execute appropriate action on cloud instance, indirectly modifying consecutive probes \cite{VaRoBu11}.
+
+\paragraph{Policies} Policy denotes a condition which, when satisfied, triggers an action that is supposed to harness cloud instance in a way that future evaluations of condition will be unsuccessful.
+Typically condition itself is accompanied by a minimal and maximal number of node instances, allowing for ensuring minimal QoS and controlling maximal costs.
+Currently, industry leaders supports two main kind of policies: \emph{expression based} and \emph{scheduled}.
+\emph{Expression-based} policy allows to define how to scale the application in response to changing conditions, which include factors such as memory, CPU usage, cost or some indirect, calculated metrics, while \emph{expression-based} scale the service according to load changes; for example, traffic increases during the weekends and decreases on working days.
+Hence, that predictable traffic patterns is used to scale application based on current time.
+
+\paragraph{Data analysis}
+Once the policies are defined, their conditions are evaluated against data acquired from sensors.
+In the simplest case this evaluation can be based on a Threshold Model \cite{LiWoZh05}, which defines a valid range.
+In cases when given metric violates that condition (i.e.
+value is either smaller than minimal or bigger than maximal acceptable) corresponding resource is properly adjusted.
+
 \section{Architecture of SAP-Cloud}
@@
   {Windows Azure}, {\url{www.windowsazure.com/}}, last access: 16/09/2013 
+\bibitem{VaRoBu11}
+  {Vaquero, L., Rodero-Merino, L. and Buyya, R.},{Dynamically Scaling Applications in the Cloud}, {ACM SIGCOMM Computer Communication Review}, vol.41, 2011
+\bibitem{LiWoZh05}
+ {Litoiu, M. and Woodside, M. and Zheng, T.}, {Hierarchical Model-based Autonomic Control of Software Systems}, {Design and evolution of autonomic application software}, 2005
+
 % Format for Journal Reference

### 2013-09-16 article/article.tex
Added architecture section

@@
 
-\section{Architecture of SAP-Cloud}
+\section{Architecture of Cloud-SAP}
+
+One can notice that elements that yields a solution for a problem stated at the begging of this article, which is ensuring that users' application provide appropriate Quality-of-Service for its customers, were introduced in previous sections:
+\begin{itemize}
+	\item scalability - ability to improve application performance by enriching 
+	\item adaptivity - ability to adapt (i.e. scale) appropriately to current usage pattern
+	\item inter-cloud awareness - ability to cooperate with different cloud provider to supply application with extra resources
+\end{itemize}
+
+To begin with, different scalability models were introduced. Noticeably, they operated at different layers: 
+\begin{itemize}
+	\item application platform tuning
+	\item single-server level (vertical scaling)
+	\item multi-server level (horizontal scaling) 
+	\item inter-cloud level (scaling out across different cloud-providers, concept introduced in previous chapter)
+\end{itemize}
+Consequently, it is expected that proposed solution's architecture reflects that multi-layered model, which hasn't been yet introduced to the marked.
+Additionally, each layer has to be characterized by an ability to adapt to a given application usage.
+As previous chapter stated, this can be achieved by adding a elasticity controller to each layer.
+This observation is a foundation of proposed architecture.
+
+One of the first models that ensured system adaptivity is an autonomic component, concept based on a feed-back loop, initially proposed by IBM \cite{IBM06}.
+Figure \ref{ch5:autonomic-component} depicts that architecture.
+
+\begin{figure}[!ht]
+  \includegraphics[width=0.47\textwidth]{drawings/autonomic-component}
+  \caption{Autonomic component}
+  \label{ch5:autonomic-component}
+\end{figure}
+
+Since designed platform operates on multiple layers, we can extend that model by using concept of a multi-hierarchical autonomic system \cite{LiWoZh05}.
+Figure \ref{ch5:hierarchical-autonomic-system} illustrates that hierarchy, were each level represents a different perspective at application scaling.
+Using terminology characteristic to a autonomic system, first three hierarchy levels (application tuning, single-server, multi-server) are centralized and controlled by single elasticity controller at each level, while the last inter-cloud level is a decentralized one - each cloud instance is fully independent.
+
+\begin{figure*}[!ht]
+  \includegraphics{drawings/hierarchical-autonomic-system}
+  \caption{Hierarchical autonomic system}
+  \label{ch5:hierarchical-autonomic-system}
+\end{figure*}
+
+As \cite{IBM06} states, each autonomic component has modules that are responsible for:
+\begin{itemize}
+	\item monitoring
+	\item analysis
+	\item planning
+	\item action execution
+\end{itemize}
+
+While the managed component has:
+\begin{itemize}
+	\item sensors
+	\item effectors
+\end{itemize}
+
+Due to the hierarchy of our architecture, each level manages an underlying autonomic component, while being managed by an upper layer at the same time.
+This lead us to observation that they are autonomic components and managed components at the same time.
 
@@
  {Litoiu, M. and Woodside, M. and Zheng, T.}, {Hierarchical Model-based Autonomic Control of Software Systems}, {Design and evolution of autonomic application software}, 2005
+\bibitem{IBM06}
+ {IBM}, {An architectural blueprint for autonomic computing, 4th edition}, 2006
 

### 2013-09-16 thesis/src/chapters/chapter-5.tex
Added architecture diagrams

@@
 
-One of the first models that ensured system adaptivity is an autonomic component, concept based on a feed-back loop, initially proposed by IBM \cite{IBM06}. Figure () depicts that architecture.
+One of the first models that ensured system adaptivity is an autonomic component, concept based on a feed-back loop, initially proposed by IBM \cite{IBM06}.
+Figure \ref{ch5:autonomic-component} depicts that architecture.
+
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics{chapter-5/autonomic-component}
+  \end{center}
+  \caption{Autonomic component}
+  \label{ch5:autonomic-component}
+\end{figure}
 
 Since designed platform operates on multiple layers, we can extend that model by using concept of a multi-hierarchical autonomic system \cite{LiWoZh05}.
-Figure () illustrates that hierarchy, were each level represents a different perspective at application scaling.
+Figure \ref{ch5:hierarchical-autonomic-system} illustrates that hierarchy, were each level represents a different perspective at application scaling.
 Using terminology characteristic to a autonomic system, first three hierarchy levels (application tuning, single-server, multi-server) are centralized and controlled by single elasticity controller at each level, while the last inter-cloud level is a decentralized one - each cloud instance is fully independent.
+
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics{chapter-5/hierarchical-autonomic-system}
+  \end{center}
+  \caption{Hierarchical autonomic system}
+  \label{ch5:hierarchical-autonomic-system}
+\end{figure}
 

### 2013-12-06 thesis/src/chapters/chapter-5.tex
Add additional chapter

@@
-\chapter{Architecture of Cloud-SAP}
+\chapter{Design of Cloud-SAP}
 
-\chapterintro{This chapter introduces architecture of Cloud-SAP, Platform-as-a-Service model that aims to satisfy requirements introduced in first chapter.}
+\chapterintro{This chapter introduces the high-level design of Cloud-SAP, a Platform-as-a-Service model that aims to satisfy requirements introduced in the first chapter.}
 

### 2013-12-06 thesis/src/chapters/chapter-7.tex
Add additional chapter

\chapter{Evaluation}

\chapterintro{
  This chapter contains results and discussions on the evaluation of tests run
  on the proposed solution.
}

### 2013-12-06 thesis/src/appendices/listings.tex
Removed listing duplicate
Added listing style
Add a draft of the first test case

@@
 \chapter{Code listings}
+
+\section{Service specifications}
+\lstinputlisting[caption=Cloud-SAP Service Specification (without scaling policies), label=lst:service-spec-test-cost]{appendices/service-spec-test-cost.json}
 
@@
 
-\lstinputlisting[caption=Auto Scaling policy\, AWS EC2]{scaling-policy.xml}
+Listings \ref{lst:scaling-ec2} illustrates XML-based policy that is used by Auto Scaling of the Amazon Web Services EC2.
+\lstinputlisting[caption=Scaling policy - AWS EC2, label=lst:scaling-ec2]{appendices/scaling-policy.xml}

### 2013-12-06 thesis/src/chapters/chapter-7.tex
Add a draft of the first test case

@@
 \chapterintro{
-  This chapter contains results and discussions on the evaluation of tests run
-  on the proposed solution.
+  This chapter contains results and discussions on the evaluation of tests run on the proposed solution.
 }
+
+\section{Introduction}
+We carried out a number of tests which aim to prove the solution be better than currently available, especially in terms of:
+\begin{enumerate}
+  \item maximum reduction of deployment cost by deploying different stacks comprising a given service on different cloud providers
+  \item reduced deployment time
+  \item TODO dopisać jeszcze kolejne
+\end{enumerate}
+
+\section{Service deployment}
+\subsection{Description and preconditions}
+The aim of this test case is to test the primary use case of the proposed solution -- deployment of a service with the emphasis of client's \textbf{cost}.
+Service specification (\ref{lst:service-spec-test-cost}) forms an input to the application.
+Its elements are different software stacks that are parts of the whole service.
+Each cloud provider has its own price for a given software stack which is shown in table \ref{tbl:test-service-deployment-cost}.
+This test should show that the platform chooses the best mapping between the stacks and cloud providers so that the client's pays the \textbf{lowest} possible price.
+
+\begin{table}
+  \centering
+  \begin{tabular}{ | c | c | c | c | }
+    \hline                        
+    & CP-1 & CP-2 & CP-3 \\
+    \hline
+    java      & 150 & 120 & 180 \\
+    ruby      & 220 & 290 & 250 \\
+    postgres  & 320 & 240 & 290 \\
+    python    & 200 & 260 & 180 \\
+    amqp      & 330 & 390 & 285 \\
+    \hline  
+  \end{tabular}
+  \caption{Price for a stack in the given cloud provider}
+  \label{tbl:test-service-deployment-cost}
+\end{table}
+
+\subsection{Results}
+Table \ref{tbl:test-service-deployment-cost-mapping} shows obtained mapping between stacks and cloud providers.
+Taking into account this result, figure \ref{ch7:service-deployment-cost} shows comparison of cost the client would have to pay with and without such a mapping.
+
+\begin{table}
+  \centering
+  \begin{tabular}{ | c | c | c | c | }
+    \hline                        
+    & CP-1 & CP-2 & CP-3 \\
+    \hline
+    java      & & x & \\
+    ruby      & x & & \\
+    postgres  & & x & \\
+    python    & & & x \\
+    amqp      & & & x \\
+    \hline  
+  \end{tabular}
+  \caption{Chosen cloud providers for the given stack}
+  \label{tbl:test-service-deployment-cost-mapping}
+\end{table}
+
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics{chapter-7/case-study-service-deployment-reduced-client-costs}
+  \end{center}
+  \caption{Comparison of the deployment cost when the service is deployed only on a selected cloud provider or a combination of cloud providers selected by Cloud-SAP}
+  \label{ch7:service-deployment-cost}
+\end{figure}
+
+\section{Auto-scaling}

### 2013-12-11 auto-scaling/cloud-controller/lib/cloud_controller.rb
[#33] Using stack as a first-class citizen for auto-scaling

@@ -13,7 +13,7 @@ require 'cloud_controller/offer_response_preparer'
 
 # Remaining CSAP stack
 require 'cloud-provider/cloud_provider'
-require 'service-controller/service_controller'
+require 'stack-controller/service_controller'
 require 'container-controller/container_controller'
 
 # common
@@ -31,7 +31,7 @@ module AutoScaling
     attr_writer :service_controller
     attr_writer :container_controller
 
-    # Handle request passed from lower layer (service-controller)
+    # Handle request passed from lower layer (stack-controller)
     #
     # * *Args* :
     # - +conclusion+ -> an action that lower layer wanted to perform
@@ -68,7 +68,7 @@ module AutoScaling
       cloud_provider = OpenNebulaClient.new(config['endpoints'][config['cloud_provider_name']])
 
       @@logger.info("Initializing ServiceController")
-      service_controller            = ServiceController.build(cloud_provider, config)
+      service_controller            = StackController.build(cloud_provider, config)
       instance.service_controller   = service_controller
       config['service_controller']  = service_controller
 

### 2013-12-11 auto-scaling/cloud-provider/lib/opennebula/opennebula_client.rb

@@ -44,8 +44,8 @@ module AutoScaling
     end
 
     # Creates an instance of a container with a predefined template in mappings
-    def instantiate_container(stack_type, container_role, service_id, mappings)
-      @frontend.instantiate_container(stack_type, container_role, service_id, mappings)
+    def instantiate_container(stack_type, container_role, mappings)
+      @frontend.instantiate_container(stack_type, container_role, mappings)
     end
 
     def monitor_container(container_id)

### 2013-12-11 auto-scaling/cloud-provider/lib/opennebula/opennebula_frontend.rb

@@ -27,7 +27,7 @@ module AutoScaling
 
     # Instantiates container with given type and role, which are used
     # to determine its template id
-    def instantiate_container(stack_type, container_role, service_id, mappings)
+    def instantiate_container(stack_type, container_role, mappings)
       @@logger.debug "Instantiating container: #{stack_type}, #{container_role}"
 
       # get template
@@ -38,7 +38,7 @@ module AutoScaling
 
       # add required variables
       raise(RuntimeError, res.message) if OpenNebula.is_error?(res)
-      t = add_required_variables(template, container_variables(service_id))
+      t = add_required_variables(template, container_variables())
 
       # instantiate it
       xml = OpenNebula::VirtualMachine.build_xml()
@@ -103,8 +103,8 @@ module AutoScaling
 
     private
     # Extra variables that needs to be added to handle base image initialization
-    def container_variables(service_id)
-      { 'SERVICE_ID' => service_id, 'VM_ID' => '$VMID' }
+    def container_variables()
+      { 'VM_ID' => '$VMID' }
     end
 
     # Modifies a template to add new variables

### 2013-12-11 auto-scaling/cloud-provider/test/integration/opennebula/opennebula_client_test.rb

@@ -78,7 +78,7 @@ eos
     end
 
     def test_shall_manage_container
-      container_info = @opennebula_client.instantiate_container('bootstrap', 'base', 'bootstrap', MAPPINGS)
+      container_info = @opennebula_client.instantiate_container('bootstrap', 'base', MAPPINGS)
       assert_equal true, container_info[:id] >= 0
       assert_equal true, container_info[:ip] != ''
 

### 2013-12-11 auto-scaling/container-controller/container_controller.rb

@@ -17,7 +17,7 @@ module AutoScaling
 
     @@logger = Logger.new(STDOUT)
 
-    # Build a service-controller instance
+    # Build a stack-controller instance
     #
     # * *Args* :
     # - +cloudLogger.new(STDOUT)

### 2013-12-11 auto-scaling/domain/service.rb

@@ -8,11 +8,6 @@ module AutoScaling
 
     property :id, Serial
 
-    property :name, String, :required => true
-    property :status, Enum[ :new, :converged ], :default => :new
-
-    has n, :stacks
-
     def to_s
       JSON.pretty_generate(self)
     end

### 2013-12-11 auto-scaling/domain/stack.rb

@@ -9,14 +9,16 @@ module AutoScaling
     property :id, Serial
     property :correlation_id, Integer
 
-    belongs_to :service
-
     has 1, :policy_set
     has n, :containers
 
     property :type, Enum[ :invalid, :java ], :default => :invalid
-    property :state, Enum[ :invalid, :queued, :pending, :deployed ], :default => :invalid
+    property :state, Enum[ :invalid, :queued, :pending, :deployed, :converged ], :default => :invalid
     property :data, String
+
+    def self.correlated(correlation_id)
+      first(:correlation_id => correlation_id)
+    end
 
     def to_s
       JSON.pretty_generate(self)

### 2013-12-11 auto-scaling/service-controller/test/unit/models/service_test.rb

@@ -1,69 +1,0 @@
-require 'rubygems'
-require 'data_mapper'
-require "test/unit"
-
-require 'utils'
-require 'domain/domain'
-
-module AutoScaling
-  class ServiceTest < Test::Unit::TestCase
-
-    def setup
-      Utils::setup_database
-    end
-
-    def test_persistence_model
-      containers = [
-        Container.create(
-          :id => 10,
-          :ip => '192.168.122.1'
-        ),
-        Container.create(
-            :id => 11,
-            :ip => '192.168.122.2'
-        ),
-        Container.create(
-            :id => 0,
-            :ip => '192.168.122.200',
-            :type => :master
-        )
-      ]
-
-      policy_set = PolicySet.create(
-        :min_vms => 1,
-        :max_vms => 3,
-        :policies => [
-          Policy.create(
-            :name => 'threshold_model',
-            :arguments => {:min => 2, :max =>3}
-          ),
-          Policy.create(
-              :name => 'threshold_model_non_existing',
-              :arguments => {:min => 2, :max =>3, :turbo => 69}
-          )
-        ]
-      )
-
-      @stack = Stack.create(
-        :type => :java,
-        :data => 'http://jenkins.com/path/to/my/app.war',
-        :containers => containers,
-        :policy_set => policy_set
-      )
-
-      @service = ::AutoScaling::Service.create(
-        :id => 110,
-        :name => 'enterprise-app',
-        :stacks => [@stack]
-      )
-
-      # test_shall_get_information_if_container_is_a_master
-      container = Container.get(0)
-      assert_equal Container.master(Stack.get(@stack.id)), container
-
-      slaves = Container.slaves(Stack.get(@stack.id))
-      assert_equal 2, slaves.size
-    end
-
-  end
-end

### 2013-12-11 auto-scaling/service-controller/test/unit/monitor/service_monitor_test.rb

@@ -1,102 +1,0 @@
-require 'rubygems'
-require 'test/unit'
-require 'mocha/setup'
-
-require 'utils'
-require 'monitor/service_monitor'
-
-module AutoScaling
-
-  class ServiceMonitorTest < Test::Unit::TestCase
-
-    def setup
-      @cloud_provider = mock()
-      Utils::setup_database
-
-      @monitor = ServiceMonitor.new @cloud_provider
-
-      instance_id = 69
-
-      @containers = [
-          Container.create(
-              :correlation_id => 10,
-              :ip => '192.168.122.1'
-          ),
-          Container.create(
-              :correlation_id => 11,
-              :ip => '192.168.122.2'
-          ),
-          Container.create(
-              :correlation_id => 12,
-              :ip => '192.168.122.200',
-              :type => :master
-          )
-      ]
-
-      @stack = Stack.create(
-          :correlation_id => instance_id,
-          :type => 'java',
-          :containers => @containers
-      )
-
-      @service = Service.create(
-          :name => 'service-name',
-          :stacks => [@stack],
-          :status => :converged
-      )
-    end
-
-    def test_shall_return_data_from_last_timestamp
-      set_1 = [["1374678040", "524288"], ["1374678083", "524288"], ["1374678113", "524288"], ["1374678155", "524288"]]
-      set_2 = [["1374678040", "524288"], ["1374678083", "524288"], ["1374678113", "524288"], ["1374678155", "524288"], ["1374678198", "524288"], ["1374678241", "524288"], ["1374678284", "524288"], ["1374678327", "524288"], ["1374678370", "524288"], ["1374678413", "524288"]]
-
-      container = Container.create(
-        :id => 100,
-        :ip => '192.168.122.1'
-      )
-
-      assert_equal set_1, @monitor.send(:last, set_1, container )
-      container.probed = "1374678155"
-      container.save
-
-      assert_equal set_2 - set_1, @monitor.send(:last, set_2, container )
-    end
-
-    def test_shall_grab_data_about_all_containers
-      data = { "CPU" => [["1", "100"], ["5", "105"]] }
-      @containers.each {|c| @cloud_provider.expects(:monitor_container).with(c.correlation_id).returns(data)}
-
-      values = { "CPU" => ["100", "105"] }
-      expected = { @stack => {@containers[0] => values, @containers[1] => values, @containers[2] => values}}
-
[…]

### 2013-12-11 auto-scaling/stack-controller/lib/analyzer/stack_analyzer.rb

@@ -6,7 +6,7 @@ require 'domain/domain'
 require 'common/common'
 
 module AutoScaling
-  class ServiceAnalyzer
+  class StackAnalyzer
 
     @@logger = Logger.new(STDOUT)
 

### 2013-12-11 auto-scaling/stack-controller/lib/executor/stack_executor.rb

@@ -8,7 +8,7 @@ require 'domain/domain'
 
 module AutoScaling
 
-  class ServiceExecutor
+  class StackExecutor
 
     @@logger = Logger.new(STDOUT)
 
@@ -20,41 +20,30 @@ module AutoScaling
     end
 
     # * *Args* :
-    # - +service+ -> description of a service. An instance of HashMap, ex:
-    #   service =
-    #  {
-    #    'name' => 'youtube',
-    #    'stacks' => [
-    #     {
-    #       'stack' => 'tomcat',
-    #       'instances' => 2,
-    #       'name' => 'enterprise-app'
-    #       'policy_set' =>
-    #          :min_vms =>  0
-    #          :max_vms =>  2
-    #          :polices => [{'name' => 'threshold_model', :parameters => {'min' => '5', 'max' => '50'}}]
-    #     }
-    #     ]
-    #  }
+    # - +stack+ -> description of a stack. An instance of HashMap, ex:
+    #   stack =
+    #   {
+    #     'stack' => 'tomcat',
+    #     'instances' => 2,
+    #     'name' => 'enterprise-app'
+    #     'policy_set' =>
+    #        :min_vms =>  0
+    #        :max_vms =>  2
+    #        :polices => [{'name' => 'threshold_model', :parameters => {'min' => '5', 'max' => '50'}}]
+    #   }
     #
-    def deploy_service(service)
-      @@logger.debug "Deploying service #{service} with mappings: #{@mappings}"
+    def deploy_stack(stack_data)
+      @@logger.debug "Deploying stack #{stack_data} with mappings: #{@mappings}"
 
-      stacks = service['stacks'].collect { |stack| create_stack(stack) }
+      stack = create_stack(stack_data)
+      update(stack)
 
-      service = Service.create(
-        :name => service['name'],
-        :stacks => stacks
-      )
-
-      update(service)
-
-      service
+      stack
     end
 
     def deploy_container(stack)
       @@logger.debug "Deploying container at #{stack} with mappings: #{@mappings}"
-      container_info = @cloud_provider.instantiate_container(stack.type.to_s, 'slave', stack.service.id, @mappings)
+      container_info = @cloud_provider.instantiate_container(stack.type.to_s, 'slave', @mappings)
 
       # persist data
       container = Container.create(
@@ -92,14 +81,27 @@ module AutoScaling
     # Updates model, so it reflects actual configuration
     #
     # * *Args* :
-    # - +service+ -> instance of domain/service
[…]

### 2013-12-11 auto-scaling/stack-controller/lib/monitor/stack_monitor.rb

@@ -5,7 +5,7 @@ require 'cloud-provider/cloud_provider'
 require 'domain/domain'
 
 module AutoScaling
-  class ServiceMonitor
+  class StackMonitor
 
     @@logger = Logger.new(STDOUT)
 

### 2013-12-11 auto-scaling/stack-controller/stack_controller.rb

@@ -4,17 +4,17 @@ require 'rubygems'
 require 'logger'
 require 'rufus-scheduler'
 
-require 'planner/service_planner'
+require 'planner/stack_planner'
 require 'planner/reservation_manager'
-require 'executor/service_executor'
-require 'monitor/service_monitor'
-require 'analyzer/service_analyzer'
+require 'executor/stack_executor'
+require 'monitor/stack_monitor'
+require 'analyzer/stack_analyzer'
 require 'analyzer/policy_evaluator'
 require 'domain/domain'
 
 module AutoScaling
 
-  class ServiceController
+  class StackController
 
     @@logger = Logger.new(STDOUT)
     attr_reader :monitor, :analyzer, :planner
@@ -57,7 +57,7 @@ module AutoScaling
       @planner.reserve?(stack_data)
     end
 
-    # Build a service-controller instance
+    # Build a stack-controller instance
     #
     # * *Args* :
     # - +cloud_provider+ -> cloud provider, ie, open nebula
@@ -70,18 +70,18 @@ module AutoScaling
       mappings = settings['mappings'][settings['cloud_provider_name']]
 
       # mapek model
-      monitor = ServiceMonitor.new(cloud_provider)
-      analyzer = ServiceAnalyzer.new(PolicyEvaluator.new)
-      executor = ServiceExecutor.new(cloud_provider, mappings)
+      monitor = StackMonitor.new(cloud_provider)
+      analyzer = StackAnalyzer.new(PolicyEvaluator.new)
+      executor = StackExecutor.new(cloud_provider, mappings)
 
       capacity = {}
       if settings['endpoints'][settings['cloud_provider_name']].key?('capacity')
         capacity = settings['endpoints'][settings['cloud_provider_name']]['capacity']
       end
       reservation_manager = ReservationManager.new(cloud_provider, capacity)
-      planner = ServicePlanner.new(executor, cloud_controller, reservation_manager)
+      planner = StackPlanner.new(executor, cloud_controller, reservation_manager)
 
-      service_controller = ServiceController.new(monitor, analyzer, planner, executor)
+      service_controller = StackController.new(monitor, analyzer, planner, executor)
       cloud_controller.service_controller = service_controller
 
       return service_controller

### 2013-12-11 auto-scaling/stack-controller/test/unit/executor/chef_renderer_test.rb

@@ -53,13 +53,6 @@ module AutoScaling
           :containers => containers
       )
 
-      service = Service.create(
-          :id => instance_id,
-          :name => 'service-name',
-          :stacks => [stack],
-          :status => :converged
-      )
-
       actual = ChefRenderer.render(stack)
 
       assert_equal expected, actual

### 2013-12-11 auto-scaling/stack-controller/test/unit/monitor/stack_monitor_test.rb

@@ -1,0 +1,14 @@
+require 'rubygems'
+require 'test/unit'
+require 'mocha/setup'
+
+require 'utils'
+require 'monitor/stack_monitor'
+
+module AutoScaling
+
+  class StackMonitorTest < Test::Unit::TestCase
+
+  end
+
+end

### 2013-12-11 auto-scaling/stack-controller/test/unit/planner/stack_planner_test.rb

@@ -3,18 +3,18 @@ require 'test/unit'
 require 'mocha/setup'
 
 require 'utils'
-require 'planner/service_planner'
+require 'planner/stack_planner'
 
 module AutoScaling
 
-  class ServicePlannerTest < Test::Unit::TestCase
+  class StackPlannerTest < Test::Unit::TestCase
 
     def setup
       @executor = mock()
       @cloud_controller = mock()
       @reservation_manager = mock()
 
-      @planner = ServicePlanner.new(@executor, @cloud_controller, @reservation_manager)
+      @planner = StackPlanner.new(@executor, @cloud_controller, @reservation_manager)
     end
 
     def test_shall_plan_service_deployment
@@ -23,7 +23,7 @@ module AutoScaling
           'instances' => 2,
           'name' => 'enterprise-app'
       }
-      @executor.expects(:deploy_service).with(service)
+      @executor.expects(:deploy_stack).with(service)
 
       @planner.plan_deployment(service)
     end

### 2013-12-11 auto-scaling/stack-controller/test/unit/stack_controller_test.rb

@@ -3,17 +3,17 @@ require 'test/unit'
 require 'mocha/setup'
 
 require 'utils'
-require 'service-controller/service_controller'
+require 'stack-controller/stack_controller'
 
 module AutoScaling
 
-  class ServiceControllerTest < Test::Unit::TestCase
+  class StackControllerTest < Test::Unit::TestCase
 
     def setup
       @monitor, @analyzer, @planner = mock(), mock(), mock()
       @scheduler = Rufus::Scheduler.new
 
-      @controller = ServiceController.new @monitor, @analyzer, @planner, @scheduler
+      @controller = StackController.new @monitor, @analyzer, @planner, @scheduler
     end
 
     def test_shall_perform_full_cycle

### 2013-12-11 auto-scaling/cloud-controller/lib/cloud_controller.rb
[#33] Using stack instead of service

@@ -13,7 +13,7 @@ require 'cloud_controller/offer_response_preparer'
 
 # Remaining CSAP stack
 require 'cloud-provider/cloud_provider'
-require 'stack-controller/service_controller'
+require 'stack-controller/stack_controller'
 require 'container-controller/container_controller'
 
 # common
@@ -28,7 +28,7 @@ module AutoScaling
     @@logger       = Logger.new(STDOUT)
     @@logger.level = Logger::DEBUG
 
-    attr_writer :service_controller
+    attr_writer :stack_controller
     attr_writer :container_controller
 
     # Handle request passed from lower layer (stack-controller)
@@ -69,7 +69,7 @@ module AutoScaling
 
       @@logger.info("Initializing ServiceController")
       service_controller            = StackController.build(cloud_provider, config)
-      instance.service_controller   = service_controller
+      instance.stack_controller   = service_controller
       config['service_controller']  = service_controller
 
       @@logger.info("Initializing Container Controller")
@@ -123,18 +123,18 @@ module AutoScaling
 
     def handle_deploy_request(metadata, payload)
       @@logger.info("Handling a deploy request")
-      service_data = JSON.parse(payload)
-      @@logger.debug("Payload: #{service_data.inspect}")
+      stack_data = JSON.parse(payload)
+      @@logger.debug("Payload: #{stack_data.inspect}")
 
       begin
-        service = {}
-        @@logger.debug "Planning deployment of: #{service_data}"
-        service = @service_controller.plan_deployment(service_data)
-        @@logger.debug "Deployed service #{service.to_json}"
+        stack = {}
+        @@logger.debug "Planning deployment of: #{stack_data}"
+        stack = @stack_controller.plan_deployment(stack_data)
+        @@logger.debug "Deployed service #{stack.to_json}"
 
-        @service_controller.converge(service)
+        @stack_controller.converge(stack)
 
-        service.stacks.containers.each do |container|
+        stack.containers.each do |container|
           @container_controller.schedule(container, config['scheduler']['interval'])
         end
 
@@ -142,7 +142,7 @@ module AutoScaling
         @@logger.error e
         raise e
       end
-      service.attributes
+      stack.attributes
     end
 
     private

### 2013-12-11 auto-scaling/cloud-controller/test/integration/cloud_controller_test.rb

@@ -32,19 +32,14 @@ module AutoScaling
     def test_shall_deploy_a_service
 
       msg = {
-        'name' => 'kwejk.pl',
-        'stacks' => [
-          {
-            'type' => 'java',
-            'instances' => 2,
-            'name' => 'stack-name',
-            'policy_set' => {
-                'min_vms' => 0,
-                'max_vms' => 2,
-                'policies' => [{'name' => 'threshold_model', 'parameters' => {'min' => '5', 'max' => '50'} }]
-            }
-          }
-        ]
+        'type' => 'java',
+        'instances' => 2,
+        'name' => 'stack-name',
+        'policy_set' => {
+            'min_vms' => 0,
+            'max_vms' => 2,
+            'policies' => [{'name' => 'threshold_model', 'parameters' => {'min' => '5', 'max' => '50'} }]
+        }
       }
 
       instance_id = 144
@@ -70,10 +65,6 @@ module AutoScaling
       FakeWeb.register_uri(:post, "http://#{master[:ip]}:4567/chef", :status => ["200", "OK"])
       FakeWeb.register_uri(:post, "http://#{slave[:ip]}:4567/chef", :status => ["200", "OK"])
 
-      service_attributes = @cloud_controller.handle_deploy_request(nil, msg.to_json)
-      service_id = service_attributes[:id]
-      service = Service.get(service_id)
-      assert_equal :converged, service.status
 
       # monitoring data
       # note that it is quite hard to mock xmlrpc (one endpoint is common to all request)
@@ -88,6 +79,12 @@ module AutoScaling
                             {:status => ["200", "OK"], :body => response, :content_type => "text/xml"},
                             {:status => ["200", "OK"], :body => master[:response], :content_type => "text/xml"},
                             {:status => ["200", "OK"], :body => slave[:response], :content_type => "text/xml"}])
+
+      stack_attributes = @cloud_controller.handle_deploy_request(nil, msg.to_json)
+      stack_id = stack_attributes[:id]
+      stack = Stack.get(stack_id)
+      assert_equal :converged, stack.state
+
 
       sleep 10
     end

### 2013-12-11 auto-scaling/stack-controller/lib/executor/stack_executor.rb

@@ -116,6 +116,7 @@ module AutoScaling
             @@logger.debug "Successfully sent configuration #{response}"
             # this is simplification - master not necessarily is the last deployed container
             container.stack.state = :converged
+            container.stack.save
             response
           else
             raise RuntimeError, "An error occurred during sending configuration, status: #{response.code}"

### 2013-12-11 auto-scaling/stack-controller/lib/monitor/stack_monitor.rb

@@ -17,8 +17,8 @@ module AutoScaling
     #
     # * *Args* :
     # - +service+ -> reference to a service. An instance of AutoScaling::Service class
-    def monitor(service)
-      @@logger.debug "Monitoring a service #{service}"
+    def monitor(stack)
+      @@logger.debug "Monitoring a service #{stack}"
 
       # check all vms that forms a service
       data = {}

### 2013-12-11 auto-scaling/stack-controller/lib/planner/stack_planner.rb

@@ -1,7 +1,7 @@
 require 'rubygems'
 require 'logger'
 require 'executor/stack_executor'
-require 'planner/reservation_manager'
+require 'common/common'
 
 module AutoScaling
 

### 2013-12-11 auto-scaling/stack-controller/stack_controller.rb

@@ -42,11 +42,9 @@ module AutoScaling
       @planner.plan_deployment(service_data)
     end
 
-    def converge(service)
-      service.stacks.each do |stack|
-        master = Container.master(stack)
-        @executor.converge(master.id)
-      end
+    def converge(stack)
+      master = Container.master(stack)
+      @executor.converge(master.id)
     end
 
     # Checks if resource reservation is possible
@@ -81,10 +79,10 @@ module AutoScaling
       reservation_manager = ReservationManager.new(cloud_provider, capacity)
       planner = StackPlanner.new(executor, cloud_controller, reservation_manager)
 
-      service_controller = StackController.new(monitor, analyzer, planner, executor)
-      cloud_controller.service_controller = service_controller
+      stack_controller = StackController.new(monitor, analyzer, planner, executor)
+      cloud_controller.stack_controller = stack_controller
 
-      return service_controller
+      return stack_controller
     end
 
   end

### 2013-12-12 thesis/src/chapters/chapter-5-design.tex
Add requirements section

\chapter{Design of Cloud-SAP}

\chapterintro{ This chapter introduces the high-level design of Cloud-SAP.
  Further elaboration on the requirements of the proposed platform is presented
  and detailed discussion about the each its layer is held.
}

\section{Requirements}

\subsection{Functional}
One can notice that elements that yields a solution for
a problem stated in the first chapter, which is ensuring that users'
application provide appropriate Quality-of-Service for its customers, were
introduced in previous chapters:

\begin{itemize}
	\item scalability           -- ability to improve application performance by enriching
	\item adaptivity            -- ability to adapt (i.e. scale) appropriately to current usage pattern
	\item inter-cloud awareness -- ability to cooperate with different cloud provider to supply application with extra resources
\end{itemize}
Having those in mind, we can make the list of functional requirements more formal:
\begin{enumerate}
  \item The user of the platform is able to:
    \begin{enumerate}
      \item deploy a service,
      \item cancel the service,
      \item check the status of the previously ordered-to-deploy service at any time. \emph{Status} means 
        \begin{inparaenum}[a)]
        \item whether or not the deployment succeeded,
        \item current uptime of the service,
        \item current cost
        \end{inparaenum}
    \end{enumerate}
  \item One of the elements of the platform is a client application that is used by the user of the platform to communicate with it,
  \item During the deployment process, the platform takes as an input a description of the service (application) that consists of:
    \begin{itemize}
    \item service name,
    \item software stacks (e.g. \emph{java}, \emph{ruby}),
    \item auto-scaling policies (per each stack) which define
      \begin{inparaenum}[i)]
      \item minimal and maximal number of VMs that are needed for the stack,
      \item name of the policy (algorithm) which is used for scaling,
      \item parameters of the policy
      \end{inparaenum}
    \end{itemize}
  \item Deployment of a service is done in a way which minimizes the cost from the client's perspective with ensuring Quality-of-Service requirements at the same time,
  \item It is assumed that the application which is going to be deployed is properly and fully tuned so that it is not possible to improve its performance by changing its or any of its components configuration(s),
  \item The platform monitors the state of the deployed services and based on the results of this process takes appropriate steps in order to meet the auto-scaling requirements. These include
    \begin{inparaenum}[a)]
    \item altering VM's parameters and configuration,
    \item vertical scaling,
    \item horizontal scaling,
    \item scaling stacks among different cloud providers
    \end{inparaenum}
\end{enumerate}

TODO Alternative scenario -- the client has a predefined budget that they cannot exceed -- it can be mentioned in the overall discussion of the solution

\subsection{Non-functional}
\begin{itemize}
  \item The platform uses \emph{OpenVZ} as a hypervisor
  \item The platform uses \emph{OpenNebula} and \emph{OneApps} as data-center management tools
  \item The platform does not confine itself to one provider, but to a \emph{ecosystem of various cloud providers} that offers deployment capabilities which vary in terms of quality of service, cost, etc.
  \item All communication between the user and the platform and among platform components should be encrypted
\end{itemize}

\section{High-level design}
\subsection{Elements of the architecture}
In the previous chapters different scalability models were introduced where each operated at a different layer of abstraction: 
\begin{itemize}
	\item application platform tuning
	\item single-server level (vertical scaling)
	\item multi-server level (horizontal scaling) 
	\item inter-cloud level (scaling out across different cloud providers)
\end{itemize}

To reduce the system design complexity it is vital that proposed solution's
architecture reflects that multi-layered model. Thus, we divided platform's components into a stack that is shown in figure \ref{design:csap-layers}.

\begin{figure}[!ht]
  \begin{center}
    \includegraphics[width=0.5\textwidth]{chapter-5/sap-layers}
  \end{center}
  \caption{Layered structure of Cloud-SAP}
  \label{design:csap-layers}
\end{figure}

\subsection{Components autonomy}
Each layer of the Cloud-SAP (shown in figure \ref{design:csap-layers}) must be characterized by an ability to adapt to a given application usage. As the previous chapter stated, this can be achieved by adding an elasticity controller to each layer. This observation is a foundation of proposed architecture.

One of the first models that ensured system adaptivity is an autonomic component, concept based on a feed-back loop, initially proposed by IBM \cite{IBM06}. Figure \ref{ch5:autonomic-component} depicts that architecture.

\begin{figure}[!ht]
  \begin{center}
    \includegraphics{chapter-5/autonomic-component}
  \end{center}
  \caption{autonomic component}
  \label{ch5:autonomic-component}
\end{figure}

Since designed platform operates on multiple layers, we can extend that model by using concept of a multi-hierarchical autonomic system \cite{LiWoZh05}. Figure \ref{ch5:hierarchical-autonomic-system} illustrates that hierarchy, were each level represents a different perspective at application scaling. Using terminology characteristic to a autonomic system, first three hierarchy levels (application tuning, single-server, multi-server) are centralized and controlled by single elasticity controller at each level, while the last inter-cloud level is a decentralized one - each cloud instance is fully independent.

\begin{figure}[!ht]
  \begin{center}
    \includegraphics{chapter-5/hierarchical-autonomic-system}
  \end{center}
  \caption{Hierarchical autonomic system}
  \label{ch5:hierarchical-autonomic-system}
\end{figure}

As \cite{IBM06} states, each autonomic component has modules that are responsible for:
\begin{itemize}
	\item monitoring
	\item analysis
	\item planning
	\item action execution
\end{itemize}

While the managed component has:
\begin{itemize}
	\item sensors
	\item effectors
\end{itemize}

Due to the hierarchy of our architecture, each level manages an underlying autonomic component, while being managed by an upper layer at the same time. This lead us to observation that they are autonomic components and managed components at the same time. Remaining of this chapter discusses in detail each layer taking into account that observation.

### 2013-12-12 thesis/src/chapters/chapter-6-implementation.tex
Add requirements section

\chapter{Implementation}

\chapterintro{
  In this chapter we outline implementation details about each component of the
  proposed solution.
}

\section{Application controller}
\section{Container controller}
\section{Service controller}
\section{Cloud controller, cloud broker and client application}

### 2013-12-16 thesis/src/chapters/chapter-7-evaluation.tex
Add deployment time test description

--- 2013-12-06 thesis/src/chapters/chapter-7.tex
@@
 
-\section{Service deployment}
+\section{Cost of service deployment}
 \subsection{Description and preconditions}
@@
 
+\section{Deployment time -- solution comparison}
+\subsection{Motivation}
+In this test we want to compare our solution to one of those available at the market which use \emph{OpenNebula} as an underlying tool for managing resources of a data center in terms of \textbf{deployment time}, one of the most important factors of products whose main purpose is to scale applications.
+\emph{Carina} \cite{Carina} can be considered a perfect match of a solution for such a comparison and tests are ran against it.
+\subsection{Description and preconditions}
+\subsubsection{Service description}
+The service comprises a simple java enterprise application, deployed in a master-slave configuration with one VM set as a load balancer and other nodes that serve as workers, which uses Tomcat as a web container.
+
+TODO dodać konfigurację każdej z maszyn wirtualnych
+\subsubsection{Hardware configuration}
+All virtual machines were deployed on one node which had 3G RAM, AMD Athlon\texttrademark 64 X2 Dual Core Processor with each core of 2000 MHz frequency and a hard drive with a capacity of 160 GB.
+\subsection{Results}
+
+\section{Deployment time -- hypervisor comparison}
+
 \section{Auto-scaling}

### 2013-12-17 thesis/src/chapters/chapter-7-evaluation.tex
Supply test case description with a figure and table of average deployment time for C-SAP

@@
 \subsection{Motivation}
-In this test we want to compare our solution to one of those available at the market which use \emph{OpenNebula} as an underlying tool for managing resources of a data center in terms of \textbf{deployment time}, one of the most important factors of products whose main purpose is to scale applications.
+In this test we want to compare our solution to one of those available at the market which use \emph{OpenNebula} as an underlying tool for managing resources of a data center and \emph{OpenVZ} as a hypervisor in terms of \textbf{deployment time}, one of the most important factors of products whose main purpose is to scale applications.
 \emph{Carina} \cite{Carina} can be considered a perfect match of a solution for such a comparison and tests are ran against it.
 \subsection{Description and preconditions}
+This test involves the steps of
+  \begin{inparaenum}[i)]
+    \item instantiating one of the tested product, i.e. Cloud-SAP or Carina,
+    \item ordering the deployment of a service whose specification is shown in listing \ref{lst:service-spec-test-deployment-time},
+    \item measuring the time needed to set up the environment of the service.
+  \end{inparaenum}
+Before performing any tests all virtual machines present at the deployment node are removed.
 \subsubsection{Service description}
@@
 \subsection{Results}
+Obtained results are shown in table \ref{tbl:test-service-deployment-time} and in figure \ref{ch7:deployment-time-test}.
+
+\begin{table}
+  \centering
+  \begin{tabular}{ c  c  c }
+    \hline
+    & \multicolumn{2}{c}{Solution} \\
+    \cline{2-3}
+    Instance no & Cloud-SAP & Carina \\
+    \hline
+    2 & 103.9s & x \\
+    3 & 166.3s & x \\
+    4 & 225.5s & x \\
+    5 & 288.3s & x \\
+    \hline
+  \end{tabular}
+  \caption{Average deployment time for the service with the various number of VMs used for the whole environment}
+  \label{tbl:test-service-deployment-time}
+\end{table}
+
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics{chapter-7/deployment-time-test}
+  \end{center}
+  \caption{Average deployment time for two competing products when the variable is the number of instances of VMs}
+  \label{ch7:deployment-time-test}
+\end{figure}
 

### 2013-12-17 auto-scaling/cloud-controller/lib/cloud_controller.rb
Change default config to point to an external AMQP queue

@@ -102,7 +102,7 @@ module AutoScaling
         amqp_conf = config['amqp']
 
         AMQP.connect(:host => amqp_conf['host']) do |connection|
-          @@logger.info('Connected to AMQP')
+          @@logger.info("Connected to AMQP (host: #{amqp_conf['host']})")
           channel = AMQP::Channel.new(connection)
 
           @publisher.exchange = channel.default_exchange

### 2013-12-22 thesis/src/chapters/chapter-5-design.tex
Thesis: Changed sap-layers drawing, moved requirements to an implementation section

@@
 \chapterintro{ This chapter introduces the high-level design of Cloud-SAP.
-  Further elaboration on the requirements of the proposed platform is presented
-  and detailed discussion about the each its layer is held.
+  Further elaboration on the layers of the proposed platform is presented
+  and detailed discussion about the each layer is held.
 }
@@
 
-\subsection{Functional}
-One can notice that elements that yields a solution for
-a problem stated in the first chapter, which is ensuring that users'
-application provide appropriate Quality-of-Service for its customers, were
-introduced in previous chapters:
-
+\section{High-level design}
+\subsection{Elements of the architecture}
+In the previous chapters, various scalability models operating at different layers of abstraction were introduced: 
 \begin{itemize}
-	\item scalability           -- ability to improve application performance by enriching
-	\item adaptivity            -- ability to adapt (i.e. scale) appropriately to current usage pattern
-	\item inter-cloud awareness -- ability to cooperate with different cloud provider to supply application with extra resources
-\end{itemize}
-Having those in mind, we can make the list of functional requirements more formal:
-\begin{enumerate}
-  \item The user of the platform is able to:
-    \begin{enumerate}
-      \item deploy a service,
-      \item cancel the service,
-      \item check the status of the previously ordered-to-deploy service at any time. \emph{Status} means 
-        \begin{inparaenum}[a)]
-        \item whether or not the deployment succeeded,
-        \item current uptime of the service,
-        \item current cost
-        \end{inparaenum}
-    \end{enumerate}
-  \item One of the elements of the platform is a client application that is used by the user of the platform to communicate with it,
-  \item During the deployment process, the platform takes as an input a description of the service (application) that consists of:
-    \begin{itemize}
-    \item service name,
-    \item software stacks (e.g. \emph{java}, \emph{ruby}),
-    \item auto-scaling policies (per each stack) which define
-      \begin{inparaenum}[i)]
-      \item minimal and maximal number of VMs that are needed for the stack,
-      \item name of the policy (algorithm) which is used for scaling,
-      \item parameters of the policy
-      \end{inparaenum}
-    \end{itemize}
-  \item Deployment of a service is done in a way which minimizes the cost from the client's perspective with ensuring Quality-of-Service requirements at the same time,
-  \item It is assumed that the application which is going to be deployed is properly and fully tuned so that it is not possible to improve its performance by changing its or any of its components configuration(s),
-  \item The platform monitors the state of the deployed services and based on the results of this process takes appropriate steps in order to meet the auto-scaling requirements. These include
-    \begin{inparaenum}[a)]
-    \item altering VM's parameters and configuration,
-    \item vertical scaling,
-    \item horizontal scaling,
-    \item scaling stacks among different cloud providers
-    \end{inparaenum}
-\end{enumerate}
-
-TODO Alternative scenario -- the client has a predefined budget that they cannot exceed -- it can be mentioned in the overall discussion of the solution
-
-\subsection{Non-functional}
-\begin{itemize}
-  \item The platform uses \emph{OpenVZ} as a hypervisor
-  \item The platform uses \emph{OpenNebula} and \emph{OneApps} as data-center management tools
-  \item The platform does not confine itself to one provider, but to a \emph{ecosystem of various cloud providers} that offers deployment capabilities which vary in terms of quality of service, cost, etc.
-  \item All communication between the user and the platform and among platform components should be encrypted
+	\item application platform: application platform tuning
+	\item container: vertical scaling
+	\item stack: horizontal scaling
+	\item inter-cloud: scaling out across different cloud providers
 \end{itemize}
 
-\section{High-level design}
-\subsection{Elements of the architecture}
-In the previous chapters different scalability models were introduced where each operated at a different layer of abstraction: 
-\begin{itemize}
-	\item application platform tuning
-	\item single-server level (vertical scaling)
-	\item multi-server level (horizontal scaling) 
-	\item inter-cloud level (scaling out across different cloud providers)
-\end{itemize}
-
-To reduce the system design complexity it is vital that proposed solution's
-architecture reflects that multi-layered model. Thus, we divided platform's components into a stack that is shown in figure \ref{design:csap-layers}.
+It is vital that proposed architecture reflects that multi-layered model. Thus, we divided platform's components into a stack that is shown in figure \ref{design:csap-layers}.
 
@@
 
-\subsection{Components autonomy}
+\subsection{Autonomic components}
 Each layer of the Cloud-SAP (shown in figure \ref{design:csap-layers}) must be characterized by an ability to adapt to a given application usage.
@@
   \caption{autonomic component}
-  \label{ch5:autonomic-component}
+  \label{design:autonomic-component}
 \end{figure}
@@
 Since designed platform operates on multiple layers, we can extend that model by using concept of a multi-hierarchical autonomic system \cite{LiWoZh05}.
-Figure \ref{ch5:hierarchical-autonomic-system} illustrates that hierarchy, were each level represents a different perspective at application scaling.
-Using terminology characteristic to a autonomic system, first three hierarchy levels (application tuning, single-server, multi-server) are centralized and controlled by single elasticity controller at each level, while the last inter-cloud level is a decentralized one - each cloud instance is fully independent.
-
-\begin{figure}[!ht]
-  \begin{center}
-    \includegraphics{chapter-5/hierarchical-autonomic-system}
-  \end{center}
-  \caption{Hierarchical autonomic system}
-  \label{ch5:hierarchical-autonomic-system}
-\end{figure}
+Figure \ref{ch5:hierarchical-autonomic-system} illustrates that hierarchy, were each level represents a different perspective on an application scaling.
+Using terminology characteristic to a autonomic system, first three hierarchy levels (application tuning, container, stack) are centralized and controlled by a single elasticity controller at each level, while the last inter-cloud level is a decentralized one - each cloud instance is fully independent.
 
@@
 
-Due to the hierarchy of our architecture, each level manages an underlying autonomic component, while being managed by an upper layer at the same time.
-This lead us to observation that they are autonomic components and managed components at the same time.
-Remaining of this chapter discusses in detail each layer taking into account that observation.
+Due to the hierarchy of our architecture, each level manages an underlying autonomic component, while being managed by an upper layer at the very same time.

### 2013-12-22 thesis/src/chapters/chapter-6-implementation.tex
Thesis: Changed sap-layers drawing, moved requirements to an implementation section

@@
 
+\section{Requirements}
+\subsection{Functional}
+One can notice that elements that yields a solution for
+a problem stated in the first chapter, which is ensuring that users'
+application provide appropriate Quality-of-Service for its customers, were
+introduced in previous chapters:
+
+\begin{itemize}
+	\item scalability           -- ability to improve application performance by enriching
+	\item adaptivity            -- ability to adapt (i.e. scale) appropriately to current usage pattern
+	\item inter-cloud awareness -- ability to cooperate with different cloud provider to supply application with extra resources
+\end{itemize}
+Having those in mind, we can make the list of functional requirements more formal:
+\begin{enumerate}
+  \item The user of the platform is able to:
+    \begin{enumerate}
+      \item deploy a service,
+      \item cancel the service,
+      \item check the status of the previously ordered-to-deploy service at any time. \emph{Status} means 
+        \begin{inparaenum}[a)]
+        \item whether or not the deployment succeeded,
+        \item current uptime of the service,
+        \item current cost
+        \end{inparaenum}
+    \end{enumerate}
+  \item One of the elements of the platform is a client application that is used by the user of the platform to communicate with it,
+  \item During the deployment process, the platform takes as an input a description of the service (application) that consists of:
+    \begin{itemize}
+    \item service name,
+    \item software stacks (e.g. \emph{java}, \emph{ruby}),
+    \item auto-scaling policies (per each stack) which define
+      \begin{inparaenum}[i)]
+      \item minimal and maximal number of VMs that are needed for the stack,
+      \item name of the policy (algorithm) which is used for scaling,
+      \item parameters of the policy
+      \end{inparaenum}
+    \end{itemize}
+  \item Deployment of a service is done in a way which minimizes the cost from the client's perspective with ensuring Quality-of-Service requirements at the same time,
+  \item It is assumed that the application which is going to be deployed is properly and fully tuned so that it is not possible to improve its performance by changing its or any of its components configuration(s),
+  \item The platform monitors the state of the deployed services and based on the results of this process takes appropriate steps in order to meet the auto-scaling requirements. These include
+    \begin{inparaenum}[a)]
+    \item altering VM's parameters and configuration,
+    \item vertical scaling,
+    \item horizontal scaling,
+    \item scaling stacks among different cloud providers
+    \end{inparaenum}
+\end{enumerate}
+
+TODO Alternative scenario -- the client has a predefined budget that they cannot exceed -- it can be mentioned in the overall discussion of the solution
+
+\subsection{Non-functional}
+\begin{itemize}
+  \item The platform uses \emph{OpenVZ} as a hypervisor
+  \item The platform uses \emph{OpenNebula} and \emph{AppFlow} as data-center management tools
+  \item The platform does not confine itself to one provider, but to a \emph{ecosystem of various cloud providers} that offers deployment capabilities which vary in terms of quality of service, cost, etc.
+  \item All communication between the user and the platform and among platform components should be encrypted
+\end{itemize}
+
 \section{Application controller}

### 2013-12-23 thesis/src/appendices/listings.tex
Supply test case description with a figure and table of average deployment time for C-SAP
New deployment time figure

@@
 \section{Service specifications}
+\lstinputlisting[caption=Carina Environment Specification which was used during tests of deployment time, label=lst:carina-config]{appendices/carina-svc-config.rb}
+
 \lstinputlisting[caption=Cloud-SAP Service Specification (without scaling policies), label=lst:service-spec-test-cost]{appendices/service-spec-test-cost.json}
+
+\lstinputlisting[caption=Cloud-SAP Service Specification used for testing deployment time, label=lst:service-spec-test-deployment-time]{appendices/service-spec-deployment-time-test.json}
 

### 2013-12-23 thesis/src/chapters/chapter-7-evaluation.tex
Thesis: Changed sap-layers drawing, moved requirements to an implementation section
New deployment time figure

@@
 
+\section{Auto-scaling}
+
 \section{Deployment time -- solution comparison}
@@
 Before performing any tests all virtual machines present at the deployment node are removed.
+\subsubsection{OpenNebula configuration}
+To ensure objectivity in tests, OpenNebula was configured in both products in the same way.
+One of the key factors that could influence the deployment time is the configuration of scheduler.
+Its parameters are shown in the listing \ref{lst:one-scheduler-config}.
+\lstinputlisting[caption=OpenNebula scheduler configuration, label=lst:one-scheduler-config]{deployment-time-test-sched-config}
 \subsubsection{Service description}
 The service comprises a simple java enterprise application, deployed in a master-slave configuration with one VM set as a load balancer and other nodes that serve as workers, which uses Tomcat as a web container.
+
+The description of a service expressed in Carina format can be found in listing \ref{lst:carina-config}.
 
@@
     \hline
-    2 & 103.9s & x \\
-    3 & 166.3s & x \\
-    4 & 225.5s & x \\
-    5 & 288.3s & x \\
+    2 & 198.0 & 158.7s \\
+    3 & 261.72 & 208.1s \\
+    4 & 294.38 & 230.6s \\
     \hline
@@
 \section{Deployment time -- hypervisor comparison}
-
-\section{Auto-scaling}

### 2013-12-23 thesis/assets/listings/appendices/carina-svc-config.rb

@@ -1,0 +1,45 @@
+ENDPOINT = {
+	'mm01' => {
+		:proxy   => 'http://192.168.0.35:2633/RPC2',
+		:oneauth => "svc:xxxxx"
+		}
+	}
+
+
+TEMPLATE = {
+	'tomcat' => {
+		:file       => "~/vm/tomcat.vm",
+		:cpu        => "0.3",
+		:memory     => 512,
+		:network_id => { 'mm01' => 7 },
+		:image_id   => { 'mm01' => 10 }
+	},
+	'haproxy' => {
+		:file       => "~/vm/haproxy.vm",
+		:cpu        => "0.3",
+		:memory     => 512,
+		:network_id => { 'mm01' => 7 },
+		:image_id   => { 'mm01' => 11 }
+	}
+}
+
+
+ENVIRONMENT = {
+	'testenv' => {
+                :type                  => "compute",
+                :endpoint              => "mm01",
+                :description           => "Example environment",
+                :master_template       => "haproxy",
+                :master_context_script => "sample-master-context-script.sh",
+                :master_setup_time     => 30,
+                :master_context_var    => "BALANCE_PORT=8080",
+                :slave_template        => "tomcat",
+                :slave_context_script  => "sample-slave-context-script.sh",
+                :slave_context_var     => "APP_PACKAGE=gwt-petstore.war",
+                :placement_policy      => "pack",
+                :num_slaves            => 3,
+                :slavedata             => "8080",
+                :adminuser             => "root",
+                :app_url               => "http://%MASTER%:8080/testapp"
+	 }
+}

### 2013-12-26 thesis/src/chapters/chapter-5-design.tex
Thesis: figure's label change

@@
 One of the first models that ensured system adaptivity is an autonomic component, concept based on a feed-back loop, initially proposed by IBM \cite{IBM06}.
-Figure \ref{ch5:autonomic-component} depicts that architecture.
+Figure \ref{design:autonomic-component} depicts that architecture.
 
@@
 Since designed platform operates on multiple layers, we can extend that model by using concept of a multi-hierarchical autonomic system \cite{LiWoZh05}.
-Figure \ref{ch5:hierarchical-autonomic-system} illustrates that hierarchy, were each level represents a different perspective on an application scaling.
+Figure \ref{design:hierarchical-autonomic-system} illustrates that hierarchy, were each level represents a different perspective on an application scaling.
 Using terminology characteristic to a autonomic system, first three hierarchy levels (application tuning, container, stack) are centralized and controlled by a single elasticity controller at each level, while the last inter-cloud level is a decentralized one - each cloud instance is fully independent.

### 2013-12-26 auto-scaling/stack-controller/lib/planner/stack_planner.rb

@@ -72,7 +72,7 @@ module AutoScaling
     end
 
     def overloaded_master(stack)
-      @@logger.warn "STACKStack #{stack} has overloaded master. Currently, nothing can be done"
+      @@logger.warn "STACK Stack #{stack} has overloaded master. Currently, nothing can be done"
     end
 
     def healthy(stack)

### 2013-12-26 thesis/src/chapters/chapter-5-design.tex
Thesis: chapter 5 structure

@@
 \section{Requirements}
+One can notice that elements that yields a solution for a problem stated in the first chapter, which is ensuring that users' application provide appropriate Quality-of-Service for its customers in a most-cost effective manner, were gradually introduced in previous chapters:
+
+\begin{itemize}
+	\item \emph{scalability} - ability to improve application performance by enriching resources
+	\item \emph{adaptivity} - ability to adapt (i.e. scale) appropriately to a current usage pattern
+	\item \emph{inter-cloud awareness} - ability to compose an application deployment using different cloud providers; cooperation with different cloud provider to supply application with extra resources while performing application scaling
+\end{itemize}
+
+Next section states the general overview of the proposed solution, while the consecutive sections presents more detailed discussion of its elements and finally the last section summarised the design choices in a context of system requirements.
 
 \section{High-level design}
-\subsection{Elements of the architecture}
+\subsection{Key elements}
 In the previous chapters, various scalability models operating at different layers of abstraction were introduced: 
@@
 Due to the hierarchy of our architecture, each level manages an underlying autonomic component, while being managed by an upper layer at the very same time.
+
+\subsubsection{Monitoring}
+
+\subsubsection{Analysis}
+
+\subsubsection{Planning}
+
+\subsubsection{Execution}
+
+\subsection{Cloud federation}
+Opis najwyzej warstwy, ze ma brokery, wykorzystuje kilka chmur.
+
+\subsection{Architecture overview}
+Taking above-mentioned points into account, we can distinguish two pivotal components responsible for auto-scaling and cooperation in intercloud respectively.
+While each auto-scaling component is entirely dedicated to a given cloud provider, the intercloud module has distributed components such as client and broker and can correspond to a variety of different auto-scaling components.
+Figure X illustrates overall architecture.
+
+(Diagram komponentowy)
+
+\section{Container controller}
+
+\section{Stack controller}
+
+\section{Cloud controller, cloud broker and client application}
+
+\section{Solution discussion}

### 2013-12-27 thesis/src/chapters/chapter-7-evaluation.tex
Thesis: restructured evaluation chapter

@@
 \begin{enumerate}
-  \item maximum reduction of deployment cost by deploying different stacks comprising a given service on different cloud providers
-  \item reduced deployment time
-  \item TODO dopisać jeszcze kolejne
+  \item deployment cost
+  \item cost of providing given Quality-of-Service
+  \item deployment time
 \end{enumerate}
@@
 \section{Cost of service deployment}
-\subsection{Description and preconditions}
+\subsection*{Description}
 The aim of this test case is to test the primary use case of the proposed solution -- deployment of a service with the emphasis of client's \textbf{cost}.
+It should show that the platform chooses the best mapping between the stacks and cloud providers so that the client's pays the \textbf{lowest} possible price.
+
+\subsection*{Preconditions}
 Service specification (\ref{lst:service-spec-test-cost}) forms an input to the application.
@@
 Each cloud provider has its own price for a given software stack which is shown in table \ref{tbl:test-service-deployment-cost}.
-This test should show that the platform chooses the best mapping between the stacks and cloud providers so that the client's pays the \textbf{lowest} possible price.
+Diagram \ref{tbl:test-service-deployment-cost-diagram} illustrates the overall configuration.
 
@@
 
-\subsection{Results}
+(JEB deployment diagram, podzial na wezly)
+
+\subsection*{Results}
 Table \ref{tbl:test-service-deployment-cost-mapping} shows obtained mapping between stacks and cloud providers.
@@
 
-\section{Auto-scaling}
+\subsection*{Conclusion}
+
+\section{Auto-scaling -- single-provider based}
+\subsection*{Description}
+\subsection*{Preconditions}
+\subsection*{Results}
+\subsection*{Conclusion}
+
+\section{Auto-scaling -- multiple-provider based}
+\subsection*{Description}
+\subsection*{Preconditions}
+\subsection*{Results}
+\subsection*{Conclusion}
 
 \section{Deployment time -- solution comparison}
-\subsection{Motivation}
+\subsection*{Description}
 In this test we want to compare our solution to one of those available at the market which use \emph{OpenNebula} as an underlying tool for managing resources of a data center and \emph{OpenVZ} as a hypervisor in terms of \textbf{deployment time}, one of the most important factors of products whose main purpose is to scale applications.
 \emph{Carina} \cite{Carina} can be considered a perfect match of a solution for such a comparison and tests are ran against it.
-\subsection{Description and preconditions}
+
 This test involves the steps of
@@
   \end{inparaenum}
+
+\subsection*{Preconditions}
 Before performing any tests all virtual machines present at the deployment node are removed.
@@
 
-TODO dodać konfigurację każdej z maszyn wirtualnych
+- TODO dodać konfigurację każdej z maszyn wirtualnych
+
 \subsubsection{Hardware configuration}
 All virtual machines were deployed on one node which had 3G RAM, AMD Athlon\texttrademark 64 X2 Dual Core Processor with each core of 2000 MHz frequency and a hard drive with a capacity of 160 GB.
-\subsection{Results}
+Diagram presents the physical configuration of nodes.
+- jeb physical diagram
+
+\subsubsection{Environment configuration}
+Components that were used during tests are shown in figure .
+- jeb logical diagram
+
+\subsection*{Results}
 Obtained results are shown in table \ref{tbl:test-service-deployment-time} and in figure \ref{ch7:deployment-time-test}.
@@
 
+\subsection*{Conclusion}
+
 \section{Deployment time -- hypervisor comparison}
+\subsection*{Description}
+\subsection*{Preconditions}
+\subsection*{Results}
+\subsection*{Conclusion}

### 2013-12-28 thesis/src/chapters/chapter-7-evaluation.tex
Add deployment diagram of Carina for deployment time test

@@
 \subsection*{Preconditions}
-Before performing any tests all virtual machines present at the deployment node are removed.
+It is assumed that \emph{Cloud-SAP} and \emph{Carina} according with \emph{OpenVZ} as an underlying virtualization technology are correctly installed and configured.
+Each test case must be run in an isolation so before performing any test all virtual machines present at the deployment node are removed.
 \subsubsection{OpenNebula configuration}
@@
 \subsubsection{Environment configuration}
-Components that were used during tests are shown in figure .
-- jeb logical diagram
+Deployment diagram for \emph{Carina} implementation is shown in figure \ref{ch7:deployment-time-test-deployment-time-carina-deployment-diagram} and for \emph{Cloud-SAP} in figure \ref{ch7:deployment-time-test-deployment-time-components}.
+
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics{chapter-7/deployment-time-carina-deployment-diagram}
+  \end{center}
+  \caption{Deployment diagram of \emph{Carina}}
+  \label{ch7:deployment-time-test-deployment-time-carina-deployment-diagram}
+\end{figure}
+
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics{chapter-7/deployment-time-components}
+  \end{center}
+  \caption{Deployment diagram of \emph{Carina}}
+  \label{ch7:deployment-time-test-deployment-time-components}
+\end{figure}
 
@@
 Obtained results are shown in table \ref{tbl:test-service-deployment-time} and in figure \ref{ch7:deployment-time-test}.
+For a given number of instances we ordered deploying a service 10 times and the values shown in those figures are an average of these runs.
 

### 2013-12-28 thesis/src/chapters/chapter-5-design.tex
Thesis: added drawings, resolved conflict

@@
 
-Next section states the general overview of the proposed solution, while the consecutive sections presents more detailed discussion of its elements and finally the last section summarised the design choices in a context of system requirements.
+Next section states the general overview of the proposed solution, while the consecutive sections presents more detailed discussion of its elements and finally the last section summarises the design choices in a context of system requirements.
 
 \section{High-level design}
-\subsection{Key elements}
-In the previous chapters, various scalability models operating at different layers of abstraction were introduced: 
+\subsection{Overview}
+
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics[width=0.8\textwidth]{chapter-5/hld-overview}
+  \end{center}
+  \caption{Cloud-SAP high level overview}
+  \label{design:hld-overview}
+\end{figure}
+
+As diagram \ref{design:hld-overview} illustrates, we can distinguish two main components of proposed solution: auto-scaling module and inter-cloud broker.
+Together, they can be seen as an hierarchical autonomic system, where each layer corresponds to a different scalability perspective:
 \begin{itemize}
-	\item application platform: application platform tuning
-	\item container: vertical scaling
-	\item stack: horizontal scaling
-	\item inter-cloud: scaling out across different cloud providers
+	\item application platform layer: application platform tuning
+	\item container layer: vertical scaling
+	\item stack layer: horizontal scaling
+	\item inter-cloud layer: scaling out across different cloud providers
 \end{itemize}
 
-It is vital that proposed architecture reflects that multi-layered model. Thus, we divided platform's components into a stack that is shown in figure \ref{design:csap-layers}.
+Diagram \ref{design:csap-layers} depicts that multi-layered structure.
 
@@
 
-\subsection{Autonomic components}
-Each layer of the Cloud-SAP (shown in figure \ref{design:csap-layers}) must be characterized by an ability to adapt to a given application usage.
+\subsection{Auto-scaling module}
+
+\subsubsection{Autonomic components}
+Each layer of the Cloud-SAP (shown in figure \ref{design:csap-layers}) must be characterised by an ability to adapt to a given application usage.
 As the previous chapter stated, this can be achieved by adding an elasticity controller to each layer.
@@
 Figure \ref{design:hierarchical-autonomic-system} illustrates that hierarchy, were each level represents a different perspective on an application scaling.
-Using terminology characteristic to a autonomic system, first three hierarchy levels (application tuning, container, stack) are centralized and controlled by a single elasticity controller at each level, while the last inter-cloud level is a decentralized one - each cloud instance is fully independent.
+Using terminology characteristic to a autonomic system, first three hierarchy levels (application tuning, container, stack) are centralised and controlled by a single elasticity controller at each level, while the last inter-cloud level is a decentralized one - each cloud instance is fully independent.
+
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics[width=0.8\textwidth]{chapter-5/hierarchical-autonomic-system}
+  \end{center}
+  \caption{Cloud-SAP as an hierarchical autonomic system}
+  \label{design:hierarchical-autonomic-system}
+\end{figure}
 
@@
 
-\subsection{Architecture overview}
-Taking above-mentioned points into account, we can distinguish two pivotal components responsible for auto-scaling and cooperation in intercloud respectively.
-While each auto-scaling component is entirely dedicated to a given cloud provider, the intercloud module has distributed components such as client and broker and can correspond to a variety of different auto-scaling components.
-Figure X illustrates overall architecture.
+\section{Auto-scaling module}
 
-(Diagram komponentowy)
-
-\section{Container controller}
-
-\section{Stack controller}
-
-\section{Cloud controller, cloud broker and client application}
+\section{Cloud federation}
 

### 2013-12-28 thesis/src/chapters/chapter-7-evaluation.tex
Thesis: added drawings, resolved conflict

@@
 Each cloud provider has its own price for a given software stack which is shown in table \ref{tbl:test-service-deployment-cost}.
-Diagram \ref{tbl:test-service-deployment-cost-diagram} illustrates the overall configuration.
+Diagram \ref{eval:deployment-cost-physical-nodes} illustrates the simplified environment setup.
 
@@
 
-(JEB deployment diagram, podzial na wezly)
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics[width=0.6\textwidth]{chapter-7/deployment-cost-physical-nodes}
+  \end{center}
+  \caption{Deployment cost: environment configuration}
+  \label{eval:deployment-cost-physical-nodes}
+\end{figure}
 
@@
 \subsection*{Preconditions}
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics[width=0.8\textwidth]{chapter-7/auto-scaling-1cp-physical-nodes}
+  \end{center}
+  \caption{Auto-scaling - single-provider based: environment configuration}
+  \label{eval:auto-scaling-1cp-physical-nodes}
+\end{figure}
+
 \subsection*{Results}
@@
 \subsection*{Preconditions}
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics[width=0.8\textwidth]{chapter-7/auto-scaling-2cp-physical-nodes}
+  \end{center}
+  \caption{Auto-scaling - multiple-provider based: environment configuration}
+  \label{eval:auto-scaling-1cp-physical-nodes}
+\end{figure}
+
 \subsection*{Results}
@@
 Diagram presents the physical configuration of nodes.
-- jeb physical diagram
+
++ fizycznych
+
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics[width=0.6\textwidth]{chapter-7/deployment-time-physical-nodes}
+  \end{center}
+  \caption{Deployment time - solution comparison: physical environment setup}
+  \label{eval:deployment-time-physical-nodes}
+\end{figure}
 
@@
   \begin{center}
-    \includegraphics{chapter-7/deployment-time-components}
+    \includegraphics[width=\textwidth]{chapter-7/deployment-time-components}
   \end{center}
-  \caption{Deployment diagram of \emph{Carina}}
+  \caption{Deployment diagram of \emph{Cloud-SAP}}
   \label{ch7:deployment-time-test-deployment-time-components}

### 2013-12-28 thesis/src/chapters/chapter-7-evaluation.tex
Change deployment diagram of csap for deployment time test

@@
 \subsubsection{Environment configuration}
-Deployment diagram for \emph{Carina} implementation is shown in figure \ref{ch7:deployment-time-test-deployment-time-carina-deployment-diagram} and for \emph{Cloud-SAP} in figure \ref{ch7:deployment-time-test-deployment-time-components}.
+Deployment diagram for \emph{Carina} implementation is shown in figure \ref{ch7:deployment-time-test-deployment-time-carina-deployment-diagram} and for \emph{Cloud-SAP} in figure \ref{ch7:deployment-time-test-cloud-sap-deployment-diagram}.
 
@@
   \begin{center}
-    \includegraphics[width=\textwidth]{chapter-7/deployment-time-components}
+    \includegraphics{chapter-7/deployment-time-test-cloud-sap-deployment-diagram}
   \end{center}
   \caption{Deployment diagram of \emph{Cloud-SAP}}
-  \label{ch7:deployment-time-test-deployment-time-components}
+  \label{ch7:deployment-time-test-cloud-sap-deployment-diagram}
 \end{figure}

### 2013-12-28 thesis/src/chapters/chapter-7-evaluation.tex
Add hardware and vms configuration in the beginning of the evaluation chapter

@@
 \end{enumerate}
+
+\subsection*{Hardware and virtual machines configuration}
+Many of the carried out tests were run on the same hardware and/or used the same virtual machines so for clarity of presentation their parameters are presented not in a description of every test, but here in table \ref{tbl:test-deployment-time-common-hardware-configuration}.
+
+\begin{table}
+  \centering
+  \begin{tabular}{ | l | l | l | l | }
+    \hline                        
+    Name & CPU & RAM & HD \\
+    \hline
+    (device) Desktop        & Intel(R) Core(TM) i5-2500 CPU @ 3.30GHz & 8GB & 1TB \\
+    (device) Node1        & AMD Athlon\texttrademark 64 X2 Dual Core, 2000MHz & 3GB & 160GB \\
+    (device) Node3        & AMD Sempron(tm) Processor 3000+ & 2.5GB & 160GB \\
+    (device) Node4        & AMD Duron(tm) Processor 1GHz& 2.5GB & 60GB \\
+    (virtual machine) Frontend\{1,2\}  & 1 CPU & 512MB & 10GB \\
+    \hline  
+  \end{tabular}
+  \caption{Configuration of hardware/virtual machines used during tests}
+  \label{tbl:test-deployment-time-common-hardware-configuration}
+\end{table}
 
@@
 \subsubsection{Hardware configuration}
-All virtual machines were deployed on one node which had 3G RAM, AMD Athlon\texttrademark 64 X2 Dual Core Processor with each core of 2000 MHz frequency and a hard drive with a capacity of 160 GB.
+All virtual machines were deployed on a host named \emph{Node1}, whose configuration can be found in table \ref{tbl:test-deployment-time-common-hardware-configuration}.
 Diagram presents the physical configuration of nodes.
@@
     \hline
-    2 & 198.0 & 158.7s \\
-    3 & 261.72 & 208.1s \\
-    4 & 294.38 & 230.6s \\
+    2 & 198.0 & 158.7 \\
+    3 & 261.72 & 208.1 \\
+    4 & 294.38 & 230.6 \\
     \hline

### 2013-12-28 thesis/src/chapters/chapter-5-design.tex
Thesis: removed redundant table

@@
 
-\subsection{Auto-scaling module}
+\subsection{Auto-scaling subsystem}
 
@@
 
-Since designed platform operates on multiple layers, we can extend that model by using concept of a multi-hierarchical autonomic system \cite{LiWoZh05}.
+\subsubsection{Monitoring}
+Monitoring module of each autonomic component aims to gather data from Sensors of a managed resource.
+While design of Cloud-SAP doesn't have any specific requirements regarding underlying monitoring mechanism there are a few aspects that should be taken into consideration.
+Table \ref{tab:monitoring-module-issues-summary} summarises main points.
+
+\begin{table}[!htbp]
+\begin{tabularx}{\textwidth}[]{ X X X }
+\specialrule{.1em}{.05em}{.05em}
+ & \textbf{Advantages} & \textbf{Disadvantages} \\ \specialrule{.1em}{.05em}{.05em}
+
+\textbf{data exchange model} & & \\ \specialrule{.1em}{.05em}{.05em}
+push based &
+-- notification driven: all changes are reflected immediately in the system
+&
+-- requires additional agent at monitored component
+\\ \hline
+pull based & 
+-- more calculations are being done at monitoring module side (doesn't involve additional load at managed component)
+& 
+-- some monitoring cycles can be redundant if scheduler is not properly tuned
+\\ \hline
+
+\textbf{data format} & & \\ \specialrule{.1em}{.05em}{.05em}
+binary &
+-- space-effective
+
+-- unreadable for human
+& \\ \hline
+character-based &
+-- human readable
+&
+-- ineffective in terms of space
+\\ \hline
+
+\textbf{communication protocol} & & \\ \specialrule{.1em}{.05em}{.05em}
+vendor specific &
+-- provides specific data to a given managed component (ie. virtual machine / container)
+& 
+-- may cause incompatibility problem while expanding cloud-sap system to a consecutive cloud instances
+\\ \hline
+standard-based &
+-- supports scaling across multiple cloud instances as long as they support standard
+&
+-- gathered data may be inadequate to a specific needs
+\\ \hline
+\end{tabularx}
+
+\caption{Monitoring - summary of issues}
+\label{tab:monitoring-module-issues-summary}
+
+\end{table}
+
+Cloud-SAP is entirely focused on further data processing, consequently whether to store monitoring data or not is entirely implementation dependant.
+
+Among the currently available monitoring systems remarkable are Cloud Watch \cite{CloudWatch}, OpenNebula (Information Manager / Virtual Machine Manager) \cite{OpenNebula} there are also open approaches such as Ganglia Monitoring System \cite{MaChCu04}.
+
+\subsubsection{Analysis}
+
+\subsubsection{Planning}
+
+\subsubsection{Execution}
+
+\subsubsection{Knowledge base}
+
+\subsubsection{Hierarchical autonomic-system}
+Due to the fact that designed platform operates on multiple layers, we can extend idea of an autonomic component by using concept of a multi-hierarchical autonomic system \cite{LiWoZh05}.
 Figure \ref{design:hierarchical-autonomic-system} illustrates that hierarchy, were each level represents a different perspective on an application scaling.
@@
 
-As \cite{IBM06} states, each autonomic component has modules that are responsible for:
-\begin{itemize}
-	\item monitoring
-	\item analysis
-	\item planning
-	\item action execution
-\end{itemize}
-
-While the managed component has:
-\begin{itemize}
-	\item sensors
-	\item effectors
-\end{itemize}
-
 Due to the hierarchy of our architecture, each level manages an underlying autonomic component, while being managed by an upper layer at the very same time.
-
-\subsubsection{Monitoring}
-
-\subsubsection{Analysis}
-
-\subsubsection{Planning}
-
-\subsubsection{Execution}
 

### 2013-12-28 thesis/src/chapters/chapter-7-evaluation.tex
Thesis: removed redundant table

@@
 \subsubsection{Hardware configuration}
+
 All virtual machines were deployed on a host named \emph{Node1}, whose configuration can be found in table \ref{tbl:test-deployment-time-common-hardware-configuration}.
 Diagram presents the physical configuration of nodes.
-
-+ fizycznych
 

### 2013-12-28 thesis/src/chapters/chapter-5-design.tex
Thesis: Restated HLD overview

@@
 
-As diagram \ref{design:hld-overview} illustrates, we can distinguish two main components of proposed solution: auto-scaling module and inter-cloud broker.
-Together, they can be seen as an hierarchical autonomic system, where each layer corresponds to a different scalability perspective:
+Diagram \ref{design:hld-overview} illustrates exemplary Cloud-SAP deployment, among depicted components we can distinguish two main parts of proposed solution: auto-scaling subsystem and inter-cloud broker.
+Auto-scaling subsystem is responsible for scaling users' application, taking into account different scalability perspectives, handled by a different layers:
 \begin{itemize}
@@
 \end{itemize}
+The last mentioned layer - inter-cloud - cooperates with different cloud providers, hence, it requires additional work that has to be done.
+In fact, these extra tasks are delegated to a inter-cloud broker, responsible for a dialog in a cloud federation.
 
@@
 -- space-effective
-
+& 
 -- unreadable for human
-& \\ \hline
+\\ \hline
 character-based &
@@
 Cloud-SAP is entirely focused on further data processing, consequently whether to store monitoring data or not is entirely implementation dependant.
-
-Among the currently available monitoring systems remarkable are Cloud Watch \cite{CloudWatch}, OpenNebula (Information Manager / Virtual Machine Manager) \cite{OpenNebula} there are also open approaches such as Ganglia Monitoring System \cite{MaChCu04}.
 
@@
 
+\subsection{Application platform layer}
+
+\subsection{Container layer}
+Among the currently available monitoring systems remarkable are Cloud Watch \cite{CloudWatch}, OpenNebula (Information Manager / Virtual Machine Manager) \cite{OpenNebula} there are also open approaches such as Ganglia Monitoring System \cite{MaChCu04}.
+
+\subsection{Stack layer}
+
+\subsection{Cloud layer}
+
 \section{Cloud federation}

### 2013-12-28 auto-scaling/cloud-controller/lib/cloud_controller.rb
Remove service id and add unique constraint on name

@@ -124,7 +124,7 @@ module AutoScaling
       service_specification = JSON.parse(payload)
       offer = @service_offer_preparer.prepare_offer(service_specification)
       if offer
-        response = @offer_response_preparer.publishify_offer(offer, :service_id => service_specification['service_id'])
+        response = @offer_response_preparer.publishify_offer(offer, :service_name => service_specification['service_name'])
         respond_with(response, service_specification['offers_routing_key'])
         @@logger.info("Respondend with an offer: #{response}")
       else

### 2013-12-28 auto-scaling/cloud-controller/lib/cloud_controller/offer_response_preparer.rb

@@ -4,10 +4,10 @@ class OfferResponsePreparer
   include Configurable
 
   def publishify_offer(offer, options)
-    raise 'Offer should have set `service id` parameter!' unless options.has_key?(:service_id)
+    raise 'Offer should have set `service name` parameter!' unless options.has_key?(:service_name)
     {
       :controller_id => config['amqp']['controller_routing_key'],
-      :service_id    => options[:service_id],
+      :service_name  => options[:service_name],
       :offers        => offer
     }.to_json
   end

### 2013-12-28 auto-scaling/cloud-controller/spec/lib/cloud_controller/offer_response_preparer_spec.rb

@@ -6,7 +6,7 @@ describe OfferResponsePreparer do
 
   describe 'when has valid input parameters' do
     let(:offer) { [] }
-    let(:options) { { :service_id => 'kubus puchatek' } }
+    let(:options) { { :service_name => 'kubus puchatek' } }
 
     subject { preparer.publishify_offer(offer, options)}
     it { should be_instance_of String }

### 2013-12-28 intercloud/cloud-broker/lib/cloud-broker-amqp/message-handler/autoscaling_message_handler.rb

@@ -10,12 +10,12 @@ class AutoscalingMessageHandler
   end
 
   def handle_message(metadata, payload)
-    @@logger.info("Handling an autoscaling message")
+    @@logger.info("Handling an autoscaling message: #{payload}")
     message        = JSON.parse(payload)
 
-    service_id     = message['service_id']
+    service_name   = message['service_name']
     stack_type     = message['type']
-    stack_to_scale = ServiceSpecification.get(service_id)
+    stack_to_scale = ServiceSpecification.get(service_name)
                                          .stacks(:type => stack_type)
                                          .first
     stack_to_scale.update(:status => :scaling)

### 2013-12-28 intercloud/cloud-broker/lib/cloud-broker-amqp/message-handler/cloud_offer_handler.rb

@@ -10,7 +10,7 @@ class CloudOfferHandler
   #
   # Offer is of the form:
   # {
-  #   'service_id': ID,
+  #   'service_name': ID,
   #   'controller_id': CID,  <- cloud controller routing key
   #   'offers': [
   #     {'cost': X, 'type': Y},
@@ -22,7 +22,7 @@ class CloudOfferHandler
 
     message = JSON.parse(payload)
     @@logger.debug("Message: #{message}")
-    service_specification = ServiceSpecification.get(message['service_id'])
+    service_specification = ServiceSpecification.get(message['service_name'])
     message['offers'].each { |offer| create_offer(service_specification, offer, message['controller_id']) }
   end
 

### 2013-12-28 intercloud/cloud-broker/lib/cloud-broker-amqp/offer_retriever.rb

@@ -24,7 +24,7 @@ class OfferRetriever
       [:type, :instances].reduce({}) { |acc, v| acc[v] = attributes[v]; acc }
     end 
     {
-      :service_id => service_specification.id,
+      :service_name => service_specification.name,
 
       # for a given stack select only limited no of attributes
       :stacks => stacks_attributes,

### 2013-12-28 intercloud/cloud-broker/lib/cloud-broker-ws/cb_ws.rb

@@ -44,6 +44,6 @@ class CloudBrokerWS < Sinatra::Base
     # create stacks for the give service spec
     create_stacks(message['stacks'], service_specification)
     fetch_cloud_offers(service_specification)
-    Rack::Response.new(service_specification.id.to_s, 201)
+    Rack::Response.new(service_specification.name, 201)
   end
 end

### 2013-12-28 intercloud/cloud-broker/lib/models/service_specification.rb

@@ -6,8 +6,8 @@ class ServiceSpecification
   include DataMapper::Resource
   include Configurable
 
-  property :id,              Serial
-  property :name,            String, :required => true
+  #property :id,              Serial
+  property :name,            String, :key => true, :required => true
   #property :broker_id,       String, :required => true, :default => ConfigUtils.load_config['broker_id']
   property :client_endpoint, String, :required => true
   property :created_at,      DateTime, :default => DateTime.now
@@ -19,7 +19,7 @@ class ServiceSpecification
   end
 
   def ready_to_deploy?
-    not deployed?    and
+    (not deployed?)  and
     stacks.count > 0 and
     stacks.reduce(false) { |status, stack| stack.ready_to_deploy? or status}
   end

### 2013-12-28 intercloud/cloud-broker/spec/lib/cloud-broker-amqp/message-handler/autoscaling_message_handler_spec.rb

@@ -1,6 +1,7 @@
 require 'spec_helper'
 require 'models/service_specification'
 require 'cloud-broker-amqp/message-handler/autoscaling_message_handler'
+require 'securerandom'
 
 
 describe AutoscalingMessageHandler do
@@ -39,7 +40,7 @@ describe AutoscalingMessageHandler do
 
   def create_service_spec
     ServiceSpecification.create(
-      :name => 'service name',
+      :name => "service name #{SecureRandom.urlsafe_base64(4)}",
       :client_endpoint => 'pussylord.com:4125',
       :stacks => stacks_attributes
     )
@@ -49,7 +50,7 @@ describe AutoscalingMessageHandler do
   let(:message_attributes) do 
     {
       :type => 'java',
-      :service_id => service_spec.id
+      :service_name => service_spec.name
     }
   end
   let(:message) { message_attributes.to_json }

### 2013-12-28 intercloud/cloud-broker/spec/lib/cloud-broker-amqp/message-handler/cloud_offer_handler_spec.rb

@@ -35,7 +35,7 @@ describe CloudOfferHandler do
 
   def create_service_spec
     ServiceSpecification.create(
-      :name => 'service name',
+      :name => "service name #{SecureRandom.urlsafe_base64(4)}",
       :client_endpoint => 'pussylord.com:4125',
       :stacks => stacks_attributes
     )
@@ -43,7 +43,7 @@ describe CloudOfferHandler do
   let(:service_spec) { create_service_spec }
   let(:message) do
     JSON.generate({
-      :service_id    => service_spec.id,
+      :service_name    => service_spec.name,
       :controller_id => 'cack-controller',
       :offers        => stacks_attributes.map { |sa| { :type => sa[:type], :cost => 24  } }
     })
@@ -51,7 +51,7 @@ describe CloudOfferHandler do
 
   it 'creates offers for stacks of a service specification' do
     subject.handle_message(nil, message)
-    expect(ServiceSpecification.get(service_spec.id).stacks.reduce(0) { |sum, s| sum + s.offers(:examined => false).count }).to be > 0
+    expect(ServiceSpecification.get(service_spec.name).stacks.reduce(0) { |sum, s| sum + s.offers(:examined => false).count }).to be > 0
   end
 
 end

### 2013-12-28 intercloud/cloud-broker/spec/lib/cloud-broker-amqp/offer_retriever_spec.rb

@@ -2,6 +2,7 @@ require 'spec_helper'
 require 'cloud-broker-amqp/offer_retriever'
 require 'models/service_specification'
 require 'models/stack'
+require 'securerandom'
 
 describe OfferRetriever do
   let(:publisher) { double(:publish => true) }
@@ -61,7 +62,7 @@ describe OfferRetriever do
 
   def create_service_spec(stacks_attributes)
     ServiceSpecification.create(
-      :name => 'service name',
+      :name => "service name #{SecureRandom.urlsafe_base64(4)}",
       :client_endpoint => 'pussylord.com:4126',
       :stacks => stacks_attributes
     )

### 2013-12-28 intercloud/cloud-broker/spec/lib/cloud-broker-amqp/service_deployer_spec.rb

@@ -1,6 +1,7 @@
 require 'spec_helper'
 require 'cloud-broker-amqp/service_deployer'
 require 'resource-mapping/offer_selector'
+require 'securerandom'
 
 describe ServiceDeployer do
   let(:publisher) { double('Publisher', :publish => true) }
@@ -8,7 +9,7 @@ describe ServiceDeployer do
 
   let(:service_specification) do
     ServiceSpecification.create(
-      :name => 'weszlo.com',
+      :name => "weszlo.com #{SecureRandom.urlsafe_base64(4)}",
       :client_endpoint => '192.168.0.166'
     )
   end

### 2013-12-28 intercloud/cloud-broker/spec/lib/cloud-broker-ws/cb_ws_spec.rb

@@ -2,6 +2,7 @@ require 'spec_helper'
 require 'rack/test'
 require 'cloud-broker-ws/cb_ws'
 require 'common/config_utils'
+require 'securerandom'
 
 
 describe 'CloudBrokerWS' do
@@ -49,7 +50,7 @@ describe 'CloudBrokerWS' do
     end
     let(:service_spec) do
       JSON.generate({
-        :name => 'sample service',
+        :name => "sample service #{SecureRandom.urlsafe_base64(4)}",
         :stacks => stacks_attributes
       })
     end 

### 2013-12-28 thesis/src/chapters/chapter-5-design.tex
Thesis: stub hld overview

@@
 
-\chapterintro{ This chapter introduces the high-level design of Cloud-SAP.
-  Further elaboration on the layers of the proposed platform is presented
-  and detailed discussion about the each layer is held.
+\chapterintro{ This chapter introduces the high-level design of Cloud-SAP, highlighting its core concepts and indicating possible implementation ideas.
 }
@@
 \end{itemize}
-The last mentioned layer - inter-cloud - cooperates with different cloud providers, hence, it requires additional work that has to be done.
+The last mentioned layer - inter-cloud - cooperates with different cloud providers, hence, it requires additional work to be done.
 In fact, these extra tasks are delegated to a inter-cloud broker, responsible for a dialog in a cloud federation.
-
-Diagram \ref{design:csap-layers} depicts that multi-layered structure.
 
@@
 Each layer of the Cloud-SAP (shown in figure \ref{design:csap-layers}) must be characterised by an ability to adapt to a given application usage.
-As the previous chapter stated, this can be achieved by adding an elasticity controller to each layer.
-This observation is a foundation of proposed architecture.
-
 One of the first models that ensured system adaptivity is an autonomic component, concept based on a feed-back loop, initially proposed by IBM \cite{IBM06}.
 Figure \ref{design:autonomic-component} depicts that architecture.
+This observation is a foundation of proposed architecture - adaptivity is achieved by enriching each layer with an elasticity controller that is in fact an autonomic component.
 
@@
 
+Due to the fact that designed platform operates on multiple layers, we can extend idea of an autonomic component by using concept of a multi-hierarchical autonomic system \cite{LiWoZh05}.
+Figure \ref{design:hierarchical-autonomic-system} illustrates that hierarchy, were each level represents a different perspective on an application scaling.
+First three hierarchy levels (application tuning, container, stack) are centralised and controlled by a single elasticity controller at each level, while the last inter-cloud level is a decentralized one - each cloud instance is fully independent.
+
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics[width=0.8\textwidth]{chapter-5/hierarchical-autonomic-system}
+  \end{center}
+  \caption{Cloud-SAP as an hierarchical autonomic system}
+  \label{design:hierarchical-autonomic-system}
+\end{figure}
+
+Considering hierarchy of our architecture, each level manages an underlying autonomic component, while being managed by an upper layer at the very same time.
+
 \subsubsection{Monitoring}
-Monitoring module of each autonomic component aims to gather data from Sensors of a managed resource.
-While design of Cloud-SAP doesn't have any specific requirements regarding underlying monitoring mechanism there are a few aspects that should be taken into consideration.
+Monitoring module of each elasticity controller aims to gather data from \textit{Sensors} of a managed resource.
+While design of Cloud-SAP doesn't have any specific requirements regarding underlying monitoring mechanism, there are a few aspects that should be taken into consideration.
 Table \ref{tab:monitoring-module-issues-summary} summarises main points.
@@
 
-Cloud-SAP is entirely focused on further data processing, consequently whether to store monitoring data or not is entirely implementation dependant.
+Cloud-SAP is entirely focused on further data processing, consequently decision whether to store monitoring data or not is entirely implementation dependent.
 
 \subsubsection{Analysis}
+Analyse module applies a specified mathematical model to a gathered data and yields a conclusion on top of that.
+There is a vast number of models that can be used for that purpose, for example \cite{LiWoZh05} enumerates:
+\begin{itemize}
+ \item Queue-based Performance Models
+ \item Dynamic models
+ \item Monotonic static models
+ \item Policy based models
+\end{itemize}
+
+It is vital that Cloud-SAP implementation use predictive approach \cite{JiPeLiCh11} and track changes \cite{ZhYaWo05} in a system, giving a solid foundation for a knowledge base that can contribute to a better data analysis in future.
 
 \subsubsection{Planning}
+Planning module is responsible for scheduling actions that have to be taken to solve problems reported by an earlier analysis. This may involve:
+\begin{itemize}
+ \item reserving some resources (for example to deploy a new virtual machine)
+ \item handling situation where multiple entities compete about the resources (for example by a prioritization)
+\end{itemize}
+
+It is possible, however, that resolving issues is not possible at a current layer. It that case, the upper-layer is responsible for handling it.
 
 \subsubsection{Execution}
+It is expected that execution module manage resource by performing various operations on resource's effector.
+Each layer is characterised by different actions that can be taken, for example: application server reconfiguration, vertical or horizontal scaling.
 
 \subsubsection{Knowledge base}
-
-\subsubsection{Hierarchical autonomic-system}
-Due to the fact that designed platform operates on multiple layers, we can extend idea of an autonomic component by using concept of a multi-hierarchical autonomic system \cite{LiWoZh05}.
-Figure \ref{design:hierarchical-autonomic-system} illustrates that hierarchy, were each level represents a different perspective on an application scaling.
-Using terminology characteristic to a autonomic system, first three hierarchy levels (application tuning, container, stack) are centralised and controlled by a single elasticity controller at each level, while the last inter-cloud level is a decentralized one - each cloud instance is fully independent.
-
-\begin{figure}[!ht]
-  \begin{center}
-    \includegraphics[width=0.8\textwidth]{chapter-5/hierarchical-autonomic-system}
-  \end{center}
-  \caption{Cloud-SAP as an hierarchical autonomic system}
-  \label{design:hierarchical-autonomic-system}
-\end{figure}
-
-Due to the hierarchy of our architecture, each level manages an underlying autonomic component, while being managed by an upper layer at the very same time.
+Knowledge base is represented as a set of rules, policies shared across autonomous components.
 

### 2013-12-28 thesis/src/chapters/chapter-5-design.tex
Thesis: hld overview part 2

@@
 
+Enforcing actions on a managed component can be done using different methods, however, all of them requires an agent installed in a managed resource.
+Cloud-SAP does not enforce any specific technology, leaving low-level decision to an implementation.
+
 \subsubsection{Knowledge base}
-Knowledge base is represented as a set of rules, policies shared across autonomous components.
+Knowledge base is represented as a set of rules, policies shared across autonomic components.
+
+-- TODO przyklad z ibma
 

### 2013-12-28 thesis/src/chapters/chapter-7-evaluation.tex
Add conclusion to cost of service deployment

@@
 \subsection*{Conclusion}
+Obtained results clearly show that our proof-of-concept product met expectations of this test case as the client was offered the cheapest deployment scheme among various cloud providers for a given software stack.
+The mapping mechanism is simple yet considerably powerful -- in this simple scenario the savings were significant since they constituted nearly 20 percent of the price offered by \emph{Cloud Provider 2}.
+This shows that implementing similar solutions in the real world could be of great benefit to cloud consumers.
 

### 2013-12-29 auto-scaling/cloud-controller/lib/cloud_controller/deployers/service_deployer.rb
Change the broken condition to send USR1 signal

@@ -40,8 +40,9 @@ module AutoScaling
     # This method is used in an integration test where it
     # sends USR1 signal to the test-runner process
     def notify_observer_process(service)
-      return if (not service.deployed?) or
-                (not (config.has_key?('test-cases') and config['test-cases'].has_key?(['runner-pid'])))
+      return if (not service.deployed?) or 
+                (not config.has_key?('test-cases')) or
+                (not config['test-cases'].has_key?('runner-pid'))
 
       pid_to_notify = config['test-cases']['runner-pid']
 

### 2014-01-02 thesis/src/chapters/chapter-7-evaluation.tex
Add description of auto-scaling test case -- first version

@@
 \subsection*{Description}
+\begin{asparaenum}
+\item[\textbf{Motivation}]The aim of this test-case is to show that introducing a multi-layered auto-scaling platform is of a great benefit in terms of cost and resource usage for cloud consumers and cloud providers respectively.
+What is more, this test should prove that once there are many levels on which scaling operations can be performed, there are significant reduction in costs paid by consumers.
+We want to prove it by comparing our proof-of-concept product to \emph{Carina} \cite{Carina}.
+\emph{Carina} embraces a whole range of various scaling policies for users' environments (environment, in this context, means an application with all its software dependencies that are to be deployed on the cloud), but there is only one way in which they are executed -- by managing the number of virtual machines (i.e.
+horizontal scaling).
+Quite on the contrary, \emph{Cloud-SAP} has mechanisms that allow to scale the environment vertically in the first place and if that turns out to be not sufficient, horizontally.
+This test shows the influence of lack/presence of this feature on price and resource consumption.
+\item[\textbf{Scenario}] The test scenario involves
+  \begin{inparaenum}[a)]
+    \item deployment of a sample environment on the cloud,
+    \item substitute the real module responsible for collecting CPU usage date for a mock one,
+    \item monitoring the scaling actions performed by each solution,
+    \item evaluation of the cost the client has to pay for the service.
+  \end{inparaenum}
+  Deployment of a service is done in a product-specific manner (up to the point where the vm deployment request is passed to OpenNebula).
+Mocking the CPU usage was possible by replacing the part in \emph{InformationManager} responsible the collecting CPU data in a host for a request to a web service which generated a time-based mocked-values.
+Choosing the appropriate function is another issue and is discussed in the next paragraph.
+\end{asparaenum}
+
+\subsubsection*{CPU usage function}
+We wanted to ensure that both solutions, \emph{Carina} and \emph{Cloud-SAP}, collected the same data regarding the CPU usage for the given time.
+What is more, the curve should resemble CPU usage in the real business scenarios as much as possible.
+Thus, we took into account the following factors:
+\begin{itemize}
+  \item every peak in CPU usage should be followed by a gradual descent as each solution performs scaling actions,
+  \item (boundary conditions) CPU usage should be between 0 and 100 and, additionally, should exceed the previously set scaling threshold value of each solutions so that it would trigger the scaling policies evaluation
+  \item once the scaling operations have completed, CPU usage should be constant at a rate which would not introduce any changes in the environment settings (such as the number of virtual machines and parameters of any virtual machine)
+\end{itemize}
+This resulted in the function whose formula is in \eqref{eq:cpu-mock-usage} and figure in \ref{eval:auto-scaling-1cp-cpu-usage-function}.
+\begin{eqnarray}
+  %f(x) = (1751 *x)/132+(3 *x**2)/8-(31* x**3)/264
+  %x <= 11.583 ? f(x) : ( x <= 21.441 ?  f(x - 10) : 25 )
+  f(x) &=& -\frac{31}{264}x^3+\frac{3}{8}x^2+\frac{1751}{132}x \nonumber \\
+  CPU\_usage(t) &=& \left\{
+  \begin{array}{l l l}
+    f(t) & \quad 0 & \le t < 11.583 \\
+    f(t-10) & \quad 11.583 & \le t < 21.441 \\
+    25 & \quad 21.441 & \le  t \\
+\end{array} \right.
+\label{eq:cpu-mock-usage}
+\end{eqnarray}
+
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics{chapter-7/cpu-usage-mock}
+  \end{center}
+  \caption{Auto-scaling - single-provider based: CPU usage function}
+  \label{eval:auto-scaling-1cp-cpu-usage-function}
+\end{figure}
+
+Because \emph{Carina} is capable of only horizontal scaling and \emph{Cloud-SAP} of both horizontal and vertical, one can question whether the descent in CPU usage can be actually the same in both cases.
+To answer this question we want to look into the description of the virtual machines comprising the deployed environment.
+It states that each VM can use up to 30\% power of the CPU of a host.
+Thus, we can be fairly sure that adding another virtual machine that uses 0.3 CPU capacity of its host is equivalent with changing the CPU usage settings of already deployed VM from 30 to 60\%.
 \subsection*{Preconditions}
+\subsubsection*{Auto-scaling policy specifications}
+\subsubsection*{Hardware and VM configuration}
+All virtual machines were deployed onto \emph{Node1}, which uses \emph{OpenVZ} as a virtualization technology and whose hardware configuration can be found in table \ref{tbl:test-deployment-time-common-hardware-configuration}.
+\emph{OpenNebula} was installed on a \emph{Frontend1} virtual machine and its configuration is in the same table.
+\emph{Carina} required the usage of 2 virtual machines which were deployed on \emph{Desktop} with KVM as a hypervisor and whose configuration is in table \ref{tbl:test-auto-scaling-1cp-hardware-configuration}.
+Diagram \ref{eval:auto-scaling-1cp-physical-nodes} illustrates the physical setup of the test-case.
+
+\begin{table}
+  \centering
+  \begin{tabular}{ | l | l | l | l | }
+    \hline                        
+    Name & CPU & RAM & HD \\
+    \hline
+    (virtual machine) carina-frontend  & 1 CPU & 1GB & 6GB \\
+    (virtual machine) carina-management-vm  & 1 CPU & 1GB & 9GB \\
+    \hline  
+  \end{tabular}
+  \caption{Configuration of hardware/virtual machines used during tests}
+  \label{tbl:test-auto-scaling-1cp-hardware-configuration}
+\end{table}
+
+\subsubsection*{Hardware and VM configuration}
 \begin{figure}[!ht]
@@
   \caption{Auto-scaling - multiple-provider based: environment configuration}
-  \label{eval:auto-scaling-1cp-physical-nodes}
+  \label{eval:auto-scaling-2cp-physical-nodes}
 \end{figure}

### 2014-01-02 thesis/src/chapters/chapter-7-evaluation.tex
Add deployment diagrams for auto-scaling test description

@@
 \subsection*{Preconditions}
-\subsubsection*{Auto-scaling policy specifications}
-\subsubsection*{Hardware and VM configuration}
+\subsubsection*{Environment specification}
+Deployment diagrams show the placement of each component in both installations.
+They are shown in figures \ref{eval:auto-scaling-1cp-carina-deployment-diagram} and \ref{eval:auto-scaling-1cp-cloud-sap-deployment-diagram}.
+
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics{chapter-7/cpu-usage-carina-deployment-diagram}
+  \end{center}
+  \caption{Auto-scaling - single-provider based: deployment diagram of Carina}
+  \label{eval:auto-scaling-1cp-carina-deployment-diagram}
+\end{figure}
+
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics{chapter-7/cpu-usage-cloud-sap-deployment-diagram}
+  \end{center}
+  \caption{Auto-scaling - single-provider based: deployment diagram of Cloud-SAP}
+  \label{eval:auto-scaling-1cp-cloud-sap-deployment-diagram}
+\end{figure}
+
+Auto-scaling policy specifications
+
+\subsubsection*{Hardware VM configuration}
 All virtual machines were deployed onto \emph{Node1}, which uses \emph{OpenVZ} as a virtualization technology and whose hardware configuration can be found in table \ref{tbl:test-deployment-time-common-hardware-configuration}.
@@
 
-\subsubsection*{Hardware and VM configuration}
 \begin{figure}[!ht]

### 2014-01-02 thesis/src/chapters/chapter-7-evaluation.tex
Add auto-scaling service specification excerpt for auto-scaling test with one cloud provider -- cloud sap

@@
 \end{itemize}
-This resulted in the function whose formula is in \eqref{eq:cpu-mock-usage} and figure in \ref{eval:auto-scaling-1cp-cpu-usage-function}.
+This resulted in the function whose formula is in \eqref{eq:cpu-mock-usage} and plot in figure \ref{eval:auto-scaling-1cp-cpu-usage-function}.
 \begin{eqnarray}
@@
 \subsubsection*{Environment specification}
-Deployment diagrams show the placement of each component in both installations.
+\begin{asparaenum}
+  \item[\textbf{Deployment diagrams}] Deployment diagrams show the placement of each component in both installations.
 They are shown in figures \ref{eval:auto-scaling-1cp-carina-deployment-diagram} and \ref{eval:auto-scaling-1cp-cloud-sap-deployment-diagram}.
@@
   \end{figure}
-
-Auto-scaling policy specifications
+\item[\textbf{Auto-scaling policy specifications}] As it is shown in figure \ref{eval:auto-scaling-1cp-cpu-usage-function}, we set in both products the threshold values of CPU usage that trigger performing auto-scaling actions to 20 and 60 percent.
+To apply this setting, the auto-scaling part of service specification looks as follows: in Carina \ldots \ldots ; in Cloud-SAP it is a matter of setting appropriate arguments to a specific policy.
+In this case the policy is \emph{threshold\_model} and values are 20 and 60 for lower and upper bound respectively.
+The auto-scaling part from service specification is shown in listing \ref{lst:auto-scaling-1cp-cloud-sap-service-spec}.
+  \lstinputlisting[caption=Cloud-SAP service specification used for testing auto-scaling with 1 cloud provider, label=lst:auto-scaling-1cp-cloud-sap-service-spec]{auto-scaling-1cp-cloud-sap-service-spec}
+\end{asparaenum}
 

### 2014-01-02 thesis/src/chapters/chapter-7-evaluation.tex
Add configuration of csap/carina/one in terms of information manager and policy evaluation interval

@@
 \begin{asparaenum}
+  \item[\textbf{Auto-scaling policy specifications}] As it is shown in figure \ref{eval:auto-scaling-1cp-cpu-usage-function}, we set in both products the threshold values of CPU usage that trigger performing auto-scaling actions to 20 and 60 percent.
+To apply this setting, the auto-scaling part of service specification looks as follows: in Carina the user has to add the parameters of an auto-scaling policy in a hash that describes the environment under key \emph{:elasticy\_policy}.
+It is possible to specify the minimal and the maximal number of virtual machines that forms the environment and, what is most important, expressions which evaluation results in scaling the application.
+The part responsible for this is shown in listing \ref{lst:auto-scaling-1cp-carina-service-spec}.\lstinputlisting[caption=Carina service specification used for testing auto-scaling with 1 cloud provider, label=lst:auto-scaling-1cp-carina-service-spec]{auto-scaling-1cp-carina-service-spec} In Cloud-SAP it is a matter of setting appropriate arguments to a specific policy.
+In this case the policy is \emph{threshold\_model} and values are 20 and 60 for lower and upper bound respectively.
+The auto-scaling part from service specification is shown in listing \ref{lst:auto-scaling-1cp-cloud-sap-service-spec}.
+\lstinputlisting[caption=Cloud-SAP service specification used for testing auto-scaling with 1 cloud provider, label=lst:auto-scaling-1cp-cloud-sap-service-spec]{auto-scaling-1cp-cloud-sap-service-spec}
+  \item[\textbf{OpenNebulaa/Carina/Cloud-SAP configuration}] Since Carina and Cloud-SAP used two different instances of OpenNebula installed on separate virtual machines, it is essential the configuration be the same on each of them.
+Listing \ref{lst:auto-scaling-1cp-one-config} shows the most important OpenNebula excerpt from settings file used in this test case -- polling interval specification, which was set to 30 seconds.
+To ensure that both products can actually use up-to-date data, they should evaluate their policies every 30 seconds or longer.
+In Carina the user cannot specify this value, because it is hard-coded in source code to 60 seconds.
+In Cloud-SAP, however, this value was set in a configuration file and was equal to 60 seconds as well as in Carina.
+\lstinputlisting[caption=OpenNebula configuration excerpt -- information manager, label=lst:auto-scaling-1cp-one-config]{auto-scaling-1cp-one-config}
   \item[\textbf{Deployment diagrams}] Deployment diagrams show the placement of each component in both installations.
@@
   \end{figure}
-\item[\textbf{Auto-scaling policy specifications}] As it is shown in figure \ref{eval:auto-scaling-1cp-cpu-usage-function}, we set in both products the threshold values of CPU usage that trigger performing auto-scaling actions to 20 and 60 percent.
-To apply this setting, the auto-scaling part of service specification looks as follows: in Carina \ldots \ldots ; in Cloud-SAP it is a matter of setting appropriate arguments to a specific policy.
-In this case the policy is \emph{threshold\_model} and values are 20 and 60 for lower and upper bound respectively.
-The auto-scaling part from service specification is shown in listing \ref{lst:auto-scaling-1cp-cloud-sap-service-spec}.
-  \lstinputlisting[caption=Cloud-SAP service specification used for testing auto-scaling with 1 cloud provider, label=lst:auto-scaling-1cp-cloud-sap-service-spec]{auto-scaling-1cp-cloud-sap-service-spec}
+
   \end{asparaenum}
 
-\subsubsection*{Hardware VM configuration}
+\subsubsection*{Hardware/VM configuration}
 All virtual machines were deployed onto \emph{Node1}, which uses \emph{OpenVZ} as a virtualization technology and whose hardware configuration can be found in table \ref{tbl:test-deployment-time-common-hardware-configuration}.

### 2014-01-03 thesis/src/chapters/chapter-7-evaluation.tex
Add expectations section to the description of an auto-scaling test case with one cloud provider

@@
 \begin{itemize}
-  \item every peak in CPU usage should be followed by a gradual descent as each solution performs scaling actions,
+  \item every peak in CPU usage should be followed by a gradual descent which would mimic the real auto-scaling actions executed by each solution,
   \item (boundary conditions) CPU usage should be between 0 and 100 and, additionally, should exceed the previously set scaling threshold value of each solutions so that it would trigger the scaling policies evaluation
@@
 
+\subsection*{Expectations}
+As auto-scaling behaviour of both products is determined by CPU usage of the deployed service, we can predict the outcome of the test by thorough analysis of the plot in figure \ref{eval:auto-scaling-1cp-cpu-usage-function}.
+
+The first conclusion is that all auto-scaling actions are performed after exceeding the upper threshold value of CPU usage, this is 60 percent in our case what happens about 5 minutes after the deployment of an environment is completed.
+CPU consumption remains high for about 2 minutes what enables both products to take appropriate, auto-scaling steps.
+We expect \emph{Carina} to deploy another virtual machine and \emph{Cloud-SAP} to scale vertically.
+Then it gradually goes down, which simulates that all taken actions have successfully completed.
+Then, for another several minutes, the cycle recurs, but this time we expect the proposed solution to scale horizontally.
+\emph{Carina} is expected to behave as in the previous cycle.
+Once the cycle has completed, after roughly 21 minutes, the value of CPU consumption remains constant at 25 percent and the test is completed.
   \subsection*{Results}

### 2014-01-03 thesis/src/chapters/chapter-7-evaluation.tex
Finish work on autoscaling with one cloud provider test case

@@
 In Cloud-SAP, however, this value was set in a configuration file and was equal to 60 seconds as well as in Carina.
-\lstinputlisting[caption=OpenNebula configuration excerpt -- information manager, label=lst:auto-scaling-1cp-one-config]{auto-scaling-1cp-one-config}
+\lstinputlisting[caption=OpenNebula configuration excerpt -- information manager, label=lst:auto-scaling-1cp-one-config]{auto-scaling-1cp-one-config}.
+Vertical scaling mechanism in \emph{Cloud-SAP} is implemented in a way that causes exponential growth of CPU resources consumption.
+When there is a need to perform a scaling action on CPU usage, current value is taken and increased by 30\%.
   \item[\textbf{Deployment diagrams}] Deployment diagrams show the placement of each component in both installations.
@@
 \subsection*{Results}
+  In the first place it is worth discussing the results of \emph{Carina} and \emph{Cloud-SAP} separately and then compare resource consumption of each solution and its influence on cost paid by the end-user.
+
+\begin{asparaenum}
+  \item[\textbf{Carina}] As \emph{Carina} is capable only of horizontal scaling, in discussion of its resource usage we can confine ourselves to counting the number of deployed virtual machines in the given time intervals.
+In the log files we can trace all actions that were triggered during the test case.
+As listing \ref{lst:auto-scaling-1cp-results-carina-jobs} shows, they completely met the expectations expressed in the previous section -- there were two ''SCALEUP'' jobs which resulted in increase of the total number of virtual machines comprising the environment.
+\lstinputlisting[caption=Carina environment manager logs with taken actions (\emph{jobs}), label=lst:auto-scaling-1cp-results-carina-jobs]{auto-scaling-1cp-results-carina-jobs}.
+Since the base configuration assumed that the environment consists of 2 virtual machines, we can say that about 10 minutes after deployment of an environment, the number of VMs increased to 3, and after another 10 minutes, to 4.
+  \item[\textbf{Cloud-SAP}] Our proof-of-concept solution behaved exactly as expected.
+To make the discussion more clear, there is an excerpt from a log file of this test run in listing \ref{lst:auto-scaling-1cp-results-csap-logs}.
+\lstinputlisting[caption=Cloud-SAP logs excerpt regarding auto-scaling actions, label=lst:auto-scaling-1cp-results-csap-logs]{auto-scaling-1cp-results-csap-logs}
+
+Deployment finished at 15:25:45 and at this time we started the mock-cpu-usage web service.
+When the platform concluded that there are insufficient resources, it tried for three times changing parameters (virtual CPU used by the vm) of the virtual machines forming the service.
+Once scaling vertically was not possible such an information was passed to cloud controller which performed scaling across different cloud providers.
+\end{asparaenum}
+
+\subsubsection*{Cost}
+In our scenario cost is roughly equivalent with the CPU usage, so the plot of a cost virtually presents the cpu usage.
+Figure \ref{eval:auto-scaling-1cp-cost-comparison} shows cpu usage by environment when deployed and configured by two competing products.
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics{chapter-7/auto-scaling-1cp-cost-comparison}
+  \end{center}
+  \caption{Auto-scaling - single-provider based: comparison of cost/CPU usage}
+  \label{eval:auto-scaling-1cp-cost-comparison}
+\end{figure}
+If we were to roughly estimate the cost in both solutions it would be equal in Carina:
+\begin{equation}
+  7\cdot 0.3 + 10\cdot 0.6 + 8\cdot 0.9 = 15.3 \quad \text{[currency unit]}
+\end{equation}
+and in Cloud-SAP:
+\begin{equation}
+  7.0166\cdot 0.3 + 2\cdot 0.39 + 2.01667\cdot0.507 + 7.05\cdot 0.659 + 6.91667\cdot 0.959 \approx 15.186483
+\end{equation}
+
 \subsection*{Conclusion}
+The obtained results shows that the proposed solution has enormous potential in terms of better utilization of available resources of the cloud and reducing cost paid by cloud consumers.
+
+Judging by the plot of cost (figure \ref{eval:auto-scaling-1cp-cost-comparison}), we can be fairly sure that it is possible to obtain better results by proper tuning of the scaling vertical mechanism.
+In particular, one can consider changing the default CPU growth from 1.3 to other or apply a different strategy, for example a growth by a constant rate.
+What is more, in order to get the best results, the policy scaling interval must be chosen with a great care with the consideration of specific parameters of a given environment.
 

### 2014-01-05 thesis/src/chapters/chapter-5-design.tex
Thesis: chapter-5
Thesis: multi-provider based autoscaling

@@
 
-Next section states the general overview of the proposed solution, while the consecutive sections presents more detailed discussion of its elements and finally the last section summarises the design choices in a context of system requirements.
+Next section states the general overview of the proposed solution, while the consecutive sections details its elements and finally the last section summarises the design choices in a context of system requirements.
 
@@
  \item reserving some resources (for example to deploy a new virtual machine)
- \item handling situation where multiple entities compete about the resources (for example by a prioritization)
+ \item handling situation where multiple entities compete about the resources (for example taking service's priority into account)
 \end{itemize}

### 2014-01-05 thesis/src/chapters/chapter-7-evaluation.tex
Thesis: multi-provider based autoscaling

@@
 \lstinputlisting[caption=Cloud-SAP service specification used for testing auto-scaling with 1 cloud provider, label=lst:auto-scaling-1cp-cloud-sap-service-spec]{auto-scaling-1cp-cloud-sap-service-spec}
-  \item[\textbf{OpenNebulaa/Carina/Cloud-SAP configuration}] Since Carina and Cloud-SAP used two different instances of OpenNebula installed on separate virtual machines, it is essential the configuration be the same on each of them.
+  \item[\textbf{OpenNebula/Carina/Cloud-SAP configuration}] Since Carina and Cloud-SAP used two different instances of OpenNebula installed on separate virtual machines, it is essential the configuration be the same on each of them.
 Listing \ref{lst:auto-scaling-1cp-one-config} shows the most important OpenNebula excerpt from settings file used in this test case -- polling interval specification, which was set to 30 seconds.
@@
 \subsection*{Description}
+
+\subsubsection{Motivation} This test case aims to prove that scaling across multiple cloud providers is vitally important while ensuring appropriate Quality-of-Service, especially in cases of increased number of resource requests.
+Such scaling scenario, known also as cloud-bursting, can be also perceived as proof of benefits arising from offloading application load to an external provider.
+In order to verify our concept we compare number of transactions per second handled by Cloud-SAP and Carina.
+While our proof of concept solution features multi-layer scaling and is cloud federation aware, Carina adopts only horizontal scaling.
+It is expected that test case prove benefits emerging from enriching application platform provider with an cloud federation awareness.
+
+\subsubsection{Scenario}
+Testing requires following steps to be done:
+\begin{enumerate}
+\item deploy exemplary service
+\item observe system behaviour under load, simulated by a resource usage mock. It is vital that resource usage exceed single provider capabilities
+\item assess system characteristics, including number of handled transactions per second and cost of application deployment
+\end{enumerate}
+
+\subsubsection*{Load simulation}
+Similarly to a previous test case, we leveraged a resource mock that allows us to precisely control monitoring information returned to an system.
+Besides, again, in order to simplify test case, we were solely focused on a CPU usage.
+The only requirement that CPU usage have to be met was that function should at some point exceed upper threshold limit and stay at that level, triggering successive auto-scaling events that at some point surpassing single cloud provider resources.
+Equation \eqref{eq:cpu-mock-usage-2cp} results in such scenario and is illustrated in figure \ref{eval:auto-scaling-2cp-cpu-usage-function}.
+\begin{equation}
+ f(x) = \frac{29}{1500}x^3-\frac{21}{20}x^2+\frac{533}{30}x
+ \label{eq:cpu-mock-usage-2cp}
+\end{equation}
+
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics{chapter-7/cpu-usage-mock-multiple}
+  \end{center}
+  \caption{Auto-scaling - multiple-provider based: CPU usage function}
+  \label{eval:auto-scaling-2cp-cpu-usage-function}
+\end{figure}
+
 \subsection*{Preconditions}
+
+\begin{asparaenum}
+  \item[\textbf{Auto-scaling policy specifications}] Policies used in this test are based on a threshold model with the following properties:
+  \begin{itemize}
+   \item value below 20 should trigger scaling down event
+   \item value grater than 60 should trigger scaling up event
+  \end{itemize}
+  However, considering the fact that we are entirely focused on scaling up, only the upper limit is relevant in our case.
+Listings \ref{lst:auto-scaling-2cp-carina-service-spec} and \ref{lst:auto-scaling-2cp-cloud-sap-service-spec} presents scaling policies for Carina and Cloud-SAP respectively.
+
+  \lstinputlisting[caption=Carina service specification used for testing auto-scaling with 2 cloud providers, label=lst:auto-scaling-2cp-carina-service-spec]{auto-scaling-2cp-carina-service-spec}
+  \lstinputlisting[caption=Cloud-SAP service specification used for testing auto-scaling with 2 cloud providers, label=lst:auto-scaling-2cp-cloud-sap-service-spec]{auto-scaling-2cp-cloud-sap-service-spec}
+
+  \item[\textbf{OpenNebula/Carina/Cloud-SAP configuration}] 
+  Common components to Carina and Cloud-SAP, namely OpenNebula instances and computing nodes, were configured in the same fashion.
+Listings \ref{lst:auto-scaling-2cp-one-config} and \ref{lst:auto-scaling-2cp-sched-config} depicts key configuration elements of OpenNebula monitoring mechanism and scheduler.
+
+  \lstinputlisting[caption=OpenNebula configuration excerpt -- virtual machine and information manager, label=lst:auto-scaling-2cp-one-config]{auto-scaling-2cp-one-config}
+  \lstinputlisting[caption=OpenNebula configuration excerpt -- virtual machine and information manager, label=lst:auto-scaling-2cp-sched-config]{auto-scaling-2cp-sched-config}
+
+  \item[\textbf{Deployment diagrams}]
+
+  \end{asparaenum}
+
+\subsubsection*{Environment specification}
+
+\subsubsection*{Hardware/VM configuration}
+
 \begin{figure}[!ht]
@@
 \end{figure}
+
+\subsection*{Expectations}
 
@@
 \subsection*{Conclusion}
-
-\section{Deployment time -- hypervisor comparison}
-\subsection*{Description}
-\subsection*{Preconditions}
-\subsection*{Results}
-\subsection*{Conclusion}

### 2014-01-06 thesis/src/chapters/chapter-7-evaluation.tex
Thesis: diagrams for a multprovider testcase

@@
 \subsection*{Preconditions}
-
+\subsubsection*{Environment specification}
 \begin{asparaenum}
@@
   \lstinputlisting[caption=OpenNebula configuration excerpt -- virtual machine and information manager, label=lst:auto-scaling-2cp-one-config]{auto-scaling-2cp-one-config}
-  \lstinputlisting[caption=OpenNebula configuration excerpt -- virtual machine and information manager, label=lst:auto-scaling-2cp-sched-config]{auto-scaling-2cp-sched-config}
+  \lstinputlisting[caption=OpenNebula configuration excerpt -- scheduler, label=lst:auto-scaling-2cp-sched-config]{auto-scaling-2cp-sched-config}
 
-  \item[\textbf{Deployment diagrams}]
+  \item[\textbf{Deployment diagrams}] Components that took part during testing are show in figures \ref{eval:auto-scaling-2cp-carina-deployment-diagram} and \ref{eval:autoscale-2cp-cloud-sap-deployment-diagram}.
 
+  \begin{figure}[!ht]
+    \begin{center}
+      \includegraphics{chapter-7/autoscale-2cp-cloud-sap-deployment-diagram}
+    \end{center}
+    \caption{Auto-scaling - multiple-provider based: deployment diagram of Carina}
+    \label{eval:autoscale-2cp-cloud-sap-deployment-diagram}
+  \end{figure}
+
+  \begin{figure}[!ht]
+    \begin{center}
+      \includegraphics{chapter-7/cpu-usage-carina-deployment-diagram}
+    \end{center}
+    \caption{Auto-scaling - multiple-provider based: deployment diagram of Carina}
+    \label{eval:auto-scaling-2cp-carina-deployment-diagram}
+  \end{figure}
 \end{asparaenum}
-
-\subsubsection*{Environment specification}
 

### 2014-01-13 thesis/src/chapters/chapter-7-evaluation.tex
Thesis: refactor multibase

--- 2014-01-03 thesis/src/chapters/chapter-7-evaluation.tex
@@
 
+\newpage
 \section{Auto-scaling -- single-provider based}
@@
 \begin{asparaenum}
+
+  \item[\textbf{Cloud-SAP}] Our proof-of-concept solution behaved exactly as expected.
+To make the discussion more clear, there is an excerpt from a log file of this test run in listing \ref{lst:auto-scaling-1cp-results-csap-logs}.
+\lstinputlisting[caption=Cloud-SAP logs excerpt regarding auto-scaling actions, label=lst:auto-scaling-1cp-results-csap-logs]{auto-scaling-1cp-results-csap-logs}
+
 \item[\textbf{Carina}] As \emph{Carina} is capable only of horizontal scaling, in discussion of its resource usage we can confine ourselves to counting the number of deployed virtual machines in the given time intervals.
@@
 Since the base configuration assumed that the environment consists of 2 virtual machines, we can say that about 10 minutes after deployment of an environment, the number of VMs increased to 3, and after another 10 minutes, to 4.
-  \item[\textbf{Cloud-SAP}] Our proof-of-concept solution behaved exactly as expected.
-To make the discussion more clear, there is an excerpt from a log file of this test run in listing \ref{lst:auto-scaling-1cp-results-csap-logs}.
-\lstinputlisting[caption=Cloud-SAP logs excerpt regarding auto-scaling actions, label=lst:auto-scaling-1cp-results-csap-logs]{auto-scaling-1cp-results-csap-logs}
 
@@
 
+\newpage
 \section{Auto-scaling -- multiple-provider based}
 \subsection*{Description}
-\subsection*{Preconditions}
+
+\subsubsection{Motivation} This test case aims to prove that scaling across multiple cloud providers is vitally important while ensuring appropriate Quality-of-Service, especially in cases of increased number of service requests.
+Such scaling scenario, known also as cloud-bursting, leverage benefits arising from offloading application load to an external provider.
+In order to verify our concept we compare number of transactions per second guaranteed by \emph{Cloud-SAP} and \emph{Carina}.
+While our proof of concept solution features multi-layer scaling and is cloud federation aware, \emph{Carina} adopts only horizontal scaling.
+It is expected that test case prove benefits emerging from enriching application platform provider with an cloud federation awareness.
+
+\subsubsection{Scenario}
+Testing requires following steps to be done:
+\begin{enumerate}
+\item deploy exemplary service
+\item observe system behaviour under load, simulated by a resource usage mock. It is vital for a resource usage to exceed a single provider capabilities
+\item assess system characteristics, including number of handled transactions per second
+\end{enumerate}
+
+\subsubsection*{Load simulation}
+Similarly to a previous test case, we leveraged a resource mock that allows us to precisely control monitoring information returned to an system.
+Besides, in order to simplify test case, we were solely focused on a CPU usage.
+CPU usage function has to exceed upper threshold limit at some point and stay at that level, triggering successive auto-scaling events.
+However, at some point single cloud provider resources will be surpassed.
+Equation \eqref{eq:cpu-mock-usage-2cp} denotes a resource usage function that is expected to fulfil above-mentioned scenario and is illustrated in figure \ref{eval:auto-scaling-2cp-cpu-usage-function}.
+\begin{equation}
+ f(x) = \frac{29}{1500}x^3-\frac{21}{20}x^2+\frac{533}{30}x
+ \label{eq:cpu-mock-usage-2cp}
+\end{equation}
+
 \begin{figure}[!ht]
   \begin{center}
-    \includegraphics[width=0.8\textwidth]{chapter-7/auto-scaling-2cp-physical-nodes}
+    \includegraphics{chapter-7/cpu-usage-mock-multiple}
+  \end{center}
+  \caption{Auto-scaling - multiple-provider based: CPU usage function}
+  \label{eval:auto-scaling-2cp-cpu-usage-function}
+\end{figure}
+
+\subsection*{Preconditions}
+\subsubsection*{Environment specification}
+\begin{asparaenum}
+  \item[\textbf{Auto-scaling policy specifications}] Policies used in this test are based on a threshold model with the following properties:
+  \begin{itemize}
+   \item value below 20 triggers scaling down event
+   \item value grater than 60 triggers scaling up event
+  \end{itemize}
+  However, considering the fact that we are entirely focused on scaling up, only the upper limit is relevant in our case.
+Listings \ref{lst:auto-scaling-2cp-carina-service-spec} and \ref{lst:auto-scaling-2cp-cloud-sap-service-spec} presents scaling policies for \emph{Carina} and \emph{Cloud-SAP} respectively.
+
+  \lstinputlisting[caption=\emph{Carina} service specification used for testing auto-scaling with 2 cloud providers, label=lst:auto-scaling-2cp-carina-service-spec]{auto-scaling-2cp-carina-service-spec}
+  \lstinputlisting[caption=\emph{Cloud-SAP} service specification used for testing auto-scaling with 2 cloud providers, label=lst:auto-scaling-2cp-cloud-sap-service-spec]{auto-scaling-2cp-cloud-sap-service-spec}
+
+  \item[\textbf{OpenNebula/Carina/Cloud-SAP configuration}] 
+  Components common to \emph{Carina} and \emph{Cloud-SAP}, namely \emph{OpenNebula} instances and computing nodes, were configured in the same fashion.
+Listing \ref{lst:auto-scaling-2cp-one-config} depicts key configuration elements of \emph{OpenNebula} monitoring mechanism.
+
+  \lstinputlisting[caption=OpenNebula configuration excerpt -- virtual machine and information manager, label=lst:auto-scaling-2cp-one-config]{auto-scaling-2cp-one-config}
+
+  \item[\textbf{Deployment diagrams}] Components that took part during testing are show in figures \ref{eval:auto-scaling-2cp-carina-deployment-diagram} and \ref{eval:autoscale-2cp-cloud-sap-deployment-diagram}.
+
+  \begin{figure}[!ht]
+    \begin{center}
+      \includegraphics{chapter-7/autoscale-2cp-cloud-sap-deployment-diagram}
+    \end{center}
+    \caption{Auto-scaling - multiple-provider based: deployment diagram of Carina}
+    \label{eval:autoscale-2cp-cloud-sap-deployment-diagram}
+  \end{figure}
+
+  \begin{figure}[!ht]
+    \begin{center}
+      \includegraphics{chapter-7/cpu-usage-carina-deployment-diagram}
+    \end{center}
+    \caption{Auto-scaling - multiple-provider based: deployment diagram of Carina}
+    \label{eval:auto-scaling-2cp-carina-deployment-diagram}
+  \end{figure}
+\end{asparaenum}
+
+\subsubsection*{Hardware/VM configuration}
+
+Figure \ref{eval:auto-scaling-2cp-physical-nodes} depicts physical configuration of the environment.
+In short, setup was as follows: all \emph{Cloud-SAP} components, apart from cloud client deployed on \emph{Laptop}, were provisioned on \emph{Desktop}.
+\emph{Cloud Proivder 1 (CP-1)} is \emph{OpenNebula} instance known as \emph{Frontend1} which uses \emph{Node1} as a computing node and is deployed on \emph{Desktop}, while \emph{CP\-2} uses \emph{Frontend2} and \emph{Node2}.
+Specification of nodes is listed in table: \ref{tbl:test-deployment-time-common-hardware-configuration}.
+Deployment cost of a java stack was 50 and 60 using \emph{CP-1} and \emph{CP-2}, respectively.
+
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics[width=0.7\textwidth]{chapter-7/auto-scaling-2cp-physical-nodes}
   \end{center}
@@
 
+\subsection*{Expectations}
+Taking CPU usage function into account, we can predict outcome of the test.
+Scaling policy is violated in 5th minute, hence, it is expected that first scaling event occur after that point.
+Moreover, since that moment, CPU usage remains beyond threshold implying successive scaling actions.
+
+Due to the fact that \emph{Cloud-SAP} favours fine-grained scaling actions such as vertical scaling, it is expected that these actions occur sooner than horizontal scaling.
+However, at some point \emph{CP-1} capacity won't be sufficient, hence, another stack instance (one master instance, two slaves) will be deployed on \emph{CP-2}.
+Subsequent scaling actions will take place solely on a \emph{CP-2} 2 and will involve further vertical scaling.
+
+Carina, on the other hand, supports solely horizontal scaling using single cloud provider.
+On top of that, it is assumed that succeeding slaves instances will be added to a service up to the point where Cloud Provider won't have enough resources to proceed with further scaling requests.
+
 \subsection*{Results}
+
+\begin{asparaenum}
+
+  \item[\textbf{Cloud-SAP}] 
+   In total, there were 13 vertical (increasing CPU limit) 2 horizontal (adding a container) scaling events.
+First action took place minute after first violation of scaling policy rule.
+Since then, slaves' CPU were successively increasing to the point were further scaling wasn't possible - 13th minute of the test.
+Therefore, \emph{Cloud-SAP} deployed two new slaves on \emph{CP-2}.
+Listing \ref{lst:auto-scaling-2cp-results-csap-logs} presents logs covering service lifecycle.
+\lstinputlisting[caption=Cloud-SAP logs excerpt regarding auto-scaling actions on multiple providers, label=lst:auto-scaling-2cp-results-csap-logs]{auto-scaling-2cp-results-csap-logs}
+
+  \item[\textbf{Carina}] 
+  Similarly to a \emph{Cloud-SAP}, first scaling event (slave addition) was noticed in 6th minute of test.
+Scaling jobs were continuously invoked every 2 minutes until capacity of a \emph{CP\-1} hasn been fully exploited.
+Listing \ref{lst:auto-scaling-2cp-results-carina-jobs} summarises jobs performed by \emph{Carina}.
+\lstinputlisting[caption=Carina environment manager logs with taken actions (\emph{jobs}), label=lst:auto-scaling-2cp-results-carina-jobs]{auto-scaling-2cp-results-carina-jobs}
+\end{asparaenum}
+
+\subsubsection*{Transactions per second}
+Chart \ref{eval:auto-scaling-2cp-tps-comparison} illustrates how service capabilities changed over the time.
+Service capability is expressed as a number of transactions per second that can be handled by a service.
+Please note, that this measure, while giving good insight into application potential, is hard to simulate, hence, we assumed that 1 VCPU provides enough resources to successfully serve 100 TPS.
+Knowing that, we roughly estimate total service capacity, which at the end of the test amounts to:
+\begin{itemize}
+ \item Cloud-SAP: 283 transactions per second
+ \item Carina: 180 transactions per second
+\end{itemize}
+
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics{chapter-7/auto-scaling-2cp-tps-comparison}
+  \end{center}
+  \caption{Auto-scaling - multiple-provider based: comparison of processed transaction per second}
+  \label{eval:auto-scaling-2cp-tps-comparison}
+\end{figure}
+
 \subsection*{Conclusion}
+Comparing expectations with test output, one can see that expectations has been fully met.
+While \emph{Carina} was solely focused on horizontal scaling and bound to a single cloud provider, \emph{Cloud-SAP} first took fine-grained actions to later leverage resources of a second cloud provider.
 
+\emph{Cloud-SAP} has been proven to be a better solution, offering greater capabilities to a service.
+This was possible by exploiting resources of two federated cloud providers, crucial concept behind \emph{Cloud-SAP} design.
+
+\newpage
 \section{Deployment time -- solution comparison}
@@
 The description of a service expressed in Carina format can be found in listing \ref{lst:carina-config}.
-
-- TODO dodać konfigurację każdej z maszyn wirtualnych
 

### 2014-01-13 auto-scaling/common/lib/planner/reservation_manager.rb

@@ -51,6 +51,8 @@ module AutoScaling
     #   :memory => 3
     # }
     def reserve?(stack_data)
+      return true
+
       needs = resources(stack_data['type'])
       needs.each {|k, v| needs[k] = v * stack_data['instances'].to_i}
 
@@ -62,6 +64,8 @@ module AutoScaling
     end
 
     def scale_up(container, parameter, requested)
+      raise InsufficientResources.new("CONTAINER Cannot reserve")
+
       @reservation.synchronize do
         host = @cloud_provider.host_by_container(container.correlation_id)
         current_state = @cloud_provider.monitor_host(host)
@@ -87,6 +91,8 @@ module AutoScaling
     #   :memory => 3
     # }
     def reserve(units)
+      raise InsufficientResources.new("STACK Cannot")
+
       @reservation.synchronize do
         backup = @capacity.clone
 

### 2014-01-13 auto-scaling/domain/service.rb

@@ -7,7 +7,8 @@ module AutoScaling
   class Service
     include DataMapper::Resource
 
-    property :name, String, :key => true, :unique => true
+    property :id, Serial, :key => true
+    property :name, String
     property :autoscaling_queue_name, String, :required => true
 
     has n, :stacks

### 2014-01-13 thesis/src/chapters/chapter-7-evaluation.tex
Thesis: deployment time conclustion

@@
 \subsection*{Conclusion}
+As one can can notice, deployment of a java stack takes using \emph{Cloud-SAP} lasts longer than when using \emph{Carina} for the same purpose.
+Moreover, the more instances are in such stack, the greater difference is.
+This difference in provisioning time is mainly driven by the fact that \emph{Cloud-SAP} uses a great deal of components in comparison to \emph{Carina}, what is a direct consequence of chosen architecture.
+Specifically, the request from a client is passed to a broker, then the best provider is chosen and finally the deployment using the select provider and its \emph{Applfow} server takes place.
+Contrary, \emph{Carina} directly operates on \emph{OpenNebula} giving much simpler flow.
+
+Noticeably, deployment time never was a crucial issue for \emph{Cloud-SAP}, hence, this issue is of a little importance.
+What is actually important for \emph{Cloud-SAP} is deployment cost and service performance.
+In other words, slightly greater deployment time is a price that is paid to be sure that resources are deployed optimally and with Quality-of-Service guarantee.

### 2014-01-13 thesis/src/chapters/chapter-7-evaluation.tex
Thesis: changed colours

@@
   \centering
-  \begin{tabular}{ | l | l | l | l | }
-    \hline                        
-    Name & CPU & RAM & HD \\
-    \hline
-    (device) Desktop        & Intel(R) Core(TM) i5-2500 CPU @ 3.30GHz & 8GB & 1TB \\
-    (device) Node1        & AMD Athlon\texttrademark 64 X2 Dual Core, 2000MHz & 3GB & 160GB \\
-    (device) Node3        & AMD Sempron(tm) Processor 3000+ & 2.5GB & 160GB \\
-    (device) Node4        & AMD Duron(tm) Processor 1GHz& 2.5GB & 60GB \\
-    (virtual machine) Frontend\{1,2\}  & 1 CPU & 512MB & 10GB \\
+  \begin{tabular}{ l l l l l }
+    \specialrule{.1em}{.05em}{.05em}                  
+    \textbf{Name} & \textbf{VM} & \textbf{CPU} & \textbf{RAM} & \textbf{HD} \\
+    \specialrule{.1em}{.05em}{.05em} 
+     Desktop   &    & Intel(R) Core(TM) i5-2500 CPU @ 3.30GHz & 8GB & 1TB \\ \hline
+     Node1   &     & AMD Athlon\texttrademark 64 X2 Dual Core, 2000MHz & 3GB & 160GB \\ \hline
+     Node3  &      & AMD Sempron(tm) Processor 3000+ & 2.5GB & 160GB \\ \hline
+     Node4  &      & AMD Duron(tm) Processor 1GHz& 2.5GB & 60GB \\ \hline
+     Frontend\{1,2\} & \checkmark & 1 CPU & 512MB & 10GB \\
     \hline  
@@
   \centering
-  \begin{tabular}{ | c | c | c | c | }
-    \hline                        
-    & CP-1 & CP-2 & CP-3 \\
-    \hline
-    java      & 150 & 120 & 180 \\
-    ruby      & 220 & 290 & 250 \\
-    postgres  & 320 & 240 & 290 \\
-    python    & 200 & 260 & 180 \\
+  \begin{tabular}{ c  c  c  c  }
+    \specialrule{.1em}{.05em}{.05em}                  
+    & \textbf{CP-1} & \textbf{CP-2} & \textbf{CP-3} \\
+    \specialrule{.1em}{.05em}{.05em}                  
+    java      & 150 & 120 & 180 \\ \hline
+    ruby      & 220 & 290 & 250 \\ \hline
+    postgres  & 320 & 240 & 290 \\ \hline
+    python    & 200 & 260 & 180 \\ \hline
     amqp      & 330 & 390 & 285 \\
@@
   \begin{tabular}{ c  c  c }
-    \hline
-    & \multicolumn{2}{c}{Solution} \\
+    \specialrule{.1em}{.05em}{.05em} 
+    & \multicolumn{2}{c}{\textbf{Solution}} \\
     \cline{2-3}
-    Instance no & Cloud-SAP & Carina \\
-    \hline
-    2 & 198.0 & 158.7 \\
-    3 & 261.72 & 208.1 \\
+    \textbf{Instance no} & Cloud-SAP & Carina \\
+    \specialrule{.1em}{.05em}{.05em} 
+    2 & 198.0 & 158.7 \\ \hline
+    3 & 261.72 & 208.1 \\ \hline
     4 & 294.38 & 230.6 \\

### 2014-01-13 thesis/src/chapters/chapter-7-evaluation.tex
Thesis: changed table view
Thesis: chapter 7 pagination

@@
   \centering
-  \begin{tabular}{ | c | c | c | c | }
-    \hline                        
-    & CP-1 & CP-2 & CP-3 \\
-    \hline
-    java      & & x & \\
-    ruby      & x & & \\
-    postgres  & & x & \\
-    python    & & & x \\
+  \begin{tabular}{ c c c  c  }
+    \specialrule{.1em}{.05em}{.05em}                  
+    & \textbf{CP-1} & \textbf{CP-2} & \textbf{CP-3} \\
+    \specialrule{.1em}{.05em}{.05em}                  
+    java      & & x & \\ \hline
+    ruby      & x & & \\ \hline
+    postgres  & & x & \\ \hline
+    python    & & & x \\ \hline
     amqp      & & & x \\
@@
 
+\newpage
 \subsection*{Conclusion}
@@
 
+\newpage
 \subsubsection*{Hardware/VM configuration}
@@
 
-  \item[\textbf{Deployment diagrams}] Components that took part during testing are show in figures \ref{eval:auto-scaling-2cp-carina-deployment-diagram} and \ref{eval:autoscale-2cp-cloud-sap-deployment-diagram}.
+  \item[\textbf{Deployment diagrams}] Components that took part during testing are show in figures \ref{eval:autoscale-2cp-cloud-sap-deployment-diagram} and \ref{eval:auto-scaling-2cp-carina-deployment-diagram}.
 
@@
     \end{center}
-    \caption{Auto-scaling - multiple-provider based: deployment diagram of Carina}
+    \caption{Auto-scaling - multiple-provider based: deployment diagram of \emph{Cloud-SAP}}
     \label{eval:autoscale-2cp-cloud-sap-deployment-diagram}
@@
     \end{center}
-    \caption{Auto-scaling - multiple-provider based: deployment diagram of Carina}
+    \caption{Auto-scaling - multiple-provider based: deployment diagram of \emph{Carina}}
     \label{eval:auto-scaling-2cp-carina-deployment-diagram}
@@
   \begin{center}
-    \includegraphics[width=0.7\textwidth]{chapter-7/auto-scaling-2cp-physical-nodes}
+    \includegraphics[width=0.6\textwidth]{chapter-7/auto-scaling-2cp-physical-nodes}
   \end{center}
@@
 
+\newpage
 \subsection*{Expectations}
@@
 
+\newpage
 \subsubsection{Environment configuration}
@@
 
+\newpage
 \subsection*{Results}

### 2014-01-17 thesis/src/chapters/chapter-state-of-the-art.tex
Thesis: chapter 5 structure

--- 2014-01-05 thesis/src/chapters/chapter-5-design.tex
@@
-\chapter{Design of Cloud-SAP}
-
-\chapterintro{ This chapter introduces the high-level design of Cloud-SAP, highlighting its core concepts and indicating possible implementation ideas.
-}
+\chapter{State of the art}
 
@@
 
-\section{High-level design}
-\subsection{Overview}
+% radek
+\section{Carina}
 
-\begin{figure}[!ht]
-  \begin{center}
-    \includegraphics[width=0.8\textwidth]{chapter-5/hld-overview}
-  \end{center}
-  \caption{Cloud-SAP high level overview}
-  \label{design:hld-overview}
-\end{figure}
+\section{OneFlow}
 
-Diagram \ref{design:hld-overview} illustrates exemplary Cloud-SAP deployment, among depicted components we can distinguish two main parts of proposed solution: auto-scaling subsystem and inter-cloud broker.
-Auto-scaling subsystem is responsible for scaling users' application, taking into account different scalability perspectives, handled by a different layers:
-\begin{itemize}
-	\item application platform layer: application platform tuning
-	\item container layer: vertical scaling
-	\item stack layer: horizontal scaling
-	\item inter-cloud layer: scaling out across different cloud providers
-\end{itemize}
-The last mentioned layer - inter-cloud - cooperates with different cloud providers, hence, it requires additional work to be done.
-In fact, these extra tasks are delegated to a inter-cloud broker, responsible for a dialog in a cloud federation.
+\section{OpenShift}
 
-\begin{figure}[!ht]
-  \begin{center}
-    \includegraphics[width=0.5\textwidth]{chapter-5/sap-layers}
-  \end{center}
-  \caption{Layered structure of Cloud-SAP}
-  \label{design:csap-layers}
-\end{figure}
-
-\subsection{Auto-scaling subsystem}
-
-\subsubsection{Autonomic components}
-Each layer of the Cloud-SAP (shown in figure \ref{design:csap-layers}) must be characterised by an ability to adapt to a given application usage.
-One of the first models that ensured system adaptivity is an autonomic component, concept based on a feed-back loop, initially proposed by IBM \cite{IBM06}.
-Figure \ref{design:autonomic-component} depicts that architecture.
-This observation is a foundation of proposed architecture - adaptivity is achieved by enriching each layer with an elasticity controller that is in fact an autonomic component.
-
-\begin{figure}[!ht]
-  \begin{center}
-    \includegraphics{chapter-5/autonomic-component}
-  \end{center}
-  \caption{autonomic component}
-  \label{design:autonomic-component}
-\end{figure}
-
-Due to the fact that designed platform operates on multiple layers, we can extend idea of an autonomic component by using concept of a multi-hierarchical autonomic system \cite{LiWoZh05}.
-Figure \ref{design:hierarchical-autonomic-system} illustrates that hierarchy, were each level represents a different perspective on an application scaling.
-First three hierarchy levels (application tuning, container, stack) are centralised and controlled by a single elasticity controller at each level, while the last inter-cloud level is a decentralized one - each cloud instance is fully independent.
-
-\begin{figure}[!ht]
-  \begin{center}
-    \includegraphics[width=0.8\textwidth]{chapter-5/hierarchical-autonomic-system}
-  \end{center}
-  \caption{Cloud-SAP as an hierarchical autonomic system}
-  \label{design:hierarchical-autonomic-system}
-\end{figure}
-
-Considering hierarchy of our architecture, each level manages an underlying autonomic component, while being managed by an upper layer at the very same time.
-
-\subsubsection{Monitoring}
-Monitoring module of each elasticity controller aims to gather data from \textit{Sensors} of a managed resource.
-While design of Cloud-SAP doesn't have any specific requirements regarding underlying monitoring mechanism, there are a few aspects that should be taken into consideration.
-Table \ref{tab:monitoring-module-issues-summary} summarises main points.
-
-\begin{table}[!htbp]
-\begin{tabularx}{\textwidth}[]{ X X X }
-\specialrule{.1em}{.05em}{.05em}
- & \textbf{Advantages} & \textbf{Disadvantages} \\ \specialrule{.1em}{.05em}{.05em}
-
-\textbf{data exchange model} & & \\ \specialrule{.1em}{.05em}{.05em}
-push based &
--- notification driven: all changes are reflected immediately in the system
-&
--- requires additional agent at monitored component
-\\ \hline
-pull based & 
--- more calculations are being done at monitoring module side (doesn't involve additional load at managed component)
-& 
--- some monitoring cycles can be redundant if scheduler is not properly tuned
-\\ \hline
-
-\textbf{data format} & & \\ \specialrule{.1em}{.05em}{.05em}
-binary &
--- space-effective
-& 
--- unreadable for human
-\\ \hline
-character-based &
--- human readable
-&
--- ineffective in terms of space
-\\ \hline
-
-\textbf{communication protocol} & & \\ \specialrule{.1em}{.05em}{.05em}
-vendor specific &
--- provides specific data to a given managed component (ie. virtual machine / container)
-& 
--- may cause incompatibility problem while expanding cloud-sap system to a consecutive cloud instances
-\\ \hline
-standard-based &
--- supports scaling across multiple cloud instances as long as they support standard
-&
--- gathered data may be inadequate to a specific needs
-\\ \hline
-\end{tabularx}
-
-\caption{Monitoring - summary of issues}
-\label{tab:monitoring-module-issues-summary}
-
-\end{table}
-
-Cloud-SAP is entirely focused on further data processing, consequently decision whether to store monitoring data or not is entirely implementation dependent.
-
-\subsubsection{Analysis}
-Analyse module applies a specified mathematical model to a gathered data and yields a conclusion on top of that.
-There is a vast number of models that can be used for that purpose, for example \cite{LiWoZh05} enumerates:
-\begin{itemize}
- \item Queue-based Performance Models
- \item Dynamic models
- \item Monotonic static models
- \item Policy based models
-\end{itemize}
-
-It is vital that Cloud-SAP implementation use predictive approach \cite{JiPeLiCh11} and track changes \cite{ZhYaWo05} in a system, giving a solid foundation for a knowledge base that can contribute to a better data analysis in future.
-
-\subsubsection{Planning}
-Planning module is responsible for scheduling actions that have to be taken to solve problems reported by an earlier analysis. This may involve:
-\begin{itemize}
- \item reserving some resources (for example to deploy a new virtual machine)
- \item handling situation where multiple entities compete about the resources (for example taking service's priority into account)
-\end{itemize}
-
-It is possible, however, that resolving issues is not possible at a current layer. It that case, the upper-layer is responsible for handling it.
-
-\subsubsection{Execution}
-It is expected that execution module manage resource by performing various operations on resource's effector.
-Each layer is characterised by different actions that can be taken, for example: application server reconfiguration, vertical or horizontal scaling.
-
-Enforcing actions on a managed component can be done using different methods, however, all of them requires an agent installed in a managed resource.
-Cloud-SAP does not enforce any specific technology, leaving low-level decision to an implementation.
-
-\subsubsection{Knowledge base}
-Knowledge base is represented as a set of rules, policies shared across autonomic components.
-
--- TODO przyklad z ibma
-
-\subsection{Cloud federation}
-Opis najwyzej warstwy, ze ma brokery, wykorzystuje kilka chmur.
-
-\section{Auto-scaling module}
-
-\subsection{Application platform layer}
-
-\subsection{Container layer}
-Among the currently available monitoring systems remarkable are Cloud Watch \cite{CloudWatch}, OpenNebula (Information Manager / Virtual Machine Manager) \cite{OpenNebula} there are also open approaches such as Ganglia Monitoring System \cite{MaChCu04}.
-
-\subsection{Stack layer}
-
-\subsection{Cloud layer}
-
-\section{Cloud federation}
-
-\section{Solution discussion}
+% radek
+\section{CloudFoundry}

### 2014-01-17 thesis/src/chapters/chapter-5-design.tex
Thesis: chapter 5 structure

--- 2013-12-28 thesis/src/chapters/chapter-5-design.tex
@@
 
-\section{Requirements}
-One can notice that elements that yields a solution for a problem stated in the first chapter, which is ensuring that users' application provide appropriate Quality-of-Service for its customers in a most-cost effective manner, were gradually introduced in previous chapters:
+\section{Motivation}
+Opis, ze wczesniejsze sa niewystarzcajace -> maja braki
+Sys. autonomiczny lata te braki + mozna go zastosowac bo sa spelnione takie a takie wymagania (ich opis jest w blueprint)
 
-\begin{itemize}
-	\item \emph{scalability} - ability to improve application performance by enriching resources
-	\item \emph{adaptivity} - ability to adapt (i.e. scale) appropriately to a current usage pattern
-	\item \emph{inter-cloud awareness} - ability to compose an application deployment using different cloud providers; cooperation with different cloud provider to supply application with extra resources while performing application scaling
-\end{itemize}
+\section{Overview}
+Przedstawienie domeny - zasobow w naszym systemie, odniesienie tego do blueprint => wynikiem jest pomysl na system / diagram
 
-Next section states the general overview of the proposed solution, while the consecutive sections presents more detailed discussion of its elements and finally the last section summarises the design choices in a context of system requirements.
+ogolny opis systemu - jak sie odnosimy do CHOP, mamy 1 kontroler orkiestrujacy, kompletny system autonomiczny
 
-\section{High-level design}
-\subsection{Overview}
+% darek
+\section{Application Platform manager}
+w kazdej sekcji opis MAPEK - co monitoruje, w jaki sposob analizuje i planuje, wykonuje
 
-\begin{figure}[!ht]
-  \begin{center}
-    \includegraphics[width=0.8\textwidth]{chapter-5/hld-overview}
-  \end{center}
-  \caption{Cloud-SAP high level overview}
-  \label{design:hld-overview}
-\end{figure}
+% darek
+\section{Autonomic container manager}
 
-Diagram \ref{design:hld-overview} illustrates exemplary Cloud-SAP deployment, among depicted components we can distinguish two main parts of proposed solution: auto-scaling subsystem and inter-cloud broker.
-Auto-scaling subsystem is responsible for scaling users' application, taking into account different scalability perspectives, handled by a different layers:
-\begin{itemize}
-	\item application platform layer: application platform tuning
-	\item container layer: vertical scaling
-	\item stack layer: horizontal scaling
-	\item inter-cloud layer: scaling out across different cloud providers
-\end{itemize}
-The last mentioned layer - inter-cloud - cooperates with different cloud providers, hence, it requires additional work to be done.
-In fact, these extra tasks are delegated to a inter-cloud broker, responsible for a dialog in a cloud federation.
+% darek
+\section{Autonomic stack manager}
 
-\begin{figure}[!ht]
-  \begin{center}
-    \includegraphics[width=0.5\textwidth]{chapter-5/sap-layers}
-  \end{center}
-  \caption{Layered structure of Cloud-SAP}
-  \label{design:csap-layers}
-\end{figure}
+% radek
+\section{Autonomic cloud instance manager}
 
-\subsection{Auto-scaling subsystem}
+% radek
+\section{Autonomic cloud federation manager}
 
-\subsubsection{Autonomic components}
-Each layer of the Cloud-SAP (shown in figure \ref{design:csap-layers}) must be characterised by an ability to adapt to a given application usage.
-One of the first models that ensured system adaptivity is an autonomic component, concept based on a feed-back loop, initially proposed by IBM \cite{IBM06}.
-Figure \ref{design:autonomic-component} depicts that architecture.
-This observation is a foundation of proposed architecture - adaptivity is achieved by enriching each layer with an elasticity controller that is in fact an autonomic component.
-
-\begin{figure}[!ht]
-  \begin{center}
-    \includegraphics{chapter-5/autonomic-component}
-  \end{center}
-  \caption{autonomic component}
-  \label{design:autonomic-component}
-\end{figure}
-
-Due to the fact that designed platform operates on multiple layers, we can extend idea of an autonomic component by using concept of a multi-hierarchical autonomic system \cite{LiWoZh05}.
-Figure \ref{design:hierarchical-autonomic-system} illustrates that hierarchy, were each level represents a different perspective on an application scaling.
-First three hierarchy levels (application tuning, container, stack) are centralised and controlled by a single elasticity controller at each level, while the last inter-cloud level is a decentralized one - each cloud instance is fully independent.
-
-\begin{figure}[!ht]
-  \begin{center}
-    \includegraphics[width=0.8\textwidth]{chapter-5/hierarchical-autonomic-system}
-  \end{center}
-  \caption{Cloud-SAP as an hierarchical autonomic system}
-  \label{design:hierarchical-autonomic-system}
-\end{figure}
-
-Considering hierarchy of our architecture, each level manages an underlying autonomic component, while being managed by an upper layer at the very same time.
-
-\subsubsection{Monitoring}
-Monitoring module of each elasticity controller aims to gather data from \textit{Sensors} of a managed resource.
-While design of Cloud-SAP doesn't have any specific requirements regarding underlying monitoring mechanism, there are a few aspects that should be taken into consideration.
-Table \ref{tab:monitoring-module-issues-summary} summarises main points.
-
-\begin{table}[!htbp]
-\begin{tabularx}{\textwidth}[]{ X X X }
-\specialrule{.1em}{.05em}{.05em}
- & \textbf{Advantages} & \textbf{Disadvantages} \\ \specialrule{.1em}{.05em}{.05em}
-
-\textbf{data exchange model} & & \\ \specialrule{.1em}{.05em}{.05em}
-push based &
--- notification driven: all changes are reflected immediately in the system
-&
--- requires additional agent at monitored component
-\\ \hline
-pull based & 
--- more calculations are being done at monitoring module side (doesn't involve additional load at managed component)
-& 
--- some monitoring cycles can be redundant if scheduler is not properly tuned
-\\ \hline
-
-\textbf{data format} & & \\ \specialrule{.1em}{.05em}{.05em}
-binary &
--- space-effective
-& 
--- unreadable for human
-\\ \hline
-character-based &
--- human readable
-&
--- ineffective in terms of space
-\\ \hline
-
-\textbf{communication protocol} & & \\ \specialrule{.1em}{.05em}{.05em}
-vendor specific &
--- provides specific data to a given managed component (ie. virtual machine / container)
-& 
--- may cause incompatibility problem while expanding cloud-sap system to a consecutive cloud instances
-\\ \hline
-standard-based &
--- supports scaling across multiple cloud instances as long as they support standard
-&
--- gathered data may be inadequate to a specific needs
-\\ \hline
-\end{tabularx}
-
-\caption{Monitoring - summary of issues}
-\label{tab:monitoring-module-issues-summary}
-
-\end{table}
-
-Cloud-SAP is entirely focused on further data processing, consequently decision whether to store monitoring data or not is entirely implementation dependent.
-
-\subsubsection{Analysis}
-Analyse module applies a specified mathematical model to a gathered data and yields a conclusion on top of that.
-There is a vast number of models that can be used for that purpose, for example \cite{LiWoZh05} enumerates:
-\begin{itemize}
- \item Queue-based Performance Models
- \item Dynamic models
- \item Monotonic static models
- \item Policy based models
-\end{itemize}
-
-It is vital that Cloud-SAP implementation use predictive approach \cite{JiPeLiCh11} and track changes \cite{ZhYaWo05} in a system, giving a solid foundation for a knowledge base that can contribute to a better data analysis in future.
-
-\subsubsection{Planning}
-Planning module is responsible for scheduling actions that have to be taken to solve problems reported by an earlier analysis. This may involve:
-\begin{itemize}
- \item reserving some resources (for example to deploy a new virtual machine)
- \item handling situation where multiple entities compete about the resources (for example by a prioritization)
-\end{itemize}
-
-It is possible, however, that resolving issues is not possible at a current layer. It that case, the upper-layer is responsible for handling it.
-
-\subsubsection{Execution}
-It is expected that execution module manage resource by performing various operations on resource's effector.
-Each layer is characterised by different actions that can be taken, for example: application server reconfiguration, vertical or horizontal scaling.
-
-\subsubsection{Knowledge base}
-Knowledge base is represented as a set of rules, policies shared across autonomous components.
-
-\subsection{Cloud federation}
-Opis najwyzej warstwy, ze ma brokery, wykorzystuje kilka chmur.
-
-\section{Auto-scaling module}
-
-\subsection{Application platform layer}
-
-\subsection{Container layer}
-Among the currently available monitoring systems remarkable are Cloud Watch \cite{CloudWatch}, OpenNebula (Information Manager / Virtual Machine Manager) \cite{OpenNebula} there are also open approaches such as Ganglia Monitoring System \cite{MaChCu04}.
-
-\subsection{Stack layer}
-
-\subsection{Cloud layer}
-
-\section{Cloud federation}
-
-\section{Solution discussion}
+\section{Summary}
+podsumowanie, wyzwania stojace przed projektem, problemy ale takze obietnica dobrobytu i spokojnej starosci

### 2014-01-17 thesis/src/chapters/chapter-adaptivity.tex
Thesis: renamed chapter

--- 2013-09-14 thesis/src/chapters/chapter-3.tex
@@
 
-\chapterintro{This chapter introduces concepts and mechanisms that enhance a platform with adaptivity capabilities, which are often achieved by fusion of rules, policies and scaling techniques.}
+\chapterintro{This chapter introduces concepts and mechanisms that enhance a platform with adaptivity capabilities achieved by a fusion of rules, policies and scaling techniques.}
 
 \section{Introduction}
-In short, platform adaptivity adds a auto-scaling features to a solution offered by Platform-as-a-Service provider.
-Key concept is to have a Elasticity Controller which gathers probes from virtual machines and uses that knowledge to execute appropriate action on cloud instance, indirectly modifying consecutive probes \cite{VaRoBu11}.
+What makes a concept of an adaptivity enticing is crucial idea behind guaranteeing Quality-of-Serivce: automatization.
+Generally, ensuring and enforcing given quality can be seen as continuous monitoring an reacting to events when necessary.
+Such idea can be expressed as following IT process:
+
+While system administrator can be directly responsible for manually performing above-mentioned steps, it is much more favourable to automatize this process leading to system self-adaptation.
+Not only does it lead to efficiency of an IT process but also to its effectiveness \cite{IBM06}.
+It is possible through:
+\begin{itemize}
+  \item \emph{rapid process initiation} - components auto-initiates actions based on information derived from a system
+  \item \emph{reduced time and skill requirements} - automatization of IT processes makes them easier and less troublesome what is especially important for skill-intensive, error-prone and long lasting tasks
+\end{itemize}
+
+In a context of delivering a computing platform, adaptivity adds auto-scaling features to a solution offered by Platform-as-a-Service provider.
+Self-managing is possible through usage of an Elasticity Controller which gathers probes from resource such as application container and uses that knowledge to execute appropriate action on that resource, indirectly modifying consecutive probes \cite{VaRoBu11}.
 This concept illustrates diagram \ref{ch3:elasticity-controller}.

### 2014-01-17 thesis/src/chapters/chapter-introduction.tex
Thesis: renamed chapter

--- 2013-08-28 thesis/src/chapters/chapter-1.tex
@@
 \section{Motivation}
-One of the keys factors that has driven transformation of computing industry in the last years is the perception of computing utilities as an ordinary property, which can be easily accessed and adjusted to a specific needs.
+One of the keys factors that has driven transformation of computing industry in the last years is the perception of computing utilities as an ordinary property\cite{BuYeVeBrBr09}, which can be easily accessed and adjusted to a specific needs.
 That point of view resulted in profusion of different services, often collectively referred as a cloud computing \cite{MeGr11}.
@@
 
-While autonomic computing has a long history, it has not been directly applied to a multi-layered problem that exists in a cloud computing environment.
+While autonomic computing has a long history \cite{Mu04}, it has not been directly applied to a multi-layered problem that exists in a cloud computing environment.
 Especially, the research area at the last layer, which sizes across different service providers, is new.
-Although, architecture known as InterCloud investigates problem of cooperation and negotiation at cloud level, it neither has been implemented nor presented in context of autonomic system.
+Although, architecture known as InterCloud \cite{BuRaCa10} investigates problem of cooperation and negotiation at cloud level, it neither has been implemented nor presented in context of autonomic system.
 
-\subsection{Business}
+\subsection{Business potential}
 The rapid growth of interest in cloud computing in recent years resulted in huge sums of money being invested in the field.

### 2014-01-20 thesis/src/chapters/chapter-state-of-the-art.tex
Thesis: State of the art -- carina, part 1

@@
 
+\subsubsection{Overview}
+Carina is an open source project, released under Apache License 2.0, built on top of OpenNebula, which aims to ``(\ldots) standardize the process for automating multi-VM deployments and setting auto-scaling and availability management policies in the cloud.'' \cite{CarinaBlog}.
+As it is stated in the requirements of the solution, \emph{Carina} should support variety of features which can be considered worth scrutinizing carefully as they are closely related to notions of adaptivity and scalability.
+To name the most relevant: \cite{CarinaBlog}
+\begin{itemize}
+  \item Collect and aggregate OS or app-specific metrics across a cluster
+  \item Drive elastic scaling of clusters based on workload or events
+  \item Support deployment and handling of failover of services across multiple datacenters
+\end{itemize}
+
+Before delving into a more detailed discussion of these features, it is essential to introduce some Carina-specific concepts and look at the product's design and its components.
+
+\subsection{Design}
+Here are some terms used by \emph{Carina}:
+\begin{itemize}
+  \item Environment -- a collection of VMs in a master-slave configuration,
+  \item Service -- a consumer of cloud resources. Each service can have its own environment configurations and create environments and control them independently of other services,
+  \item Pools -- various clusters or virtual data centers in OpenNebula that can be targets for creating an environment
+\end{itemize}
+
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics[width=0.9\textwidth]{state-of-the-art/carina-design}
+  \end{center}
+  \caption{Carina -- components interaction}
+  \label{sota:carina-design}
+\end{figure}
+
+Adaptivity
+
+Scaling
+- vertical
+- horizontal
+
+Limitations
+
 \section{OneFlow}
@@
 \section{CloudFoundry}
+
+% radek
+\section{Providers comparison}

### 2014-01-20 thesis/src/chapters/chapter-5-design.tex
Thesis: chapter 5

@@
 \section{Autonomic container manager}
+It is expected that autonomic container manager supervises lifecycle of a container, modifying its properties accordingly to a current demand.
+
+\subsection{Managed resource}
+Container is an entity that provides an execution environment for an application platform.
+Whilst Cloud-SAP is a hypervisor agnostic, it is recommended to use operating system level virtualization technique instead of a full virtualization as a underlying mechanism for a container provisioning.
+Lightweight containers are more suitable for our needs especially because we rely entirely on Linux based operating system.
+Hence, it not necessary to use hardware-based hypervisor that supports other operating systems.
+Moreover, lightweight containers have been proved more effective in some scenarios: \cite{RaHiSj13}.
+
+Container uses variety of underlying resources such as storage or network connection. However, for a simplicity, Cloud-SAP doesn't require autonomic container manager to manage them.
+
+\subsubsection{Touchpoint}
+Container's touchpoint externalises hypervisor and operating system APIs. Similarly to all touchpoints, two main parts can be distinguished:
+\begin{itemize}
+ \item \emph{sensors} - set of properties that expose system metrics such as CPU, memory, disk usage, network bandwidth
+ \item \emph{effectors} - a collection of 'set' operations that aims to change container state: increase or decrease CPU or memory, for example.
+In fact, these operations directly corresponds to hypervisor capabilities.
+\end{itemize}
+
+It is highly recommended that sensors and effectors to be linked together, forming manageability capabilities.
+In other words, property (i.e.
+CPU) can be read only when it is possible to set it as well.
+
+\subsection{Autonomic controller}
+As one can see in XXX, autonomic controller has 4
+
+\subsubsection{Monitoring}
+
+\subsubsection{Analysis}
+\subsubsection{Planning}
+\subsubsection{Execution}
+\subsubsection{Knowledge}
 

### 2014-01-20 thesis/src/chapters/chapter-adaptivity.tex
Thesis: chapter 5

@@
 What makes a concept of an adaptivity enticing is crucial idea behind guaranteeing Quality-of-Serivce: automatization.
-Generally, ensuring and enforcing given quality can be seen as continuous monitoring an reacting to events when necessary.
-Such idea can be expressed as following IT process:
+Generally, ensuring and enforcing given quality can be seen as continuous monitoring and reacting to events when necessary.
+Such idea can be expressed as IT process shown in figure \ref{ch3:it-process}.
 
-While system administrator can be directly responsible for manually performing above-mentioned steps, it is much more favourable to automatize this process leading to system self-adaptation.
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics{chapter-3/it-process}
+  \end{center}
+  \caption{Exemplary IT process}
+  \label{ch3:it-process}
+\end{figure}
+
+While system administrator can be directly responsible for manually performing above-mentioned steps, it is much more favourable to automatize this process leading to a system self-management or an adaptable system.
 Not only does it lead to efficiency of an IT process but also to its effectiveness \cite{IBM06}.

### 2014-01-20 thesis/src/chapters/chapter-adaptivity.tex
Thesis: refactoring

@@
 Generally, ensuring and enforcing given quality can be seen as continuous monitoring and reacting to events when necessary.
-Such idea can be expressed as IT process shown in figure \ref{ch3:it-process}.
+Such idea can be expressed as IT process shown in figure \ref{fig:it-process}.
 
@@
   \begin{center}
-    \includegraphics{chapter-3/it-process}
+    \includegraphics{chapter-adaptivity/it-process}
   \end{center}
   \caption{Exemplary IT process}
-  \label{ch3:it-process}
+  \label{fig:it-process}
 \end{figure}
@@
 Self-managing is possible through usage of an Elasticity Controller which gathers probes from resource such as application container and uses that knowledge to execute appropriate action on that resource, indirectly modifying consecutive probes \cite{VaRoBu11}.
-This concept illustrates diagram \ref{ch3:elasticity-controller}.
+This concept illustrates diagram \ref{fig:elasticity-controller}.
 
@@
   \begin{center}
-    \includegraphics{chapter-3/elasticity-controller}
+    \includegraphics{chapter-adaptivity/elasticity-controller}
   \end{center}
   \caption{Elasticity controller}
-  \label{ch3:elasticity-controller}
+  \label{fig:elasticity-controller}
 \end{figure}
@@
 In cases when given metric violates that condition (i.e.
-value is either smaller than minimal or bigger than maximal acceptable) corresponding resource is properly adjusted - figure \ref{ch3:threshold-model} illustrates that idea.
+value is either smaller than minimal or bigger than maximal acceptable) corresponding resource is properly adjusted - figure \ref{fig:threshold-model} illustrates that idea.
 
@@
   \begin{center}
-    \includegraphics{chapter-3/threshold-model}
+    \includegraphics{chapter-adaptivity/threshold-model}
   \end{center}
   \caption{Threshold model}
-  \label{ch3:threshold-model}
+  \label{fig:threshold-model}
 \end{figure}

### 2014-01-20 thesis/src/chapters/chapter-evaluation.tex
Thesis: refactoring

--- 2014-01-13 thesis/src/chapters/chapter-7-evaluation.tex
@@
 \chapterintro{
-  This chapter contains results and discussions on the evaluation of tests run on the proposed solution.
+  This chapter contains results and discussions on the figuation of tests run on the proposed solution.
 }
@@
 \subsection*{Hardware and virtual machines configuration}
-Many of the carried out tests were run on the same hardware and/or used the same virtual machines so for clarity of presentation their parameters are presented not in a description of every test, but here in table \ref{tbl:test-deployment-time-common-hardware-configuration}.
+Many of the carried out tests were run on the same hardware and/or used the same virtual machines so for clarity of presentation their parameters are presented not in a description of every test, but here in table \ref{tab:test-deployment-time-common-hardware-configuration}.
 
@@
   \caption{Configuration of hardware/virtual machines used during tests}
-  \label{tbl:test-deployment-time-common-hardware-configuration}
+  \label{tab:test-deployment-time-common-hardware-configuration}
 \end{table}
@@
 Its elements are different software stacks that are parts of the whole service.
-Each cloud provider has its own price for a given software stack which is shown in table \ref{tbl:test-service-deployment-cost}.
-Diagram \ref{eval:deployment-cost-physical-nodes} illustrates the simplified environment setup.
+Each cloud provider has its own price for a given software stack which is shown in table \ref{tab:test-service-deployment-cost}.
+Diagram \ref{fig:deployment-cost-physical-nodes} illustrates the simplified environment setup.
 
@@
   \caption{Price for a stack in the given cloud provider}
-  \label{tbl:test-service-deployment-cost}
+  \label{tab:test-service-deployment-cost}
 \end{table}
@@
   \begin{center}
-    \includegraphics[width=0.6\textwidth]{chapter-7/deployment-cost-physical-nodes}
+    \includegraphics[width=0.6\textwidth]{chapter-evaluation/deployment-cost-physical-nodes}
   \end{center}
   \caption{Deployment cost: environment configuration}
-  \label{eval:deployment-cost-physical-nodes}
+  \label{fig:deployment-cost-physical-nodes}
 \end{figure}
@@
 \subsection*{Results}
-Table \ref{tbl:test-service-deployment-cost-mapping} shows obtained mapping between stacks and cloud providers.
+Table \ref{tab:test-service-deployment-cost-mapping} shows obtained mapping between stacks and cloud providers.
 Taking into account this result, figure \ref{ch7:service-deployment-cost} shows comparison of cost the client would have to pay with and without such a mapping.
@@
   \caption{Chosen cloud providers for the given stack}
-  \label{tbl:test-service-deployment-cost-mapping}
+  \label{tab:test-service-deployment-cost-mapping}
 \end{table}
@@
   \begin{center}
-    \includegraphics{chapter-7/case-study-service-deployment-reduced-client-costs}
+    \includegraphics{chapter-evaluation/case-study-service-deployment-reduced-client-costs}
   \end{center}
@@
     \item monitoring the scaling actions performed by each solution,
-    \item evaluation of the cost the client has to pay for the service.
+    \item figuation of the cost the client has to pay for the service.
   \end{inparaenum}
@@
   \item every peak in CPU usage should be followed by a gradual descent which would mimic the real auto-scaling actions executed by each solution,
-  \item (boundary conditions) CPU usage should be between 0 and 100 and, additionally, should exceed the previously set scaling threshold value of each solutions so that it would trigger the scaling policies evaluation
+  \item (boundary conditions) CPU usage should be between 0 and 100 and, additionally, should exceed the previously set scaling threshold value of each solutions so that it would trigger the scaling policies figuation
   \item once the scaling operations have completed, CPU usage should be constant at a rate which would not introduce any changes in the environment settings (such as the number of virtual machines and parameters of any virtual machine)
 \end{itemize}
-This resulted in the function whose formula is in \eqref{eq:cpu-mock-usage} and plot in figure \ref{eval:auto-scaling-1cp-cpu-usage-function}.
+This resulted in the function whose formula is in \eqref{eq:cpu-mock-usage} and plot in figure \ref{fig:auto-scaling-1cp-cpu-usage-function}.
 \begin{eqnarray}
@@
   \begin{center}
-    \includegraphics{chapter-7/cpu-usage-mock}
+    \includegraphics{chapter-evaluation/cpu-usage-mock}
   \end{center}
   \caption{Auto-scaling - single-provider based: CPU usage function}
-  \label{eval:auto-scaling-1cp-cpu-usage-function}
+  \label{fig:auto-scaling-1cp-cpu-usage-function}
 \end{figure}
@@
 \begin{asparaenum}
-  \item[\textbf{Auto-scaling policy specifications}] As it is shown in figure \ref{eval:auto-scaling-1cp-cpu-usage-function}, we set in both products the threshold values of CPU usage that trigger performing auto-scaling actions to 20 and 60 percent.
+  \item[\textbf{Auto-scaling policy specifications}] As it is shown in figure \ref{fig:auto-scaling-1cp-cpu-usage-function}, we set in both products the threshold values of CPU usage that trigger performing auto-scaling actions to 20 and 60 percent.
 To apply this setting, the auto-scaling part of service specification looks as follows: in Carina the user has to add the parameters of an auto-scaling policy in a hash that describes the environment under key \emph{:elasticy\_policy}.
-It is possible to specify the minimal and the maximal number of virtual machines that forms the environment and, what is most important, expressions which evaluation results in scaling the application.
+It is possible to specify the minimal and the maximal number of virtual machines that forms the environment and, what is most important, expressions which figuation results in scaling the application.
 The part responsible for this is shown in listing \ref{lst:auto-scaling-1cp-carina-service-spec}.\lstinputlisting[caption=Carina service specification used for testing auto-scaling with 1 cloud provider, label=lst:auto-scaling-1cp-carina-service-spec]{auto-scaling-1cp-carina-service-spec} In Cloud-SAP it is a matter of setting appropriate arguments to a specific policy.
@@
 Listing \ref{lst:auto-scaling-1cp-one-config} shows the most important OpenNebula excerpt from settings file used in this test case -- polling interval specification, which was set to 30 seconds.
-To ensure that both products can actually use up-to-date data, they should evaluate their policies every 30 seconds or longer.
+To ensure that both products can actually use up-to-date data, they should figuate their policies every 30 seconds or longer.
 In Carina the user cannot specify this value, because it is hard-coded in source code to 60 seconds.
@@
   \item[\textbf{Deployment diagrams}] Deployment diagrams show the placement of each component in both installations.
-They are shown in figures \ref{eval:auto-scaling-1cp-carina-deployment-diagram} and \ref{eval:auto-scaling-1cp-cloud-sap-deployment-diagram}.
+They are shown in figures \ref{fig:auto-scaling-1cp-carina-deployment-diagram} and \ref{fig:auto-scaling-1cp-cloud-sap-deployment-diagram}.
 
@@
     \begin{center}
-      \includegraphics{chapter-7/cpu-usage-carina-deployment-diagram}
+      \includegraphics{chapter-evaluation/cpu-usage-carina-deployment-diagram}
     \end{center}
     \caption{Auto-scaling - single-provider based: deployment diagram of Carina}
-    \label{eval:auto-scaling-1cp-carina-deployment-diagram}
+    \label{fig:auto-scaling-1cp-carina-deployment-diagram}
   \end{figure}
@@
     \begin{center}
-      \includegraphics{chapter-7/cpu-usage-cloud-sap-deployment-diagram}
+      \includegraphics{chapter-evaluation/cpu-usage-cloud-sap-deployment-diagram}
     \end{center}
     \caption{Auto-scaling - single-provider based: deployment diagram of Cloud-SAP}
-    \label{eval:auto-scaling-1cp-cloud-sap-deployment-diagram}
+    \label{fig:auto-scaling-1cp-cloud-sap-deployment-diagram}
   \end{figure}
@@
 \subsubsection*{Hardware/VM configuration}
-All virtual machines were deployed onto \emph{Node1}, which uses \emph{OpenVZ} as a virtualization technology and whose hardware configuration can be found in table \ref{tbl:test-deployment-time-common-hardware-configuration}.
+All virtual machines were deployed onto \emph{Node1}, which uses \emph{OpenVZ} as a virtualization technology and whose hardware configuration can be found in table \ref{tab:test-deployment-time-common-hardware-configuration}.
 \emph{OpenNebula} was installed on a \emph{Frontend1} virtual machine and its configuration is in the same table.
-\emph{Carina} required the usage of 2 virtual machines which were deployed on \emph{Desktop} with KVM as a hypervisor and whose configuration is in table \ref{tbl:test-auto-scaling-1cp-hardware-configuration}.
-Diagram \ref{eval:auto-scaling-1cp-physical-nodes} illustrates the physical setup of the test-case.
+\emph{Carina} required the usage of 2 virtual machines which were deployed on \emph{Desktop} with KVM as a hypervisor and whose configuration is in table \ref{tab:test-auto-scaling-1cp-hardware-configuration}.
+Diagram \ref{fig:auto-scaling-1cp-physical-nodes} illustrates the physical setup of the test-case.
 
@@
   \caption{Configuration of hardware/virtual machines used during tests}
-  \label{tbl:test-auto-scaling-1cp-hardware-configuration}
+  \label{tab:test-auto-scaling-1cp-hardware-configuration}
 \end{table}
@@
   \begin{center}
-    \includegraphics[width=0.8\textwidth]{chapter-7/auto-scaling-1cp-physical-nodes}
+    \includegraphics[width=0.8\textwidth]{chapter-evaluation/auto-scaling-1cp-physical-nodes}
   \end{center}
   \caption{Auto-scaling - single-provider based: environment configuration}
-  \label{eval:auto-scaling-1cp-physical-nodes}
+  \label{fig:auto-scaling-1cp-physical-nodes}
 \end{figure}
@@
 \subsection*{Expectations}
-As auto-scaling behaviour of both products is determined by CPU usage of the deployed service, we can predict the outcome of the test by thorough analysis of the plot in figure \ref{eval:auto-scaling-1cp-cpu-usage-function}.
+As auto-scaling behaviour of both products is determined by CPU usage of the deployed service, we can predict the outcome of the test by thorough analysis of the plot in figure \ref{fig:auto-scaling-1cp-cpu-usage-function}.
 
@@
 In our scenario cost is roughly equivalent with the CPU usage, so the plot of a cost virtually presents the cpu usage.
-Figure \ref{eval:auto-scaling-1cp-cost-comparison} shows cpu usage by environment when deployed and configured by two competing products.
+Figure \ref{fig:auto-scaling-1cp-cost-comparison} shows cpu usage by environment when deployed and configured by two competing products.
 \begin{figure}[!ht]
   \begin{center}
-    \includegraphics{chapter-7/auto-scaling-1cp-cost-comparison}
+    \includegraphics{chapter-evaluation/auto-scaling-1cp-cost-comparison}
   \end{center}
   \caption{Auto-scaling - single-provider based: comparison of cost/CPU usage}
-  \label{eval:auto-scaling-1cp-cost-comparison}
+  \label{fig:auto-scaling-1cp-cost-comparison}
 \end{figure}
@@
 
-Judging by the plot of cost (figure \ref{eval:auto-scaling-1cp-cost-comparison}), we can be fairly sure that it is possible to obtain better results by proper tuning of the scaling vertical mechanism.
+Judging by the plot of cost (figure \ref{fig:auto-scaling-1cp-cost-comparison}), we can be fairly sure that it is possible to obtain better results by proper tuning of the scaling vertical mechanism.
 In particular, one can consider changing the default CPU growth from 1.3 to other or apply a different strategy, for example a growth by a constant rate.
@@
 However, at some point single cloud provider resources will be surpassed.
-Equation \eqref{eq:cpu-mock-usage-2cp} denotes a resource usage function that is expected to fulfil above-mentioned scenario and is illustrated in figure \ref{eval:auto-scaling-2cp-cpu-usage-function}.
+Equation \eqref{eq:cpu-mock-usage-2cp} denotes a resource usage function that is expected to fulfil above-mentioned scenario and is illustrated in figure \ref{fig:auto-scaling-2cp-cpu-usage-function}.
 \begin{equation}
@@
   \begin{center}
-    \includegraphics{chapter-7/cpu-usage-mock-multiple}
+    \includegraphics{chapter-evaluation/cpu-usage-mock-multiple}
   \end{center}
   \caption{Auto-scaling - multiple-provider based: CPU usage function}
-  \label{eval:auto-scaling-2cp-cpu-usage-function}
+  \label{fig:auto-scaling-2cp-cpu-usage-function}
 \end{figure}
@@
 
-  \item[\textbf{Deployment diagrams}] Components that took part during testing are show in figures \ref{eval:autoscale-2cp-cloud-sap-deployment-diagram} and \ref{eval:auto-scaling-2cp-carina-deployment-diagram}.
+  \item[\textbf{Deployment diagrams}] Components that took part during testing are show in figures \ref{fig:autoscale-2cp-cloud-sap-deployment-diagram} and \ref{fig:auto-scaling-2cp-carina-deployment-diagram}.
 
@@
     \begin{center}
-      \includegraphics{chapter-7/autoscale-2cp-cloud-sap-deployment-diagram}
+      \includegraphics{chapter-evaluation/autoscale-2cp-cloud-sap-deployment-diagram}
     \end{center}
     \caption{Auto-scaling - multiple-provider based: deployment diagram of \emph{Cloud-SAP}}
-    \label{eval:autoscale-2cp-cloud-sap-deployment-diagram}
+    \label{fig:autoscale-2cp-cloud-sap-deployment-diagram}
   \end{figure}
@@
     \begin{center}
-      \includegraphics{chapter-7/cpu-usage-carina-deployment-diagram}
+      \includegraphics{chapter-evaluation/cpu-usage-carina-deployment-diagram}
     \end{center}
     \caption{Auto-scaling - multiple-provider based: deployment diagram of \emph{Carina}}
-    \label{eval:auto-scaling-2cp-carina-deployment-diagram}
+    \label{fig:auto-scaling-2cp-carina-deployment-diagram}
   \end{figure}
@@
 
-Figure \ref{eval:auto-scaling-2cp-physical-nodes} depicts physical configuration of the environment.
+Figure \ref{fig:auto-scaling-2cp-physical-nodes} depicts physical configuration of the environment.
 In short, setup was as follows: all \emph{Cloud-SAP} components, apart from cloud client deployed on \emph{Laptop}, were provisioned on \emph{Desktop}.
 \emph{Cloud Proivder 1 (CP-1)} is \emph{OpenNebula} instance known as \emph{Frontend1} which uses \emph{Node1} as a computing node and is deployed on \emph{Desktop}, while \emph{CP\-2} uses \emph{Frontend2} and \emph{Node2}.
-Specification of nodes is listed in table: \ref{tbl:test-deployment-time-common-hardware-configuration}.
+Specification of nodes is listed in table: \ref{tab:test-deployment-time-common-hardware-configuration}.
 Deployment cost of a java stack was 50 and 60 using \emph{CP-1} and \emph{CP-2}, respectively.
@@
   \begin{center}
-    \includegraphics[width=0.6\textwidth]{chapter-7/auto-scaling-2cp-physical-nodes}
+    \includegraphics[width=0.85\textwidth]{chapter-evaluation/auto-scaling-2cp-physical-nodes}
   \end{center}
   \caption{Auto-scaling - multiple-provider based: environment configuration}
-  \label{eval:auto-scaling-2cp-physical-nodes}
+  \label{fig:auto-scaling-2cp-physical-nodes}
 \end{figure}
@@
 \subsubsection*{Transactions per second}
-Chart \ref{eval:auto-scaling-2cp-tps-comparison} illustrates how service capabilities changed over the time.
+Chart \ref{fig:auto-scaling-2cp-tps-comparison} illustrates how service capabilities changed over the time.
 Service capability is expressed as a number of transactions per second that can be handled by a service.
@@
   \begin{center}
-    \includegraphics{chapter-7/auto-scaling-2cp-tps-comparison}
+    \includegraphics{chapter-evaluation/auto-scaling-2cp-tps-comparison}
   \end{center}
   \caption{Auto-scaling - multiple-provider based: comparison of processed transaction per second}
-  \label{eval:auto-scaling-2cp-tps-comparison}
+  \label{fig:auto-scaling-2cp-tps-comparison}
 \end{figure}
@@
 
-\newpage
 \section{Deployment time -- solution comparison}
@@
 
-All virtual machines were deployed on a host named \emph{Node1}, whose configuration can be found in table \ref{tbl:test-deployment-time-common-hardware-configuration}.
+All virtual machines were deployed on a host named \emph{Node1}, whose configuration can be found in table \ref{tab:test-deployment-time-common-hardware-configuration}.
 Diagram presents the physical configuration of nodes.
@@
   \begin{center}
-    \includegraphics[width=0.6\textwidth]{chapter-7/deployment-time-physical-nodes}
+    \includegraphics[width=0.6\textwidth]{chapter-evaluation/deployment-time-physical-nodes}
   \end{center}
   \caption{Deployment time - solution comparison: physical environment setup}
-  \label{eval:deployment-time-physical-nodes}
+  \label{fig:deployment-time-physical-nodes}
 \end{figure}
@@
   \begin{center}
-    \includegraphics{chapter-7/deployment-time-carina-deployment-diagram}
+    \includegraphics{chapter-evaluation/deployment-time-carina-deployment-diagram}
   \end{center}
@@
   \begin{center}
-    \includegraphics{chapter-7/deployment-time-test-cloud-sap-deployment-diagram}
+    \includegraphics{chapter-evaluation/deployment-time-test-cloud-sap-deployment-diagram}
   \end{center}
@@
 \subsection*{Results}
-Obtained results are shown in table \ref{tbl:test-service-deployment-time} and in figure \ref{ch7:deployment-time-test}.
+Obtained results are shown in table \ref{tab:test-service-deployment-time} and in figure \ref{ch7:deployment-time-test}.
 For a given number of instances we ordered deploying a service 10 times and the values shown in those figures are an average of these runs.
@@
   \caption{Average deployment time for the service with the various number of VMs used for the whole environment}
-  \label{tbl:test-service-deployment-time}
+  \label{tab:test-service-deployment-time}
 \end{table}
@@
   \begin{center}
-    \includegraphics{chapter-7/deployment-time-test}
+    \includegraphics{chapter-evaluation/deployment-time-test}
   \end{center}

### 2014-01-20 thesis/src/chapters/chapter-interoperability.tex
Thesis: refactoring

--- 2013-09-13 thesis/src/chapters/chapter-4.tex
@@
 Complexity related to the management of cloud components, Vendor lock-in, interoperability and reliability were among the most frequent obstacles to adoption in 2013 for they constituted 46\%, 35\%, 27\% and 22.3\% of responders' votes respectively.
-Total results are shown in the figure \ref{ch4:cloud-computing-adoption-obstacles-2013}.
+Total results are shown in the figure \ref{fig:cloud-computing-adoption-obstacles-2013}.
 \begin{figure}[!ht]
   \begin{center}
-    \includegraphics{chapter-4/cloud-computing-adoption-obstacles-2013}
+    \includegraphics{chapter-interoperability/cloud-computing-adoption-obstacles-2013}
   \end{center}
   \caption{Major obstacles to cloud adoption in 2013 according to \cite{NBSurvey13}}
-  \label{ch4:cloud-computing-adoption-obstacles-2013}
+  \label{fig:cloud-computing-adoption-obstacles-2013}
 \end{figure}
@@
 The second one is about the pros of adopting the hybrid cloud  -- potential users find more control (59\%) and better security (54\%) top benefits of using this deployment model.
-The other most responded benefits are shown in the figure \ref{ch4:hybrid-cloud-top-benefits}.
+The other most responded benefits are shown in the figure \ref{fig:hybrid-cloud-top-benefits}.
 
@@
   \begin{center}
-    \includegraphics{chapter-4/hybrid-cloud-top-benefits}
+    \includegraphics{chapter-interoperability/hybrid-cloud-top-benefits}
   \end{center}
   \caption{Top benefits of using the hybrid model according to \cite{RackspaceSurvey13}}
-  \label{ch4:hybrid-cloud-top-benefits}
+  \label{fig:hybrid-cloud-top-benefits}
 \end{figure}

### 2014-01-20 thesis/src/chapters/chapter-introduction.tex
Thesis: refactoring

@@
 The rapid growth of interest in cloud computing in recent years resulted in huge sums of money being invested in the field.
-Figure \ref{ch1:public-cloud-services-market-size} shows the size of the public cloud services market in 2012 and the forecast of its nearly two times growth in 2016.
+Figure \ref{chapter-fig:public-cloud-services-market-size} shows the size of the public cloud services market in 2012 and the forecast of its nearly two times growth in 2016.
 This data suggests that the subject is attractive for IT industry from the economic point of view.
@@
   \begin{center}
-    \includegraphics{chapter-1/public-cloud-services-market-size}
+    \includegraphics{chapter-introduction/public-cloud-services-market-size}
   \end{center}
   \caption{Public Cloud Services Market Size, 2010-2016 (forecast). Source: \textit{Gartner, 08/2012}}
-  \label{ch1:public-cloud-services-market-size}
+  \label{chapter-fig:public-cloud-services-market-size}
 \end{figure}

### 2014-01-20 thesis/src/chapters/chapter-scalability.tex
Thesis: renamed chapter
Thesis: refactoring

--- 2013-09-14 thesis/src/chapters/chapter-2.tex
@@
 While importance of application tuning cannot be underestimated, its detailed analysis lies outside of the scope of this dissertation.
-Figure \ref{ch2:scalability-layers} presents different scalability layers and actions that can be taken at each level to improve application performance.
+Figure \ref{fig:scalability-layers} presents different scalability layers and actions that can be taken at each level to improve application performance.
 
@@
   \begin{center}
-    \includegraphics{chapter-2/scalability-layers}
+    \includegraphics{chapter-scalability/scalability-layers}
   \end{center}
   \caption{Scalability layers}
-  \label{ch2:scalability-layers}
+  \label{fig:scalability-layers}
 \end{figure}
@@
 With all that said, there is no silver bullet - not matter what underlying mechanism platform provider decides to use, the application developer is still responsible for creating an application with scaling in-mind.
-This statement has been already proven in 1967 by Amdahl law, which in short states that sequential component of a parallel algorithm impacts efficiency for a sufficiently large number processors \cite{Am67} as shown in Figure \ref{ch2:amdahl-law}.
+This statement has been already proven in 1967 by Amdahl law, which in short states that sequential component of a parallel algorithm impacts efficiency for a sufficiently large number processors \cite{Am67} as shown in Figure \ref{fig:amdahl-law}.
 In other words, adding supplementary resources to a poorly written application (i.e.
@@
   \begin{center}
-    \includegraphics{chapter-2/amdahl-law}
+    \includegraphics{chapter-scalability/amdahl-law}
   \end{center}
   \caption{Amdahl's law}
-  \label{ch2:amdahl-law}
+  \label{fig:amdahl-law}
 \end{figure}
@@
 
-Table \ref{tab:cloud-providers-scaling} presents a summary of cloud provider scaling capabilities.
+Table \ref{tab:cloud-providers-scaling} presents a summary of cloud providers auto-scaling capabilities.
 Interestingly, all of them are focused solely on horizontal scaling, ignoring advantages offered by a fine-grained approach to scaling that leverage scaling up and application tuning.
@@
 
-Heroku & \checkmark & $\times$ & $\times$ \\ \hline
+Heroku & $\times$ & $\times$ & $\times$ \\ \hline
 \end{tabularx}

### 2014-01-20 thesis/src/chapters/chapter-state-of-the-art.tex
Thesis: refactoring
Thesis: openshift description

@@
 \chapter{State of the art}
+
+\chapterintro{This chapter presents recent achievements in a field of providing platform as a service by examining four vital products: Carina, OneFlow, CloudFoundry and OpenShift.
+It fouces mainly on aspects of adaptivity, scalability and cloud federation awareness.}
 
@@
 \begin{itemize}
+	\item \emph{adaptivity} - ability to adapt (i.e. scale) appropriately to a current usage pattern
 	\item \emph{scalability} - ability to improve application performance by enriching resources
-	\item \emph{adaptivity} - ability to adapt (i.e. scale) appropriately to a current usage pattern
 	\item \emph{inter-cloud awareness} - ability to compose an application deployment using different cloud providers; cooperation with different cloud provider to supply application with extra resources while performing application scaling
@@
   \begin{center}
-    \includegraphics[width=0.9\textwidth]{state-of-the-art/carina-design}
+    \includegraphics[width=0.9\textwidth]{chapter-state-of-the-art/carina-design}
   \end{center}
@@
 \section{OpenShift}
+The section aims to answer the question: \emph{Why OpenShift is good, if at all, at what it does?} Following sections give insight into \emph{OpenShift} \cite{OpenShift} features and technical concepts, especially covering: multi-tenacy, auto-scaling and platform deployment.
+
+\subsection{Introduction}
+\emph{OpenShift} is a developed by \emph{RedHat} Platform-as-a-Service for applications. It is available in two different flavours:
+\begin{itemize}
+ \item OpenShift Enterprise, (private cloud)
+ \item OpenShift Online / FreeShift, hosted by RedHat (public cloud)
+\end{itemize}
+
+The latter is freely available, but it is limited to only 3 gears (a 'gear' is a resource maintained by \emph{OpenShift}, such as a application server or database).
+Beside this, source code of this platform is hosted on github as a \emph{OpenShift Origin}, licensed with \emph{Apache License 2.0}.
+
+\subsection{Features}
+Key, high level features that distinguishes \emph{OpenShift} are:
+\begin{itemize}
+  \item \emph{Accelerated Application Service Delivery} by on-demand and self-provision application stack access
+  \item \emph{Minimised Vendor Lock-in} by portability (there is no proprietary APIs, technologies)
+  \item \emph{Automatic Application Scaling}
+\end{itemize}
+
+Beside this, main features include:
+\begin{itemize}
+  \item \emph{Polyglot} - there is a number of built in application stacks (jee, ruby, python, databases); user is allowed to defined one himself through a concept of 'cartridge'
+  \item \emph{One Click Deploy} - application are managed by OpenShift through git repository. Deploying new version of application comes down to pushing new version application's source code.
+  \item \emph{SELinux-based Secure Containers for multi-tenancy}
+  \item \emph{Automatic Application Stack Provisioning}, it is only necessary to specify required cartridge (application stack) and all the dependencies are provided by platform
+\end{itemize}
+
+\subsection{Key technical concepts}
+
+\subsubsection{Multi-tenancy}
+In \emph{OpenShift}, each application is represented as a set of gears.
+For example, gear is a database or application server.
+Such gears are hosted on \emph{OpenShift Nodes}.
+There is a many-to-one relationship between gear and \emph{OpenShift Nodes}.
+What is leveraged to achieve this multi-tenancy is:
+\begin{itemize}
+ \item SELinux, isolation between application running at the same node
+ \item Control Groups (cgroups), fine-grained control over the memory / CPU / IO utlization / networking on per process basis
+ \item Kernel namespaces, groups of processes are separated, so that they cannot see resources in other groups
+\end{itemize}
+
+\subsubsection{Auto-scaling}
+While creating new application, it is possible to specify whether application should scale (later this setting cannot be changed).
+Provided that scaling is enabled, additional gear is allocated to host \emph{HAProxy} \cite{HAProxy}.
+
+In order to scale, \emph{OpenShift} uses a relatively straightforward algorithm \cite{OpenShiftScaling}:
+\emph{
+\begin{quote}
+The algorithm for scaling up and scaling down is based on the number of concurrent requests to your application.
+OpenShift allocates 10 connections per gear - if HAProxy sees that you're sustaining 90% of your peak capacity, it adds another gear.
+If your demand falls to 50% of your peak capacity for several minutes, HAProxy removes that gear.
+\end{quote}}
+
+\subsubsection{Cloud federation}
+Before getting into discussion about possible work in a cloud federation environment, it is necessary to recap \emph{OpenShift's} deployment process.
+
+\begin{asparaenum}
+  \item[\textbf{Deployment}] As one may notice in 'Multi-tenancy' section, \emph{OpenShift} does not provision application as virtual machines.
+Instead they are simply processes running on \emph{OpenShift Nodes}, which themselves are instances of \emph{RedHat Enterprise Linux}.
+The \emph{OpenShift Enterprise (private cloud )} is a flavour that allows for choosing underlying infrastructure running \emph{RHEL}.
+It can be a physical node as well as virtual machine, managed, for example, by OpenStack.
+
+  \item[\textbf{Hybrid configuration}] As previous point highlights, \emph{OpenShift} doesn't use any notion of a datacenter / service provider.
+Consequently, it not possible to \emph{OpenShift} to cooperate in an intra-datacenter / service provider environment.
+\end{asparaenum}
+
+\subsection{Summary}
+Summing up, key advantage of the \emph{OpenShift} seems to be it simplicity: creating and deploying applications using \emph{OpenShift} is as simple as it gets.
+Beside this, built in support for most popular technologies and simple API for OpenShift attractive platform.
+Having that said, auto-scaling and being constraint to a single cloud provider may be not sufficient in complex scenarios.
 

### 2014-01-20 thesis/src/chapters/chapter-state-of-the-art.tex
Thesis: carina description in state of the art

@@
 	\item \emph{scalability} - ability to improve application performance by enriching resources
-	\item \emph{inter-cloud awareness} - ability to compose an application deployment using different cloud providers; cooperation with different cloud provider to supply application with extra resources while performing application scaling
+	\item \emph{cloud-federation awareness} - ability to compose an application deployment using different cloud providers; cooperation with different cloud provider to supply application with extra resources while performing application scaling
 \end{itemize}
@@
 Carina is an open source project, released under Apache License 2.0, built on top of OpenNebula, which aims to ``(\ldots) standardize the process for automating multi-VM deployments and setting auto-scaling and availability management policies in the cloud.'' \cite{CarinaBlog}.
+The project is used by the authors at their work at RIM in an OpenNebula-based private cloud.
 As it is stated in the requirements of the solution, \emph{Carina} should support variety of features which can be considered worth scrutinizing carefully as they are closely related to notions of adaptivity and scalability.
@@
 
-Before delving into a more detailed discussion of these features, it is essential to introduce some Carina-specific concepts and look at the product's design and its components.
-
-\subsection{Design}
-Here are some terms used by \emph{Carina}:
+Before delving into more detailed description it is vital to introduce most important terms used in \emph{Carina} documentation to depict this product:
 \begin{itemize}
@@
 
+\subsubsection{Scaling}
+Chapter 3 introduced and discussed various types of application scaling.
+\emph{Carina} offers only one type of scaling and this is \emph{horizontal} scaling -- user's capabilities are limited only to add/remove a virtual machine to/from a given environment.
+Unfortunately there are no means to modify the parameters of a given or a set of virtual machines.
+
+In this paper we strongly advocate spanning cloud resources across multiple vendors/providers.
+Despite having in its requirements such a position, \emph{Carina} \textbf{does not} support deployment or scaling of services across different providers.
+In the environment configuration we define only \textbf{one} endpoint -- this is can be thought of a reference to a cloud provider.
+As it can be only one per environment, it is not possible to deploy or scale some artifacts across many providers.
+\subsubsection{Adaptivity}
+In \emph{Carina} there are mechanisms which can perform application scaling, both in manual and automatic fashion.
+It is possible for a system administrator to directly modify the existing application and its environment by changing its parameters manually.
+Automatic management of an environment is done by defining \emph{scaling policies} which can be in two flavours: \emph{time-based} and \emph{load-based}.
+Those are called by the authors \emph{elasticy policies}.
+
+\emph{Time-based} policies defines how the system should react in the given window frame(s).
+In each frame we can specify the minimum number of virtual machines that comprises an environment.
+On the other hand, \emph{load-based} policies define the way the system reacts to \textbf{average cpu usage} by the environment.
+This is the simplest \emph{threshold-model} as introduced in chapter 2.
+The user enters predicates which are evaluated against data gathered by OpenNebula and execute scaling actions  (\emph{scaleup} and \emph{scaledown}) if they are true.
+What is more, the minimal and maximal number of virtual machines must be specified.
+At present the only parameter that is taken into account while performing actions triggered by \emph{load-based} policies is cpu usage.
+
+To increase the application availability, \emph{Carina} introduces the notion of \emph{availability policies}.
+Their main function is to take a recovery action in response to a deletion or errors during deployment of a virtual machine.
+Recreation of a virtual machine is the most fundamental and atomic operation under this model.
+
+\subsubsection{Design}
+To complete the discussion about \emph{Carina} it is essential to include its architecture overview. This is shown in figure \ref{sota:carina-design}.
+
 \begin{figure}[!ht]
   \begin{center}
-    \includegraphics[width=0.9\textwidth]{chapter-state-of-the-art/carina-design}
+    \includegraphics[width=\textwidth]{chapter-state-of-the-art/carina-design}
   \end{center}
@@
 
-Adaptivity
+There are number of things that cause concern:
+\begin{asparaenum}
+\item[\textbf{Persistency}] No one knows for what reason, but \emph{Carina} uses two databases.
+What makes things even worse, they completely differ in types, as \emph{MySQL} is a relational and \emph{redis} is a nosql database.
+This places additional burden on technical requirements of the platform.
+\item[\textbf{Usage of Apache}] To work properly, \emph{Carina} needs Apache http server.
+In the server cgi-bin directory there are placed bash scripts which are invoked by contextualization scripts executed in different stages of virtual machines life cycle.
+Instead, we recommend setting proper web services, environment-agnostic, in a central node which can be run next to global scheduler.
+\item[\textbf{User accounts}] \emph{Carina} forces the creation of a system-wide user for each service. These concepts should be totally independent.
+\item[\textbf{Provisioning}] User has to use in the environment specification previously defined OpenNebula images.
+Their further adjusting happens in the contextualization phase.
+This forces the user to modify the state of a virtual machine when there is a need to perform any changes, for example in the software stack.
+Instead we suggest using some provisioning software such as puppet or chef -- this would enable users to change software stack only in configuration files (recipes) not by altering the virtual machine state.
+\end{asparaenum}
 
-Scaling
-- vertical
-- horizontal
-
-Limitations
+\subsubsection{Summary}
+\emph{Carina} emerged as a promising platform built on top of OpenNebula, which lacked most of its functionalities at that time.
+The good point is the fact that this product is used in a business context and judging by the authors' description it actually works.
+However, due to deficiencies mentioned in the previous sections, \emph{Carina} cannot be considered a mature and generic solution for more demanding cloud providers.
 

### 2014-01-20 thesis/src/chapters/chapter-state-of-the-art.tex
Thesis: change subsub to sub

@@
 
-\chapterintro{This chapter presents recent achievements in a field of providing platform as a service by examining four vital products: Carina, OneFlow, CloudFoundry and OpenShift.
-It fouces mainly on aspects of adaptivity, scalability and cloud federation awareness.}
+\chapterintro{This chapter presents recent achievements in a field of platform-as-a-service model by examining: Carina, OneFlow, CloudFoundry and OpenShift.
+Especially, it focuses on aspects of adaptivity, scalability and cloud federation awareness.}
 
 \section{Requirements}
-One can notice that elements that yields a solution for a problem stated in the first chapter, which is ensuring that users' application provide appropriate Quality-of-Service for its customers in a most-cost effective manner, were gradually introduced in previous chapters:
+One can notice that elements that yields a solution to a problem stated in the first chapter, which is ensuring that users' application provide appropriate Quality-of-Service for its customers in a most-cost effective manner, were gradually introduced in previous chapters:
 
@@
 
-\subsubsection{Overview}
+\subsection{Overview}
 Carina is an open source project, released under Apache License 2.0, built on top of OpenNebula, which aims to ``(\ldots) standardize the process for automating multi-VM deployments and setting auto-scaling and availability management policies in the cloud.'' \cite{CarinaBlog}.
@@
 
-\subsubsection{Scaling}
+\subsection{Scaling}
 Chapter 3 introduced and discussed various types of application scaling.
@@
 As it can be only one per environment, it is not possible to deploy or scale some artifacts across many providers.
-\subsubsection{Adaptivity}
+
+\subsection{Adaptivity}
 In \emph{Carina} there are mechanisms which can perform application scaling, both in manual and automatic fashion.
@@
 
-\subsubsection{Design}
+\subsection{Design}
 To complete the discussion about \emph{Carina} it is essential to include its architecture overview. This is shown in figure \ref{sota:carina-design}.
@@
 
-\subsubsection{Summary}
+\subsection{Summary}
 \emph{Carina} emerged as a promising platform built on top of OpenNebula, which lacked most of its functionalities at that time.
@@
 
+\subsection{Introduction}
+\emph{OneFlow} \cite{OneFlow} is a management system integrated into \emph{OpenNebula} that aspires to enhance it with multi-tiered applications (service) deployment and auto-scaling capabilities.
+It is available as a core part of \emph{OpenNebula} since 4.2 version and released under \emph{Apache License, Version 2.0}.
+
+\subsection{Features}
+In a context of this thesis, \emph{OneFlow} enriches \emph{OpenNebula} with only one, yet powerful feature: application scaling based on policies.
+Besides this, benefits of using \emph{OneFlow} are as follows:
+\begin{itemize}
+\item defining multi-tiered applications (services) as a collection of applications
+\item providing configurable services from a catalogue and self-service portal
+\item enabling tight, efficient administrative control
+\item fine-grained access control for the secure sharing of services with other users
+\end{itemize}
+
+\subsection{Key technical concepts}
+\subsubsection{Adaptivity}
+Adaptivity is achieved by specifying policies, that violated, trigger events impacting service environment. There are two types of policies available:
+\begin{itemize}
+ \item \emph{Auto-scaling based on metrics} - policy defines an expression that triggers scaling adjustments.
+Expressions can use performance data sent from virtual machine using \emph{OneGate} as well as virtual machine data such as CPU, memory, network bandwidth
+ \item \emph{Auto-scaling based on schedule} - policy specifies a time or a time recurrence and correlates that information to appropriate service adjustments
+\end{itemize}
+
+Apart from scaling policies, scaling is controlled by a minimal and maximal number of virtual machines that can compose a service.
+
+\subsubsection{Scaling}
+\emph{OneFlow} is capable of \textbf{only} horizontal scaling: it adds or removes virtual machine using \emph{OpenNebula API}, when appropriate event is triggered.
+Violating given policy results in virtual machine's state transition into \texttt{SCALING}.
+After proper action is finished, virtual machines converges to a \texttt{COOLDOWN} state.
+What this simple state transition does is allows an environment to accommodate changes and adjust to a recent changes, disabling scaling for a given period of time.
+
+\subsubsection{Cloud federation awareness}
+\emph{OneFlow} itself \textbf{does not} cooperate with different cloud instance.
+Hence, one is limited to \emph{OpenNebula} hybrid or public cloud capabilities that employs oZones and \emph{AWS}, respectively.
+Although, they expand single \emph{OpenNebula} instance capacity it is limited in a way that discriminates it as possible solution.
+Firstly, the problem with oZones is that they require centralised management, while cloud federation is by definition distributed and independent.
+Secondly, cloud bursting into \emph{AWS} limits instance monitoring capabilities.
+Consequently, it is not possible to supervise provisioned instances.
+
+\subsection{Summary}
+Taking everything into consideration, OneFlow is a valuable addition to OpenNebula ecosystem.
+Undoubtedly, its auto-scaling feature with elastic policy mechanism cannot be underestimated.
+However, actions taken by scaling mechanism are coarse-grained by being limited to a horizontal scaling.
+What makes things worse, OneFlow can not leverage resources of multiple providers associated within a cloud federation.
+
 \section{OpenShift}
-The section aims to answer the question: \emph{Why OpenShift is good, if at all, at what it does?} Following sections give insight into \emph{OpenShift} \cite{OpenShift} features and technical concepts, especially covering: multi-tenacy, auto-scaling and platform deployment.
 
 \subsection{Introduction}
-\emph{OpenShift} is a developed by \emph{RedHat} Platform-as-a-Service for applications. It is available in two different flavours:
+\emph{OpenShift} is developed by \emph{RedHat} Platform-as-a-Service solution. It is available in two different flavours:
 \begin{itemize}
- \item OpenShift Enterprise, (private cloud)
- \item OpenShift Online / FreeShift, hosted by RedHat (public cloud)
+ \item \emph{OpenShift Enterprise} (private cloud)
+ \item \emph{OpenShift Online / FreeShift} hosted by RedHat (public cloud)
 \end{itemize}
 
-The latter is freely available, but it is limited to only 3 gears (a 'gear' is a resource maintained by \emph{OpenShift}, such as a application server or database).
-Beside this, source code of this platform is hosted on github as a \emph{OpenShift Origin}, licensed with \emph{Apache License 2.0}.
+The latter is freely distributed, but it is limited to only 3 gears (a 'gear' is a resource maintained by \emph{OpenShift}, such as application server or database).
+Beside this, source code is hosted on github and licensed under \emph{Apache License 2.0}.
 
 \subsection{Features}
-Key, high level features that distinguishes \emph{OpenShift} are:
+High level features that distinguishes \emph{OpenShift} are as follows:
 \begin{itemize}
-  \item \emph{Accelerated Application Service Delivery} by on-demand and self-provision application stack access
-  \item \emph{Minimised Vendor Lock-in} by portability (there is no proprietary APIs, technologies)
-  \item \emph{Automatic Application Scaling}
+  \item accelerated application service delivery by on-demand and self-provision application stack access
+  \item minimised vendor lock-in by portability (there are no proprietary APIs, technologies)
+  \item automatic application scaling
 \end{itemize}
 
-Beside this, main features include:
+Beside this, key qualities include:
 \begin{itemize}
-  \item \emph{Polyglot} - there is a number of built in application stacks (jee, ruby, python, databases); user is allowed to defined one himself through a concept of 'cartridge'
-  \item \emph{One Click Deploy} - application are managed by OpenShift through git repository. Deploying new version of application comes down to pushing new version application's source code.
-  \item \emph{SELinux-based Secure Containers for multi-tenancy}
-  \item \emph{Automatic Application Stack Provisioning}, it is only necessary to specify required cartridge (application stack) and all the dependencies are provided by platform
+  \item polyglot stacks: there is a number of built in application stacks (Java, Ruby, Python, databases); user is allowed to defined one himself through a concept of 'cartridge'
+  \item one click deployment: application are managed by \emph{OpenShift} through git repository. Deploying new version of application comes down to pushing new version application's source code.
+  \item SELinux-based secure containers for multi-tenancy
+  \item automatic application stack provisioning, it is only necessary to specify required cartridge (application stack) and all the dependencies are provided by platform
 \end{itemize}
@@
 
+\subsubsection{Adaptivity}
+\emph{OpenShift}'s adaptivity capabilities are constrained to a simple case: while creating a new application, it is possible to specify whether application should scale.
+As a consequence, additional gear is allocated to serve \emph{HAProxy} \cite{HAProxy}.
+There is only \textbf{one built-in} scaling policy that is based on a relatively straightforward algorithm \cite{OpenShiftScaling}:
+\emph{\begin{quote}
+The algorithm for scaling up and scaling down is based on the number of concurrent requests to your application.
+OpenShift allocates 10 connections per gear - if HAProxy sees that you're sustaining 90\% of your peak capacity, it adds another gear.
+If your demand falls to 50\% of your peak capacity for several minutes, HAProxy removes that gear.
+\end{quote}}
+
+\subsubsection{Scalability}
+As previous point implies, OpenShift \textbf{solely} uses horizontal scaling: it adds additional gears if number of requests increases beyond certain level.
+
+\subsubsection{Cloud federation awareness}
+OpenShift does not use any notion of a cloud provider or datacenter.
+However, since cartridges are in fact Linux processes running on OpenShift Nodes, it is possible to deploy  OpenShift Enterprise on any infrastructure running RHEL.
+It can consist of physical nodes as well as virtual machines, managed, for example, by OpenStack.
+With all that said, is still requires a centralised, OpenShift controlled environment.
+Hence, OpenShift \textbf{is not} capable to work in a cloud federation.
+
 \subsubsection{Multi-tenancy}
-In \emph{OpenShift}, each application is represented as a set of gears.
+In OpenShift, each application is represented as a set of gears.
 For example, gear is a database or application server.
-Such gears are hosted on \emph{OpenShift Nodes}.
-There is a many-to-one relationship between gear and \emph{OpenShift Nodes}.
+Gears are hosted on OpenShift Nodes.
+There is a many-to-one relationship between gear and OpenShift Nodes.
 What is leveraged to achieve this multi-tenancy is:
 \begin{itemize}
- \item SELinux, isolation between application running at the same node
- \item Control Groups (cgroups), fine-grained control over the memory / CPU / IO utlization / networking on per process basis
- \item Kernel namespaces, groups of processes are separated, so that they cannot see resources in other groups
+ \item SELinux, isolation between applications running at the same node
+ \item control groups (cgroups), fine-grained control over the memory / CPU / IO utilisation / networking on per process basis
+ \item kernel namespaces, groups of processes are separated, so that they cannot see resources in other groups
 \end{itemize}
 
-\subsubsection{Auto-scaling}
-While creating new application, it is possible to specify whether application should scale (later this setting cannot be changed).
-Provided that scaling is enabled, additional gear is allocated to host \emph{HAProxy} \cite{HAProxy}.
-
-In order to scale, \emph{OpenShift} uses a relatively straightforward algorithm \cite{OpenShiftScaling}:
-\emph{
-\begin{quote}
-The algorithm for scaling up and scaling down is based on the number of concurrent requests to your application.
-OpenShift allocates 10 connections per gear - if HAProxy sees that you're sustaining 90% of your peak capacity, it adds another gear.
-If your demand falls to 50% of your peak capacity for several minutes, HAProxy removes that gear.
-\end{quote}}
-
-\subsubsection{Cloud federation}
-Before getting into discussion about possible work in a cloud federation environment, it is necessary to recap \emph{OpenShift's} deployment process.
-
-\begin{asparaenum}
-  \item[\textbf{Deployment}] As one may notice in 'Multi-tenancy' section, \emph{OpenShift} does not provision application as virtual machines.
-Instead they are simply processes running on \emph{OpenShift Nodes}, which themselves are instances of \emph{RedHat Enterprise Linux}.
-The \emph{OpenShift Enterprise (private cloud )} is a flavour that allows for choosing underlying infrastructure running \emph{RHEL}.
-It can be a physical node as well as virtual machine, managed, for example, by OpenStack.
-
-  \item[\textbf{Hybrid configuration}] As previous point highlights, \emph{OpenShift} doesn't use any notion of a datacenter / service provider.
-Consequently, it not possible to \emph{OpenShift} to cooperate in an intra-datacenter / service provider environment.
-\end{asparaenum}
-
 \subsection{Summary}
-Summing up, key advantage of the \emph{OpenShift} seems to be it simplicity: creating and deploying applications using \emph{OpenShift} is as simple as it gets.
-Beside this, built in support for most popular technologies and simple API for OpenShift attractive platform.
-Having that said, auto-scaling and being constraint to a single cloud provider may be not sufficient in complex scenarios.
+Summing up, key advantage of the OpenShift seems to be it simplicity: creating and deploying applications using OpenShift is as simple as it gets.
+Beside this, built in support for most popular technologies and simple API makes OpenShift attractive platform.
+Having that said, oversimplified auto-scaling and being constrained to a single cloud provider may be not sufficient in complicated scenarios.
 

### 2014-01-21 thesis/src/chapters/chapter-state-of-the-art.tex
Thesis: oneflow description

@@
 
-\subsection{Overview}
+\subsubsection{Overview}
 Carina is an open source project, released under Apache License 2.0, built on top of OpenNebula, which aims to ``(\ldots) standardize the process for automating multi-VM deployments and setting auto-scaling and availability management policies in the cloud.'' \cite{CarinaBlog}.
@@
 
-\subsection{Scaling}
+\subsubsection{Scaling}
 Chapter 3 introduced and discussed various types of application scaling.
@@
 
-\subsection{Adaptivity}
+\subsubsection{Adaptivity}
 In \emph{Carina} there are mechanisms which can perform application scaling, both in manual and automatic fashion.
@@
 
-\subsection{Design}
+\subsubsection{Design}
 To complete the discussion about \emph{Carina} it is essential to include its architecture overview. This is shown in figure \ref{sota:carina-design}.
@@
 
-\subsection{Summary}
+\subsubsection{Summary}
 \emph{Carina} emerged as a promising platform built on top of OpenNebula, which lacked most of its functionalities at that time.
@@
 \subsection{Summary}
-Taking everything into consideration, OneFlow is a valuable addition to OpenNebula ecosystem.
-Undoubtedly, its auto-scaling feature with elastic policy mechanism cannot be underestimated.
-However, actions taken by scaling mechanism are coarse-grained by being limited to a horizontal scaling.
-What makes things worse, OneFlow can not leverage resources of multiple providers associated within a cloud federation.
 

### 2014-01-22 thesis/src/chapters/chapter-state-of-the-art.tex
Thesis: summary of oneflow section

@@
 \subsection{Summary}
+Taking everything into consideration, OneFlow is a valuable addition to OpenNebula ecosystem.
+Undoubtedly, its auto-scaling feature with elastic policy mechanism cannot be underestimated.
+However, actions taken by scaling mechanism are coarse-grained by being limited to a horizontal scaling.
+What makes things worse, OneFlow can not leverage resources of multiple providers associated within a cloud federation.
 

### 2014-01-22 thesis/src/chapters/chapter-state-of-the-art.tex
Thesis: Add history of cloud foundry

--- 2014-01-20 thesis/src/chapters/chapter-state-of-the-art.tex
@@
 \section{CloudFoundry}
+\emph{Cloud Foundry} is an open-source platform-as-a-service solution which can be installed on local or off-premises infrastructure, such as Amazon Web Services, OpenStack or vSphere.
+
+\subsection{History}
+The project was initially developed by VMWare and its first release was in 2011, It was then that VMware decided to make it available to the general public and released it under Apache License 2.0.
+For the next consecutive two years, there were maintained two versions of \emph{Cloud Foundry} -- as a hosted solution owned by VMWare and an open-source project.
+However, in December 2012 VMware and EMC shared information on the \emph{Pivotal Initiative} \cite{PivotalInitiative} -- a virtual organization of people from both companies with background in \emph{big data} and \emph{cloud application platforms}.
+With the advent of the this new entity, \emph{Pivotal} \cite{GoPivotal}, it was announced the launch of the new product -- Pivotal One, which is powered by \emph{Pivotal CF}, the enterprise version of \emph{Cloud Foundry}.
+Now users can choose between the open-source and commercial solutions.
 

### 2014-01-22 thesis/src/chapters/chapter-state-of-the-art.tex
Thesis: add adaptivity and scaling sections for CF

@@
 \section{CloudFoundry}
-\emph{Cloud Foundry} is an open-source platform-as-a-service solution which can be installed on local or off-premises infrastructure, such as Amazon Web Services, OpenStack or vSphere.
+\subsection{Overview}
+\emph{Cloud Foundry} is a platform-as-a-service solution which can be installed on local or off-premises infrastructure, such as Amazon Web Services, OpenStack or vSphere.
+It is available in two versions as
+\begin{itemize}
+  \item an open-source project
+  \item a commercial solution, \emph{Pivotal CF}
+\end{itemize}
 
@@
 
+\subsection{Features}
+
+\subsection{Key technical concepts}
+
+\subsubsection{Adaptivity}
+\emph{Cloud foundry} does not support auto-scaling -- it is only possible for the cloud consumer to scale the application manually.
+However, there can be additional tools which can add this functionality, such as \emph{RightScale}.
+
+In terms of monitoring the state of applications, the platform uses \emph{Health Manager} to collect relevant data and then it notifies \emph{Cloud Controller}, which in turn takes appropriate actions.
+This enables, for example, to spawn a new node if an existing one needs a replacement.
+
+\subsubsection{Scalability}
+\emph{Cloud foundry} support only horizontal scaling.
+
 % radek

### 2014-01-24 thesis/src/chapters/chapter-state-of-the-art.tex
Thesis: add features section of CF evaluation

@@
 \end{itemize}
+In this section we want to elaborate only on an open-source project.
 
@@
 \subsection{Features}
+Below there are listed some of the main functionalities offered by \emph{Cloud Foundry}:
+\begin{itemize}
+  \item Support for multi-provider ecosystem (\emph{Multi-Cloud})
+  \item Reduced the management burden of the life-cycle of an application by a simplified CLI that enables, e.g. faster and instant deployment,
+  \item Dynamic routing which enables and enhances horizontal scaling and updating applications,
+  \item Usage of own-developed lightweight containers (\emph{Warden}), which ensures the proper isolation level and improves the speed of movement of application instances between virtual machines (utilizing \emph{aufs} -- advanced multi layered unification filesystem)
+  \item Health management -- continually monitoring application instances and taking according recovery actions if needed
+  \item Standards-based authorization and authentication system which supports various standards such as LDAP, SAML or OAuth 2.0
+\end{itemize}
 
@@
 \subsubsection{Scalability}
-\emph{Cloud foundry} support only horizontal scaling.
+\emph{Cloud foundry} supports only horizontal scaling.
 

### 2014-01-24 thesis/src/chapters/chapter-state-of-the-art.tex
Thesis: add key technical concepts section of CF evaluation

@@
 \subsubsection{Scalability}
-\emph{Cloud foundry} supports only horizontal scaling.
+\emph{Cloud foundry} supports only horizontal scaling not only within one provider, but also across many. This feature is called \emph{Multi-Cloud} in Cloud Foundry nomenclature.
+
+\subsubsection{Cloud-federation awareness}
+As it was stated in previous sections, Cloud Foundry is \emph{cloud-federation aware}, i.e.
+it allows users to span their applications across many providers, which results in a hybrid solution that comes in many flavours: \emph{public-public}, \emph{public-private} or \emph{private-private}.
+
+% do kminy - dopisac Bosha/DEA/Warden ?
 

### 2014-01-24 thesis/src/chapters/chapter-state-of-the-art.tex
Thesis: add cf-specific technological concepts and summary

@@
 
+\subsubsection{Cloud Foundry-specific}
+\begin{asparaenum}
+\item[\textbf{Warden}] The main goal of \emph{Warden} is to manage isolated and resource-controller environments called containers.
+It is responsible for management of the whole lifecycle of an environment and provides API for actually performing any operation on a container.
+At first, \emph{Warden} used LXC technology under the hood, but now containers are an own product of \emph{Cloud Foundry} community.
+\item[\textbf{Droplet Execution Agent}] \emph{DEA} manages \emph{Warden} containers, stages user's application and runs droplets -- wrappers around already deployed applications.
+\item[\textbf{BOSH}] \emph{BOSH} is a tool chain for release, deployment and lifecycle management of distributed services. Its functionalities can be compared to those offered by \emph{Chef}.
+\end{asparaenum}
+
 \subsubsection{Adaptivity}
@@
 
-% do kminy - dopisac Bosha/DEA/Warden ?
+\subsection{Summary}
+Owing to the variety of useful features, great support from community and significant IT enterprises such as IBM, VMware to name a few, \emph{Cloud Foundry} can be considered a mature product which can compete with current leaders in the fields, such as Amazon AWS.
+Additionally, it is highly probable that acquiring the hosted solution by \emph{Pivotal} may be of a great benefit to the whole cloud community as the market will be more competitive.
 

### 2014-01-24 thesis/src/chapters/chapter-adaptivity.tex
Thesis: Move providers comparison to one place

@@
 Actions that are being triggered by a elasticity controller are focused on application scaling. Previous chapter described that problem in detail.
-
-\section{Providers comparison}
-
-Table \ref{tab:cloud-providers-adaptivity} summaries chapter with approaches to adaptivity taken by different cloud providers.
-Section 'triggered action' is omitted for brevity - it was described extensively in previous chapter.
-
-\begin{table}[!htbp]
-\begin{tabularx}{\textwidth}[]{ X  X X }
-\specialrule{.1em}{.05em}{.05em}
-
-  & \textbf{Policies} & \textbf{Data analysis} \\
-\specialrule{.1em}{.05em}{.05em}
-
-\multicolumn{3}{ l }{\textbf{Infrastructure provider}} \\
-\specialrule{.1em}{.05em}{.05em}
-
-Carina & 
--- time frame based
-
--- expression based (only for CPU)
-&
--- threshold model that takes into account minimal and maximal permitted instances of an application as well as application priority
-
-\\ \hline
-
-OneFlow 4.2 & 
--- time frame based with customizable padding
-
--- expression based build on custom language, where all vm's metrics are supported
-
--- customizable adjustment padding, cooldown time
-&
--- threshold model
-
-\\ \hline
-
-AWS EC2 & 
--- time frame based
-
--- expression based, where expressions corresponds to a AutoScalingGroup
-
--- actions are triggered by a CloudWatch alarms
-
--- customizable adjustments paddings, types, cooldown time
-&
--- threshold model, takes into account minimal and maximal permitted instances of an application as well as application priority
-\\ \hline
-
-\multicolumn{3}{ l }{\textbf{Platform provider}} \\
-\specialrule{.1em}{.05em}{.05em}
-
-CloudFoundry & $\times$ & $\times$ \\ \hline
-
-OpenShift & 
--- single built-in policy
-
- &
--- single built-in threshold model that scales an application when CPU load is greater than 50\% for a given period
-\\ \hline
-
-AppEngine &
-
--- built-in policy based on request queue length
-
--- adjustable minimal, maximal number of application instances, pending latency
-
- &
--- queue-based, new instance is provisioned if queue length got too long 
-  \\ \hline
-
-Azure & 
--- time frame based
-
--- expression based, where expression can involve either CPU usage or Queue length
-
--- customizable adjustments paddings, types, cooldown time
-&
--- threshold model, that takes into account minimal and maximal allowed instances
- \\ \hline
-
-Heroku & $\times$ & $\times$ \\ \hline
-\end{tabularx}
-
-\caption{Comparison of cloud providers approach to adaptivity}
-\label{tab:cloud-providers-adaptivity}
-
-\end{table}

### 2014-01-24 thesis/src/chapters/chapter-scalability.tex
Thesis: Move providers comparison to one place

@@
 Since this is a basic operation, all above-mentioned hypervisors supports this scenario as long as required resources are available.
-
-\section{Providers comparison}
-
-Table \ref{tab:cloud-providers-scaling} presents a summary of cloud providers auto-scaling capabilities.
-Interestingly, all of them are focused solely on horizontal scaling, ignoring advantages offered by a fine-grained approach to scaling that leverage scaling up and application tuning.
-
-\begin{table}[!htbp]
-\begin{tabularx}{\textwidth}[]{ X  X  X  X}
-\specialrule{.1em}{.05em}{.05em}
-
-  & \textbf{Horizontal scaling} & \textbf{Vertical scaling} & \textbf{Application tuning} \\
-\specialrule{.1em}{.05em}{.05em}
-
-\multicolumn{4}{ l }{\textbf{Infrastructure provider}} \\
-\specialrule{.1em}{.05em}{.05em}
-
-Carina & \checkmark & $\times$ & $\times$ \\ \hline
-
-OneFlow & \checkmark & $\times$ & $\times$ \\ \hline
-
-AWS EC2 & \checkmark & $\times$ & $\times$ \\ \hline
-
-\multicolumn{4}{ l }{\textbf{Platform provider}} \\
-\specialrule{.1em}{.05em}{.05em}
-
-CloudFoundry & $\times$ & $\times$ & $\times$ \\ \hline
-
-OpenShift & \checkmark & $\times$ & $\times$ \\ \hline
-
-AppEngine & \checkmark & $\times$ & $\times$ \\ \hline
-
-Azure & \checkmark & $\times$ & $\times$ \\ \hline
-
-Heroku & $\times$ & $\times$ & $\times$ \\ \hline
-\end{tabularx}
-
-\caption{Comparison of cloud providers scaling capabilities}
-\label{tab:cloud-providers-scaling}
-
-\end{table}

### 2014-01-24 thesis/src/chapters/chapter-state-of-the-art.tex
Thesis: Move providers comparison to one place

@@
 \section{Providers comparison}
+In this section we would like to compare all depicted providers as well as other solution in terms of \emph{adaptivity} and \emph{scalability}.
+
+Table \ref{tab:cloud-providers-scaling} presents a summary of cloud providers auto-scaling capabilities.
+Interestingly, all of them are focused solely on horizontal scaling, ignoring advantages offered by a fine-grained approach to scaling that leverage scaling up and application tuning.
+
+\begin{table}[!htbp]
+\begin{tabularx}{\textwidth}[]{ X  X  X  X}
+\specialrule{.1em}{.05em}{.05em}
+
+  & \textbf{Horizontal scaling} & \textbf{Vertical scaling} & \textbf{Application tuning} \\
+\specialrule{.1em}{.05em}{.05em}
+
+\multicolumn{4}{ l }{\textbf{Infrastructure provider}} \\
+\specialrule{.1em}{.05em}{.05em}
+
+Carina & \checkmark & $\times$ & $\times$ \\ \hline
+
+OneFlow & \checkmark & $\times$ & $\times$ \\ \hline
+
+AWS EC2 & \checkmark & $\times$ & $\times$ \\ \hline
+
+\multicolumn{4}{ l }{\textbf{Platform provider}} \\
+\specialrule{.1em}{.05em}{.05em}
+
+CloudFoundry & $\times$ & $\times$ & $\times$ \\ \hline
+
+OpenShift & \checkmark & $\times$ & $\times$ \\ \hline
+
+AppEngine & \checkmark & $\times$ & $\times$ \\ \hline
+
+Azure & \checkmark & $\times$ & $\times$ \\ \hline
+
+Heroku & $\times$ & $\times$ & $\times$ \\ \hline
+\end{tabularx}
+
+\caption{Comparison of cloud providers scaling capabilities}
+\label{tab:cloud-providers-scaling}
+
+\end{table}
+On the other hand, table \ref{tab:cloud-providers-adaptivity} is a summary with approaches to adaptivity taken by different cloud providers.
+
+\begin{table}[!htbp]
+\begin{tabularx}{\textwidth}[]{ X  X X }
+\specialrule{.1em}{.05em}{.05em}
+
+  & \textbf{Policies} & \textbf{Data analysis} \\
+\specialrule{.1em}{.05em}{.05em}
+
+\multicolumn{3}{ l }{\textbf{Infrastructure provider}} \\
+\specialrule{.1em}{.05em}{.05em}
+
+Carina & 
+-- time frame based
+
+-- expression based (only for CPU)
+&
+-- threshold model that takes into account minimal and maximal permitted instances of an application as well as application priority
+
+\\ \hline
+
+OneFlow 4.2 & 
+-- time frame based with customizable padding
+
+-- expression based build on custom language, where all vm's metrics are supported
+
+-- customizable adjustment padding, cooldown time
+&
+-- threshold model
+
+\\ \hline
+
+AWS EC2 & 
+-- time frame based
+
+-- expression based, where expressions corresponds to a AutoScalingGroup
+
+-- actions are triggered by a CloudWatch alarms
+
+-- customizable adjustments paddings, types, cooldown time
+&
+-- threshold model, takes into account minimal and maximal permitted instances of an application as well as application priority
+\\ \hline
+
+\multicolumn{3}{ l }{\textbf{Platform provider}} \\
+\specialrule{.1em}{.05em}{.05em}
+
+CloudFoundry & $\times$ & $\times$ \\ \hline
+
+OpenShift & 
+-- single built-in policy
+
+ &
+-- single built-in threshold model that scales an application when CPU load is greater than 50\% for a given period
+\\ \hline
+
+AppEngine &
+
+-- built-in policy based on request queue length
+
+-- adjustable minimal, maximal number of application instances, pending latency
+
+ &
+-- queue-based, new instance is provisioned if queue length got too long 
+  \\ \hline
+
+Azure & 
+-- time frame based
+
+-- expression based, where expression can involve either CPU usage or Queue length
+
+-- customizable adjustments paddings, types, cooldown time
+&
+-- threshold model, that takes into account minimal and maximal allowed instances
+ \\ \hline
+
+Heroku & $\times$ & $\times$ \\ \hline
+\end{tabularx}
+
+\caption{Comparison of cloud providers approach to adaptivity}
+\label{tab:cloud-providers-adaptivity}
+
+\end{table}

### 2014-01-27 thesis/src/chapters/chapter-design.tex
Thesis: autonomic container, stub

--- 2014-01-20 thesis/src/chapters/chapter-5-design.tex
@@
 
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics{chapter-design/autonomic-component}
+  \end{center}
+  \caption{autonomic component}
+  \label{img:autonomic-component}
+\end{figure}
+
 % darek
-\section{Application Platform manager}
+\section{Application platform manager}
 w kazdej sekcji opis MAPEK - co monitoruje, w jaki sposob analizuje i planuje, wykonuje
@@
 
-It is highly recommended that sensors and effectors to be linked together, forming manageability capabilities.
+It is highly recommended for sensors and effectors to be linked, forming together manageability capabilities.
 In other words, property (i.e.
@@
 \subsection{Autonomic controller}
-As one can see in XXX, autonomic controller has 4
+Generally speaking, autonomic container controller aims to automate container's management function and externalise its configuration through its interfaces.
+In order to do achieve that, it implements a control loop that fully covers container life cycle: gathering information about it, analysing it, planning and executing actions on top of that.
+These four functions along with knowledge necessary to perform them designate modular structure of a controller as shown in figure \ref{img:autonomic-component}.
 
 \subsubsection{Monitoring}
+As it was already mentioned, monitoring module aims to gather data from container's sensors.
+While design of Cloud-SAP doesn't have any specific requirements regarding underlying monitoring mechanism, there are a few aspects that should be taken into consideration:
+\begin{asparaenum}
+  \item[\textbf{Data filters}]
+
+  \item[\textbf{Persistence}]
+
+  \item[\textbf{Standard compatibility}] Whilst there is no single, industry accepted standard of virtual machine monitoring, there were initiatives such as OCCI \cite{OCCI} that aims to close that gap.
+Having scaling across multiple cloud instances in mind, it is vital to provide compatibility at a container monitoring level.
+However, in some cases, standard-defined API may no be sufficient for a specific needs
+\end{asparaenum}
 

### 2014-01-27 thesis/src/abstract.tex
Thesis: motivation for a cloudsap

@@
 
-The main contributions of this thesis are as follows: a) proposition of an architecture that enables aforementioned scenarios, b) implementation of that architecture c) simulation and laboratory tests.
+The main contributions of this thesis are as follows: a) reference architecture that enables aforementioned scenarios, b) implementation of that architecture c) simulation and laboratory tests.

### 2014-01-27 thesis/src/chapters/chapter-design.tex
Thesis: motivation for a cloudsap

@@
-\chapter{Design of Cloud-SAP}
+\chapter{Cloud-SAP}
 
-\chapterintro{ This chapter introduces the high-level design of Cloud-SAP, highlighting its core concepts and indicating possible implementation ideas.
+\chapterintro{ This chapter introduces the reference architecture known as Cloud-SAP, highlights its core concepts and indicates possible implementation ideas.
 }
@@
 \section{Motivation}
-Opis, ze wczesniejsze sa niewystarzcajace -> maja braki
-Sys. autonomiczny lata te braki + mozna go zastosowac bo sa spelnione takie a takie wymagania (ich opis jest w blueprint)
+As one can see, latest advances in providing platform-as-a-service still do not yield a comprehensive solution that embraces a range of fine-grained actions relevant to a given situation in an environment, as it was summarised in a previous chapter.
+All of the presented solutions solely engage horizontal scaling, while other IT processes and corresponding actions such as vertical scaling, cloud bursting or application platform restart remains neglected.
+What is more, none of the investigated systems is a self-managing one, taking proactive approach in enforcing given QoS.
+To make matter worse, some platforms such as Carina or OneFlow are provider-specific, hence, requiring vast knowledge and begin time-consuming.
+
+What system has all of the above-mentioned features is an autonomic one \cite{IBM06}, characterised by a self-management ability and driven by a four crucial attributes:
+\begin{itemize}
+ \item \emph{self-configuration} - dynamic adaptation to changing environment such as provisioning new application instances, application tuning, traffic delegation to an external provider
+  \item \emph{self-healing} - discovering, diagnosing and reacting to environment disruption such as container outage
+  \item \emph{self-optimising} - monitoring and tuning resources: migrating containers, offloading traffic, for example
+  \item \emph{self-protecting} - detecting and protecting against threats such as Deny-of-Service by provisioning extra resources
+\end{itemize}
+Beside this, key customer values such as IT process execution cost or time needed to complete it are enhanced through rapid process initiation and reduced time and skill requirements \cite{IBM06}.
+As specified in blueprint, applying autonomic system to manage application environment is possible because following conditions are met:
+\begin{itemize}
+  \item tasks involved in associated IT process needs to be automated
+  \item it is possible to initiate these processes based on observable and detectable situations
+  \item autonomic manager posses sufficient knowledge
+\end{itemize}
+
+Cloud Self Adaptive Platform (Cloud-SAP) is a reference architecture aspiring to supply application providers with an autonomic computing environment.
+Figure X shows Cloud-SAP place in an exemplary application provisioning request.
+
+Service consumers can be both used by an end-user or other system interface, Cloud-SAP is suitable for both cases.
+Cloud-SAP is entirely cloud provider and resources agnostic as long as they externalise management interfaces.
+
+Successive sections defines core layers of a Cloud-SAP and gives some insight into possible implementation, whilst next chapter presents proof-of-concept implementation.
 
@@
 % darek
-\section{Application platform manager}
-w kazdej sekcji opis MAPEK - co monitoruje, w jaki sposob analizuje i planuje, wykonuje
+%\section{Application platform manager}
+%w kazdej sekcji opis MAPEK - co monitoruje, w jaki sposob analizuje i planuje, wykonuje
 

### 2014-01-27 thesis/src/chapters/chapter-design.tex
Thesis: motivation for a cloudsap, v2.

@@
 \section{Motivation}
-As one can see, latest advances in providing platform-as-a-service still do not yield a comprehensive solution that embraces a range of fine-grained actions relevant to a given situation in an environment, as it was summarised in a previous chapter.
+As one can see, latest advances in providing platform-as-a-service still do not yield a comprehensive solution that embraces a range of fine-grained actions relevant to a given situation in an environment, as it was summarised in previous chapter.
 All of the presented solutions solely engage horizontal scaling, while other IT processes and corresponding actions such as vertical scaling, cloud bursting or application platform restart remains neglected.
-What is more, none of the investigated systems is a self-managing one, taking proactive approach in enforcing given QoS.
-To make matter worse, some platforms such as Carina or OneFlow are provider-specific, hence, requiring vast knowledge and begin time-consuming.
+What is more, none of the investigated systems is a self-managing one, taking passive approach in enforcing given QoS.
+To make matter worse, some platforms such as Carina or OneFlow are provider-specific, hence, requiring vast knowledge and time.
 
@@
   \item tasks involved in associated IT process needs to be automated
-  \item it is possible to initiate these processes based on observable and detectable situations
+  \item it is possible to initiate processes based on observable and detectable situations
   \item autonomic manager posses sufficient knowledge
@@
 
-Service consumers can be both used by an end-user or other system interface, Cloud-SAP is suitable for both cases.
+Service consumers can be both end-users or other system interfaces, Cloud-SAP is suitable for both cases.
 Cloud-SAP is entirely cloud provider and resources agnostic as long as they externalise management interfaces.
 
-Successive sections defines core layers of a Cloud-SAP and gives some insight into possible implementation, whilst next chapter presents proof-of-concept implementation.
+Successive sections define core layers of a Cloud-SAP and give insights into possible implementation decisions, while next chapter presents proof-of-concept implementation.
 

### 2014-01-28 thesis/src/chapters/chapter-design.tex
Thesis: refinements in motivation of csap: introducing a control loop and adding an architecture context

@@
 \end{itemize}
+
+Another indispensable function of any autonomic entity is its ability to detect an undesirable state and take appropriate actions.
+This is achieved by the presence of a \emph{control loop} which collects data about the environment, analyzes it and takes all needed steps.
+Not only can the introduced attributes be bound to an autonomic system, but they can also be thought of main and broad categories of a control loop \cite{IBM06}.
+
 Beside this, key customer values such as IT process execution cost or time needed to complete it are enhanced through rapid process initiation and reduced time and skill requirements \cite{IBM06}.
@@
 Cloud Self Adaptive Platform (Cloud-SAP) is a reference architecture aspiring to supply application providers with an autonomic computing environment.
-Figure X shows Cloud-SAP place in an exemplary application provisioning request.
+Figure \ref{img:architecture-context} shows Cloud-SAP place in an exemplary application provisioning request.
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics[width=0.9\textwidth]{chapter-design/architecture-context}
+  \end{center}
+  \caption{Cloud-SAP place in an exemplary application provisioning request}
+  \label{img:architecture-context}
+\end{figure}
 
@@
 
-Successive sections define core layers of a Cloud-SAP and give insights into possible implementation decisions, while next chapter presents proof-of-concept implementation.
+Successive sections define core layers of a Cloud-SAP and give insights into possible implementation decisions, while the next chapter presents proof-of-concept implementation.
 
@@
   \label{img:autonomic-component}
+\end{figure}
+
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics[width=\textwidth]{chapter-design/csap-layers}
+  \end{center}
+  \caption{autonomic component}
+  \label{img:csap-layers}
 \end{figure}

### 2014-01-28 thesis/src/chapters/chapter-design.tex
Thesis: Overview, part I

@@
 
-What system has all of the above-mentioned features is an autonomic one \cite{IBM06}, characterised by a self-management ability and driven by a four crucial attributes:
+What system has all of the above-mentioned features is an autonomic one \cite{IBM06}, characterised by a self-management ability and driven by four crucial attributes:
 \begin{itemize}
@@
 \end{itemize}
-
-Another indispensable function of any autonomic entity is its ability to detect an undesirable state and take appropriate actions.
-This is achieved by the presence of a \emph{control loop} which collects data about the environment, analyzes it and takes all needed steps.
-Not only can the introduced attributes be bound to an autonomic system, but they can also be thought of main and broad categories of a control loop \cite{IBM06}.
 
@@
 \section{Overview}
-Przedstawienie domeny - zasobow w naszym systemie, odniesienie tego do blueprint => wynikiem jest pomysl na system / diagram
+The previous section dealt with all premises that led to a conclusion that the proposed architecture can rely heavily on the concept of an autonomic system.
+In this section the further elaboration on the solution is presented.
+Figure \ref{img:architecture-context} showed the context which the architecture applies to.
+As one can see, Cloud-SAP can be seen as a proxy between the user and cloud instances and as a coordinator of the latter.
+Although cloud instance itself consists of a cloud manager and additional resources, in general, it is perfectly legitimate to consider it a resource as well.
 
-ogolny opis systemu - jak sie odnosimy do CHOP, mamy 1 kontroler orkiestrujacy, kompletny system autonomiczny
+Before the more detailed diagram of the solution is presented, it is essential to introduce another indispensable function that lies in the heart of any autonomic entity -- its ability to detect an undesirable state and take appropriate actions.
+This is achieved by the presence of a \emph{control loop} which collects data about the environment, analyzes it and takes all needed steps to recover the system.
+Not only can the attributes introduced in the previous section be bound to an autonomic system, but they can also be thought of main and broad categories of a control loop \cite{IBM06}.
+
+Having said so, we are ready to give a layered overview of a proposed solution, which is shown in figure \ref{img:csap-layered-structure}.
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics[width=\textwidth]{chapter-design/csap-layers}
+  \end{center}
+  \caption{Cloud-SAP components}
+  \label{img:csap-layered-structure}
+\end{figure}
+
+The building blocks of the architecture are:
+\begin{itemize}
+  \item Managed resources
+  \item Touchpoints
+  \item Touchpoint autonomic managers
+  \item Orchestrating autonomic managers
+\end{itemize}
+As resources were identified on various levels of abstraction, starting from the application that is to be deployed and ending on a cloud instance that aggregates different applications and services, we can think of the last three layers (managed resources, touchpoints and touchpoint autonomic managers) as a one logical component which can aggregate and manage other components.
+This recursive definition applies to identified managed resources described in the next section, apart from application as we consider it the most fundamental and indivisible resource.
+
+\subsection{Managed resources}
+\emph{Managed resources} form the lowest layer of a stack. When it comes to application provisioning in a cloud environment we can think of the following resources:
+\begin{itemize}
+  \item \emph{Application} -- application itself can be considered resources which can be managed by an external entity, e.g. by restarting/killing it,
+  \item \emph{Execution environment of an application} forms an inherent part of an application lifecycle and its proper tuning can greatly influence its performance,
+  \item \emph{Container} -- an isolated and controlled environment in which applications are deployed,
+  \item \emph{Stack} -- every application has a technology stack associated with it; what is more, if we confine ourselves to map one instance of a container with only one application, we can think of stack as a composition of application (and its execution environment) and a container,
+  \item \emph{Service} -- more complex solutions/applications require interactions with various components that are not necessarily implemented using the same technology stack; service denotes a logical entity that consists of a list of stacks that should form one, indivisible component,
+  \item \emph{Cloud instance} -- an entity able to deploy \emph{services}
+\end{itemize}
+What is more, the listed resources are nested within one another and one's ability to work/run smoothly highly depends on the condition and state of the resource it forms part of.
+This is shown in figure \ref{img:managed-resources-aggregation}.
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics{chapter-design/csap-resources-aggregation}
+  \end{center}
+  \caption{\emph{Managed resources} identified by Cloud-SAP and their aggregation}
+  \label{img:managed-resources-aggregation}
+\end{figure}
+
+As it is shown in figure \ref{img:csap-layered-structure}, each \emph{managed resource} can have embedded self-management control loop.
+The control loop may or may not be externally visible by the manageability interface.
+For example, an application can have \emph{self-configuring} control loop which performs appropriate runtime tuning of its configuration files, while not being visible (by not providing any API) and manageable by any entity.
+
+\subsection{Touchpoints}
+\emph{Touchpoint} forms the layer above managed resources. It has two main functions \cite{IBM06}:
+\begin{inparaenum}[1)]
+\item provides manageability interface and
+\item implements \emph{sensor} and \emph{effector} behaviour.
+\end{inparaenum}
+
+By means of manageability interface, external entities can control managed resources.
+It can be done by, e.g.
+leveraging API, configuration files or logs.
+Additionally it is possible to obtain information via touchpoint about the state of the resource.
+\emph{Sensor} and \emph{effector} behaviours refers to mechanisms that allows collecting data and change the behaviour of the resource, respectively.
+
+\subsection{Touchpoint Autonomic Manager}
+\emph{Touchpoint Autonomic Manager} is a component that implements the control loop behaviour and, in effect, manages the managed resources through exposed touchpoints.
+As it was stated in the previous sections, only four types of control loops are of an interest in the proposed architecture, i.e.
+\emph{self-configuring}, \emph{self-healing}, \emph{self-optimising} and \emph{self-protecting}.
+
+The actions executed by each control loop on the assigned managed resource are defined in a \emph{policy} that is associated with the given autonomic manager.
+Thus, there is a one-to-one mapping between managed resource and touchpoint, and touchpoint and autonomic manager.
+
+Each autonomic manager implements the control loop which can be split into four consecutive blocks, which share knowledge among one another.
+The output from one block forms an input to another.
+However, it is perfectly possible for a block to perform an action for its side effects, e.g.
+an action can result in producing relevant information contained in log files, which, in turn, comprises a piece of shared knowledge.
+Each block implements different behaviour based on the abstraction level and resource they operate on (see figure \ref{img:csap-layered-structure}).
+Nevertheless it is possible to define their generic behaviour, which embraces the following phases that correspond to a given block:
+\begin{itemize}
+  \item \emph{monitor} -- this function of autonomic manager is responsible for collecting and aggregating data about the resource it manages,
+  \item \emph{analyse} -- this part takes the data from \emph{monitor} phase and perform in-depth analysis of it, which can result in, e.g.
+prediction plans.
+It is possible to incorporate more complex mechanisms, such as machine-learning, to fully use the obtained information for better prediction models,
+  \item \emph{plan} -- this block is responsible for planning the appropriate action based on the output of the previous block and policies,
+  \item \emph{execute} -- the goal of this function of autonomic manager is to execute the planned action
+\end{itemize}
+What is more, an autonomic manager exposes \emph{sensor} and \emph{effector} manageability interfaces which enables other entities (including other autonomic managers) to manage it in a similar manner as the component itself manages a resource.
+The image of an autonomic manager with its control loop and provided interface is shown in figure \ref{img:autonomic-component}.
+%The loop which is drawn in the given component denotes a \emph{control loop} that is associated with it.
+
+%Przedstawienie domeny - zasobow w naszym systemie, odniesienie tego do blueprint => wynikiem jest pomysl na system / diagram
+
+%ogolny opis systemu - jak sie odnosimy do CHOP, mamy 1 kontroler orkiestrujacy, kompletny system autonomiczny
 
@@
   \end{center}
-  \caption{autonomic component}
+  \caption{Autonomic manager}
   \label{img:autonomic-component}
@@
 
-\begin{figure}[!ht]
-  \begin{center}
-    \includegraphics[width=\textwidth]{chapter-design/csap-layers}
-  \end{center}
-  \caption{autonomic component}
-  \label{img:csap-layers}
-\end{figure}
+\subsection{Orchestrating autonomic manager}
 

### 2014-01-28 thesis/src/chapters/chapter-design.tex
Thesis: added arrows to figure

@@
 
-Service consumers can be both end-users or other system interfaces, Cloud-SAP is suitable for both cases.
-Cloud-SAP is entirely cloud provider and resources agnostic as long as they externalise management interfaces.
+Application providers can be both end-users or other system interfaces, Cloud-SAP is suitable for both cases.
+Cloud-SAP is entirely cloud providers and resources agnostic as long as they externalise management interfaces.
 
-Successive sections define core layers of a Cloud-SAP and give insights into possible implementation decisions, while the next chapter presents proof-of-concept implementation.
+Successive sections define core layers of a Cloud-SAP and give insight into possible implementation decisions, while the next chapter presents a proof-of-concept implementation.
 
@@
 \subsection{Orchestrating autonomic manager}
+While an autonomic manager which manages a single resource may be satisfactory enough in many business cases, there are situations when additional coordination among managers is essential to attain the certain goal.
+This can result in incorporating autonomic behaviour within an entity in a system-wide scope.
+
+Such a coordination can be achieved by introducing another autonomic manager whose sole purpose is to harmonize the work of dependent ones by extensive use of their \emph{sensor} and \emph{effector} interfaces.
+These managers are called \emph{orchestrating managers}.
+
+When it comes to the complex task of ensuring the Quality of Service requirements for a client of a cloud computing environment, it is vital to ensure proper mechanisms that would allow seamless cooperation among different cloud providers.
+Such cooperation could result in a migration of an application between different clouds, market-driven choice of the cloud providers and so on.
+
+As the exemplary actions can be regarded complex ones, they cannot be associated with actions taken by autonomic managers of only one type.
+Thus, an orchestrating autonomic manager should coordinate other managers of any type -- an arbitrary mixture of self-healing, self-configuring, self-optimizing and self-protecting managers.
+Employment of these concepts in Cloud-SAP will be discussed later in detail.
 
@@
 \section{Autonomic cloud federation manager}
+\emph{Autonomic cloud federation manager} is a building block of the proposed architecture that forms the highest layer of its model (see figure \ref{img:csap-layered-structure}).
+Additionally, it is an element which directly cooperates with a client by capturing and handling their requests.
+
+This manager has two main responsibilities, which include
+\begin{inparaenum}[1)]
+\item handling deployment of a service requests from the clients and subsequent requests directly applied to that service (e.g. queries about its state), 
+\item scaling the previously-deployed service across different cloud vendors.
+\end{inparaenum}
+The former involves capturing the request from a client, determining the best cloud vendor or vendors for the described service and its deployment.
+This behaviour is strictly related with \emph{plan} and \emph{execute} functions of an autonomic manager.
+The latter, on the other hand, involves \emph{monitoring} the dependant autonomic managers that manage in an autonomic fashion cloud instances, \emph{analysing} obtained knowledge and deciding if taking any auto-scaling action is actually necessary to ensure satisfying Quality-of-Service requirements.
+If such an action is needed, it must be \emph{planned} and \emph{executed}.
+
+The use case of scaling a service withing a federation of cloud vendors can be considered an action which coordinates cloud instance managers.
+It is perfectly legitimate to view the cloud federation manager as an \emph{orchestrating} one.
+What is more, such a coordination requires cooperation with not only autonomic managers of a single type, but with the ones that are a mixture of self-healing, self-configuring, self-optimizing and self-protecting.
+Taking this into account, the orchestration can be classified as an \emph{orchestration across disciplines} \cite{IBM06}.
+
+\subsection{Structure}
+It is clear from the description of requirements of this autonomic manager in the previous section that \emph{plan} and \emph{execution} functions are common for both of them.
+Therefore the \emph{cloud federation manager} is composed of two internal autonomic managers, which perform complete different functions, the first one -- \emph{monitor} and \emph{analyse} and the second -- \emph{plan} and \emph{execute}, yet they cooperate with each other to realize the full control loop.
+
+To be able to handle incoming requests, the manager exposes API a client communicates with.
+The complete structure of this component is shown in figure \ref{img:orchestrating-autonomic-manager}.
+In the next sections a detail description of each internal autonomic manager will be presented.
+
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics[width=0.7\textwidth]{chapter-design/csap-orchestrating-autonomic-manager}
+  \end{center}
+  \caption{Design of an orchestrating autonomic manager}
+  \label{img:orchestrating-autonomic-manager}
+\end{figure}
+
+\subsection{Managed resource}
+As it was stated in the previous sections and can be seen in figure \ref{img:orchestrating-autonomic-manager}, \emph{cloud instance autonomic managers} are managed resources of this manager.
+Each cloud instance is a deployment environment for a service or a stack, i.e.
+it is possible to span the service across different clouds by deploying stacks which make up the given service to different providers.
+Of course it is legitimate to deploy the whole service to only one provider.
+\subsubsection*{Touchpoints}
+As the cloud federation manager has to gain information about the state of a given cloud instance and be able to order the deployment of a stack or a service, the cloud instance exposes its manageability interface and its two types:
+\begin{asparaenum}
+\item[\textbf{Sensors}] -- set of attributes which forms the state of a given cloud instance.
+Attributes should include inter alia overall resources usage, pricing, geographical location.
+The more thorough elaboration will be held in the section devoted to \emph{plan and execution manager}.
+\item[\textbf{Effectors}] -- set of stack/service management operations, such as deploy, migrate or delete.
+\end{asparaenum}
+
+\subsection{Plan and Execution Autonomic Manager}
+\subsection{Monitor and Analyze Autonomic Manager}
 

### 2014-01-28 thesis/src/chapters/chapter-design.tex
Thesis: comment structure

@@
 
+Beside this, key customer values such as IT process execution cost or time needed to complete it are enhanced through rapid process initiation and reduced time and skill requirements \cite{IBM06}.
+As specified in blueprint, applying autonomic system to manage application environment is possible because following conditions are met:
+
+These four system capabilities combined together yield a promising architecture from QoS ensuring perspective.
 Beside this, key customer values such as IT process execution cost or time needed to complete it are enhanced through rapid process initiation and reduced time and skill requirements \cite{IBM06}.
@@
 \section{Autonomic container manager}
-It is expected that autonomic container manager supervises lifecycle of a container, modifying its properties accordingly to a current demand.
+It is expected that autonomic container manager supervise lifecycle of a container, modifying its properties accordingly to a current demand.
 
@@
 \begin{itemize}
- \item \emph{sensors} - set of properties that expose system metrics such as CPU, memory, disk usage, network bandwidth
+ \item \emph{sensors} - set of properties that expose system metrics such as CPU, memory, disk usage, network bandwidth - lista propertisow, charakterystyka (czy wymagane, co daja, dobra metryka?)
  \item \emph{effectors} - a collection of 'set' operations that aims to change container state: increase or decrease CPU or memory, for example.
 In fact, these operations directly corresponds to hypervisor capabilities.
+- lista wymagancych, + hypervisor, metryka
 \end{itemize}
@@
 As it was already mentioned, monitoring module aims to gather data from container's sensors.
-While design of Cloud-SAP doesn't have any specific requirements regarding underlying monitoring mechanism, there are a few aspects that should be taken into consideration:
+While Cloud-SAP does not have any specific requirements regarding underlying monitoring mechanism, there are a few aspects that should be taken into consideration:
 \begin{asparaenum}
-  \item[\textbf{Data filters}]
+  \item[\textbf{Metrics}] rozne opcje: cpu, memory, co wybrac przynajmniej cpu
 
-  \item[\textbf{Persistence}]
+  \item[\textbf{Data filters}] co robic z danymi, agregowac, jesli tak to jak, trzeba niwelowac zaburzenia (art? o monitoringu)
+
+  \item[\textbf{Persistence}] czy przechowywac dane, jesli tak to jakie, potrzebne do analizy, wyciagania wnioskow (dane, reakcja -> odpowiedz)
 
@@
 \subsubsection{Analysis}
+%modele kolejkowe, obliczanie jakiegos scierwa, przynajmniej threshold
+%uwzgledniac cooldown
+
 \subsubsection{Planning}
+%zarzadanie zasobami, czy wystarczajaco
+%opcjonalnie: priorytety (wtedy moze rezerwacja tak jak bylo w sieciach)
+
 \subsubsection{Execution}
+%akcje wykonywane czyli skalowanie up
+
 \subsubsection{Knowledge}
+%???
+%na pewno jest wbudowana: plan -> wykonanie (scaling +/-)
 
@@
 \section{Autonomic stack manager}
+\subsection{Managed resource}
+%stos - jako homogeniczna grupa (moze + load balancer)
+%touchpoint - cloud provider:
+
+%sensory - ruch jaki na ktore
+%efectory - + / - instancja
 

### 2014-01-29 thesis/src/chapters/chapter-design.tex
Thesis: add introduction section of orchestrating autonomic manager

--- 2014-01-28 thesis/src/chapters/chapter-design.tex
@@
 
-Application providers can be both end-users or other system interfaces, Cloud-SAP is suitable for both cases.
-Cloud-SAP is entirely cloud providers and resources agnostic as long as they externalise management interfaces.
+Service consumers can be both end-users or other system interfaces, Cloud-SAP is suitable for both cases.
+Cloud-SAP is entirely cloud provider and resources agnostic as long as they externalise management interfaces.
 
-Successive sections define core layers of a Cloud-SAP and give insight into possible implementation decisions, while the next chapter presents a proof-of-concept implementation.
+Successive sections define core layers of a Cloud-SAP and give insights into possible implementation decisions, while the next chapter presents proof-of-concept implementation.
 
@@
 
-\subsection{Managed resource}
-As it was stated in the previous sections and can be seen in figure \ref{img:orchestrating-autonomic-manager}, \emph{cloud instance autonomic managers} are managed resources of this manager.
-Each cloud instance is a deployment environment for a service or a stack, i.e.
-it is possible to span the service across different clouds by deploying stacks which make up the given service to different providers.
-Of course it is legitimate to deploy the whole service to only one provider.
-\subsubsection*{Touchpoints}
-As the cloud federation manager has to gain information about the state of a given cloud instance and be able to order the deployment of a stack or a service, the cloud instance exposes its manageability interface and its two types:
-\begin{asparaenum}
-\item[\textbf{Sensors}] -- set of attributes which forms the state of a given cloud instance.
-Attributes should include inter alia overall resources usage, pricing, geographical location.
-The more thorough elaboration will be held in the section devoted to \emph{plan and execution manager}.
-\item[\textbf{Effectors}] -- set of stack/service management operations, such as deploy, migrate or delete.
-\end{asparaenum}
-
 \subsection{Plan and Execution Autonomic Manager}

### 2014-01-29 thesis/src/chapters/chapter-design.tex
Thesis: add managed resources section to the description of orchestrating manager

@@
 
+\subsection{Managed resource}
+As it was stated in the previous sections and can be seen in figure \ref{img:orchestrating-autonomic-manager}, \emph{cloud instance autonomic managers} are managed resources of this manager.
+Each cloud instance is a deployment environment for a service or a stack, i.e.
+it is possible to span the service across different clouds by deploying stacks which make up the given service to different providers.
+Of course it is legitimate to deploy the whole service to only one provider.
+\subsubsection*{Touchpoints}
+As the cloud federation manager has to gain information about the state of a given cloud instance and be able to order the deployment of a stack or a service, the cloud instance exposes its manageability interface and its two types:
+\begin{asparaenum}
+\item[\textbf{Sensors}] -- set of attributes which forms the state of a given cloud instance.
+Attributes should include inter alia overall resources usage, pricing, geographical location.
+The more thorough elaboration will be held in the section devoted to \emph{plan and execution manager}.
+\item[\textbf{Effectors}] -- set of stack/service management operations, such as deploy, migrate or delete.
+\end{asparaenum}
+
 \subsection{Plan and Execution Autonomic Manager}

### 2014-01-30 thesis/src/chapters/chapter-design.tex
Thesis: knowledge source

@@
 What is more, none of the investigated systems is a self-managing one, taking passive approach in enforcing given QoS.
-To make matter worse, some platforms such as Carina or OneFlow are provider-specific, hence, requiring vast knowledge and time.
+To make matter worse, some platforms such as Carina or OneFlow are provider-specific, hence, requiring vast knowledge and time of entities responsible for its supervision.
 
@@
   \item Orchestrating autonomic managers
+  \item Knowledge source
 \end{itemize}
@@
 leveraging API, configuration files or logs.
-Additionally it is possible to obtain information via touchpoint about the state of the resource.
+Additionally, it is possible to obtain information via touchpoint about the state of the resource.
 \emph{Sensor} and \emph{effector} behaviours refers to mechanisms that allows collecting data and change the behaviour of the resource, respectively.
@@
 
+\subsection{Knowledge source}
+\emph{Knowledge source} provides access to knowledge stored for example in a registry, dictionary or database.
+Depending on implementation, it can be represented as a registry, database or dictionary.
+It is recommended that knowledge source share knowledge among multiple managers, consequently extending their capabilities.
+
+Knowledge is expressed as data such as symptoms, policies, change requests or history logs. There are three main ways to obtain a knowledge:
+\begin{itemize}
+ \item dynamic
+
+\end{itemize}
+
 % darek
+%\section{Application manager}
+
 %\section{Application platform manager}
@@
 \section{Autonomic container manager}
-It is expected that autonomic container manager supervises lifecycle of a container, modifying its properties accordingly to a current demand.
+It is expected that autonomic container manager supervise lifecycle of a container, modifying its properties accordingly to a given condition.
 
 \subsection{Managed resource}
-Container is an entity that provides an execution environment for an application platform.
-Whilst Cloud-SAP is a hypervisor agnostic, it is recommended to use operating system level virtualization technique instead of a full virtualization as a underlying mechanism for a container provisioning.
-Lightweight containers are more suitable for our needs especially because we rely entirely on Linux based operating system.
-Hence, it not necessary to use hardware-based hypervisor that supports other operating systems.
-Moreover, lightweight containers have been proved more effective in some scenarios: \cite{RaHiSj13}.
+To recap, \emph{container} is an entity that provides an execution environment for an application platform.
+There is a variety of technologies that intend to provide an isolated, secure execution environment.
+Table \ref{tab:containeristation-technologies} groups them into three main categories: full virtualization, os-level virtualization, operating system process.
 
-Container uses variety of underlying resources such as storage or network connection. However, for a simplicity, Cloud-SAP doesn't require autonomic container manager to manage them.
+\begin{table}
+  \centering
+  \begin{tabular}{ c  c  c  c }
+    \specialrule{.1em}{.05em}{.05em} 
+    & \textbf{Full virtualization} & \textbf{OS-level virtualization} & \textbf{OS process} \\
+    \specialrule{.1em}{.05em}{.05em} 
+    Features &  &  & \\ \hline
+    Examples &  &  & \\
+    \hline
+  \end{tabular}
+  \caption{Containerisation technologies}
+  \label{tab:containeristation-technologies}
+\end{table}
+
+While choosing containerisation technique is entirely left to an implementation, Cloud-SAP recommends using os-level virtualization or operating system processes as they involve lesser resources and are more flexible, especially in terms of dynamic adjustment.
+Moreover, lightweight containers have been proved more effective than full virtualization in some scenarios: \cite{RaHiSj13}.
+
+Not only can container be a managed resource but also it can have embedded control loop, depending on chosen mechanism.
+For example, Xen uses mechanism called 'memory ballooning' to intelligently distribute memory resources among guest systems what is in fact an example of self-optimising control loop.
+Though, embedded container loops are interesting they does not lie in Cloud-SAP area of interests and as a consequence are not further discussed.
+
+Container uses variety of underlying resources such as storage or network connection.
+However, for simplicity, Cloud-SAP does not take them into consideration and therefore they are represented merely as a container properties not first-class resources.
 
@@
 \subsection{Autonomic controller}
-Generally speaking, autonomic container controller aims to automate container's management function and externalise its configuration through its interfaces.
+\emph{Autonomic container controller} aims to automate container's management function and externalise its configuration through its interfaces.
 In order to do achieve that, it implements a control loop that fully covers container life cycle: gathering information about it, analysing it, planning and executing actions on top of that.

### 2014-01-30 thesis/src/chapters/chapter-design.tex
Thesis: knowledge source

@@
 The image of an autonomic manager with its control loop and provided interface is shown in figure \ref{img:autonomic-component}.
-%The loop which is drawn in the given component denotes a \emph{control loop} that is associated with it.
-
-%Przedstawienie domeny - zasobow w naszym systemie, odniesienie tego do blueprint => wynikiem jest pomysl na system / diagram
-
-%ogolny opis systemu - jak sie odnosimy do CHOP, mamy 1 kontroler orkiestrujacy, kompletny system autonomiczny
 
@@
 \subsection{Knowledge source}
-\emph{Knowledge source} provides access to knowledge stored for example in a registry, dictionary or database.
-Depending on implementation, it can be represented as a registry, database or dictionary.
-It is recommended that knowledge source share knowledge among multiple managers, consequently extending their capabilities.
+\emph{Knowledge source} provides access to knowledge stored, for example, in a registry, dictionary or database.
+It is recommended that knowledge source share knowledge among multiple managers, consequently extending their capabilities by performing additional tasks covering recognising particular symptoms or applying specified policies, for instance.
 
-Knowledge is expressed as data such as symptoms, policies, change requests or history logs. There are three main ways to obtain a knowledge:
+Whilst knowledge can be expressed as symptoms, policies, change requests or history logs, Cloud-SAP differentiates its two main types:
 \begin{itemize}
- \item dynamic
+ \item \emph{policy} - set of constraints that, when evaluated, influence system behaviour and possibly trigger further actions.
+Cloud-SAP does not specify policy format, however, it is vital to employ only one format in whole system so it can be freely exchanged
+ \item \emph{problem determination knowledge} - correlation of data, symptoms and decision trees determining actions that can be taken to eliminate symptoms
+\end{itemize}
 
+An autonomic manager can obtain knowledge in one of three ways:
+\begin{itemize}
+ \item \emph{receive it dynamically} - a common way for passing a policy to a manager. 
+ \item \emph{retrieve it from external knowledge source} - an approach for obtaining symptom definitions or historical knowledge.
+For example, detailed history of a resource, its problems and actions taken can be retrieved from a log file.
+ \item \emph{create it dynamically} - monitoring and execution phase are strictly correlated and consequently data received from sensors, actions taken by effectors and their result can be stored as a knowledge and provide valuable information during future problem investigations
 \end{itemize}

### 2014-01-30 thesis/src/chapters/chapter-design.tex
Thesis: comments removal
Thesis: add the description of autonomic plan and execute manager

@@
 
-% darek
-%\section{Application manager}
+\section{Application manager}
 
-%\section{Application platform manager}
+\section{Application platform manager}
 %w kazdej sekcji opis MAPEK - co monitoruje, w jaki sposob analizuje i planuje, wykonuje
@@
 \subsubsection{Analysis}
-%modele kolejkowe, obliczanie jakiegos scierwa, przynajmniej threshold
-%uwzgledniac cooldown
 
 \subsubsection{Planning}
-%zarzadanie zasobami, czy wystarczajaco
-%opcjonalnie: priorytety (wtedy moze rezerwacja tak jak bylo w sieciach)
 
 \subsubsection{Execution}
-%akcje wykonywane czyli skalowanie up
 
 \subsubsection{Knowledge}
-%???
-%na pewno jest wbudowana: plan -> wykonanie (scaling +/-)
 
@@
 \subsection{Managed resource}
-%stos - jako homogeniczna grupa (moze + load balancer)
-%touchpoint - cloud provider:
-
-%sensory - ruch jaki na ktore
-%efectory - + / - instancja
 
@@
 The complete structure of this component is shown in figure \ref{img:orchestrating-autonomic-manager}.
-In the next sections a detail description of each internal autonomic manager will be presented.
+In the next sections a detailed description of each internal autonomic manager will be presented.
 
@@
 \subsection{Plan and Execution Autonomic Manager}
-\subsection{Monitor and Analyze Autonomic Manager}
+The main purpose of this component is to select, based on knowledge provided directly to this manager from \emph{monitor and analyze manager} or a user (i.e.
+the specification of a service), cloud providers for a given service and carry out the planned action.
+In the following sections a detailed description of each function is given.
+
+\subsubsection{Plan}
+The aim of this function is to give a detailed plan for the action to be performed.
+In the context of this dissertation the first output produced by this component, regarding the use cases of deploying a service and dynamically scaling it, is a \emph{service deployment plan}, which is a mapping between available resources (cloud instances in this case) and concrete stacks making up the service.
+The second, definitely less complex, is a queries scheme about the state of the already-deployed services which is based on user's requests.
+In this case the component only propagates the request to the execute function.
+As this is only a mediation, there will be no further elaboration on this aspect.
+
+\begin{asparaenum}
+  \item[\textbf{Problem description}]
+The problem of resource management in cloud computing environments is hard and complex.
+This is because of several aspects.
+The first one is the characteristic of resources available in a cloud -- not only are they geographically distributed, but they can be under different administrative domains and can vary in accessibility.
+This places a heavy burden on cooperation possibilities across cloud providers, both in a technical and non-technical way as providers need to maintain the high level of trust in order to effectively  deliver its products.
+Additionally, different entities forms the resources -- starting from individual devices, through virtual machines and whole workstations, ending with clusters and data centers.
+The second one are the increasing Quality-of-Service requirements from the clients.
+They want resources they use to be reliable, flexible (able to scale according to workload), fault-tolerant, energy-efficient and, of course, cheap.
+It is clear that clients and cloud vendors have different objectives and supply-and-demand patterns.
+
+There are two main approaches to tackle the problem of effective resource management and scheduling:
+  \begin{itemize}
+    \item \textbf{conventional style}, in which the mapping decision is determined by some \emph{cost function}.
+The downside of this method is the fact that in many cases the decision is a function of system-centric parameters \cite{buyya2002economic} and because of it the outcome may lead to better utilization of cloud resources, not necessarily the user's application.
+What is more, this model treats all resources as if their cost was the same and applications are considered equally important what not always is the case.
+
+      The positive side of this method is the fact that it involves only one entity that governs the scheduling and allocating the resources.
+What is more, it is up to the platform how many factors should be incorporated in such a function.
+Depending on the needs, a more sophisticated or simpler matching algorithm be applied, e.g.
+an algorithm that takes into account only available budget of a client.
+
+      Legion \cite{chapin1999legion} can be an example of a solution which uses this method.
+    \item \textbf{economics-paradigm based} \cite{buyya2001case}, in which the decision is not made statically by one entity, but is driven by the users' requirements.
+The model tries to adopt the real-world economic solution that involves markets and brokers on the computational ground.
+In this paradigm, clients (representing \emph{demand}) want to solve their problems at the lowest possible price while assuring required Quality-of-Service requirements and constraints, such as time frame.
+Resource providers (representing \emph{supply}) want to maximise the utilization of their resources while keeping their prices at the level that is attractive for other customers.
+
+      Here are the most common economic models which can be applied to resource scheduling and management \cite{buyya2002economic}:
+      \begin{itemize}
+        \item the commodity market model;
+        \item the posted price model;
+        \item the bargaining model;
+        \item the tendering/contract-net model;
+        \item the auction model;
+        \item the bid-based proportional resource sharing model;
+        \item the community/coalition/bartering model;
+        \item the monopoly and oligopoly
+      \end{itemize}
+  \end{itemize}
+
+\item[\textbf{Function input}] Depending on the actual request, this function can have access to different knowledge that makes its input:
+
+  \begin{asparaenum}
+  \item[\textbf{Service deployment}] When the user wants to deploy a service on a cloud infrastructure, they have to prepare its detailed description containing technical (stacks) and Quality-of-Service requirements.
+  \item[\textbf{Service scaling}] When an application needs to be scaled, there is available additional knowledge that this component should use of -- detailed information about service's performance up to that point, e.g.
+service resource utilization level (CPU, I/O), response time, its users profile including their geographical location and so on.
+This knowledge in a form of a prediction model should be passed to the component by the \emph{analyse} function.
+  \end{asparaenum}
+
+\item[\textbf{Function output}] As it was stated in the introductory section, this function produces \emph{service deployment plan}.
+In the same section we conducted a quick survey of the most common approaches to the resource allocation and mapping problem.
+The proposed architecture places no restrictions on the chosen approach, as both paradigms can be fitted into it.
+In the case of \emph{conventional approach}, this component should be equipped with a matching function that takes as an input service description (and the information of its so far execution in the case of an auto-scaling action), offers from cloud providers and make a final decision based on these arguments.
+  In the other case, this component can be thought of a cloud exchange \cite{InterCloud10}, that brings together service producers and consumers.
+\end{asparaenum}
+
+\subsubsection{Execute}
+This function takes either a query about the state of a service or a deployment plan of the given service and passes execution requests to the proper cloud instance autonomic managers.
+\subsection{Monitor and Analyse Autonomic Manager}
 
 \section{Summary}
-podsumowanie, wyzwania stojace przed projektem, problemy ale takze obietnica dobrobytu i spokojnej starosci
+%podsumowanie, wyzwania stojace przed projektem, problemy ale takze obietnica dobrobytu i spokojnej starosci

### 2014-01-30 thesis/src/chapters/chapter-design.tex
Thesis: refinements in the last commit

@@
   In the other case, this component can be thought of a cloud exchange \cite{InterCloud10}, that brings together service producers and consumers.
+Autonomic cloud instance managers could be considered bid-makers and would compete with one another for the client.
+What is more, if the model is bid-based, the client can have implemented behaviour that would also allow them to participate in auctions.
+This could be done by introducing another component that would act in user's name.
 \end{asparaenum}

### 2014-01-30 thesis/src/chapters/chapter-design.tex
Thesis: container as a resource

@@
 
-\section{Application manager}
+\section{Autonomic application manager}
 
-\section{Application platform manager}
+\section{Autonomic application platform manager}
 %w kazdej sekcji opis MAPEK - co monitoruje, w jaki sposob analizuje i planuje, wykonuje
@@
 
-\begin{table}
+\begin{table}[!htbp]
   \centering
-  \begin{tabular}{ c  c  c  c }
+  \begin{tabularx}{\textwidth}[]{ X X X X }
     \specialrule{.1em}{.05em}{.05em} 
@@
     \specialrule{.1em}{.05em}{.05em} 
-    Features &  &  & \\ \hline
-    Examples &  &  & \\
-    \hline
-  \end{tabular}
+    \textbf{Features} & 
+-- complete simulation of machine's hardware
+
+-- full isolation
+
+-- host system is not shared among guests
+
+    &
+-- isolation is based on user-spaces
+
+-- shared kernel
+
+-- lesser overhead than full virutalization
+
+-- effective i/o operations
+    & 
+-- custom solution that uses processes, cgroups, SELinux, for example
+
+-- lesser overhead possible
+
+\\ \hline
+\textbf{Examples} &
+-- KVM
+
+-- Xen
+
+-- VMware &
+
+-- LXC
+
+-- OpenVZ
+&
+-- OpenShift containers
+\\ \hline
+  \end{tabularx}
   \caption{Containerisation technologies}
@@
 \subsubsection{Touchpoint}
-Container's touchpoint externalises hypervisor and operating system APIs. Similarly to all touchpoints, two main parts can be distinguished:
-\begin{itemize}
- \item \emph{sensors} - set of properties that expose system metrics such as CPU, memory, disk usage, network bandwidth - lista propertisow, charakterystyka (czy wymagane, co daja, dobra metryka?)
- \item \emph{effectors} - a collection of 'set' operations that aims to change container state: increase or decrease CPU or memory, for example.
-In fact, these operations directly corresponds to hypervisor capabilities.
-- lista wymagancych, + hypervisor, metryka
-\end{itemize}
-
-It is highly recommended for sensors and effectors to be linked, forming together manageability capabilities.
+Container's touchpoint externalises hypervisor and operating system APIs.
+Similarly to all touchpoints, sensors for gathering data and effectors for influencing it can be distinguished.
+It is highly recommended for them to be linked, forming together manageability capabilities.
 In other words, property (i.e.
 CPU) can be read only when it is possible to set it as well.
+For instance, free / used CPU, free / used memory, network bandwidth, disk usage can be measured.
+Cloud-SAP requires free CPU to be at least supported.
 

### 2014-01-31 thesis/src/chapters/chapter-design.tex
Thesis: add introduction to the monitor and analyse manager

@@
 This function takes either a query about the state of a service or a deployment plan of the given service and passes execution requests to the proper cloud instance autonomic managers.
+
 \subsection{Monitor and Analyse Autonomic Manager}
+\subsubsection*{Introduction}
+In many cases designers of applications cannot in advance foresee the detailed requirements of resources of their products, e.g.
+the number of virtual machines comprising it.
+This is especially true in case of \emph{start-ups} when it is not known whether the product will catch public attention and who exactly matches the target audience.
+In such circumstances the cloud provider should be able to quickly adapt to new conditions to prevent the service from being unable to work properly, e.g.
+take into account the geographical location of the application's users and deploy a service in close proximity to them so as to minimize the latency.
+
+As it was discussed in earlier chapters, the above-mentioned use case refers to \emph{scaling} capabilities of a cloud provider.
+However, even if a cloud provider has \emph{auto-scaling} mechanisms, there can be unacceptable delays in application response time because of the \emph{reactive} nature of the feature -- the system decides to scale the application based on the \emph{current} resource consumption and utilization rate.
+Since it involves instantiating and provisioning virtual machines there is always a time overhead associated with it.
+Hence, there is a need that a system could \emph{predict} demands and behaviours of a hosted service in a \emph{proactive} way.
+
+The goal of this component is to gather data about current performance of a deployed service and based on it perform an in-depth analysis which results in a resource usage prediction model and auto-scaling schemes.
+This is the objective of the \emph{monitor} and \emph{analyse} functions, respectively.
+
+\subsubsection*{Prediction models}
+\subsubsection{Monitor}
+\subsubsection{Analyse}
 

### 2014-01-31 thesis/src/chapters/chapter-design.tex
Thesis: monitoring of a container

@@
 \section{Autonomic container manager}
-It is expected that autonomic container manager supervise lifecycle of a container, modifying its properties accordingly to a given condition.
+It is expected that autonomic container manager:
+\begin{inparaenum}[1)]
+ \item supervises lifecycle of a container: provisions, monitors, migrates and destroys a container
+ \item modifies container its properties (i.e. increasing CPU, memory) accordingly to a given condition to ensure that service request are served with a sufficient Quality-of-Service
+\end{inparaenum}
 
@@
 While choosing containerisation technique is entirely left to an implementation, Cloud-SAP recommends using os-level virtualization or operating system processes as they involve lesser resources and are more flexible, especially in terms of dynamic adjustment.
-Moreover, lightweight containers have been proved more effective than full virtualization in some scenarios: \cite{RaHiSj13}.
+Moreover, lightweight containers have been proved more effective than full virtualization in some scenarios \cite{RaHiSj13}.
 
@@
 \emph{Autonomic container controller} aims to automate container's management function and externalise its configuration through its interfaces.
-In order to do achieve that, it implements a control loop that fully covers container life cycle: gathering information about it, analysing it, planning and executing actions on top of that.
-These four functions along with knowledge necessary to perform them designate modular structure of a controller as shown in figure \ref{img:autonomic-component}.
+In order to do achieve that, it implements a control loop that fully covers container life cycle.
+It is suggested for a controller to fully support self-management by implementing \emph{self-configuring, self-healing, self-optimising, self-healing} control loops, however, \emph{self-configuring} is the only one required.
+Modular structure of a controller is designated by its main four functions: data monitoring, data analysis, action planning and execution along with a knowledge necessary to perform these operations as shown in figure \ref{img:autonomic-component}.
 
@@
 \begin{asparaenum}
-  \item[\textbf{Metrics}] rozne opcje: cpu, memory, co wybrac przynajmniej cpu
+  \item[\textbf{Metrics}] There is a variety of metrics describing container state: CPU, memory, disk usage, swap, network bandwidth, to name a few.
+Significantly, there is no point in monitoring data that client is not interested in.
+In case when his policies relies entirely on CPU measurements, memory or bandwidth is irrelevant.
 
-  \item[\textbf{Data filters}] co robic z danymi, agregowac, jesli tak to jak, trzeba niwelowac zaburzenia (art? o monitoringu)
+  \item[\textbf{Data filters}] Provided that data is gathered from a container, not all of it needs to be analysed.
+Specifically, there may be some random peaks in a measurements, caused by a temporary distortions such as garbage collecting.
+Therefore, Cloud-SAP recommends to transform data using aggregation functions like average, median or apply low or high-pass filters.
 
-  \item[\textbf{Persistence}] czy przechowywac dane, jesli tak to jakie, potrzebne do analizy, wyciagania wnioskow (dane, reakcja -> odpowiedz)
+  \item[\textbf{Persistence}] Should data be used in knowledge creation and proactive approach, it is persisted, especially along with processes applied by an executor.
+As a consequence, future control loop iterations can be more effective by taking into account past symptoms and configuration adjustments.
+Decision whether to store data in databases, files, external location is entirely left to an implementation.
 
@@
 \subsubsection{Execution}
+Execution module is simple yet it plays a role that cannot be underestimated: carrying procedure scheduled beforehand during planning.
+To do so, it leverages effectors and API exposed by a container.
+In case of a self-learning system, change plan and its effects should be recorded to serve as a reference point in future.
 

### 2014-02-01 thesis/src/chapters/chapter-design.tex
Thesis: container monitoring, planning, execution

@@
 
-\section{Autonomic application platform manager}
-%w kazdej sekcji opis MAPEK - co monitoruje, w jaki sposob analizuje i planuje, wykonuje
+\section{Autonomic execution environment manager}
 
@@
 \subsubsection{Planning}
+During planning phase, procedure to enact desired state in resource is created.
+Depending on problem it may involve only one step on a complex workflow.
+Before implementing planning module, following issues should be taken into consideration:
+\begin{asparaenum}
+ \item[\textbf{Priorities}] One of key benefits of a containerisation is multi-tenancy.
+However, as a consequence, containers contends again each other to get access to resources.
+This issue can be mitigated by one of the following: \begin{inparaenum}[1)]
+    \item avoid overprovisioning, each tenant has guaranteed resource pool
+    \item prioritise tenant, each tenant has assigned priority that indicates its business value; that priority is used to favour one tenant at the expense of the others.
+ \end{inparaenum}
+
+ \item[\textbf{Reservations}] As previous point indicates, different containers may have different priorities.
+Therefore, it is suggested to guarantee that most valuable containers have their QoS requirements satisfied by resource reservation.
+Inspired by a computer networks, there are at least to ways to retain a certain resource level:
+\begin{inparaenum}[1)]
+    \item use parameters such as priority to favour given containers when it is not possible to content every single tenant \cite{nichols1998definition}
+    \item secure in advance to retain resource throughout container lifecycle, possibly by using common history patterns \cite{braden1997resource}
+ \end{inparaenum}
+\end{asparaenum}
 
 \subsubsection{Execution}
-Execution module is simple yet it plays a role that cannot be underestimated: carrying procedure scheduled beforehand during planning.
+Execution module role is simple yet it plays a role that cannot be underestimated: carrying procedure scheduled beforehand during planning.
 To do so, it leverages effectors and API exposed by a container.

### 2014-02-01 thesis/src/chapters/chapter-design.tex
Thesis: drawing update

@@
 Cloud Self Adaptive Platform (Cloud-SAP) is a reference architecture aspiring to supply application providers with an autonomic computing environment.
-Figure \ref{img:architecture-context} shows Cloud-SAP place in an exemplary application provisioning request.
+Figure \ref{img:csap-usage-context} shows Cloud-SAP place in an exemplary application provisioning request.
 \begin{figure}[!ht]
   \begin{center}
-    \includegraphics[width=0.9\textwidth]{chapter-design/architecture-context}
+    \includegraphics[width=0.9\textwidth]{chapter-design/csap-usage-context}
   \end{center}
   \caption{Cloud-SAP place in an exemplary application provisioning request}
-  \label{img:architecture-context}
+  \label{img:csap-usage-context}
 \end{figure}
@@
 In this section the further elaboration on the solution is presented.
-Figure \ref{img:architecture-context} showed the context which the architecture applies to.
+Figure \ref{img:csap-usage-context} showed the context which the architecture applies to.
 As one can see, Cloud-SAP can be seen as a proxy between the user and cloud instances and as a coordinator of the latter.
@@
 \emph{Sensor} and \emph{effector} behaviours refers to mechanisms that allows collecting data and change the behaviour of the resource, respectively.
+Structure of touchpoint is shown in figure \ref{img:csap-touchpoint}.
+
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics[width=0.7\textwidth]{chapter-design/csap-touchpoint}
+  \end{center}
+  \caption{Touch point component}
+  \label{img:csap-touchpoint}
+\end{figure}
 
@@
 What is more, an autonomic manager exposes \emph{sensor} and \emph{effector} manageability interfaces which enables other entities (including other autonomic managers) to manage it in a similar manner as the component itself manages a resource.
-The image of an autonomic manager with its control loop and provided interface is shown in figure \ref{img:autonomic-component}.
+The image of an autonomic manager with its control loop and provided interface is shown in figure \ref{img:csap-manager}.
 
@@
   \begin{center}
-    \includegraphics{chapter-design/autonomic-component}
+    \includegraphics[width=0.7\textwidth]{chapter-design/csap-manager}
   \end{center}
   \caption{Autonomic manager}
-  \label{img:autonomic-component}
+  \label{img:csap-manager}
 \end{figure}

### 2014-02-01 thesis/src/chapters/chapter-design.tex
Thesis: container analysis

@@
 \subsubsection{Analysis}
+The analysis module processes data received from monitoring component.
+It aims to evaluate it against various models and possibly determine if container require further intervention.
+The analyse function can employ variety of models, performing complex simulations or predictions.
+As \cite{LiWoZh05} proposes, it analysis can be based on:
+\begin{itemize}
+ \item Queue-based Performance Models
+ \item Dynamic models
+ \item Monotonic static models
+ \item Policy based models
+\end{itemize}
+Self-learning container manager should also use a predictive approach \cite{JiPeLiCh11} and track changes made in a container \cite{ZhYaWo05}.
+
+Cloud-SAP requires to at least support policy-based model such as one threshold-model \cite{tong1978threshold}.
 
 \subsubsection{Planning}
-During planning phase, procedure to enact desired state in resource is created.
-Depending on problem it may involve only one step on a complex workflow.
+During planning phase, procedure to enact desired state of a container is created.
+Depending on problem it may involve only one step such as increasing CPU or a more complex one like container migration.
 Before implementing planning module, following issues should be taken into consideration:
@@
 To do so, it leverages effectors and API exposed by a container.
+Collectively, these actions scales container vertically, for example:
+\begin{itemize}
+ \item increase / decrease CPU
+ \item increase / decrease memory
+ \item increase / decrease disk space
+ \item increase / decrease network bandwidth
+\end{itemize}
+Not every hypervisor supports dynamic vertical scaling, some requires a container to be restared. Noticeably, lightweight containers are flexible in th
+
 In case of a self-learning system, change plan and its effects should be recorded to serve as a reference point in future.
-
-\subsubsection{Knowledge}
 

### 2014-02-01 thesis/src/chapters/chapter-design.tex
Thesis: add prediction models to monitor and analyse manager

@@
 \subsubsection*{Prediction models}
+One way to divide prediction models into categories is to classify each one as a \emph{reactive} or a \emph{proactive} one.
+Since in the introductory section there were given reasons why the reactive approach is inefficient, here is a short overview of proactive approaches to the issue.
+
+Using pattern-matching algorithms is one method to tackle the problem and this approach was presented in \cite{caron2010forecasting}.
+In their work, the algorithm analyses past usage data of an application and tries to identify the most similar patterns to current resource usage characteristics by using a pattern-matching algorithm.
+Then the obtained patterns are interpolated and the final prediction model is evaluated.
+
+A performance prediction model and only its statistical evaluation is presented in \cite{islam2012empirical}.
+In that work, the usage of two machine-learning algorithms, \emph{Error Correction Neural Network} and \emph{Linear Regression}, is proposed.
+Additionally, the authors analyse the problem from an application's provider point of view contrary to most of the work in the field.
+
+An analysis regarding the effectiveness of usage a proactive approach rather than static allocation methods can be found in \cite{kupferman2009scaling}, where a set of scoring metrics is proposed as an efficiency evaluator.
 \subsubsection{Monitor}

### 2014-02-01 thesis/src/chapters/chapter-adaptivity.tex
Thesis: container analysis

@@
 
-While trivial in its form, cases of AWS, OpenShift, Carina, OneCloud proves it is useful in a real-world scenarios. Having that said, more sophisticated algorithms also exists:
-\begin{itemize}
-  \item \textit{integer programming} - auto scaling is reduced to server integer programming problems, which aims to minimize the cost or maximize the computing power with either computing power constraints or
-budget constraints \cite{MaLiHu10}
-  \item \textit{burst based padding} - employs a signal processing technique based on fast Fourier transform, burst pattern is extracted and used to calculate a padding value.
-Coefficients that represents the amplitude of each frequency component are used to calculate burst density.
-Depending of that value (i.e.
-is higher than 50\%) appropriate percentile of the burst values are used \cite{ShSuGuWi11}
-  \item \textit{remedial padding} - padding errors are being recorded and used in successive padding evaluations.
-In other words, let $e_1, e_2, ...
-e_k$ denote the recent prediction errors, next the weighted moving average is calculated.
-Actual applied padding is either padding itself or weighted average, depending which one is greater \cite{ShSuGuWi11}
-  \item \textit{linearised dynamic control} - linearised correction model is based on control equations \cite{AbShBh02}:
-
-    \begin{equation}
-      x(k+1)=Ax(k)+Bu(k)
-    \end{equation}
-
-    \begin{equation}
-       y(k) = C x(k) + D u(k) + E z(k)
-    \end{equation}
-    where x denotes the state variable vector and coefficient matrices: A B C D E are fitted to historical data as a regression model - \cite{DiLuFrHePa03}
-
-  \item \textit{markov decision process model} - computes optimal reaction to state changes by using observation on the system with assumptions about the rate of changes expected in the future \cite{AbWo02}
-\end{itemize}
+While trivial in its form, cases of AWS, OpenShift, Carina, OneCloud proves it is useful in a real-world scenarios.
 

### 2014-02-01 thesis/src/chapters/chapter-design.tex
Thesis: container analysis

@@
 The analyse function can employ variety of models, performing complex simulations or predictions.
-As \cite{LiWoZh05} proposes, it analysis can be based on:
+Analysis can be based on one or more of the followings models \cite{LiWoZh05}:
 \begin{itemize}
- \item Queue-based Performance Models
- \item Dynamic models
- \item Monotonic static models
- \item Policy based models
+ \item[\textbf{Queue-based performance models}] This models aims to predict QoS measures using resource states and environmental disturbances such as number of service requests.
+Then a configuration satisfying given QoS is found by a search technique using queues and layered queues.
+
+ \item[\textbf{Dynamic models}] Aims to represent QoS as a cost function of the difference between the measured value from sensors and set by effectors.
+Model aspires to minimize that function.
+Linearized dynamic control is an example of a dynamic model.
+
+ \item[\textbf{Monotonic static models}] Sought resource state is calculated by discretising a performance model or performing heuristic principles in performance engineering
+
+ \item[\textbf{Error correction}]
+
+ Analysis precision can be improved by applying paddings based on historical data:
+    \begin{itemize}
+      \item \textit{Burst based padding} - employs a signal processing technique based on fast Fourier transform, burst pattern is extracted and used to calculate a padding value.
+Coefficients that represents the amplitude of each frequency component are used to calculate burst density.
+Depending of that value (i.e.
+is higher than 50\%) appropriate percentile of the burst values are used \cite{ShSuGuWi11}
+      \item \textit{Remedial padding} - padding errors are being recorded and used in successive padding evaluations.
+In other words, let $e_1, e_2, ...
+e_k$ denote the recent prediction errors, next the weighted moving average is calculated.
+Actual applied padding is either padding itself or weighted average, depending which one is greater \cite{ShSuGuWi11}
+    \end{itemize}
+
+ \item[\textbf{Policy based models}] Policy is expressed as a set of constraints that characterise needed resource state. Policies can be categories into two types:
+    \begin{itemize}
+     \item \emph{Threshold control} - it was already discussed and illustrated in figure \ref{fig:threshold-model}.
+For example, CPU load can be monitored and when it exceeds a given value, CPU share is increased.
+     \item \emph{Policy function control} - a mutli-dimensional extension of a threshold model.
+It uses multiple variables and control levels \cite{abdeen2002seeking}.
+It can be further optimised by using Markovian model for a system behaviour \cite{ye2000markov}.
+    \end{itemize}
+
 \end{itemize}
 Self-learning container manager should also use a predictive approach \cite{JiPeLiCh11} and track changes made in a container \cite{ZhYaWo05}.
-
-Cloud-SAP requires to at least support policy-based model such as one threshold-model \cite{tong1978threshold}.
+Cloud-SAP requires to at least threshold-model \cite{tong1978threshold} to be supported to enable simple auto-scaling.
 

### 2014-02-01 thesis/src/chapters/chapter-design.tex
Thesis: complete the description of cloud federation manager

@@
 An analysis regarding the effectiveness of usage a proactive approach rather than static allocation methods can be found in \cite{kupferman2009scaling}, where a set of scoring metrics is proposed as an efficiency evaluator.
+
 \subsubsection{Monitor}
+The main goal of this function is to collect all data regarding usage, performance, resource utilization, capacity and availability of every cloud instance that forms a federation.
+Further elaboration on the function can be done by aligning use cases of this component:
+\begin{asparaenum}
+  \item[\textbf{Deployment of a new service}] When a new service is about to be deployed on a cloud federation, the system has to a priori aggregate information about resource capacity, utilization rates, geographical location and so on from cloud providers to be able to perform the best match of a cloud provider with the given service.
+  \item[\textbf{Scaling of a deployed service}] When a service which had been hosted on a cloud needs to be scaled, the similar knowledge as in the previous case is required to be passed to \emph{analyse} function to perform the perfect match of a cloud instance.
+\end{asparaenum}
+
+The knowledge that the component obtains is about current \emph{topology} (which cloud instances make the cloud federation, what is their geographical location, how are they related, etc.) and \emph{policies} (data required to be compared against policies of deployed serviced).
+What is more, for the system to be able to perform auto-scaling, the gathered information can be classified as \emph{problem determination knowledge} \cite{IBM06}.
+
+We do not strive to list the data which is required to be collected by this component in a system that tries to conform to the proposed architecture as it completely depends on the system needs.
+We want to give an overview of possibilities instead and leave the decision to the implementers.
+The same applies to the components and technology of a persistence layer.
+
 \subsubsection{Analyse}
+This function takes knowledge about the cloud federation and a service as an input and outputs a \emph{prediction plan}, which should be helpful in choosing the cloud instance for an application.
+
+In the introductory section a brief overview of some possible techniques and algorithms that can be used in the analyse component was presented.
+We do not impose a particular method on this component -- it is up to the system complexity and design, which solution would be more suitable to the given needs.
 

### 2014-02-01 thesis/src/chapters/chapter-design.tex
Thesis: drawings update

@@
   \begin{center}
-    \includegraphics{chapter-design/csap-resources-aggregation}
+    \includegraphics[width=\textwidth]{chapter-design/csap-resources-aggregation}
   \end{center}
@@
 \section{Autonomic cloud instance manager}
+The main responsibility of this manager is the management of a life cycle of resources of a given cloud provider.
+It could be thought of an autonomic entity which automates the process of managing resources of data centers that forms the services of a cloud instance.
+The cloud instance manager governs \emph{stacks} of the user-deployed services and its functions include
+\begin{inparaenum}[a)]
+\item monitoring the state of cloud provider's infrastructure,
+\item providing prediction models in terms of resource utilization levels and possible violations of Service Level Agreements,
+\item determining the best place (e.g. a physical host) the stack is to be deployed to,
+\item ordering a deployment of a stack and
+\item migration or eviction of the stacks that do not conform to Quality-of-Service policies provided by the user.
+\end{inparaenum}
+Additionally, it exposes the manageability interface for taking request from the orchestrating manager.
+
+\subsection{Managed resource}
+The resources of a cloud provider can be associated with a data center comprising physical nodes, i.e. clusters, hosts, etc.
+
+\subsubsection*{Touchpoint}
+At this level of abstraction, operating system mechanisms and utilities are used to gather information about the state of a given host, e.g.
+its memory consumption, network throughput and usage statistics, forming the \emph{sensor} behaviour.
+Additionally, they are means of performing any changes in the resource what makes them an \emph{effector}.
+We want to stress that these features are entirely os-specific and are out of scope of this thesis.
+For example, there are completely different ways to measure memory usage in Unix/Linux and Windows\texttrademark  environments.
+
+\subsubsection{Monitor}
+This function collects information about the data center of a cloud provider. The requirements for this component fall into two categories: resource discovery and information collection.
+
+\begin{asparaenum}
+\item[\textbf{Resource discovery}] When a new node is attached to the environment, the function should be able to quickly recognize it and takes it into account in the whole topology of resources comprising the data center.
+\item[\textbf{Information collection}] This function gathers and collects nodes' parameters, such as:
+  \begin{itemize}
+    \item memory (RAM) consumption rate and its attributes (e.g. frequency),
+    \item available disk space and other storage parameters (e.g. SSD/HDD),
+    \item network type, usage statistics and its attributes such as throughput,
+    \item CPU(s) attributes (the number of cores, etc.)
+  \end{itemize}
+  These parameters are passed to the next function for further analysis.
+\end{asparaenum}
+
+\subsubsection{Analyse}
+Predictions mechanisms depicted in the \emph{analyse} function of autonomic cloud federation manager applies here as well, so for the clarity of presentation we do not describe it again.
+
+\subsubsection{Plan}
+This function aims to solve the problem of an efficient mapping services to available resources so it can be thought of a scheduler of virtual machines within a data center.
+As in case of other elements of the proposed architecture we do not impose an implementation of a particular solution on an implementer, but we provide a quick overview of available mechanisms and let the final decision be driven by current needs.
+
+One of the approach to this problem is by implementing a Decentralized Online Clustering algorithm \cite{quiroz2009towards}.
+This solution is an aggregation of virtual machines' provisioning with a workload modeling technique called Quadrature Response Surface Model.
+It is based on an observation that requests made by a client can be inaccurate and may lead to over-provisioning.
+The model takes into account also application-specific QoS and SLA requirements.
+
+Constraint Programming can be another method that can be implemented \cite{nguyen2009autonomic} to tackle this problem.
+This is done by trying to optimize the global utility function of resource provisioning costs and application-specific requirements and constraints.
+In effect, the management of infrastructure is done in an automated fashion with the effort to reach the maximum performance of hosted applications, keeping the provider's cost at the minimum level at the same time.
+
+\subsubsection{Execute}
+The function takes as an input the deployment scheme which has mapped every stack of a service with the given host, irrespective of the use case that triggered the action, i.e.
+scaling or deployment.
+The request is delegated to the appropriate stack managers which are responsible for its further processing.
 

### 2014-02-02 thesis/src/chapters/chapter-design.tex
Thesis: add autonomic cloud instance manager
Thesis: added stack controller

@@
 \section{Autonomic stack manager}
+Autonomic stack manager is responsible for a management of a homogeneous resource - containers that together form a stack. More specifically, stack is subjected to:
+\begin{inparaenum}[a)]
+ \item monitoring its containers
+ \item horizontal scaling
+ \item changing applied load-balancing algorithms
+ \item component replication
+\end{inparaenum}
+
 \subsection{Managed resource}
+As it was stated in an overview, a stack is a group of containers configured to serve a common purpose (i.e.
+as a server cluster).
+In practise, this group consists of a master node and slaves: loadbalancer and java application workers or a master and slave databases, for example.
+Each stack type may leverage different mechanisms to enable that master-slave relation.
+Although specific decisions are left to an implementation, chapter X summarises common load-balancing mechanisms.
+
+\subsubsection{Touchpoint}
+Stack's touchpoint aims to provide information about the state and condition of a whole group of nodes: master as well as its slaves.
+As a consequence, from a structural perspective, it is decentralise and consists of a multiple agents.
+Key properties that should be manageable by a touch point include:
+\begin{itemize}
+ \item number of service requests passed to a master node
+ \item cluster's response time
+ \item average cpu / memory usage in cluster
+ \item packet loss from all nodes
+\end{itemize}
+
+\subsection{Autonomic controller}
+
+\subsubsection{Monitor}
+During monitoring phase, controller gathers data from a multiple agents located in stacks' nodes: master and slaves.
+It is important to ensure that received data is synchronised in time, consequently it can be analysed collectively to find a symptoms covering all of the nodes.
+
+\subsubsection{Analysis}
+Analysis phase can be based on one of the models mentioned in an overview. Beside this, following stack and loadbalancing specific models can be taken into consideration:
+\begin{asparaenum}
+ \item[\textbf{Discrete event simulation model}] It evaluates different balancing algorithms (e.g.
+round robin, least connection, etc ) to predict which produces least Load-Balancing-Metric (LBM) and packet loss \cite{idziorek2010discrete}.
+ \item[\textbf{Optimal static load balancing}] It aims to minimize mean response time of load distributed among servers \cite{tantawi1985optimal}.
+This model takes into account node's processing time as well as network delays - mean delay time is expressed as weighted sum of them.
+ \item[\textbf{Power usage minimisation}] This model is focused on saving computing power by removing computing node when its resource are not necessary.
+Key assumption is that there is a very little difference in cost of keeping node alive and keeping it 100\% busy.
+As a consequence nodes should be shut down whenever it is possible \cite{pinheiro2001load}.
+\end{asparaenum}
+
+\subsubsection{Planning}
+Plan module creates a sequence of changes enacting required QoS.
+Importantly, produced workflow should minimize or eliminate time during which stack becomes an unavailable due to incorporating additonal slaves or changing load-balancing mechanism, for instance.
+This can be done using redundant stack or its components to serve requests for a given period of time.
+
+\subsubsection{Execution}
+Stack autonomic controller executes actions such as:
+\begin{itemize}
+ \item horizontal scaling: adding slave nodes
+ \item changing load-balancing algorithm
+ \item changing network-topology to optimize time needed to reach a stack
+ \item replicating master node and hence eliminating SPOF
+\end{itemize}
 

### 2014-02-02 thesis/src/chapters/chapter-state-of-the-art.tex
Thesis: motivation for a cloudsap
Thesis: carina refactor

@@
-\chapter{State of the art}
+\chapter{Related work}
 
@@
 
-\subsection{Overview}
+\subsection{Introduction}
 Carina is an open source project, released under Apache License 2.0, built on top of OpenNebula, which aims to ``(\ldots) standardize the process for automating multi-VM deployments and setting auto-scaling and availability management policies in the cloud.'' \cite{CarinaBlog}.
 The project is used by the authors at their work at RIM in an OpenNebula-based private cloud.
+
+\subsection{Features}
 As it is stated in the requirements of the solution, \emph{Carina} should support variety of features which can be considered worth scrutinizing carefully as they are closely related to notions of adaptivity and scalability.
@@
 
-\subsection{Scaling}
-Chapter 3 introduced and discussed various types of application scaling.
-\emph{Carina} offers only one type of scaling and this is \emph{horizontal} scaling -- user's capabilities are limited only to add/remove a virtual machine to/from a given environment.
-Unfortunately there are no means to modify the parameters of a given or a set of virtual machines.
+\subsection{Key technical concepts}
 
-In this paper we strongly advocate spanning cloud resources across multiple vendors/providers.
-Despite having in its requirements such a position, \emph{Carina} \textbf{does not} support deployment or scaling of services across different providers.
-In the environment configuration we define only \textbf{one} endpoint -- this is can be thought of a reference to a cloud provider.
-As it can be only one per environment, it is not possible to deploy or scale some artifacts across many providers.
-
-\subsection{Adaptivity}
+\subsubsection{Adaptivity}
 In \emph{Carina} there are mechanisms which can perform application scaling, both in manual and automatic fashion.
@@
 Recreation of a virtual machine is the most fundamental and atomic operation under this model.
+
+\subsubsection{Scaling}
+Chapter 3 introduced and discussed various types of application scaling.
+\emph{Carina} offers only one type of scaling and this is \emph{horizontal} scaling -- user's capabilities are limited only to add/remove a virtual machine to/from a given environment.
+Unfortunately there are no means to modify the parameters of a given or a set of virtual machines.
+
+\subsubsection{Cloud federation awareness}
+In this paper we strongly advocate spanning cloud resources across multiple vendors/providers.
+Despite having in its requirements such a position, \emph{Carina} \textbf{does not} support deployment or scaling of services across different providers.
+In the environment configuration we define only \textbf{one} endpoint -- this is can be thought of a reference to a cloud provider.
+As it can be only one per environment, it is not possible to deploy or scale some artifacts across many providers.
 

### 2014-02-02 thesis/src/chapters/chapter-adaptivity.tex
Thesis: fixes after review

@@
-\chapter{Platform adaptivity}
+\chapter{Problem domain}
+\chapterintro{This chapter details mechanisms and practises that ensures a requested Quality-of-Service level is met by a means of a self-adaptive system that scales across multiple service providers.}
 
-\chapterintro{This chapter introduces concepts and mechanisms that enhance a platform with adaptivity capabilities achieved by a fusion of rules, policies and scaling techniques.}
+\section{Self-adaptive system}
 
-\section{Introduction}
+\subsection{Introduction}
 What makes a concept of an adaptivity enticing is crucial idea behind guaranteeing Quality-of-Serivce: automatization.
@@
 
-\section{Policies}
+\subsection{Policies}
 While auto-scaling is offered by a vast amount of cloud providers (e.g AWS, OpenShift, OpenNebula) it often lacks a sophisticated mechanisms allowing for specific scaling policies, being limited to only one predefined rule as it is in case of OpenShift for example.
@@
 
-\section{Data analysis}
+\subsection{Data analysis}
 Having policies defined, their conditions are evaluated against data acquired from sensors.
@@
 
-\section{Triggered actions}
+\subsection{Triggered actions}
 Actions that are being triggered by a elasticity controller are focused on application scaling. Previous chapter described that problem in detail.

### 2014-02-02 thesis/src/chapters/chapter-interoperability.tex
Thesis: fixes after review

@@
-\chapter{Interoperability of clouds}
+\newpage
+\section{Interoperability of clouds}
 
-\chapterintro{This chapter introduces the notion of a hybrid cloud and explains its role in IT industry.
-On top of this deployment model, the concept of InterCloud is presented and elaborated with the emphasis on its application in ensuring scalability of users' services.}
-
-\section{Introduction}
+\subsection{Introduction}
 From the perspective of a user of PaaS services it is vital that they are able to deploy seamlessly their applications using libraries, tools and services supported by the cloud provider\cite{MeGr11}.
@@
 
-\section{Hybrid cloud}
+\subsection{Hybrid cloud}
 One can imagine scenarios in which customers of cloud services know their applications are vulnerable to sudden variations in demand and their responsiveness must be kept at the same level all the time.
@@
 
-\subsection{Deployment models}
+\subsubsection{Deployment models}
 The informal introduction to the concept of a \emph{hybrid cloud} in the previous section requires a strict definition, but it is virtually impossible without defining other deployment models:
@@
 
-\subsection{Current usage and trends}
-\subsubsection*{Cloud -- clients' view}
-Before digging into the details of current usage and popularity of the hybrid model, it is worth discussing the general attitude of clients towards cloud computing.
+\subsubsection{Current usage and trends}
+\begin{asparaenum}
+ \item[\textbf{Cloud -- clients' view}] Before digging into the details of current usage and popularity of the hybrid model, it is worth discussing the general attitude of clients towards cloud computing.
 As the recent survey \cite{NBSurvey13} shows, the major factor that prevents companies from adopting cloud solutions is their concern over security -- in 2012 as much as 52\% responders considered it as a main concern with the regard to cloud in general.
@@
 \end{figure}
-
 The same survey shows that the cloud adoption growth rate is high -- 75 percent of responders stated usage of some sort of a cloud platform.
@@
 
-\subsubsection*{View on hybrid cloud}
-When it comes to the application of a hybrid model in industry, in most cases the definition introduced in the previous chapter now becomes a 'public-private' composition.
+ \item[\textbf{View on hybrid cloud}] When it comes to the application of a hybrid model in industry, in most cases the definition introduced in the previous chapter now becomes a 'public-private' composition.
 And this is how the term should be understood when discussing the results of the surveys which aimed to provide insights onto the view on a hybrid cloud from the customers' perspective.
@@
 
-\section{Federation of clouds -- InterCloud}
+\end{asparaenum}
+
+\subsection{Federation of clouds -- InterCloud}
 As stated in the previous sections, one of the major obstacles that prevents consumers from adopting cloud solutions is reliability.
@@
 
-\subsection{Usage in industry}
+\subsubsection{Usage in industry}
 The depicted model has not yet been adopted in the industry, yet some simulations were carried out on a \emph{CloudSim} platform and the obtained results showed that this concept has ``immense potential'' \cite{InterCloud10}.

### 2014-02-02 thesis/src/chapters/chapter-introduction.tex
Thesis: fixes after review

@@
 For example it is nearly impossible to create an autoscaling cloud federation with Amazon Web Services (current leader in providing cloud services\cite{GartnerMagicQuadrantSep2013}) and another provider.
-What is more, Amazon AWS users cannot use more lightweight virtualization methods, such as linux containers.
-Nevertheless, when compared to other companies especially in terms of autoscaling capabilities, Amazon really shines.
+Amazon EC2, when compared to other companies especially in terms of autoscaling capabilities, really shines.
 OpenShift, RedHat PaaS solution, ensures application scaling but with very limited possibilities of customisation of the process -- the user can only choose if their application should scale and the whole algorithm is solely based on the number of concurrent requests to the application.

### 2014-02-02 thesis/src/chapters/chapter-scalability.tex
Thesis: fixes after review

@@
-\chapter{Scaling applications}
+\newpage
+\section{Scaling applications}
 
-\chapterintro{This chapter is devoted to the concept of scaling users' application from the perspective of a cloud platform provider.
-To achieve that, it presents attainments of research groups working in that area as well as it considers mechanisms used in products currently available on the market.}
-
-\section{Introduction}
+\subsection{Introduction}
 
@@
 
-\section{Horizontal scaling}
+\subsection{Horizontal scaling}
 As outlined in previous section, horizontal scaling is about adding supplementary nodes to a system.
@@
 
-\subsection{Load-balancing algorithms}
+\subsubsection{Load-balancing algorithms}
 Generally, there are two types of load-balancers: hardware and software based.
@@
 
-\subsection{Load-balancing scalability}
+\subsubsection{Load-balancing scalability}
 Although, it may seem that balancing workloads eliminates problem of a single point of failure (SPOF) among different servers, it is in fact shifted to load-balancing layer.
@@
 
-\subsection{Load-balancer comparision}
+\subsubsection{Load-balancer comparision}
 
@@
 
-HAProxy & 
+HAProxy \cite{HAProxy} & 
 -- a single-process, event-driven model reduces the cost of context switch and the memory usage
@@
 
-BIG-IP Local Traffic Manager &
+BIG-IP Local Traffic Manager \cite{LTM} &
 -- managing at application services level rather than at individual devices and objects
@@
 
-Apache HTTPD & 
+Apache HTTPD \cite{ApacheHTTPD} & 
 -- support for session stickiness by using cookies and URL encoding.
@@
 
+Zeus Load Balancer \cite{Zeus} &
+
+-- Layer-7 traffic management
+
+-- HTTP content compression
+
+-- connection concurrency control
+
+-- server health monitoring
+
+& 
+-- content-aware algorithms
+\\ \hline
+
+Zeus Global Load Balancer \cite{ZeusGlobal} &
+
+-- service health monitoring
+
+-- active - passive failover
+
+-- balancing is based on observed performance
+
+& 
+-- geographic proximity
+
+-- adaptive load balancing (based on proximity and load)
+
+\\ \hline
+
 \end{tabularx}
@@
 \newpage
-\section{Vertical scaling}
+\subsection{Vertical scaling}
 Essentially, vertical scaling is concentrated upon increasing capacity of single node.
@@
 
-\subsection{Virtual machine resizing}
+\subsubsection{Virtual machine resizing}
 
@@
 
-\subsection{Virtual machine replacement}
+\subsubsection{Virtual machine replacement}
 As it was highlighted in previous section, reasoning behind virtual machine replacement is that, in case when dynamic resizing is not possible, a new virtual machine with a desired configuration can be provisioned and replace the old one.

### 2014-02-02 thesis/src/chapters/chapter-state-of-the-art.tex
Thesis: fixes after review

@@
 \section{Providers comparison}
-In this section we would like to compare all depicted providers as well as other solution in terms of \emph{adaptivity} and \emph{scalability}.
-
+In this section we would like to compare all depicted providers as well as other solution in terms of \emph{adaptivity} and \emph{scalability} as none of the examined providers support \emph{cloud federated} deployments.
 Table \ref{tab:cloud-providers-scaling} presents a summary of cloud providers auto-scaling capabilities.
 Interestingly, all of them are focused solely on horizontal scaling, ignoring advantages offered by a fine-grained approach to scaling that leverage scaling up and application tuning.
+Table \ref{tab:cloud-providers-adaptivity} summarises platforms approach to adaptivity: their data analysis mechanisms and policies support.
 
@@
 \end{table}
-On the other hand, table \ref{tab:cloud-providers-adaptivity} is a summary with approaches to adaptivity taken by different cloud providers.
 

### 2014-02-06 thesis/src/chapters/chapter-design.tex
Thesis: summary of cloud-sap and container impl

@@
 \begin{itemize}
-  \item \emph{Application} -- application itself can be considered resources which can be managed by an external entity, e.g. by restarting/killing it,
+  \item \emph{Application} - application itself can be considered resources which can be managed by an external entity, e.g.
+by restarting/killing it.
+Interestingly, an application itself can be an autonomic component with an embedded control loop.
+However, such setting is completely transparent to Cloud-SAP and therefore cannot be managed by it.
   \item \emph{Execution environment of an application} forms an inherent part of an application lifecycle and its proper tuning can greatly influence its performance,
@@
 Each block implements different behaviour based on the abstraction level and resource they operate on (see figure \ref{img:csap-layered-structure}).
-Nevertheless it is possible to define their generic behaviour, which embraces the following phases that correspond to a given block:
+Nevertheless it is possible to define their generic behaviour, which embraces the four phases: monitoring, analysis, planning, execution.
+
+\subsubsection{Monitoring}
+This function of autonomic manager is responsible for collecting and aggregating data about the resource it manages.
+While Cloud-SAP does not have any specific requirements regarding underlying monitoring mechanism, there are a few aspects that should be taken into consideration:
+\begin{asparaenum}
+  \item[\textbf{Metrics}] Whilst there is a variety of metrics describing resource state, there is no point in monitoring data that client is not interested in.
+For example, in case when client-defined policies rely entirely on CPU measurements, memory or bandwidth is irrelevant.
+
+  \item[\textbf{Data filters}] Provided that data is gathered from a resource, not all of it should be analysed as some portion of it may be distorted.
+More specifically, there may be some random spikes in measurements, caused by a temporary process such as garbage collecting.
+Therefore, Cloud-SAP recommends to transform data using aggregation functions like average, median or filter it by apply low or high-pass filters.
+
+  \item[\textbf{Persistence}] Should collected data be used as a reference point during future control loop iterations, it is persisted, especially along with measures applied by an executor.
+As a consequence, successive assessments can be more effective by taking into account past symptoms and configuration adjustments.
+Decision whether to store data in databases, files, external location is entirely left to an implementation.
+
+  \item[\textbf{Standard compatibility}] Whilst there is no single, industry accepted standard of monitoring above-mentioned resources ranging from a single application to whole cloud instance, there were initiatives such as OCCI \cite{OCCI} that aims to close that gap and provide unified view of common appliances.
+Having scaling across multiple cloud instances in mind, it is vital to provide compatibility at a monitoring level.
+However, in some cases, standard-defined API may no be sufficient for a specific needs.
+\end{asparaenum}
+
+\subsubsection{Analysis}
+This part takes the data from \emph{monitor} phase and performs in-depth analysis of it, which can result in, e.g.
+prediction plans.
+It is possible to incorporate more complex mechanisms, such as machine-learning, to fully use the obtained information for better prediction models.
+For example one or more of the following models can be leveraged \cite{LiWoZh05}:
+\begin{asparaenum}
+ \item[\textbf{Queue-based performance models}] This models aims to predict QoS measures using resource states and environmental disturbances such as number of service requests.
+Then a configuration satisfying given QoS is found by a search technique using queues and layered queues.
+ \item[\textbf{Dynamic models}] Aims to represent QoS as a cost function of the difference between the measured value from sensors and set by effectors.
+Model aspires to minimize that function.
+Linearized dynamic control is an example of a dynamic model.
+ \item[\textbf{Monotonic static models}] Sought resource state is calculated by discretising a performance model or performing heuristic principles in performance engineering. 
+ \item[\textbf{Error correction}] Analysis precision can be improved by applying paddings based on historical data:
     \begin{itemize}
-  \item \emph{monitor} -- this function of autonomic manager is responsible for collecting and aggregating data about the resource it manages,
-  \item \emph{analyse} -- this part takes the data from \emph{monitor} phase and perform in-depth analysis of it, which can result in, e.g.
-prediction plans.
-It is possible to incorporate more complex mechanisms, such as machine-learning, to fully use the obtained information for better prediction models,
-  \item \emph{plan} -- this block is responsible for planning the appropriate action based on the output of the previous block and policies,
-  \item \emph{execute} -- the goal of this function of autonomic manager is to execute the planned action
+      \item \textit{Burst based padding} - employs a signal processing technique based on fast Fourier transform, burst pattern is extracted and used to calculate a padding value.
+Coefficients that represents the amplitude of each frequency component are used to calculate burst density.
+Depending of that value (i.e.
+is higher than 50\%) appropriate percentile of the burst values are used \cite{ShSuGuWi11}
+      \item \textit{Remedial padding} - padding errors are being recorded and used in successive padding evaluations.
+In other words, let $e_1, e_2, ...
+e_k$ denote the recent prediction errors, next the weighted moving average is calculated.
+Actual applied padding is either padding itself or weighted average, depending which one is greater \cite{ShSuGuWi11}
     \end{itemize}
+ \item[\textbf{Policy based models}] Policy is expressed as a set of constraints that characterise needed resource state. Policies can be categories into two types:
+    \begin{itemize}
+     \item \emph{Threshold control} - it was already discussed and illustrated in figure \ref{fig:threshold-model}.
+For example, CPU load can be monitored and when it exceeds a given value, CPU share is increased.
+     \item \emph{Policy function control} - a mutli-dimensional extension of a threshold model.
+It uses multiple variables and control levels \cite{abdeen2002seeking}.
+It can be further optimised by using Markovian model for a system behaviour \cite{ye2000markov}.
+    \end{itemize}
+\end{asparaenum}
+Self-learning manager should also use a predictive approach \cite{JiPeLiCh11} and track changes made in a container \cite{ZhYaWo05}.
+Cloud-SAP requires to at least threshold-model \cite{tong1978threshold} to be supported to enable simple auto-scaling scenario.
+
+\subsubsection{Planning}
+During planning phase, procedure to enact desired state of a resource is created.
+Depending on problem, it may involve only one step such as adjusting one parameter or a more complex like migrating a resource.
+Before implementing planning module, following issues should be taken into consideration:
+\begin{asparaenum}
+ \item[\textbf{Priorities}] Multi-tenancy is indispensably correlated with cloud services.
+As a consequence, resource contends again each other to access to lower-level resources.
+This can be mitigated by one of the following: \begin{inparaenum}[1)]
+    \item avoiding over-provisioning, each tenant has guaranteed resource pool
+    \item prioritise tenants, each tenant has assigned priority that indicates its business value; that priority is used to favour one tenant at the expense of the others.
+ \end{inparaenum}
+
+ \item[\textbf{Reservations}] As previous point indicates, different resources may have different priorities.
+Therefore, it is suggested to guarantee that most valuable have their QoS requirements satisfied by resource reservation.
+Inspired by a computer networks, there are at least to ways to retain a certain resource level:
+\begin{inparaenum}[1)]
+    \item use parameters such as priority to favour given resource when it is not possible to content every single tenant \cite{nichols1998definition}
+    \item secure in advance to retain resource throughout container lifecycle, possibly by using common history patterns \cite{braden1997resource}
+ \end{inparaenum}
+\end{asparaenum}
+
+\subsubsection{Execution}
+The goal of this function of autonomic manager is to execute the planned action
+
 What is more, an autonomic manager exposes \emph{sensor} and \emph{effector} manageability interfaces which enables other entities (including other autonomic managers) to manage it in a similar manner as the component itself manages a resource.
@@
 
-\section{Autonomic application manager}
+\section{Autonomic execution environment manager}
+Autonomic execution environment manager supervises lifecycle of a deployed application's environment by tuning it appropriately to a given usage. More specifically it :
+\begin{inparaenum}[1)]
+ \item monitors application execution environment i.e. application server
+ \item optimises resources used by an application such as thread pool, cache size, databases connections number
+ \item restarts application when necessary
+\end{inparaenum}
 
-\section{Autonomic execution environment manager}
+\subsection{Managed resource}
+Application execution environment is a set of tools, APIs and frameworks that provides a generalized approach to creating and running an application.
+It is strictly application specific.
+In case of java, for example, it can be represented by an application server such as Oracle Weblogic or IBM WebSphere.
+
+\subsubsection{Touchpoint}
+Due to the fact that execution environment is so closely related to a given application type, it is not possible to enlist properties that may be exposed by a touchpoint becouse they vary from application to application.
+In case of Java EE application touchpoint would externalise management of thread pools, cache and JVM settings such as heap size.
+
+\subsection{Autonomic controller}
+Although autonomic controller manages a single resource - an execution environment - it in fact conducts a variety of homogeneous resources - each application type has different properties and consequently vary in actions that can be taken to optimise them.
+Nevertheless, following subsections presents some common issues and concerns.
+
+\subsubsection{Monitoring}
+During monitoring phase, controller collects data from execution environment's touchpoint.
+Depending on application type it can involve different protocols and APIs.
+Should one monitor application execution environment, issues and concerns specified in overview are taken into consideration.
+
+\subsubsection{Analysis}
+Applying mathematical models to optimise application environment is an exceptionally difficult challenge due to non-linearity of a function denoting dependencies between various parameters.
+To make matter worse, space of possible solution is unknown and optimal solution heavily depends on a workflow that is tested against.
+With all that said, models and theories enumerated in an overview can be employed during analysis process.
+Below is presented a group of such attempts:
+\begin{asparaenum}
+  \item[\textbf{Smart hill-climbing}] As proposed in \cite{xi2004smart}, application server tuning can be seen as a black-box optimisation problem that can be solved using smart-hill climbing algorithm based on ideas of importance sampling and gradient-based optimisation.
+
+  \item[\textbf{Active Harmony}] This research project uses a simplex method to find the optimal application server parameter settings.
+Algorithm was validated against a TPC-W benchmark and was proven to optimise application response time by 16\% \cite{chung2004automated}.
+\end{asparaenum}
+
+\subsubsection{Planning}
+Planning module creates a workflow enacting application environment.
+It should take into account that different environment requires different procedures and some actions may involve additional tasks such as server restart to be introduced.
+Besides, some actions interfere with each other and their execution should be planned cautiously.
+For example changing cache size and cache invalidation interfere with each other and consequently applying this change at same time may not be a good idea.
+
+\subsubsection{Execution}
+Finally, controller applies a change plan that tunes the execution environment involving operations such as:
+\begin{itemize}
+ \item adjusting thread pool
+ \item adjusting buffer size
+ \item adjusting cache size and its parameters
+ \item specifying maximal simultaneous connections
+ \item restarting application server
+\end{itemize}
 
@@
 \subsubsection{Monitoring}
-As it was already mentioned, monitoring module aims to gather data from container's sensors.
-While Cloud-SAP does not have any specific requirements regarding underlying monitoring mechanism, there are a few aspects that should be taken into consideration:
-\begin{asparaenum}
-  \item[\textbf{Metrics}] There is a variety of metrics describing container state: CPU, memory, disk usage, swap, network bandwidth, to name a few.
-Significantly, there is no point in monitoring data that client is not interested in.
-In case when his policies relies entirely on CPU measurements, memory or bandwidth is irrelevant.
-
-  \item[\textbf{Data filters}] Provided that data is gathered from a container, not all of it needs to be analysed.
-Specifically, there may be some random peaks in a measurements, caused by a temporary distortions such as garbage collecting.
-Therefore, Cloud-SAP recommends to transform data using aggregation functions like average, median or apply low or high-pass filters.
-
-  \item[\textbf{Persistence}] Should data be used in knowledge creation and proactive approach, it is persisted, especially along with processes applied by an executor.
-As a consequence, future control loop iterations can be more effective by taking into account past symptoms and configuration adjustments.
-Decision whether to store data in databases, files, external location is entirely left to an implementation.
-
-  \item[\textbf{Standard compatibility}] Whilst there is no single, industry accepted standard of virtual machine monitoring, there were initiatives such as OCCI \cite{OCCI} that aims to close that gap.
-Having scaling across multiple cloud instances in mind, it is vital to provide compatibility at a container monitoring level.
-However, in some cases, standard-defined API may no be sufficient for a specific needs
-\end{asparaenum}
+During monitoring phase, controller collects data from container's touchpoint. Issues such as data filtering, aggregation should be taken into consideration as it was stated in overview.
 
 \subsubsection{Analysis}
-The analysis module processes data received from monitoring component.
-It aims to evaluate it against various models and possibly determine if container require further intervention.
-The analyse function can employ variety of models, performing complex simulations or predictions.
-Analysis can be based on one or more of the followings models \cite{LiWoZh05}:
-\begin{itemize}
- \item[\textbf{Queue-based performance models}] This models aims to predict QoS measures using resource states and environmental disturbances such as number of service requests.
-Then a configuration satisfying given QoS is found by a search technique using queues and layered queues.
-
- \item[\textbf{Dynamic models}] Aims to represent QoS as a cost function of the difference between the measured value from sensors and set by effectors.
-Model aspires to minimize that function.
-Linearized dynamic control is an example of a dynamic model.
-
- \item[\textbf{Monotonic static models}] Sought resource state is calculated by discretising a performance model or performing heuristic principles in performance engineering
-
- \item[\textbf{Error correction}]
-
- Analysis precision can be improved by applying paddings based on historical data:
-    \begin{itemize}
-      \item \textit{Burst based padding} - employs a signal processing technique based on fast Fourier transform, burst pattern is extracted and used to calculate a padding value.
-Coefficients that represents the amplitude of each frequency component are used to calculate burst density.
-Depending of that value (i.e.
-is higher than 50\%) appropriate percentile of the burst values are used \cite{ShSuGuWi11}
-      \item \textit{Remedial padding} - padding errors are being recorded and used in successive padding evaluations.
-In other words, let $e_1, e_2, ...
-e_k$ denote the recent prediction errors, next the weighted moving average is calculated.
-Actual applied padding is either padding itself or weighted average, depending which one is greater \cite{ShSuGuWi11}
-    \end{itemize}
-
- \item[\textbf{Policy based models}] Policy is expressed as a set of constraints that characterise needed resource state. Policies can be categories into two types:
-    \begin{itemize}
-     \item \emph{Threshold control} - it was already discussed and illustrated in figure \ref{fig:threshold-model}.
-For example, CPU load can be monitored and when it exceeds a given value, CPU share is increased.
-     \item \emph{Policy function control} - a mutli-dimensional extension of a threshold model.
-It uses multiple variables and control levels \cite{abdeen2002seeking}.
-It can be further optimised by using Markovian model for a system behaviour \cite{ye2000markov}.
-    \end{itemize}
-
-\end{itemize}
-Self-learning container manager should also use a predictive approach \cite{JiPeLiCh11} and track changes made in a container \cite{ZhYaWo05}.
-Cloud-SAP requires to at least threshold-model \cite{tong1978threshold} to be supported to enable simple auto-scaling.
+Analysis should incorporate prediction techniques and mathematical models specified in overview, producing change requests optimising container behaviour.
 
 \subsubsection{Planning}
-During planning phase, procedure to enact desired state of a container is created.
-Depending on problem it may involve only one step such as increasing CPU or a more complex one like container migration.
-Before implementing planning module, following issues should be taken into consideration:
-\begin{asparaenum}
- \item[\textbf{Priorities}] One of key benefits of a containerisation is multi-tenancy.
-However, as a consequence, containers contends again each other to get access to resources.
-This issue can be mitigated by one of the following: \begin{inparaenum}[1)]
-    \item avoid overprovisioning, each tenant has guaranteed resource pool
-    \item prioritise tenant, each tenant has assigned priority that indicates its business value; that priority is used to favour one tenant at the expense of the others.
- \end{inparaenum}
-
- \item[\textbf{Reservations}] As previous point indicates, different containers may have different priorities.
-Therefore, it is suggested to guarantee that most valuable containers have their QoS requirements satisfied by resource reservation.
-Inspired by a computer networks, there are at least to ways to retain a certain resource level:
-\begin{inparaenum}[1)]
-    \item use parameters such as priority to favour given containers when it is not possible to content every single tenant \cite{nichols1998definition}
-    \item secure in advance to retain resource throughout container lifecycle, possibly by using common history patterns \cite{braden1997resource}
- \end{inparaenum}
-\end{asparaenum}
+This part schedules change requests submitted by an analysis model producing change procedures.
+Depending on implementation, it may use reservation and priorities mechanisms to enforce given Quality-of-Service level.
+It vital for a planning phase to be well suited for an underlying hypervisor since not all its types supports dynamic vertical scaling some requires container to be restarted.
+In that case, container should be cloned and then resized, restarted and finally substitute old, non-scaled instance.
 
@@
 \section{Summary}
-%podsumowanie, wyzwania stojace przed projektem, problemy ale takze obietnica dobrobytu i spokojnej starosci
+Cloud-SAP fills the gap in existing platforms, providing a standardised approach to scaling applications that uses a fine-grained actions, operating on different platform levels: starting from execution environment to end on a cloud federation.
+It is based on a concept of an autonomic system, providing methodical yet flexible foundation for designed system.
+Cloud-SAP leverages core benefits of such system: promise of a self-configuring, self-optimising, self-healing, self-protecting system driven by a customer-defined policies.
+
+% TODO Further work?, wady

### 2014-02-10 thesis/src/chapters/chapter-adaptivity.tex
Thesis: adaptivity section has been rewritten

@@
 
-\section{Self-adaptive system}
+\section{Self-manageable system}
 
 \subsection{Introduction}
-What makes a concept of an adaptivity enticing is crucial idea behind guaranteeing Quality-of-Serivce: automatization.
-Generally, ensuring and enforcing given quality can be seen as continuous monitoring and reacting to events when necessary.
-Such idea can be expressed as IT process shown in figure \ref{fig:it-process}.
+Guaranteeing Quality-of-Service, a critical concept in this paper, can be seen as continuous monitoring and reacting to undesirable conditions when necessary.
+In fact, this process is as an IT procedure and traditionally belongs to the duties of a system administrator.
+Figure \ref{fig:it-process} illustrates exemplary IT procedure.
 
@@
 
-While system administrator can be directly responsible for manually performing above-mentioned steps, it is much more favourable to automatize this process leading to a system self-management or an adaptable system.
-Not only does it lead to efficiency of an IT process but also to its effectiveness \cite{IBM06}.
-It is possible through:
+While system administrator can be directly responsible for manually performing steps depicted in figure \ref{fig:it-process}, it is much more favourable to automatize this process leading to a self-manageable system.
+Not only does it lead to efficiency of an IT process but also to its effectiveness \cite{IBM06} achieved through:
 \begin{itemize}
@@
 
-In a context of delivering a computing platform, adaptivity adds auto-scaling features to a solution offered by Platform-as-a-Service provider.
-Self-managing is possible through usage of an Elasticity Controller which gathers probes from resource such as application container and uses that knowledge to execute appropriate action on that resource, indirectly modifying consecutive probes \cite{VaRoBu11}.
-This concept illustrates diagram \ref{fig:elasticity-controller}.
+As  \cite{murch2004autonomic} explains, there are four levels of a system manageability maturity. The higher the level, the better it manages itself leading to a adaptive or autonomic system at best:
+\begin{itemize}
+ \item \emph{Basic} - IT professionals manage each system element. Activities related to configuration, optimisation, healing, and protection of IT infrastructure components are performed manually.
+ \item \emph{Managed} - System management technologies can be used to collect information from different systems and analyze this information.
+In this scenario, IT professionals do most of the analysis, but it is a starting point for the automation of IT tasks.
+ \item \emph{Predictive} - Individual components monitor themselves, analyze changes and offer advice. Therefore, dependency on people is reduced and decision-making is improved.
+ \item \emph{Adaptive} - IT components can, individually or group-wise, monitor, analyze operations and offer advice with minimal human intervention.
+ \item \emph{Autonomic} - System operations are managed by business policies established by the administrator.
+In fact, business policy drives overall IT management, while at the adaptive level; there is an interaction between humans and the system.
+\end{itemize}
+
+In heart of the every self-adaptive systems, according to \cite{brun2009engineering}, lies a feedback loop, concept that originates from a control engineering. As paper's authors advocate:
+\begin{quote}
+Feedback loops provide the generic mechanism for self-adaptation.
+Positive feedback occurs when an initial change in a system is reinforced, which leads toward an amplification of the change.
+In contrast, negative feedback triggers a response that counteracts a perturbation.
+\end{quote}
+
+What control loop is made of is collection of data, its analysis, making a decisions and enforcing them in the environment.
+Figure \ref{fig:feedback-loop} depicts a generic feedback loop, giving context associated with each phase.
 
@@
   \begin{center}
-    \includegraphics{chapter-adaptivity/elasticity-controller}
+    \includegraphics[width=\textwidth]{chapter-adaptivity/feedback-loop}
   \end{center}
-  \caption{Elasticity controller}
-  \label{fig:elasticity-controller}
+  \caption{Feedback loop \cite{brun2009engineering}}
+  \label{fig:feedback-loop}
 \end{figure}
 
-The remaining of this chapter describes crucial elements that compose elasticity controller: policies, data analysis, triggered actions and presents a comparison of cloud providers.
+Successive sections details aforementioned phases and gives an examples of systems leveraging feedback loops.
 
-\subsection{Policies}
-While auto-scaling is offered by a vast amount of cloud providers (e.g AWS, OpenShift, OpenNebula) it often lacks a sophisticated mechanisms allowing for specific scaling policies, being limited to only one predefined rule as it is in case of OpenShift for example.
-
-Policy denotes a condition which, when satisfied, triggers an action that is supposed to harness cloud instance in a way that future evaluations of condition will be unsuccessful.
-Typically condition itself is accompanied by a minimal and maximal number of node instances, allowing for ensuring minimal QoS and controlling maximal costs.
-Currently, industry leaders supports two main kind of policies \cite{AmazonAutoScaling}:
+\subsection{Collecting data}
+During this phase, relevant data is collected from sensors.
+After that, data can be aggregated, filtered or stored to provide a reference point during future feedback loop evaluation.
+During this phase, there are a few concerns that should be investigated \cite{brun2009engineering}:
 \begin{itemize}
- \item \textit{expression based} - allows to define how you to scale application in response to changing conditions, which include factors such as memory, CPU usage, cost or some indirect, calculated metrics
- \item \textit{scheduled} - allows to scale an application in response to predictable load changes.
-For example, traffic increases during the weekends and decreases on working days.
-Hence, that predictable traffic patterns is used to scale application based on current time.
+ \item sample rate
+ \item probe reliability
+ \item data format
 \end{itemize}
 
-Technically, policies are expressed in some human-readable format such as JSON, XML as it is in case of AWS EC2 or custom expression used for example by Carina.
-Appendix \ref{app:scaling-policies} presents example configuration used by AWS E2 Auto-Scaling.
+Almost any cloud platform providers incorporates into its solution monitoring component, responsible for data collection.
+Mechanisms used vary from providers to provider and may involve mixture of protocols and standards.
+For example, Carina \cite{Carina} uses OpenNebula IM and VMM drivers relaying on a plaintext data, transferred through a SSH connection.
 
-\subsection{Data analysis}
-Having policies defined, their conditions are evaluated against data acquired from sensors.
-In a simplest case this evaluation can be based on a Threshold Model \cite{LiWoZh05}, which defines a valid range.
-In cases when given metric violates that condition (i.e.
+\subsection{Analysis}
+Analysis part applies mathematical models to reason about data gathered during previous phase. It can leverage variety of approaches and be driven by a customer-defined policies.
+
+As \cite{brun2009engineering} specifies, it is important to take into account following concerns during analysis phase:
+\begin{itemize}
+ \item applying historical data and existing patterns to current situation
+ \item archiving data
+ \item applying adequate model to current situation
+ \item model's stability
+\end{itemize}
+
+Currently, service providers employs only simplified models such as threshold model \cite{LiWoZh05}, that defines a valid range for a given metric.
+In cases when it is violated (i.e.
 value is either smaller than minimal or bigger than maximal acceptable) corresponding resource is properly adjusted - figure \ref{fig:threshold-model} illustrates that idea.
@@
 
-While trivial in its form, cases of AWS, OpenShift, Carina, OneCloud proves it is useful in a real-world scenarios.
+Threshold model, while trivial in its form, has been proved useful in a real-world scenarios as it is in case of AWS, OpenShift, Carina or OneCloud.
 
-\subsection{Triggered actions}
-Actions that are being triggered by a elasticity controller are focused on application scaling. Previous chapter described that problem in detail.
+As it was already mentioned, models are often controlled by a policies that denotes a condition which, when satisfied, triggers an action, harnessing environment's disturbance.
+Currently, industry leaders supports \cite{AmazonAutoScaling} two kind of policies:
+\begin{itemize}
+ \item \textit{expression based} - allows to define how to scale application in response to changing conditions, which include factors such as memory, CPU usage, cost or some indirect, calculated metrics
+ \item \textit{scheduled} - allows to scale an application in response to predictable load changes.
+For example, traffic increases during the weekends and decreases on working days.
+Hence, that predictable traffic patterns is used to scale application based on current time.
+\end{itemize}
+
+Technically, policies are expressed in a human-readable format such as JSON, XML as it is in case of AWS EC2 or custom expression used for example by Carina.
+Appendix \ref{app:scaling-policies} presents XML format adopted by AWS E2 Auto-Scaling.
+
+\subsection{Decision arrangement}
+Before implementing changes in an environment, decision what resources have to be tuned in order to reach a desirable state is made. Decision should be made on the basis of:
+\begin{itemize}
+ \item risk analysis - analysing how change effects whole environment
+ \item prioritising identified problems and resources
+\end{itemize}
+
+In a case of platform-as-a-service providers, none of them explicitly specifies how a change plan is composed.
+
+\subsection{Actions}
+Finally, a change plan is implemented in a system. Actions can cover a variety of scaling and resource tuning operations. Next chapter investigates problem of scaling applications in detail.

### 2014-02-10 thesis/src/chapters/chapter-design.tex
Thesis: adaptivity section has been rewritten

@@
 It is based on a concept of an autonomic system, providing methodical yet flexible foundation for designed system.
-Cloud-SAP leverages core benefits of such system: promise of a self-configuring, self-optimising, self-healing, self-protecting system driven by a customer-defined policies.
+What is more, Cloud-SAP leverages core benefits of such system: promise of a self-configuring, self-optimising, self-healing, self-protecting system driven by a customer-defined policies and consequently ensuring that customer is served with required Quality-of-Service level with optimised costs.
 

### 2014-02-11 thesis/src/chapters/chapter-introduction.tex
Thesis: drawings update

@@
 
-\subsection{Business potential}
+\subsection*{Business potential}
 The rapid growth of interest in cloud computing in recent years resulted in huge sums of money being invested in the field.

### 2014-02-13 thesis/src/chapters/chapter-design.tex
Thesis: drawings update
Thesis: implementation draft

--- 2014-02-06 thesis/src/chapters/chapter-design.tex
@@
   \end{center}
-  \caption{Cloud-SAP components}
+  \caption{Cloud-SAP components seen as IBM autonomic system \cite{IBM06}}
   \label{img:csap-layered-structure}
@@
 \section{Summary}
-Cloud-SAP fills the gap in existing platforms, providing a standardised approach to scaling applications that uses a fine-grained actions, operating on different platform levels: starting from execution environment to end on a cloud federation.
-It is based on a concept of an autonomic system, providing methodical yet flexible foundation for designed system.
-Cloud-SAP leverages core benefits of such system: promise of a self-configuring, self-optimising, self-healing, self-protecting system driven by a customer-defined policies.
+Cloud-SAP fills the gap in existing platforms by providing a standardised approach to scaling applications using a fine-grained actions and operating on different platform levels: starting from execution environment and ending on a cloud federation.
+It is based on a concept of an autonomic system, providing methodical but flexible foundation for its implementation.
+What is more, Cloud-SAP leverages core benefits of such a system: promise of a self-configuration, self-optimisation, self-health, self-protection.
+It is driven by a customer-defined policies and consequently ensures that customer is served with demanded Quality-of-Service and with optimised costs.
 

### 2014-02-13 thesis/src/chapters/chapter-implementation.tex
Thesis: refactoring
Thesis: implementation draft

--- 2013-12-12 thesis/src/chapters/chapter-6-implementation.tex
@@
 \chapterintro{
-  In this chapter we outline implementation details about each component of the
-  proposed solution.
+  This chapter details a proof-of-concept implementation of Cloud-SAP architecture.
 }
 
-\section{Application controller}
-\section{Container controller}
-\section{Service controller}
-\section{Cloud controller, cloud broker and client application}
+\section{Introduction}
+Previous chapters detailed requirements that a self-manageable platform oriented toward Quality-of-Service assurance should comply with.
+Beside this, reference architecture known as Cloud-SAP was introduced.
+This chapter in turn highlights key elements of a proof-of-concept implementation of Cloud-SAP.
+Noticeably, our implementation is by all means not exhaustive and merely intends to prove that presented architecture successfully tackles raised challenges.
+Hence, we implemented only following subset of managers specified by Cloud-SAP:
+\begin{itemize}
+  \item Autonomic container manager
+  \item Autonomic stack manager
+  \item Autonomic cloud instance manager
+  \item Autonomic cloud federation manager
+\end{itemize}
+
+Apart from that, we based analysis solely on threshold model, discarding more advanced techniques using prediction mechanisms.
+
+\section{Overview}
+As previous section indicates, key elements of discussed implementation are as follows: autonomic container manager, autonomic stack manager, autonomic cloud instance manager, autonomic cloud federation manager.
+Taking their role in service deployment into account, we grouped them into auto-scaling and cloud brokerage subsystems.
+Figure XXX illustrates relation between Cloud-SAP managers and above-mentioned subsystems and figure YYY depicts high level overview of a system.
+
+As one can see, our implementation is composed by following parts:
+\begin{asparaenum}
+ \item[\textbf{Auto-scaling subsystem}]
+
+ \item[\textbf{Cloud brokerage subsystem}]
+
+ \item[\textbf{Cloud provider}]
+
+ \item[\textbf{Customer}] 
+\end{asparaenum}
+
+With those information in mind, we can illustrate overall architecture, components and communication protocols on deployment diagram ZZZ. Successive sections portrays in detail specified elements.
+
+\section{Auto-scaling subsystem}
+
+\section{Cloud brokerage subsystem}
+
+\section{Cloud provider}
+
+\section{Exemplary scenarios}
+
+\subsection{Service deployment}
+
+\subsection{Scaling application}
+
+\subsection{Scaling application across multiple cloud providers}
+
+\section{Summary}

### 2014-02-13 thesis/src/chapters/chapter-state-of-the-art.tex
Thesis: implementation draft

@@
 \section{Requirements}
-One can notice that elements that yields a solution to a problem stated in the first chapter, which is ensuring that users' application provide appropriate Quality-of-Service for its customers in a most-cost effective manner, were gradually introduced in previous chapters:
+One can notice that elements that yields a solution to a problem stated in the first chapter, which is ensuring that users' application provide appropriate Quality-of-Service for its customers in a most-cost effective manner, were gradually introduced in previous chapter:
 
@@
 
-Next section states the general overview of the proposed solution, while the consecutive sections details its elements and finally the last section summarises the design choices in a context of system requirements.
+In the next sections we examine latest advancements used by a platform-as-a-service providers that is: Carina, OneFlow, CloudFoundry and OpenShift.
+Apart from getting into detail regarding above-mentioned factors, we scrutinise whether given product satisfies following functional and non-functional requirements.
+
+\subsection{Functional requirements}
+\begin{enumerate}
+  \item The user of the platform is able to:
+    \begin{enumerate}
+      \item deploy a service,
+      \item cancel the service,
+      \item check the status of the previously ordered-to-deploy service at any time. \emph{Status} means 
+        \begin{inparaenum}[a)]
+        \item whether or not the deployment succeeded,
+        \item current uptime of the service,
+        \item current cost
+        \end{inparaenum}
+    \end{enumerate}
+  \item One of the elements of the platform is a client application that is used by the user of the platform to communicate with it,
+  \item During the deployment process, the platform takes as an input a description of the service (application) that consists of:
+    \begin{itemize}
+    \item service name,
+    \item software stacks (e.g. \emph{java}, \emph{ruby}),
+    \item scaling policies (per each stack) which define
+      \begin{inparaenum}[i)]
+      \item minimal and maximal number of VMs that are needed for the stack,
+      \item name of the policy (algorithm) which is used for scaling,
+      \item parameters of the policy
+      \end{inparaenum}
+    \end{itemize}
+  \item Deployment of a service is done in a way which minimizes the cost from the client's perspective with ensuring Quality-of-Service requirements at the same time,
+  \item The platform monitors the state of the deployed services and based on the results of this process takes appropriate steps in order to meet the auto-scaling requirements. These include
+    \begin{inparaenum}[a)]
+    \item altering VM's parameters and configuration,
+    \item vertical scaling,
+    \item horizontal scaling,
+    \item scaling stacks among different cloud providers
+    \end{inparaenum}
+\end{enumerate}
+
+\subsection{Non-functional}
+\begin{itemize}
+  \item The platform does not confine itself to one provider, but to a \emph{ecosystem of various cloud providers} that offers deployment capabilities which vary in terms of quality of service, cost, etc.
+\end{itemize}
 

### 2014-02-13 thesis/src/chapters/chapter-implementation.tex
Thesis: added deployment diagrams, overview

@@
 
-Apart from that, we based analysis solely on threshold model, discarding more advanced techniques using prediction mechanisms.
+Apart from that, we implemented minimal viable modules that is monitoring, analysis, planning and execution.
+For example, we based analysis solely on threshold model, discarding more advanced techniques that uses prediction mechanisms.
 
@@
 Taking their role in service deployment into account, we grouped them into auto-scaling and cloud brokerage subsystems.
-Figure XXX illustrates relation between Cloud-SAP managers and above-mentioned subsystems and figure YYY depicts high level overview of a system.
+High level overview of system, including its internal and external elements is depicted in figure \ref{fig:hlo-implementation}.
+Diagram \ref{fig:csap-layers-subsystem} illustrates thee relationship between Cloud-SAP managers and above-mentioned subsystems.
+
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics[width=\textwidth]{chapter-implementation/hlo-implementation}
+  \end{center}
+  \caption{High level system overview}
+  \label{fig:hlo-implementation}
+\end{figure}
+
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics{chapter-implementation/csap-layers-subsystem}
+  \end{center}
+  \caption{System parts and their relation with autonomic managers}
+  \label{fig:csap-layers-subsystem}
+\end{figure}
 
@@
 \begin{asparaenum}
- \item[\textbf{Auto-scaling subsystem}]
+ \item[\textbf{Auto-scaling subsystem}] Auto-scaling subsystem supervises service's life cycle, monitoring it and enacting scaling actions when necessary.
+It is concentrated on all service's aspects: containers, stacks and application instances thus it is in fact a group of autonomic managers composed by container, stack and cloud instance managers.
 
- \item[\textbf{Cloud brokerage subsystem}]
+ \item[\textbf{Cloud brokerage subsystem}] Brokerage subsystem groups components that together mediates between cloud providers. In particular, it consists of:
+  \begin{itemize}
+   \item Cloud broker - probes cloud providers and selects best offer according to a given policy.
+In Cloud-SAP model, it plays a role of orchestrating autonomic manager, that is, autonomic cloud federation manager.
+   \item Cloud client - delegates service provisioning request to a cloud broker.
+  \end{itemize}
 
- \item[\textbf{Cloud provider}]
+ \item[\textbf{Cloud provider}] External system that manages a group of resource such as computing nodes, storage and networking typologies.
+Particularly, it is able to deploy, shutdown, migrate and monitor containers.
+Although Cloud-SAP is utterly cloud provider independent, our implementation is solely focused on OpenNebula that uses OpenVZ as a hypervisor.
+We selected OpenNebula due to its simplicity, flexibility and our expertise in managing it.
+Choosing hypervisor, we were compelled to select one that is based on lightweight containers due to theirs flexibility in scaling as previously advocated.
+OpenVZ was a natural choice due to its maturity and our familiarity within it.
 
- \item[\textbf{Customer}] 
+ \item[\textbf{Application provider}] An entity that is interested in application scaling and deployment. It can be represented by a human being as well as by an external system.
 \end{asparaenum}
 
-With those information in mind, we can illustrate overall architecture, components and communication protocols on deployment diagram ZZZ. Successive sections portrays in detail specified elements.
+With those information in mind, we can illustrate overall architecture, components and communication protocols on deployment diagram \ref{fig:hlo-deployment}.
+Successive sections portrays in detail specified elements.
+
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics{chapter-implementation/hlo-deployment}
+  \end{center}
+  \caption{Deployment diagram of a Cloud-SAP implementation}
+  \label{fig:hlo-deployment}
+\end{figure}
 
 \section{Auto-scaling subsystem}
+
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics[width=\textwidth]{chapter-implementation/auto-scaling-subsystem-deployment}
+  \end{center}
+  \caption{Deployment diagram of a auto-scaling subsystem}
+  \label{fig:auto-scaling-subsystem-deployment}
+\end{figure}
 

### 2014-02-14 thesis/src/chapters/chapter-implementation.tex
Thesis: add a technical overview stub

@@
 
+\section{Technology stack overview}
+This section aims to give a brief outline of technologies that were used in this implementation with an endeavour to justify our choice.
+The chosen solutions are grouped according to their presence in appropriate subsystems.
+
+\begin{itemize}
+  \item Cloud brokerage subsystem
+    \begin{asparaenum}
+    \item[\textbf{Web Services/HTTP}] Communication between a client and a cloud broker is done by the usage of \emph{web services} over \emph{http} protocol.
+The cloud broker for a given client exposes a \emph{RESTful} API for the deployment of a service.
+The web service accepts \emph{JSON}-encoded messages, which comprises the name of a service and specification of each stack comprising it.
+The example of such a message can be found in listing \ref{lst:service-spec-test-deployment-time}.
+
+      The reason we chose this technology is because of its simplicity, maturity and great support from Ruby platform.
+The other solutions that could be used in place of this one include some message-oriented middleware standards such as AMQP or JMS, other standards that facilitate communications among systems/components such as CORBA, and other such as RMI or RPC.
+Some of the aforementioned technologies had been eliminated once we chose Ruby as the language the platform would be implemented in.
+This is because of their platform-specific nature, e.g.
+JMS requires the system to be written in Java.
+
+    \item[\textbf{AMQP}] Being able to communicate in an asynchronous, more scalable and loosely coupled way between a cloud broker and different cloud instance managers representing cloud providers requires the usage of message-oriented middleware.
+In our case we decided to use \emph{Advanced Message Queuing Protocol}as this standard is mature and its support in Ruby is solid.
+
+      This communication channel is used in the implementation to 
+      \begin{inparaenum}[a)]
+      \item get offers from cloud providers for a given service and
+      \item commission the cloud provider to deploy a service.
+      \end{inparaenum}
+
+    \item[\textbf{SQLite}] For the persistence layer there was a need for a lightweight database engine that would have good support in Ruby platform.
+Possible choices included some nonrelational solutions, such as Redis, and more traditional, relational ones, such as SQLite.
+We chose the latter as the drivers for \emph{DataMapper}, an ORM library, have better support in Ruby.
+    \end{asparaenum}
+
+  \item Auto-scaling subsystem
+    \begin{asparaenum}
+    \item[\textbf{OpenNebula}]
+    \item[\textbf{OpenNebula AppFlow/AppStage ?}]
+    \item[\textbf{Chef}]
+    \item[\textbf{OpenVZ (drivers)}]
+    \end{asparaenum}
+\end{itemize}
+
 \begin{figure}[!ht]

### 2014-02-18 thesis/src/chapters/chapter-design.tex
Thesis: Cloud-SAP as self-adaptive system

@@
 All of the presented solutions solely engage horizontal scaling, while other IT processes and corresponding actions such as vertical scaling, cloud bursting or application platform restart remains neglected.
-What is more, none of the investigated systems is a self-managing one, taking passive approach in enforcing given QoS.
+What is more, none of the investigated systems takes proactive approach in enforcing given QoS.
 To make matter worse, some platforms such as Carina or OneFlow are provider-specific, hence, requiring vast knowledge and time of entities responsible for its supervision.
 
-What system has all of the above-mentioned features is an autonomic one \cite{IBM06}, characterised by a self-management ability and driven by four crucial attributes:
+What system has all of the above-mentioned features is self-adaptive one.
+As \cite{brun2009engineering} advocates, one of the first architectures of a self-adaptive system was presented in IBM's blueprint for autonomic computing \cite{IBM06}.
+Since IBM's autonomic system is widely recognised and respected model, we decided to leverage it as a foundation of our work.
+A concept of autonomic system is alluring due to its four crucial attributes:
 \begin{itemize}
@@
 \end{itemize}
-
-Beside this, key customer values such as IT process execution cost or time needed to complete it are enhanced through rapid process initiation and reduced time and skill requirements \cite{IBM06}.
-As specified in blueprint, applying autonomic system to manage application environment is possible because following conditions are met:
 
@@
 
-Cloud Self Adaptive Platform (Cloud-SAP) is a reference architecture aspiring to supply application providers with an autonomic computing environment.
+Cloud Self Adaptive Platform (Cloud-SAP) is a reference architecture aspiring to supply application providers with a self-adaptive computing environment.
 Figure \ref{img:csap-usage-context} shows Cloud-SAP place in an exemplary application provisioning request.

### 2014-02-18 thesis/src/chapters/chapter-adaptivity.tex
Thesis: Adaptivity

@@
 
-\section{Self-manageable system}
+\section{Self-adaptive system}
 
@@
   \item \emph{reduced time and skill requirements} - automatization of IT processes makes them easier and less troublesome what is especially important for skill-intensive, error-prone and long lasting tasks
-\end{itemize}
-
-As  \cite{murch2004autonomic} explains, there are four levels of a system manageability maturity. The higher the level, the better it manages itself leading to a adaptive or autonomic system at best:
-\begin{itemize}
- \item \emph{Basic} - IT professionals manage each system element. Activities related to configuration, optimisation, healing, and protection of IT infrastructure components are performed manually.
- \item \emph{Managed} - System management technologies can be used to collect information from different systems and analyze this information.
-In this scenario, IT professionals do most of the analysis, but it is a starting point for the automation of IT tasks.
- \item \emph{Predictive} - Individual components monitor themselves, analyze changes and offer advice. Therefore, dependency on people is reduced and decision-making is improved.
- \item \emph{Adaptive} - IT components can, individually or group-wise, monitor, analyze operations and offer advice with minimal human intervention.
- \item \emph{Autonomic} - System operations are managed by business policies established by the administrator.
-In fact, business policy drives overall IT management, while at the adaptive level; there is an interaction between humans and the system.
 \end{itemize}

### 2014-02-18 thesis/src/chapters/chapter-adaptivity.tex
Thesis: bibliography fix

@@
 
-In heart of the every self-adaptive systems, according to \cite{brun2009engineering}, lies a feedback loop, concept that originates from a control engineering. As paper's authors advocate:
+In the heart of the every self-adaptive systems, according to \cite{brun2009engineering}, lies a feedback loop, concept that originates from a control engineering. As paper's authors advocate:
 \begin{quote}

### 2014-02-18 thesis/src/chapters/chapter-scalability.tex
Thesis: adaptivity section has been rewritten
Thesis: scalability fix

@@
 As a consequence, only mod\_proxy\_balancer was taken into account during comparision.
+ \item \textbf{Zeus Load Balancer} \cite{Zeus} - load balancer offered by layer47 \cite{Layer47}, it incorporates techniques such as Layer-7 traffic management and content aware algorithms
+ \item \textbf{Zeus Global Load Balancer} \cite{ZeusGlobal} - load balancer offered by layer47 \cite{Layer47}, it is specialised in balancing load among globally distributed data centres
+
 \end{itemize}

### 2014-02-18 thesis/src/chapters/chapter-design.tex
Thesis: spellings

@@
 All of the presented solutions solely engage horizontal scaling, while other IT processes and corresponding actions such as vertical scaling, cloud bursting or application platform restart remains neglected.
-What is more, none of the investigated systems takes proactive approach in enforcing given QoS.
-To make matter worse, some platforms such as Carina or OneFlow are provider-specific, hence, requiring vast knowledge and time of entities responsible for its supervision.
+What is more, none of the investigated systems take proactive approach in enforcing given QoS.
 
-What system has all of the above-mentioned features is self-adaptive one.
+What has all of the above-mentioned features is a self-adaptive system that has a holistic approach to application scaling.
 As \cite{brun2009engineering} advocates, one of the first architectures of a self-adaptive system was presented in IBM's blueprint for autonomic computing \cite{IBM06}.

### 2014-02-19 thesis/src/chapters/chapter-implementation.tex
Thesis: Cloud-SAP as self-adaptive system
Thesis: autoscaling subsystem goals

@@
 
+\newpage
 \begin{figure}[!ht]
@@
   \end{center}
-  \caption{Deployment diagram of a Cloud-SAP implementation}
+  \caption{System's deployment diagram}
   \label{fig:hlo-deployment}
@@
 
+\section{Cloud brokerage subsystem}
+
 \section{Auto-scaling subsystem}
+
+\subsection{Objectives and structure}
+The onus is on auto-scaling subsystem to:
+\begin{itemize}
+ \item scale applications in the most cost effective way
+ \item handle application deployment requests and pass them to a cloud provider 
+\end{itemize}
+
+As overview states, this subsystem is in fact a loosely coupled group of three autonomic managers, cooperating together to achieve aforementioned subsystem's goals.
+System is structured (figure \ref{fig:auto-scaling-subsystem-deployment}) on top of that observation:
+
+\begin{asparaenum}
+ \item[\textbf{Container manager}]
+ \item[\textbf{Stack manager}]
+ \item[\textbf{Cloud instance manager}]
+ \item[\textbf{Cloud provider client}]
+ \item[\textbf{Cloud broker}]
+ \item[\textbf{Cloud provider (OpenNebula and OpenVZ nodes)}]
+\end{asparaenum}
 
@@
   \end{center}
-  \caption{Deployment diagram of a auto-scaling subsystem}
+  \caption{Deployment diagram of auto-scaling subsystem}
   \label{fig:auto-scaling-subsystem-deployment}
@@
 
-\section{Cloud brokerage subsystem}
+\subsection{Container manager}
+
+\subsubsection{Objectives}
+\subsubsection{Implementation details}
+\subsubsection{Control loop iteration}
+
+\subsection{Stack manager}
+
+\subsubsection{Objectives}
+\subsubsection{Implementation details}
+
+\subsection{Cloud instance manager}
+
+\subsubsection{Objectives}
+\subsubsection{Implementation details}
 

### 2014-02-19 thesis/src/chapters/chapter-implementation.tex
Thesis: description of ass compontents

@@
 
-\subsection{Objectives and structure}
+\subsection{Introduction}
 The onus is on auto-scaling subsystem to:
@@
 As overview states, this subsystem is in fact a loosely coupled group of three autonomic managers, cooperating together to achieve aforementioned subsystem's goals.
-System is structured (figure \ref{fig:auto-scaling-subsystem-deployment}) on top of that observation:
-
-\begin{asparaenum}
- \item[\textbf{Container manager}]
- \item[\textbf{Stack manager}]
- \item[\textbf{Cloud instance manager}]
- \item[\textbf{Cloud provider client}]
- \item[\textbf{Cloud broker}]
- \item[\textbf{Cloud provider (OpenNebula and OpenVZ nodes)}]
-\end{asparaenum}
+System is structured (figure \ref{fig:auto-scaling-subsystem-deployment}) on top of that observation.
 
@@
 \end{figure}
+
+\begin{asparaenum}
+ \item[\textbf{Container manager}] Manager responsible for supervising container lifecycle and taking apropriate actions when necessary.
+It is a most fine-grained controller in our implementation.
+It delegates a problem resolution to a stack manager when necessary.
+
+ \item[\textbf{Stack manager}] It supervises a group of homogeneous resources, that is containers, known as a stack.
+Similarly to a container manager, execution may be passed to cloud instance manager when there is no choice left.
+
+ \item[\textbf{Cloud instance manager}] Supervisor of a whole cloud instance correlated to a given cloud provider.
+It participates in dialogue with cloud broker by advertising cloud provider capabilities and handling provisioning requests.
+Besides, it is eligible of most coarse-grained scaling action: horizontal scaling across multiple providers.
+
+ \item[\textbf{Cloud broker}] Is a part of cloud brokerage subsystem responsible for mediating between different auto-scaling subsystems and cloud client. Previous section covers it in more details.
+
+ \item[\textbf{Cloud provider}] External system that provisions and manages container and underlying resources.
+In case of this implementation it is an OpenNebula frontend and AppFlow server, accessible through a Cloud provider client.
+Apart from that, OpenVZ exposes its manageability capabilities by OpenVZ agent and its REST interfaces.
+Next section details cloud provider.
+
+ \item[\textbf{Cloud provider client}] Generic component that brings cloud provider features to a auto-scaling subsystem.
+Noticeably, it may be a facade to different systems, freeing subsystem from accessing cloud provider's internal structure as it is in case of our implementation.
+For example, it delegates service provisioning requests to an Applfow server while monitoring and scaling are passed to an OpenNebula frontend.
+\end{asparaenum}
 

### 2014-02-19 thesis/src/chapters/chapter-implementation.tex
Thesis: Finish technical overview in the implementation chapter

@@
     \begin{asparaenum}
-    \item[\textbf{OpenNebula}]
-    \item[\textbf{OpenNebula AppFlow/AppStage ?}]
-    \item[\textbf{Chef}]
-    \item[\textbf{OpenVZ (drivers)}]
+    \item[\textbf{OpenNebula}] There was a need for a tool that would be used for efficient management of the resources of a data center.
+OpenNebula is an open source respond to this need -- first released in 2008, with good support from the community, seemed a good candidate for a tool.
+Our familiarity with it was an additional factor which influenced our final choice of OpenNebula as the tool for the job.
+    \item[\textbf{Chef}] As manual virtual machines provisioning is a tedious and error-prone task, this is a perfect match for a tool that would automate the process.
+We chose \emph{Chef} as it is written in the language of the whole platform (i.e.
+Ruby), mature (initially released in 2009) and widely used tool.
+\footnote{According to the website of the product (\url{http://www.getchef.com/customers/}), it is used by such companies as Facebook, BlueKai or BookRenter.
+More can be found at the provided URL.}
+    \item[\textbf{OpenVZ (drivers)}] The significance of choosing linux containers as a virtualisation technology was fully described in the preceding chapters.
+However, the drivers for the OpenVZ are not supported by the OpenNebula maintainers, so we had to use our own-developed tool while taking a course at the university.
+This tools forms an official add-on for OpenNebula 3.8.
     \end{asparaenum}

### 2014-02-21 thesis/src/chapters/chapter-implementation.tex
Thesis: container manager details

@@
 \subsubsection{Objectives}
+Container manager key responsibilities are as follows:
+\begin{itemize}
+ \item supervising container and adjusting its settings to a current usage demand, i.e. scaling it vertically
+ \item delegating scaling operations to a stack manager when vertical scaling is not possible
+\end{itemize}
+
 \subsubsection{Implementation details}
+Container manager is strictly related to an autonomic container manager, component specified by Cloud-SAP.
+Consequently, its structure reflects a control loop, cyclically invoked by a scheduler and involves:
+\begin{itemize}
+ \item monitoring - manager collects data from OpenNebula frontend using a cloud provider client. Currently, implementation supports following metrics:
+    \begin{itemize}
+     \item used cpu
+     \item used memory
+    \end{itemize}
+    Beside this, manager aggregates collected data - it calculates its average, in case when there is more than one measurement in a cycle.
+
+ \item analysis - Aggregated data is analysed against client defined policy.
+Policies (e.g.
+listing Y) are based on a threshold model, denoting a valid range for a resource measurement.
+Hence, its evaluation results in \texttt{FITS, LESSER, GREATER}.
+
+ \item planning - During this phase, analysis conclusion is mapped to an action.
+Table \ref{tab:container-manager-planning} presents such mapping.
+As one can see, only the CPU is supported for the time being.
+In case of increasing CPU, resource manager is first probed to determine whether there are enough resources to perform requested action.
+Unless cloud provider is capable of handling request, execution is delegated to an upper layer - stack manager.
+
+ \item execution - Manager adjusts container settings using an OpenVZ agent, located on computing node where container is deployed.
+As for OpenVZ hypervisor, container's properties are in fact User Bean Counter (UBC), exposed through a REST interface.
+\end{itemize}
+
+\begin{table}[!htbp]
+\begin{tabularx}{\textwidth}{ l  X  X }
+\specialrule{.1em}{.05em}{.05em} 
+\textbf{Description} & \textbf{Symptom} & \textbf{Action} \\
+\specialrule{.1em}{.05em}{.05em}
+
+CPU exceeds upper usage limit & \texttt{GREATER\_CPU} & Increase CPU \\ \hline
+CPU is in legal range & \texttt{LESSER\_CPU} & Decrease CPU \\ \hline
+CPU is lesser than allowed limit & \texttt{FITS\_CPU} & None \\ \hline
+
+\end{tabularx}
+\caption{Container manager's analysis conclusion mappings}
+\label{tab:container-manager-planning}
+\end{table}
+
 \subsubsection{Control loop iteration}
@@
 
+\subsection{Cloud provider client}
+
 \section{Cloud provider}

### 2014-02-21 thesis/src/chapters/chapter-implementation.tex
Thesis: control loop sequence diagram

@@
 \end{figure}
+\newpage
 
+% radek
 \section{Cloud brokerage subsystem}
 
+% darek
 \section{Auto-scaling subsystem}
@@
 
+% darek
 \subsection{Container manager}
@@
  \item execution - Manager adjusts container settings using an OpenVZ agent, located on computing node where container is deployed.
-As for OpenVZ hypervisor, container's properties are in fact User Bean Counter (UBC), exposed through a REST interface.
+As for OpenVZ hypervisor, container's properties are User Bean Counter (UBC), exposed through a REST interface.
 \end{itemize}
@@
 CPU exceeds upper usage limit & \texttt{GREATER\_CPU} & Increase CPU \\ \hline
-CPU is in legal range & \texttt{LESSER\_CPU} & Decrease CPU \\ \hline
-CPU is lesser than allowed limit & \texttt{FITS\_CPU} & None \\ \hline
+CPU is in legal range & \texttt{FITS\_CPU} & None \\ \hline
+CPU is lesser than allowed & \texttt{LESSER\_CPU} & Decrease CPU \\ \hline
 
@@
 
-\subsubsection{Control loop iteration}
+Manager constantly loops through all aforementioned phases as depicted in figure \ref{fig:contanier-lifecycle-seq}.
 
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics[width=\textwidth]{chapter-implementation/contanier-lifecycle-seq}
+  \end{center}
+  \caption{Sequence diagram illustrating container manager control loop}
+  \label{fig:contanier-lifecycle-seq}
+\end{figure}
+
+% darek
 \subsection{Stack manager}
@@
 \subsubsection{Objectives}
+What lies in duties of a stack manager is:
+\begin{itemize}
+ \item supervising and scaling out stack
+ \item delegating problem to a cloud instance manager, when not capable of resolving it
+\end{itemize}
+
 \subsubsection{Implementation details}
 
+% radek
 \subsection{Cloud instance manager}
@@
 
+% darek
 \subsection{Cloud provider client}
 
+% darek
 \section{Cloud provider}
 
+% ?
 \section{Exemplary scenarios}

### 2014-02-22 thesis/src/chapters/chapter-implementation.tex
Thesis: add implementation of cloud instance manager

@@
 
+% napisac o zyskach ze stosowania amqp:
+% asynchronicznosc, idealne do topologii (fan-out), publish/subscribe
+
+% format wiadomosci JSON - zarowno client-broker jak i broker-coordinator
+
 % darek
@@
 \subsubsection{Objectives}
+\begin{itemize}
+  \item Processing of the deployment requests of stacks from cloud brokers
+  \item Delegating the horizontal-scaling requests to the cloud broker responsible for carrying out the whole process
+  \item Advertising the existence of a given cloud provider to the cloud broker
+\end{itemize}
+
 \subsubsection{Implementation details}
+As all of the aforementioned objectives require a communication medium between the cloud broker and cloud instance manager, we decided to employ a message-oriented middleware solution, Advanced Message Queuing Protocol (AMQP), for this purpose.
+The detailed justification of this choice can be found in the section devoted to \emph{Cloud brokerage subsystem}.
+
+The types and formats of messages this component can process and send clearly reflects its objectives and are as follows:
+\begin{asparaenum}
+\item[\textbf{Deployment request message}] The message sent by a cloud broker that orders the given cloud provider to deploy a service.
+Data contained in the message includes the name of a service and attributes of each stack comprising it.
+The attributes of a stack are as follows:
+  \begin{compactitem}
+  \item type (e.g. java, ruby)
+  \item instances (the number of virtual machines that makes the given stack)
+  \item policies (scaling policies)
+  \end{compactitem}
+\item[\textbf{Offer request message}] The message sent by a cloud broker when it wants to retrieve the deployment capabilities of a cloud provider.
+It contains the attributes of the service to be deployed.
+They match those described in the \emph{deployment request message}.
+
+  This manager creates an offer for a service that consists of the cost of each stack.
+This cost is not dynamically computed, but fetched from the configuration file that contains mapping between the stack and its cost.
+The offer is sent back to the cloud broker for further processing.
+\item[\textbf{Horizontal-scaling message}] This message is send when it is not possible to continue having deployed a stack on the given cloud instance by the lack of sufficient resources or violation of a policy.
+It contains the stack specification as the platform needs this information to redeploy the stack to another cloud instance.
+\end{asparaenum}
+
+The behaviour required from the Cloud-SAP architecture regarding the control loop within the scope of \emph{monitoring} resources of a cloud provider is delegated to underlying components, stack and container managers.
+As this manager merely mediates between the cloud provider and broker, it does not employ any \emph{analysis} and \emph{plan} mechanisms.
+Conveying information forms the \emph{execute} attribute.
 

### 2014-02-22 thesis/src/chapters/chapter-implementation.tex
Thesis: add cloud brokerage subsystem description in the implementation section

@@
 \section{Cloud brokerage subsystem}
+\subsection{Introduction}
+This subsystem consists of components which are directly used by a client or acts on their behalf.
+These are \emph{cloud client} and \emph{cloud broker} respectively.
+The placement of each component with annotated communication protocols can be seen in figure \ref{fig:csap-cloud-brokerage-deployment-diagram}.
+To annotate the cooperating components, the diagram embeds an auto-scaling subsystem node, which is not a part of this component.
 
-% napisac o zyskach ze stosowania amqp:
-% asynchronicznosc, idealne do topologii (fan-out), publish/subscribe
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics{chapter-implementation/csap-cloud-brokerage-deployment-diagram}
+  \end{center}
+  \caption{Deployment diagram of a Cloud brokerage subsystem}
+  \label{fig:csap-cloud-brokerage-deployment-diagram}
+\end{figure}
 
-% format wiadomosci JSON - zarowno client-broker jak i broker-coordinator
+\subsection{Cloud client}
+The client is a command line application that allows a user to communicate with a cloud broker which will further process the message and act on behalf the user.
+By means of this application, the user is able to deploy a service by preparing its specification in a JSON format and request its deployment to the cloud broker.
+Listing \ref{lst:service-spec-test-deployment-time} is an example of a service description.
+
+The application is written purely in Ruby and uses http protocol for communication.
+
+\subsection{Cloud broker}
+The cloud broker mediates between a client and cloud providers.
+Its main responsibility is to, during the deployment or scaling of a service, gather information and offers from cloud providers forming a cloud federation and choose the best, in terms of cost, providers and orders them the deployment of specific stacks comprising the service.
+
+The broker uses Advanced Message Queuing Protocol (AMQP) for communication between itself and cloud providers.
+It was considered the best choice for this purpose as it ensure asynchronous message processing and sending messages according to fan-out paradigm which is ideal for this use case.
+
+The message format is the same as in the case of a client application.
+
+\subsubsection{Cloud providers selection}
+One of the main responsibilities of the Cloud broker is to map the stacks of a given service to cloud providers.
+This happens when a service is to be deployed or horizontally scaled.
+As this implementation merely wants to be a minimal viable product, we took only the cost attribute into consideration when performing this function.
+
+To formalize the previous statements, the problem is as follows: given a set whose elements (representing stacks) can take an non-negative values (they come from offers from cloud providers), find the mapping so as the overall sum (total cost) be minimal.
+
+The solution is to take the minimal available price for a given stack.
+The simplest proof is by contradiction -- we can assume that we computed our sum using a non-minimal value for some elements.
+But in this case, taking the minimal value for any element causes the sum to be less than previously computed -- contradiction.
 

### 2014-02-22 thesis/src/chapters/chapter-implementation.tex
Thesis: added stack manager

@@
 \begin{itemize}
+\item deploying stack
  \item supervising and scaling out stack
@@
 \subsubsection{Implementation details}
+
+\begin{asparaenum}
+ \item[\textbf{Stack supervision}] Stack manager's does not implement full control loop, as specified by Cloud-SAP. Instead it is solely focused on planning and execution:
+\begin{itemize}
+ \item planning - Stack manager retrieves problem resolution requests from container manager.
+It internally maps diagnosed problems to actions as stated in table \ref{tab:stack-manager-planning}.
+Similarly to a container manager, it attempts to reserve resources needed to implement desired change.
+In case when it is not possible, it problem is further delegated - to a cloud instance manager.
+
+ \item execution - Manager horizontally scales container invoking OpenNebula frontend API through cloud provider client.
+Noticeably, currently only adding and removing slave instances is possible.
+Hence, attempts to re-scale master instance (i.e.
+load balancer) are not supported.
+\end{itemize}
+
+\begin{table}[!htbp]
+\begin{tabularx}{\textwidth}{ l  l l l }
+\specialrule{.1em}{.05em}{.05em} 
+\textbf{Description} & \textbf{Container's symptom} & \textbf{Stack's symptom} & \textbf{Action} \\
+\specialrule{.1em}{.05em}{.05em}
+
+Not enough slaves to handle load & \texttt{GREATER\_CPU} (slave) & \texttt{INSUFFICIENT\_SLAVES} &  Add slave \\ \hline
+Slaves are not fully occupied & \texttt{LESSER\_CPU} (slave) & \texttt{REDUNDANT} &  Remove slave \\ \hline
+Master is overly engaged & \texttt{GREATER\_CPU} (master) & \texttt{OVERLOADED\_MASTER} &  None \\ \hline
+Stack is healthy & \texttt{FITS\_CPU} & \texttt{HEALTHY} & None \\ \hline
+
+\end{tabularx}
+\caption{Stack manager's symptoms mappings}
+\label{tab:stack-manager-planning}
+\end{table}
+
+Figure X portrays above-mentioned sequence of steps.
+
+ \item[\textbf{Stack deployment}] Stack deployment is done in two phases:
+ \begin{inparaenum}
+  \item stacks's resource are reserved apriori
+  \item cloud provider client is leveraged to create service's AppFlow template and then instantiating it.
+ \end{inparaenum}
+
+\end{asparaenum}
 
@@
 \subsection{Cloud provider client}
+\subsubsection{Objectives}
+Cloud provider client is obliged to:
+
+\begin{itemize}
+ \item provide a unified access to a subset of cloud provider's features, required from an auto-scaling subsystem perspective
+\end{itemize}
+
+\subsubsection{Implementation details}
+As for OpenNebula, there are two APIs that have to be supported:
+\begin{itemize}
+ \item OpenNebula frontend - client is based on OpenNebula's ruby bindings providing access to operations such as virtual machine provisioning, container monitoring.
+XMLRPC is used as an underlying protocol.
+ \item AppFlow - client exposes AppFlow capabilities: creating service template and instantiating it. AppFlow is a REST web service.
+\end{itemize}
 
@@
 \section{Cloud provider}
+\subsection{Introduction}
+What cloud provider is responsible for is:
+\begin{itemize}
+ \item deploying, destroying, monitoring virtual machine
+ \item deploying, destroying a stack (group of correlated virtual machines)
+\end{itemize}
+
+As it was previously mentioned, we selected OpenNebula with AppFlow addition as a cloud provider.
+Besides, we used OpenVZ as a hypervisor.
+What is more, we had to supply OpenNebula's computing nodes with an OpenVZ agent - this is to enable vertical scaling, not available by default.
+
+Figure XXX shows exemplary deployment of OpenNebula along with involved components.
 

### 2014-02-22 thesis/src/chapters/chapter-implementation.tex
Thesis: implementation scenarios

@@
 
-\newpage
 \begin{figure}[!ht]
@@
 \end{figure}
-\newpage
 
@@
 Policies (e.g.
-listing Y) are based on a threshold model, denoting a valid range for a resource measurement.
+listing ref{lst:service-spec-test-deployment-time}) are based on a threshold model, denoting a valid range for a resource measurement.
 Hence, its evaluation results in \texttt{FITS, LESSER, GREATER}.
@@
 Not enough slaves to handle load & \texttt{GREATER\_CPU} (slave) & \texttt{INSUFFICIENT\_SLAVES} &  Add slave \\ \hline
-Slaves are not fully occupied & \texttt{LESSER\_CPU} (slave) & \texttt{REDUNDANT} &  Remove slave \\ \hline
-Master is overly engaged & \texttt{GREATER\_CPU} (master) & \texttt{OVERLOADED\_MASTER} &  None \\ \hline
+Slaves are not fully occupied & \texttt{LESSER\_CPU} (slave) & \texttt{REDUNDANT\_SLAVE} &  Remove slave \\ \hline
+Master is overly occupied & \texttt{GREATER\_CPU} (master) & \texttt{OVERLOADED\_MASTER} &  None \\ \hline
+Master is not occupied & \texttt{LESSER\_CPU} (master) & \texttt{UNOCCUPIED\_MASTER} &  None \\ \hline
 Stack is healthy & \texttt{FITS\_CPU} & \texttt{HEALTHY} & None \\ \hline
@@
 
-Figure X portrays above-mentioned sequence of steps.
+Figure \ref{fig:stack-lifecycle-seq} portrays above-mentioned sequence of steps.
+
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics{chapter-implementation/stack-lifecycle-seq}
+  \end{center}
+  \caption{Sequence diagram illustrating stack manager control loop}
+  \label{fig:stack-lifecycle-seq}
+\end{figure}
 
@@
 \section{Cloud provider}
-\subsection{Introduction}
 What cloud provider is responsible for is:
@@
 
-Figure XXX shows exemplary deployment of OpenNebula along with involved components.
+Figure \ref{fig:cloud-provider-deployment} shows exemplary deployment of OpenNebula along with involved components.
 
-% ?
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics{chapter-implementation/cloud-provider-deployment}
+  \end{center}
+  \caption{Cloud provider's deployment diagram}
+  \label{fig:cloud-provider-deployment}
+\end{figure}
+
+\newpage
 \section{Exemplary scenarios}
@@
 \subsection{Service deployment}
+Diagram \ref{fig:case-deployment-seq} depicts sequence of steps involved in service deployment.
+Such request is triggered by a cloud client and involves multiple cloud provider.
+Consequently, cloud broker selects best offer for its client.
+
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics[width=\textwidth]{chapter-implementation/case-deployment-seq}
+  \end{center}
+  \caption{Deployment service - sequence diagram}
+  \label{fig:case-deployment-seq}
+\end{figure}
 
 \subsection{Scaling application}
+Figure \ref{fig:scaling-1cp-seq} illustrates control loop that engages container and stack manager and results in horizontal scaling.
+This is due to the fact that computing node, where container is deployed, does not have enough resources to perform vertical scaling.
+
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics[width=\textwidth]{chapter-implementation/scaling-1cp-seq}
+  \end{center}
+  \caption{Scaling within a single cloud provider - sequence diagram}
+  \label{fig:scaling-1cp-seq}
+\end{figure}
 
 \subsection{Scaling application across multiple cloud providers}
+That scenario employs all layers of implemented system: container, stack, cloud instance managers and cloud broker.
+Similarly to a previous case, policy is violated what results in attempt to scale vertically.
+This is, however, not possible and hence stack container checks if there are enough resources for a horizontal scaling.
+As this is not the case, request is passed to cloud instance manager and then cloud broker to deploy container on another cloud.
+This scenario is also known as a cloud bursting.
 
+\begin{figure}[!ht]
+  \begin{center}
+    \includegraphics[width=\textwidth]{chapter-implementation/scaling-2cp-seq}
+  \end{center}
+  \caption{Scaling across two cloud provider - sequence diagram}
+  \label{fig:scaling-2cp-seq}
+\end{figure}
+
+\newpage
 \section{Summary}

### 2014-02-23 thesis/src/chapters/chapter-summary.tex
Thesis: add draft of a summary

\chapter{Summary}

\chapterintro{This chapter provides the summery of the dissertation and gives information about future work and research directions.}

\section{Conclusions}
In this dissertation we proposed an architecture of a platform that ensures satisfying Quality of Service requirements of users' applications deployed in a cloud computing environment. The key attributes of the proposed system include self-adaptability (in terms of the ability to conduct self-monitoring, analysing, planning and executing) and cloud-federation awareness. Additionally, we implemented a minimal viable product which conforms to the devised architecture and made its evaluation by performing tests which compare it to some of the currently available solutions.

The architecture proposal and its implementation clearly indicates that the goals of the dissertation stated in the first chapter have been achieved. Since the implementation is merely a proof-of-concept product, it cannot be seen as a complete or reference solution, though. Irrespective of the simplicity of the employed prediction, planning or resource mapping algorithms, promising test results were obtained. However, in order to get better results, more sophisticated algorithms and methods should be implemented in place of them.

All tests of the platform were not simulated on any software, but conducted on the real hardware forming a home-made data center. Tests were chosen and design so as to reflect at the highest possible rate real business use cases. Their results show considerable potential of the proposed solution.

TODO co z OpenVZ--KVM?

\section{Future work}
As the provided implementation cannot be regarded as a fully fledged one, there is a need to implement a system with the advanced algorithms regarding actions in a control loop of autonomic managers. This will enable to conduct more insightful tests of the architecture and prove its usefulness.

TODO za mało ..

### 2014-02-24 thesis/src/chapters/chapter-summary.tex
Thesis: summary v2

@@
 In this dissertation we proposed an architecture of a platform that ensures satisfying Quality of Service requirements of users' applications deployed in a cloud computing environment.
-The key attributes of the proposed system include self-adaptability (in terms of the ability to conduct self-monitoring, analysing, planning and executing) and cloud-federation awareness.
+The key attribute of the proposed system is holistic approach to self-adaptability achieved by insightful understanding of platform-as-a-service model.
+As a consequence, designed architecture manages variety of resources that users' applications incorporate leading to a system that is capable of vertical, horizontal and most importantly cloud-federation aware scaling.
 Additionally, we implemented a minimal viable product which conforms to the devised architecture and made its evaluation by performing tests which compare it to some of the currently available solutions.
 
+% tutaj mozna by wymienic krotko rozwiazania czyli ze autonomiczny, vertykalny na konetenerach - i chyba tyle
+Cloud-SAP derives ability to conduct self-monitoring, analysing, planning and executing from mature and widely respected model of an autonomic system.
+
 The architecture proposal and its implementation clearly indicates that the goals of the dissertation stated in the first chapter have been achieved.
-Since the implementation is merely a proof-of-concept product, it cannot be seen as a complete or reference solution, though.
+Since the implementation is merely a proof-of-concept, it cannot be seen as a complete or reference solution, though.
 Irrespective of the simplicity of the employed prediction, planning or resource mapping algorithms, promising test results were obtained.
@@
 
-TODO co z OpenVZ--KVM?
+TODO co z OpenVZ--KVM?, hmm to jest wpisane pod vertical scaling, mozna by ewentualnie jakos wydzielic te poziomy skalowania i je opisac
 
 \section{Future work}
-As the provided implementation cannot be regarded as a fully fledged one, there is a need to implement a system with the advanced algorithms regarding actions in a control loop of autonomic managers.
+As the provided implementation cannot be regarded as a fully fledged one, there is a need to implement a system with the advanced algorithms regarding actions in a control loop of autonomic managers such as:
+\begin{itemize}
+  \item prediction
+  \item prioritization
+  \item application self-tuning
+\end{itemize}
 This will enable to conduct more insightful tests of the architecture and prove its usefulness.
 
-TODO za mało ..
+Having more mature platform implemented, next step should be to engage communities working on Open Source cloud platform such as OpenNebula, OpenStack or Eucalyptus.
+It is expected that such cooperation broaden pool of supported cloud providers and thus truly enable diversified federation of clouds.
+
+% moze cos o wyzwaniach z tym zwiazanych, ze trzeba sie dogadac, ustalic wspolny model, format

### 2014-02-24 thesis/src/chapters/chapter-summary.tex
Thesis: summary v3

@@
 
-% tutaj mozna by wymienic krotko rozwiazania czyli ze autonomiczny, vertykalny na konetenerach - i chyba tyle
-Cloud-SAP derives ability to conduct self-monitoring, analysing, planning and executing from mature and widely respected model of an autonomic system.
-
 The architecture proposal and its implementation clearly indicates that the goals of the dissertation stated in the first chapter have been achieved.
@@
 Their results show considerable potential of the proposed solution.
-
-TODO co z OpenVZ--KVM?, hmm to jest wpisane pod vertical scaling, mozna by ewentualnie jakos wydzielic te poziomy skalowania i je opisac
+Nevertheless, we did not manage to evaluate implementation in a real-world scenario that entails using enterprise-level application such as the ones based on Java Enterprise Edition.
 
@@
 As the provided implementation cannot be regarded as a fully fledged one, there is a need to implement a system with the advanced algorithms regarding actions in a control loop of autonomic managers such as:
-\begin{itemize}
+\begin{inparaenum}[1)]
   \item prediction
-  \item prioritization
+  \item service prioritization
   \item application self-tuning
-\end{itemize}
+\end{inparaenum}
 This will enable to conduct more insightful tests of the architecture and prove its usefulness.
 
-Having more mature platform implemented, next step should be to engage communities working on Open Source cloud platform such as OpenNebula, OpenStack or Eucalyptus.
+Beside this, the implementation should became a business-ready product instead of being simply a proof-of-concept. What should be done to enact it is to:
+\begin{itemize}
+ \item enrich Cloud-SAP architecture with a number of financial aspects such as service maintenance cost, policy and customer oriented billings (i.e.
+fee should take into account customer-specified policy or customer himself).
+ \item evaluate solution against real-world benchmarks such as SPECweb96, SPECweb99 or TPC-C
+\end{itemize}
+Due to Cloud-SAP flexibility, above-mentioned features should be easily and seamlessly incorporated in proposed model.
+
+Having more mature platform implemented and tested, next step should be to engage communities working on Open Source cloud platform such as OpenNebula, OpenStack or Eucalyptus.
 It is expected that such cooperation broaden pool of supported cloud providers and thus truly enable diversified federation of clouds.
-
-% moze cos o wyzwaniach z tym zwiazanych, ze trzeba sie dogadac, ustalic wspolny model, format
+However, for cloud providers to cooperate, there is a need for discussion and consideration regarding aspects such as authentication, authorization, resource
+ access, resource discovery, to name a few.

### 2014-02-27 thesis/src/chapters/chapter-implementation.tex
Thesis: implementation summary

@@
 \section{Summary}
+Summing up, design of a proof-of-concept implementation of Cloud-SAP was presented.
+Despite the fact that it is not exhaustive, more specifically, it does not harness any prediction or advanced analysis mechanisms, it comprehensive enough to validate key points of Cloud-SAP and provide an entry point for discussions and future implementations.
+Although, it supports solely OpenNebula, it remains open for any cloud provider.
+Similarly, platform currently supports vertical scaling of CPU, horizontal scaling and scaling across multiple cloud provider, however, it is expected that next versions support wider range of actions (i.e.
+vertical scaling of memory, application platform tuning).

### 2014-02-27 thesis/src/chapters/chapter-introduction.tex
Thesis: summary v3
Thesis: implementation summary
Thesis: motivation spellings

@@
 \section{Motivation}
-One of the keys factors that has driven transformation of computing industry in the last years is the perception of computing utilities as an ordinary property\cite{BuYeVeBrBr09}, which can be easily accessed and adjusted to a specific needs.
-That point of view resulted in profusion of different services, often collectively referred as a cloud computing \cite{MeGr11}.
-Similarly to services common to traditional markets, customers expect them to be accessible on demand and in easy manner, while paying only for the consumed goods.
-Furthermore, customers are interested in a given service provider only when it is eligible to guarantee appropriate quality of service.
+One of the keys factors that has driven transformation of computing industry in the last years is the perception of computing utilities as a commodity\cite{BuYeVeBrBr09}, easily accessible and adjustable to specific needs.
+As a consequence, we observe a profusion of different services, often collectively referred as a cloud computing \cite{MeGr11}.
+Similarly to services known from traditional markets, customers expect them to be accessible on demand and in easy manner, while paying only for the consumed goods.
+Furthermore, customers are interested in a given service only when its provider is eligible to guarantee appropriate quality of service.
 
-The particular service providers that are addressed by this paper are the ones that supply users with an application execution platform, what is widely known as providing Platform-as-a-Service.
-In that case, a customer is an entity that has developed application and is eager to deploy it on an application platform that is able to fulfil his specific requirements, both in terms of quality and cost.
+The particular service providers that are addressed by this thesis are the ones that supply users with an application execution platform, a model widely known as providing Platform-as-a-Service.
+In that case, a customer is an entity that has developed an application and is eager to deploy it on an application platform provided that it is able to fulfil his specific requirements, both in terms of quality and cost.
 
-Having customer requirements in mind, it is crucial that service provider is able to adapt itself to meet them.
+Having customer requirements in mind, it is crucial that a service provider is able to adapt itself to fulfil customer's contract.
 For example, such adaptation can be triggered by a sudden spike in resource demand and may result in provisioning additional application platforms.
-However, due to the complexity of a system under consideration, there are different levels where adaptation is possible:
+However, due to the problem complexity and number of concerns involved, there are different levels where adaptation is possible:
 \begin{itemize}
@@
 
-While autonomic computing has a long history \cite{Mu04}, it has not been directly applied to a multi-layered problem that exists in a cloud computing environment.
+While self-adaptive systems have a long history \cite{Mu04}, it have not been directly applied to a multi-layered problem that exists in a cloud computing environment.
 Especially, the research area at the last layer, which sizes across different service providers, is new.
-Although, architecture known as InterCloud \cite{BuRaCa10} investigates problem of cooperation and negotiation at cloud level, it neither has been implemented nor presented in context of autonomic system.
+Although, architecture known as InterCloud \cite{BuRaCa10} investigates problem of cooperation and negotiation at a cloud level, it neither has been implemented nor presented in a context of self-adaptive system.
 
@@
   \item The implementation of the proposed architecture using OpenNebula technology stack
+  \item The evaluation of implemented proof-of-concept
 \end{itemize}

### 2014-02-28 thesis/src/appendices/glossary.tex
Thesis: add glossary style

\newglossaryentry{perl}{name={Perl},  description={A scripting language}}  
\newglossaryentry{tex}{name={\TeX},  description={A typesetting language},sort={TeX}}  
\newglossaryentry{html}{name={html},  description={A mark up language}}

### 2014-02-28 thesis/src/abstract.tex
Thesis: abstract, impact

@@
 
-Cloud computing has become an attractive model for provisioning on demand computing resources as services to end-users.
+Cloud computing has become an attractive model for on demand provisioning of computing resources as services to end-users.
 It is based on the assumption that almost anything can be viewed as a service, starting from applications delivered over Internet, through hardware in the data centers and ending on computing power.
@@
 they support sudden spikes in demand followed by underutilization at other times.
-An architecture of a cloud computing system, that meets these requirements, has a characteristic of a multi-hierarchical autonomic system, where different layers corresponds to different levels where cloud operates: starting from an application layer, through the application platform, infrastructure to end on a cloud instance.
+An architecture of a cloud computing system, that meets these requirements, has a characteristic of a multi-layered self-adaptive system, where different layers corresponds to different levels where cloud operates: starting from an application layer, through the application platform, infrastructure to end on a cloud instance.
 
-Problem complexity raises challenges in a variety of aspects, especially in terms of providing cooperation and mutual sharing of resources that may belong to different kind of cloud providers.
-Therefore, an architecture that enables seamless cooperation among cloud providers and takes into account various QoS requirements of end-users be developed is absolutely vital.
+Problem complexity raises challenges in a variety of aspects, especially in terms of providing cooperation and mutual sharing of resources that may belong to different cloud providers.
+Therefore, an architecture that enables seamless cooperation among cloud providers and takes into account various QoS requirements of end-users is absolutely vital.
 The InterCloud architecture is one of the first attempts that had been made in this direction.
-Having characteristic of an application platform in mind, we propose a variation of that architecture that supports cooperation among these application platform and fulfils needs of a decentralized environment.
+Having characteristic of an application platform in mind, we propose a variation of that architecture.
 

### 2014-02-28 thesis/src/chapters/chapter-introduction.tex
Thesis: abstract, impact

@@
 \end{itemize}
-What is more, the fact that single service provider is constrained by his finite amount of resources poses a risk that it may not be able to serve customer all the time.
+What is more, the fact that a single service provider is constrained by his finite amount of resources poses a risk that it may not be able to serve customer all the time.
 Consequently, it is expected that adaptation at a service provider level is also possible, i.e.
@@
 \begin{itemize}
-  \item A proposal of an architecture of federated cloud computing environment, which is based on and can be viewed as a simplified version of \emph{InterCloud}
-  \item The notion of considering each service model as an autonomic system
+  \item A proposal of an architecture of a cloud computing federation, which is based on and can be viewed as a simplified version of \emph{InterCloud}
   \item The implementation of the proposed architecture using OpenNebula technology stack
@@
 \section{Impact}
-We hope that the concept of representing each level of an autoscaling subsystem as an autonomic one can be thought-provoking for cloud computing scientists.
+We hope that the concept of self-adaptation at each level of cloud can be thought-provoking for computing scientists.
 What is more, we believe that our successful attempt to implement a simplified variant of an InterCloud architecture will cause its gain in interest and popularity.
-Finally, we consider the ideas contained in this work be beneficial to the OpenNebula ecosystem as they provide insights into the ways Quality of Service can be ensured:
+Finally, we consider the ideas contained in this work be beneficial to the OpenNebula ecosystem as they provide insights into the ways Quality of Service can be assured:
 \begin{itemize}
-  \item implementing autoscaling capabilities
+  \item implementing auto-scaling capabilities
   \item designing \emph{cloud infrastructure} in accordance with InterCloud architecture

### 2014-02-28 thesis/src/appendices/glossary.tex
Thesis: glossary part a

@@
-\newglossaryentry{perl}{name={Perl},  description={A scripting language}}  
-\newglossaryentry{tex}{name={\TeX},  description={A typesetting language},sort={TeX}}  
-\newglossaryentry{html}{name={html},  description={A mark up language}}
+\newglossaryentry{auto-scaling}{name={auto-scaling},  description={System ability to dynamically scale, i.e. add additional resources}}
+\newglossaryentry{self-adaptation}{name={self-adaptation},  description={Ability of a system to adapt its behaviour in order to react towards the environment dynamic}}
+\newglossaryentry{cloud federation}{name={cloud federation},  description={A concept that comprises services from different
+providers aggregated in a single pool supporting three basic interoperability features - resource migration, resource
+redundancy and combination of complementary resources
+resp. services}}
+\newglossaryentry{InterCloud}{name={InterCloud},  description={An example of a federated cloud computing environment architecture}}
+\newglossaryentry{PaaS}{name={PaaS},  description={Service that enables deployment onto the cloud infrastructure consumer-created or acquired applications created using programming languages, libraries, services, and tools supported by the provider.}}
+\newglossaryentry{IaaS}{name={IaaS},  description={Service that enables provisioning of processing, storage, networks, and other fundamental computing resources where the consumer is able to deploy and run arbitrary software, which can include operating
+systems and applications}}
+\newglossaryentry{QoS}{name={QoS},  description={Totality of characteristics of a service that bear on its ability to satisfy stated and implied needs of the user of the service.}}
+\newglossaryentry{SLA}{name={SLA},  description={A document which defines the relationship between two parties: the provider and the recipient. It should
+\begin{inparaenum}[1)]
+ \item Identify and define the customer’s needs
+ \item Provide a framework for understanding
+\end{inparaenum}
+}}
+\newglossaryentry{policy}{name={policy},  description={A principle or protocol to guide decisions and achieve rational outcomes.}}

### 2014-02-28 thesis/src/appendices/glossary.tex
Thesis: glossary part 2

@@
 \newglossaryentry{InterCloud}{name={InterCloud},  description={An example of a federated cloud computing environment architecture}}
-\newglossaryentry{PaaS}{name={PaaS},  description={Service that enables deployment onto the cloud infrastructure consumer-created or acquired applications created using programming languages, libraries, services, and tools supported by the provider.}}
-\newglossaryentry{IaaS}{name={IaaS},  description={Service that enables provisioning of processing, storage, networks, and other fundamental computing resources where the consumer is able to deploy and run arbitrary software, which can include operating
+\newglossaryentry{PaaS}{name={Platform-as-a-Service (PaaS)},  description={A service that enables deployment onto the cloud infrastructure consumer-created or acquired applications created using programming languages, libraries, services, and tools supported by the provider.}}
+\newglossaryentry{IaaS}{name={Infrastructure-as-a-Service (IaaS)},  description={A service that enables provisioning of processing, storage, networks, and other fundamental computing resources where the consumer is able to deploy and run arbitrary software, which can include operating
 systems and applications}}
-\newglossaryentry{QoS}{name={QoS},  description={Totality of characteristics of a service that bear on its ability to satisfy stated and implied needs of the user of the service.}}
-\newglossaryentry{SLA}{name={SLA},  description={A document which defines the relationship between two parties: the provider and the recipient. It should
+\newglossaryentry{QoS}{name={Quality-of-Service (QoS)},  description={Totality of characteristics of a service that bear on its ability to satisfy stated and implied needs of the user of the service.}}
+\newglossaryentry{SLA}{name={Service Level Agreement (SLA)},  description={A document which defines the relationship between two parties: the provider and the recipient. It should
 \begin{inparaenum}[1)]

### 2014-02-28 thesis/src/appendices/glossary.tex
Thesis: add the part of a glossary

@@
 \newglossaryentry{cloud federation}{name={cloud federation},  description={A concept that comprises services from different
-providers aggregated in a single pool supporting three basic interoperability features - resource migration, resource
-redundancy and combination of complementary resources
-resp. services}}
+providers aggregated in a single pool supporting three basic interoperability features - resource migration, resource redundancy and combination of complementary resources resp. services}}
 \newglossaryentry{InterCloud}{name={InterCloud},  description={An example of a federated cloud computing environment architecture}}
@@
 \newglossaryentry{policy}{name={policy},  description={A principle or protocol to guide decisions and achieve rational outcomes.}}
+
+\newglossaryentry{adaptable system}{name={adaptable system},  description={
+An adaptable system is a system that has the following attributes:
+\begin{inparaenum}[1)]
+\item it can monitor its state and take appropriate actions,
+\item IT staff is responsible for managing performance against SLAs and
+\item the human interaction should be at the minimal level.
+\end{inparaenum}
+}}
+
+\newglossaryentry{autonomic system}{name={autonomic system},  description={ A
+system with the ability to manage itself and dynamically adapt to change in
+accordance with business policies and objectives. Self-managing environments
+can perform such activities based on situations they observe or sense in the IT
+environment rather than requiring IT professionals to initiate the task.  These
+environments are self-configuring, self-healing, self-optimizing, and
+self-protecting.  }}
+
+\newglossaryentry{lightweight virtualization}{name={lightweight
+virtualization}, description={ Lightweight or operating system–level
+  virtualization is a server virtualization method where the kernel of an
+  operating system allows for multiple isolated user-space instances, instead
+  of just one. Such instances (often called containers, virtualization engines
+(VE), virtual private servers (VPS) or jails) may look and feel like a real
+server, from the point of view of its owner. }}
+
+\newglossaryentry{horizontal scalability}{name={horizontal scaling},
+description={ The ability to connect multiple hardware or software entities,
+such as virtual machines, so that they work as a single logical unit.  }}
+
+\newglossaryentry{vertical scalability}{name={vertical scaling}, description={
+The ability to increase the capacity of existing hardware or software by adding
+resources - for example, adding processing power to a server to make it faster
+}}
+
+\newglossaryentry{hybrid cloud}{name={hybrid cloud},  description={
+The cloud infrastructure is a composition of two or more distinct cloud
+infrastructures (private, community, or public) that remain unique entities,
+but are bound together by standardized or proprietary technology that enables
+data and application portability (e.g., cloud bursting for load balancing
+between clouds)
+}}
+
+\newglossaryentry{cloud bursting}{name={cloud bursting},  description={
+A deployment model which enables seamless resource infrastructure scaling for
+a cloud customer which has its own private cloud. It relies on the concept of
+using resources provided by an external provider for a given time interval 
+when necessary, e.g. in case of a sudden spike in demand of a web application
+}}

### 2014-03-01 thesis/src/chapters/chapter-adaptivity.tex
Thesis: adaptivity fix

@@
 \chapter{Problem domain}
-\chapterintro{This chapter details mechanisms and practises that ensures a requested Quality-of-Service level is met by a means of a self-adaptive system that scales across multiple service providers.}
+\chapterintro{This chapter details mechanisms and practises that ensures a Quality-of-Service level is met by means of a self-adaptive system that scales across multiple service providers.}
 
@@
 \subsection{Introduction}
-Guaranteeing Quality-of-Service, a critical concept in this paper, can be seen as continuous monitoring and reacting to undesirable conditions when necessary.
-In fact, this process is as an IT procedure and traditionally belongs to the duties of a system administrator.
-Figure \ref{fig:it-process} illustrates exemplary IT procedure.
+Guaranteeing Quality-of-Service can be seen as continuous monitoring and reacting to undesirable conditions when necessary.
+In fact, this process is an IT procedure performed by system administrator.
+Figure \ref{fig:it-process} illustrates an example of such procedure.
+While system administrator can be directly responsible for manually performing steps depicted in figure \ref{fig:it-process}, it is much more favourable to automatize this process leading to a self-manageable system.
+Not only does it lead to efficiency of an IT process but also to its effectiveness \cite{IBM06}.
+It is achieved through:
+\begin{itemize}
+  \item \emph{rapid process initiation} - components auto-initiate actions based on information derived from a system
+  \item \emph{reduced time and skill requirements} - automatization of IT processes makes them easier and less troublesome what is especially important for skill-intensive, error-prone and long lasting tasks
+\end{itemize}
 
@@
 
-While system administrator can be directly responsible for manually performing steps depicted in figure \ref{fig:it-process}, it is much more favourable to automatize this process leading to a self-manageable system.
-Not only does it lead to efficiency of an IT process but also to its effectiveness \cite{IBM06} achieved through:
-\begin{itemize}
-  \item \emph{rapid process initiation} - components auto-initiates actions based on information derived from a system
-  \item \emph{reduced time and skill requirements} - automatization of IT processes makes them easier and less troublesome what is especially important for skill-intensive, error-prone and long lasting tasks
-\end{itemize}
-
 In the heart of the every self-adaptive systems, according to \cite{brun2009engineering}, lies a feedback loop, concept that originates from a control engineering. As paper's authors advocate:
@@
 
-What control loop is made of is collection of data, its analysis, making a decisions and enforcing them in the environment.
-Figure \ref{fig:feedback-loop} depicts a generic feedback loop, giving context associated with each phase.
+What control loop is made of is:
+\begin{inparaenum}[1)]
+ \item collection of data
+ \item data analysis
+ \item making a decisions
+ \item enforcing actions
+\end{inparaenum}
+Figure \ref{fig:feedback-loop} depicts a generic feedback loop, giving context associated with each phase. Successive sections details aforementioned phases.
 
@@
 \end{figure}
-
-Successive sections details aforementioned phases and gives an examples of systems leveraging feedback loops.
 

### 2014-03-01 thesis/src/chapters/chapter-introduction.tex
Thesis: adaptivity fix

@@
 \begin{itemize}
-  \item A proposal of an architecture of a cloud computing federation, which is based on and can be viewed as a simplified version of \emph{InterCloud}
+  \item A proposal of an architecture of a self-adaptive platform that ensures satisfying Quality of Service requirements of deployed users' applications in a cloud computing environment by leveraging horizontal, vertical and cloud-federation-aware scaling
   \item The implementation of the proposed architecture using OpenNebula technology stack
-  \item The evaluation of implemented proof-of-concept
+  \item The evaluation of the implemented proof-of-concept system
 \end{itemize}
@@
 We hope that the concept of self-adaptation at each level of cloud can be thought-provoking for computing scientists.
-What is more, we believe that our successful attempt to implement a simplified variant of an InterCloud architecture will cause its gain in interest and popularity.
+What is more, we believe that our successful attempt to implement the proposed architecture will cause a greater gain in interest in models that enable scaling applications across different cloud providers, e.g.
+\emph{InterCloud}.
 Finally, we consider the ideas contained in this work be beneficial to the OpenNebula ecosystem as they provide insights into the ways Quality of Service can be assured:
+
 \begin{itemize}
   \item implementing auto-scaling capabilities
-  \item designing \emph{cloud infrastructure} in accordance with InterCloud architecture
+  \item designing cloud-federation-aware cloud infrastructures
 \end{itemize}
@@
 \section{Thesis structure}
+This dissertation is organised as follows:
+
+\begin{enumerate}[\bfseries {Chapter} I]
+    \setcounter{enumi}{1}
+  \item presents an introduction to the problem domain giving an outline and definition of concepts of \emph{self-adaptive systems}, \emph{application scaling} and \emph{interoperability of clouds}
+  \item describes and compares similar solutions in the field with the emphasis on their auto-scaling capabilities
+  \item presents our proposal of an architecture, giving an in depth description of each component which constitutes it
+  \item is devoted to our proof-of-concept implementation of the architecture
+  \item provides information regarding tests of the implementation and their results
+  \item presents conclusions and points further research directions
+\end{enumerate}

### 2014-03-01 thesis/src/chapters/chapter-design.tex
Thesis: refinements

@@
 As one can see, Cloud-SAP can be seen as a proxy between the user and cloud instances and as a coordinator of the latter.
-Although cloud instance itself consists of a cloud manager and additional resources, in general, it is perfectly legitimate to consider it a resource as well.
+Although the cloud instance itself consists of a cloud manager and additional resources, in general, it is perfectly legitimate to consider it a resource as well.
 
@@
 The output from one block forms an input to another.
-However, it is perfectly possible for a block to perform an action for its side effects, e.g.
+However, it is perfectly possible for a block to perform an action for obtaining side effects, e.g.
 an action can result in producing relevant information contained in log files, which, in turn, comprises a piece of shared knowledge.
 Each block implements different behaviour based on the abstraction level and resource they operate on (see figure \ref{img:csap-layered-structure}).
-Nevertheless it is possible to define their generic behaviour, which embraces the four phases: monitoring, analysis, planning, execution.
+Nevertheless it is possible to define their generic behaviour, which embraces the four phases: monitoring, analysis, planning and execution.
 
@@
 \subsection{Structure}
-It is clear from the description of requirements of this autonomic manager in the previous section that \emph{plan} and \emph{execution} functions are common for both of them.
-Therefore the \emph{cloud federation manager} is composed of two internal autonomic managers, which perform complete different functions, the first one -- \emph{monitor} and \emph{analyse} and the second -- \emph{plan} and \emph{execute}, yet they cooperate with each other to realize the full control loop.
+
+In the previous chapter requirements for this component were presented.
+It is absolutely correct to fulfill them using only the internal control loop within this manager, but when we dissect the flow of actions taken during service deployment and service scaling we can see that during the deployment only \emph{planning} and \emph{executing} take place opposed to scaling, when the whole control loop must be executed.
+This is because the phase of \emph{analysis} is responsible for creating a plan \emph{what} should be done based on data passed from the \emph{monitoring} function.
+In case of deployment, it is clearly known which action should be executed.
+Additionally, if we do not consider handling an explicit deployment request a requirement for \emph{monitoring} function, we can see that this function does not participate in the service deployment as well.
+
+Therefore we can separate the concerns into two different, internal autonomic managers performing completely different functions -- the first one -- \emph{monitor} and \emph{analyse} and the second -- \emph{plan} and \emph{execute}, yet they cooperate with each other to realize the full control loop.
+It is proper to see that the \emph{orchestrating autonomic manager} is \textbf{composed} of \textbf{two}, aforementioned managers.
 
@@
   \item[\textbf{Problem description}]
-The problem of resource management in cloud computing environments is hard and complex.
-This is because of several aspects.
+The problem of resource management in cloud computing environments is hard and complex. This is because of several aspects, but we will highlight only two reasons.
+
 The first one is the characteristic of resources available in a cloud -- not only are they geographically distributed, but they can be under different administrative domains and can vary in accessibility.
@@
 Additionally, different entities forms the resources -- starting from individual devices, through virtual machines and whole workstations, ending with clusters and data centers.
+
 The second one are the increasing Quality-of-Service requirements from the clients.
@@
   \item[\textbf{Service deployment}] When the user wants to deploy a service on a cloud infrastructure, they have to prepare its detailed description containing technical (stacks) and Quality-of-Service requirements.
+Additionally, this component should have access to shared knowledge regarding the information about every cloud instance that forms the federation.
+This should include resource capacity and utilization rates.
   \item[\textbf{Service scaling}] When an application needs to be scaled, there is available additional knowledge that this component should use of -- detailed information about service's performance up to that point, e.g.
@@
 \item[\textbf{Deployment of a new service}] When a new service is about to be deployed on a cloud federation, the system has to a priori aggregate information about resource capacity, utilization rates, geographical location and so on from cloud providers to be able to perform the best match of a cloud provider with the given service.
+This component should collect this information and make it available to other components by placing it in \emph{shared knowledge}.
   \item[\textbf{Scaling of a deployed service}] When a service which had been hosted on a cloud needs to be scaled, the similar knowledge as in the previous case is required to be passed to \emph{analyse} function to perform the perfect match of a cloud instance.

### 2014-03-01 thesis/src/chapters/chapter-evaluation.tex
Thesis: openshift description
Thesis: refinements

@@
 \chapterintro{
-  This chapter contains results and discussions on the figuation of tests run on the proposed solution.
+  This chapter contains results and discussions on the evaluation of tests run on the proposed solution.
 }
@@
     \item monitoring the scaling actions performed by each solution,
-    \item figuation of the cost the client has to pay for the service.
+    \item evaluation of the cost the client has to pay for the service.
   \end{inparaenum}
@@
   \item every peak in CPU usage should be followed by a gradual descent which would mimic the real auto-scaling actions executed by each solution,
-  \item (boundary conditions) CPU usage should be between 0 and 100 and, additionally, should exceed the previously set scaling threshold value of each solutions so that it would trigger the scaling policies figuation
+  \item (boundary conditions) CPU usage should be between 0 and 100 and, additionally, should exceed the previously set scaling threshold value of each solutions so that it would trigger the scaling policies evaluation
   \item once the scaling operations have completed, CPU usage should be constant at a rate which would not introduce any changes in the environment settings (such as the number of virtual machines and parameters of any virtual machine)
 \end{itemize}
-This resulted in the function whose formula is in \eqref{eq:cpu-mock-usage} and plot in figure \ref{fig:auto-scaling-1cp-cpu-usage-function}.
+To find the equation of a function that would satisfy the previously-mentioned requirements we used \emph{polynomial interpolation}.
+As we chose 3 points which were the input to the algorithm, we got a cubic polynomial as an output.
+Adding further constraints to mimic two spikes in CPU usage resulted in the function whose formula is in \eqref{eq:cpu-mock-usage} and plot in figure \ref{fig:auto-scaling-1cp-cpu-usage-function}.
 \begin{eqnarray}
@@
 To apply this setting, the auto-scaling part of service specification looks as follows: in Carina the user has to add the parameters of an auto-scaling policy in a hash that describes the environment under key \emph{:elasticy\_policy}.
-It is possible to specify the minimal and the maximal number of virtual machines that forms the environment and, what is most important, expressions which figuation results in scaling the application.
+It is possible to specify the minimal and the maximal number of virtual machines that forms the environment and, what is most important, expressions whose evaluation results in scaling the application.
 The part responsible for this is shown in listing \ref{lst:auto-scaling-1cp-carina-service-spec}.\lstinputlisting[caption=Carina service specification used for testing auto-scaling with 1 cloud provider, label=lst:auto-scaling-1cp-carina-service-spec]{auto-scaling-1cp-carina-service-spec} In Cloud-SAP it is a matter of setting appropriate arguments to a specific policy.
@@
 As listing \ref{lst:auto-scaling-1cp-results-carina-jobs} shows, they completely met the expectations expressed in the previous section -- there were two ''SCALEUP'' jobs which resulted in increase of the total number of virtual machines comprising the environment.
-\lstinputlisting[caption=Carina environment manager logs with taken actions (\emph{jobs}), label=lst:auto-scaling-1cp-results-carina-jobs]{auto-scaling-1cp-results-carina-jobs}.
+The one-hour shift in the listing is caused by the wrong clock setting on a virtual machine.
+\lstinputlisting[caption=Carina environment manager logs with taken actions (\emph{jobs}), label=lst:auto-scaling-1cp-results-carina-jobs]{auto-scaling-1cp-results-carina-jobs}
 Since the base configuration assumed that the environment consists of 2 virtual machines, we can say that about 10 minutes after deployment of an environment, the number of VMs increased to 3, and after another 10 minutes, to 4.
@@
 \begin{asparaenum}
-
   \item[\textbf{Cloud-SAP}] 

### 2014-03-01 thesis/src/chapters/chapter-implementation.tex
Thesis: refinements

@@
 The other solutions that could be used in place of this one include some message-oriented middleware standards such as AMQP or JMS, other standards that facilitate communications among systems/components such as CORBA, and other such as RMI or RPC.
-Some of the aforementioned technologies had been eliminated once we chose Ruby as the language the platform would be implemented in.
+Some of the aforementioned technologies had been eliminated once we chose Ruby as the language which the platform would be implemented in.
 This is because of their platform-specific nature, e.g.
@@
     \item[\textbf{AMQP}] Being able to communicate in an asynchronous, more scalable and loosely coupled way between a cloud broker and different cloud instance managers representing cloud providers requires the usage of message-oriented middleware.
-In our case we decided to use \emph{Advanced Message Queuing Protocol}as this standard is mature and its support in Ruby is solid.
+In our case we decided to use \emph{Advanced Message Queuing Protocol} as this standard is mature and its support in Ruby is solid.
 
@@
 
-To formalize the previous statements, the problem is as follows: given a set whose elements (representing stacks) can take an non-negative values (they come from offers from cloud providers), find the mapping so as the overall sum (total cost) be minimal.
+To formalize the previous statements, the problem is as follows: given a set whose elements (representing stacks) can take an non-negative values (they come from offers from cloud providers), find the mapping so as the overall sum of its elements (total cost) be minimal.
 
-The solution is to take the minimal available price for a given stack.
+The solution is to take the minimal available price for every stack.
 The simplest proof is by contradiction -- we can assume that we computed our sum using a non-minimal value for some elements.

### 2014-03-01 thesis/src/chapters/chapter-interoperability.tex
Thesis: refinements

@@
 From the perspective of a user of PaaS services it is vital that they are able to deploy seamlessly their applications using libraries, tools and services supported by the cloud provider\cite{MeGr11}.
-Judging by such factors as the popularity of Heroku -- currently one of the most popular PaaS providers which does not offer more advanced features which would enable management of the infrastructure underpinning the deployment platform, the fact that Microsoft added auto-scaling to its Azure platform as late as in June 2013, it is perfectly possible most PaaS users are satisfied with the current offers of their providers and do need another, more sophisticated functionalities.
+Judging by such factors as the popularity of Heroku -- currently one of the most popular PaaS provider which does not offer more advanced features which enable management of the infrastructure underpinning the deployment platform, the fact that Microsoft added auto-scaling to its Azure platform as late as in June 2013, it is perfectly possible most PaaS users are satisfied with the current offers of their providers and do not need another, more sophisticated functionalities.
 However, there are more complex applications and systems whose requirements regarding technology stack, availability and scalability are considerably more demanding.
@@
 However, the tendency is that more and more enterprises do not find it a major issue as in 2012 the number declined to 46\%.
-Complexity related to the management of cloud components, Vendor lock-in, interoperability and reliability were among the most frequent obstacles to adoption in 2013 for they constituted 46\%, 35\%, 27\% and 22.3\% of responders' votes respectively.
+Complexity related to the management of cloud components, vendor lock-in, interoperability and reliability were among the most frequent obstacles to adoption in 2013 for they constituted 46\%, 35\%, 27\% and 22.3\% of responders' votes respectively.
 Total results are shown in the figure \ref{fig:cloud-computing-adoption-obstacles-2013}.

### 2014-03-01 thesis/src/chapters/chapter-state-of-the-art.tex
Thesis: refinements

@@
 
-Before delving into more detailed description it is vital to introduce most important terms used in \emph{Carina} documentation to depict this product:
+Before delving into a more detailed description it is vital to introduce most important terms used in \emph{Carina} documentation to depict this product:
 \begin{itemize}
@@
 On the other hand, \emph{load-based} policies define the way the system reacts to \textbf{average cpu usage} by the environment.
-This is the simplest \emph{threshold-model} as introduced in chapter 2.
+This is the simplest \emph{threshold-model} as introduced in section 2.1.3.
 The user enters predicates which are evaluated against data gathered by OpenNebula and execute scaling actions  (\emph{scaleup} and \emph{scaledown}) if they are true.
@@
 \subsubsection{Scaling}
-Chapter 3 introduced and discussed various types of application scaling.
+Section 2.2 introduced and discussed various types of application scaling.
 \emph{Carina} offers only one type of scaling and this is \emph{horizontal} scaling -- user's capabilities are limited only to add/remove a virtual machine to/from a given environment.
@@
 \item[\textbf{Provisioning}] User has to use in the environment specification previously defined OpenNebula images.
-Their further adjusting happens in the contextualization phase.
+Their further adjustment happens in the contextualization phase.
 This forces the user to modify the state of a virtual machine when there is a need to perform any changes, for example in the software stack.
@@
 \subsubsection{Scalability}
-\emph{Cloud foundry} supports only horizontal scaling not only within one provider, but also across many. This feature is called \emph{Multi-Cloud} in Cloud Foundry nomenclature.
+\emph{Cloud foundry} supports horizontal scaling not only within one provider, but also across many. This feature is called \emph{Multi-Cloud} in Cloud Foundry nomenclature.
 

### 2014-03-01 thesis/src/chapters/chapter-adaptivity.tex
Thesis: problem domain

@@
 Guaranteeing Quality-of-Service can be seen as continuous monitoring and reacting to undesirable conditions when necessary.
-In fact, this process is an IT procedure performed by system administrator.
+In fact, this process is an IT procedure performed by a system administrator.
 Figure \ref{fig:it-process} illustrates an example of such procedure.
@@
  \item data analysis
- \item making a decisions
- \item enforcing actions
-\end{inparaenum}
-Figure \ref{fig:feedback-loop} depicts a generic feedback loop, giving context associated with each phase. Successive sections details aforementioned phases.
+ \item planning decisions
+ \item enforcing actions.
+\end{inparaenum} Figure \ref{fig:feedback-loop} depicts such a generic feedback loop. Successive sections detail aforementioned phases.
 
@@
 During this phase, relevant data is collected from sensors.
-After that, data can be aggregated, filtered or stored to provide a reference point during future feedback loop evaluation.
-During this phase, there are a few concerns that should be investigated \cite{brun2009engineering}:
+After that, data can be aggregated, filtered or stored to provide a reference point for future feedback loop evaluations.
+During this phase, there are a few matters that should be examined \cite{brun2009engineering}:
 \begin{itemize}
@@
 
-Almost any cloud platform providers incorporates into its solution monitoring component, responsible for data collection.
-Mechanisms used vary from providers to provider and may involve mixture of protocols and standards.
-For example, Carina \cite{Carina} uses OpenNebula IM and VMM drivers relaying on a plaintext data, transferred through a SSH connection.
+Mechanisms used by a monitoring component vary from provider to provider and may involve a mixture of protocols and standards.
+For example, Carina \cite{Carina} uses OpenNebula IM and VMM drivers that relays on a plaintext data transferred through a SSH connection.
 
 \subsection{Analysis}
-Analysis part applies mathematical models to reason about data gathered during previous phase. It can leverage variety of approaches and be driven by a customer-defined policies.
-
-As \cite{brun2009engineering} specifies, it is important to take into account following concerns during analysis phase:
+Analysis part applies mathematical models to reason about data gathered during previous phase.
+It can leverage a variety of approaches and be driven by a customer-defined policies.
+As \cite{brun2009engineering} specifies, it is vital to take into account following concerns during analysis phase:
 \begin{itemize}
- \item applying historical data and existing patterns to current situation
+ \item applying historical data and existing patterns to a current situation
  \item archiving data
- \item applying adequate model to current situation
+ \item applying adequate model to a current situation
  \item model's stability
@@
 
-As it was already mentioned, models are often controlled by a policies that denotes a condition which, when satisfied, triggers an action, harnessing environment's disturbance.
+As it was already mentioned, models are often controlled by policies that denote a condition which, when satisfied, triggers an action harnessing environment's disturbance.
 Currently, industry leaders supports \cite{AmazonAutoScaling} two kind of policies:
 \begin{itemize}
- \item \textit{expression based} - allows to define how to scale application in response to changing conditions, which include factors such as memory, CPU usage, cost or some indirect, calculated metrics
- \item \textit{scheduled} - allows to scale an application in response to predictable load changes.
+ \item \textit{expression based policy} - it allows to define how to scale application in response to changing conditions, which include factors such as memory, CPU usage, cost or some indirect, calculated metrics
+ \item \textit{scheduled policy} - it allows to scale an application in response to predictable load changes.
 For example, traffic increases during the weekends and decreases on working days.
@@
 \begin{itemize}
- \item risk analysis - analysing how change effects whole environment
+ \item risk analysis - analysing how a change affects environment
  \item prioritising identified problems and resources
@@
 
-In a case of platform-as-a-service providers, none of them explicitly specifies how a change plan is composed.
-
 \subsection{Actions}
-Finally, a change plan is implemented in a system. Actions can cover a variety of scaling and resource tuning operations. Next chapter investigates problem of scaling applications in detail.
+Finally, a change plan is implemented in a system. Actions can cover a variety of scaling and resource tuning operations. Next chapter investigates problem of scaling an application in detail.

### 2014-03-01 thesis/src/chapters/chapter-scalability.tex
Thesis: problem domain

@@
 
-The reason why scaling application lies in our area of interest is the fact that it is widely accepted measure for improving application performance, consequently increasing offered Quality-of-Service.
+Scaling is a widely accepted measure of improving application performance leading to increase of offered Quality-of-Service.
 Enriching system with capability to scale entails avoiding additional costs that are related to coping with excessive traffic.
@@
 Hence, it is necessary to clarify this term before going into further discussion.
-Instinctively, scalability is perceived as ability of a system to accommodate an increasing number of elements or objects to process.
+Instinctively, scalability is perceived as an ability of a system to accommodate an increasing number of elements or objects to process.
 In particular, we can point out different types of scalability that are affected by increased number of requests: \cite{Bo00}:
 \begin{itemize}
-	\item \textit{load scalability} - ability to work without delays and unproductive resource consumption at light, moderate, or heavy loads while making good use of available resources.
+	\item \textit{load scalability} - an ability to work without delays and unproductive resource consumption at light, moderate, or heavy loads while making good use of available resources.
 Factors that may hinder load scalability include: scheduling shared resource, self-expansion, inadequate exploitation of parallelism
@@
 Although, all of the aforementioned aspects are vital for any application, our work focuses solely on the first type of scalability.
-The reasoning behind this statement is that, while all of these scalability types lies in direct responsibility of an application developer, the load scalability can be additionally improved by adding additional resources to a system.
-This brings us to a question what kind of resources are used by an application or more appropriately in context of this dissertation: \textit{what kind of resources can we add to improve application performance?} Required resources varies from an application to an application.
-However, among the most common ones we can distinguish:
+The reasoning behind this is that while all of these scalability types lies in direct responsibility of an application developer, the load scalability can be additionally improved by adding additional resources to a system.
+This brings us to a question what kind of resources are used by an application or more appropriately in context of this dissertation: \textit{what kind of resources can we add to improve application performance?} Required resources vary from application to application.
+However, among the most common one we can distinguish:
 \begin{itemize}
@@
 
-It is commonly agreed that there are two main possible ways the resource can be added:
+It is commonly agreed that there are two ways of adding a resource:
 \begin{itemize}
@@
 
-What makes scaling application particularly interesting are the benefits offered by a cloud computing, especially the illusion of a virtually infinite computing infrastructure \cite{VaRoBu11}.
-Making use of virtualization technologies, which often underpins cloud computing platform, allows for resource manipulation in a dynamic, on-demand manner.
-Although, cloud computing offers additional scaling capabilities, it increases solution complexity since they operate in different layers: server, platform, network as stated in \cite{VaRoBu11}.
+Cloud computing makes scaling application especially interesting due to the illusion of a virtually infinite computing infrastructure \cite{VaRoBu11}.
+Virtualization technologies, which often underpins cloud computing platform, allows for resource manipulation in a dynamic, on-demand manner.
+Although cloud computing offers additional scaling capabilities, it increases solution complexity because it operates in different layers: server, platform and network as stated in \cite{VaRoBu11}.
 However, since platform containers are often represented either as virtual machines or another isolated environment (e.g.
-OpenShift leverages SELinux and cgroups) they are similar in nature to server scaling and supports both scaling up and out.
+OpenShift leverages SELinux and cgroups) they are similar to server scaling and supports both scaling up and out.
 Therefore, the remaining of this chapter is focused solely on server scaling, omitting network scaling as it lays outside of scope of this dissertation.
 
-Having that said, common sense dictates that adding resources is only a part of the success - it should be accompanied by tuning application platform configuration.
-For example, adding supplementary CPUs without increasing thread pool size that handle requests makes a little sense.
+Adding resources is only a part of the success - it should be accompanied by tuning application platform's configuration.
+For example, adding supplementary CPUs without increasing thread pool size makes a little sense.
 Similarly, in context of a Java application, we have to increase heap size, to make a good use of extra memory.
-While importance of application tuning cannot be underestimated, its detailed analysis lies outside of the scope of this dissertation.
+While importance of application tuning cannot be underestimated, its detailed analysis lays outside of the scope of this dissertation.
 Figure \ref{fig:scalability-layers} presents different scalability layers and actions that can be taken at each level to improve application performance.
@@
 As outlined in previous section, horizontal scaling is about adding supplementary nodes to a system.
-As it is common to cloud computing, nodes are represented as virtual machines and this assumption is used in further discussion.
+In cloud computing, it is common to represent nodes as virtual machines and this assumption is used in further discussion.
 Consequently, adding server comes down to cloning a new virtual machine from a template and possibly installing additional software and reconfiguring it later.
@@
 Provisioning new server is only a first step in scaling an application, it is required to configure load balancing mechanism to make use of additional node.
-The two important aspects that have to be consider are: load-balancing algorithms and scalability.
+The two important aspects that have to be consider are: load-balancing algorithms and load-balancing scalability.
 
@@
 Generally, there are two types of load-balancers: hardware and software based.
-Due to the dynamic nature of system under consideration, we focus only on the latter as it offers a greater deal of flexibility.
+Due to the dynamic nature of application requirements, we focus only on the latter as it offers a greater deal of flexibility.
 Among the most common algorithms we can distinguish \cite{HaProxyDoc}:
 \begin{itemize}
- \item \textit{round-robin scheduling} - request are sent to successive nodes, according to their weights.
-This algorithm is fairest when the server's processing time remains equally distributed \cite{HaProxyDoc}
+ \item \textit{round-robin scheduling} - request are sent to successive nodes, according to their weights. This algorithm is fairest when processing time remains equally distributed \cite{HaProxyDoc}
  \item \textit{least connection} - the server with the lowest number of connections receives the connection
@@
  \item \textit{URI hashing} - URI that designates resource is hashed and divided by the total weight of the running servers.
-Such hash designates which server that receives the request.
+Such hash designates which server receives the request.
 In practice, this algorithm is commonly used with proxy caches and anti-virus proxies in order to maximize the cache hit rate.
- \item \textit{request counting algorithm} - load is distributed the requests among the various workers, ensuring that each gets their configured share of the number of requests
- \item \textit{weighted traffic counting algorithm} - variation of above-mentioned algorithm with a difference that it is focused on bytes rather than number of request
- \item \textit{pending request counting algorithm} - scheduler keeps track of how many requests each worker is assigned at present.
+ \item \textit{request counting algorithm} - load is distributed the requests among the various workers, ensuring that each gets its configured share of the number of requests
+ \item \textit{weighted traffic counting algorithm} - variation of the above-mentioned algorithm. However, it is focused on bytes rather than number of request
+ \item \textit{pending request counting algorithm} - scheduler keeps track of how many requests is assigned to each worker.
 A new request is automatically assigned to the worker with the lowest number of active requests
@@
 
-Situation gets further complicated when considering real-world web application that sends user information using cookies, what imposes requirement on load-balancer for session stickiness \cite{StBaMa11}.
+Situation gets further complicated when considering real-world web application that sends user information using cookies, what imposes on load-balancer requirement for session stickiness \cite{StBaMa11}.
 
@@
 Therefore, in cases where high availability is required, multi-tiered load balancing architecture should be considered.
-This, however, seems not to be a case among IaaS or PaaS providers - none of them unequivocally specifies whether their provide redundancy at load-balancer level.
+This, however, seems not to be a case among current IaaS or PaaS providers - none of them unequivocally specifies whether their provide redundancy at load-balancer level.
 
@@
 
-While there are many load-balancers available on the market, following are credited to be most popular:
+While there are many load-balancers available on the market, following are credited to be the most popular:
 \begin{itemize}
  \item \textbf{HAProxy} \cite{HAProxy} - load-balancer initially written by Willy Tarreau. Noticeably, it's used by OpenShift \cite{OpenShift} to distribute load among gears \cite{OpenShiftScaling}
- \item \textbf{BIG-IP Local Traffic Manager (LTM)} - solution offered by F5 \cite{F5}. Although LTM is a hardware solution, omitted in this section, it also has also its virtualized counterpart.
+ \item \textbf{BIG-IP Local Traffic Manager (LTM)} - solution offered by F5 \cite{F5}. Although LTM is a hardware solution, omitted in this section, it also has its virtualized counterpart.
  \item \textbf{Apache HTTPD} \cite{ApacheHTTPD} - popular HTTP server.
@@
 
--- optimized HTTP header analysis: headers are parsed an interpreted on the fly
+-- optimized HTTP header analysis: headers are parsed and interpreted on the fly
 
@@
 BIG-IP Local Traffic Manager \cite{LTM} &
--- managing at application services level rather than at individual devices and objects
+-- managing application services level rather than individual devices and objects
 
@@
 \subsection{Vertical scaling}
-Essentially, vertical scaling is concentrated upon increasing capacity of single node.
+Essentially, vertical scaling is concentrated upon increasing capacity of a single node.
 Again, when considering technical advancements that comes with cloud computing and virtualization, we can differ two categories of scaling: virtual machine resizing and virtual machines replacement.
-This distinction is dictated by limitation hypervisors - not all of them are able to resize virtual machine without shutting it down.
+This distinction is dictated by limitation of hypervisors - not all of them are able to resize virtual machine without shutting it down.
 
@@
 As it was highlighted in previous section, reasoning behind virtual machine replacement is that, in case when dynamic resizing is not possible, a new virtual machine with a desired configuration can be provisioned and replace the old one.
-Since this is a basic operation, all above-mentioned hypervisors supports this scenario as long as required resources are available.
+This is a basic operation and therefore all above-mentioned hypervisors support it as long as required resources are available.

### 2014-03-01 thesis/src/chapters/chapter-evaluation.tex
Thesis: add test case duration
Thesis: chapter 3

@@
 To find the equation of a function that would satisfy the previously-mentioned requirements we used \emph{polynomial interpolation}.
-As we chose 3 points which were the input to the algorithm, we got a cubic polynomial as an output.
+As we chose 4 points which were the input to the algorithm, we got a cubic polynomial as an output.
 Adding further constraints to mimic two spikes in CPU usage resulted in the function whose formula is in \eqref{eq:cpu-mock-usage} and plot in figure \ref{fig:auto-scaling-1cp-cpu-usage-function}.
@@
 Diagram \ref{fig:auto-scaling-1cp-physical-nodes} illustrates the physical setup of the test-case.
+
+\subsubsection*{Test case duration}
+The test case lasts 25 minutes as the cpu usage is constant and equals 25\% after 21 minutes and 44 seconds.
 
@@
 
+\subsubsection*{Test case duration}
+The test case lasts 25 minutes as the cpu usage is constant and equals 25\% after 21 minutes and 44 seconds.
+
 \newpage

### 2014-03-01 thesis/src/chapters/chapter-state-of-the-art.tex
Thesis: problem domain
Thesis: chapter 3

@@
 \chapterintro{This chapter presents recent achievements in a field of platform-as-a-service model by examining: Carina, OneFlow, CloudFoundry and OpenShift.
-Especially, it focuses on aspects of adaptivity, scalability and cloud federation awareness.}
+It is especially focused on aspects of adaptivity, scalability and cloud federation awareness.}
 
@@
 
+Consequently, it is expected that mature and industry-leading solution has all above-mentioned features.
 In the next sections we examine latest advancements used by a platform-as-a-service providers that is: Carina, OneFlow, CloudFoundry and OpenShift.
@@
 
-Apart from scaling policies, scaling is controlled by a minimal and maximal number of virtual machines that can compose a service.
+Apart from scaling policies, scaling is controlled by a minimal and maximal number of virtual machines that compose a service.
 
 \subsubsection{Scaling}
-\emph{OneFlow} is capable of \textbf{only} horizontal scaling: it adds or removes virtual machine using \emph{OpenNebula API}, when appropriate event is triggered.
+\emph{OneFlow} is capable of \textbf{only} horizontal scaling: it adds or removes virtual machine using \emph{OpenNebula API}.
 Violating given policy results in virtual machine's state transition into \texttt{SCALING}.
 After proper action is finished, virtual machines converges to a \texttt{COOLDOWN} state.
-What this simple state transition does is allows an environment to accommodate changes and adjust to a recent changes, disabling scaling for a given period of time.
+What this state transition does is to allow an environment to accommodate changes and adjust to a recent changes, disabling scaling for a given period of time.
 
@@
 Hence, one is limited to \emph{OpenNebula} hybrid or public cloud capabilities that employs oZones and \emph{AWS}, respectively.
-Although, they expand single \emph{OpenNebula} instance capacity it is limited in a way that discriminates it as possible solution.
+Although it expands single \emph{OpenNebula} instance capacity, it is limited in a way that discriminates it as possible solution.
 Firstly, the problem with oZones is that they require centralised management, while cloud federation is by definition distributed and independent.
@@
 \subsection{Introduction}
-\emph{OpenShift} is developed by \emph{RedHat} Platform-as-a-Service solution. It is available in two different flavours:
+\emph{OpenShift} is a developed by \emph{RedHat} Platform-as-a-Service solution. It is available in two different flavours:
 \begin{itemize}
@@
 \subsubsection{Scalability}
-As previous point implies, OpenShift \textbf{solely} uses horizontal scaling: it adds additional gears if number of requests increases beyond certain level.
+As previous point implies, OpenShift \textbf{solely} uses horizontal scaling: it adds additional gears if number of requests increases beyond a certain level.
 

### 2014-03-01 thesis/src/chapters/chapter-design.tex
Thesis: cloud-sap fixes

@@
 
-What has all of the above-mentioned features is a self-adaptive system that has a holistic approach to application scaling.
+What has all of the above-mentioned features is a self-adaptive system that has a holistic approach to an application scaling.
 As \cite{brun2009engineering} advocates, one of the first architectures of a self-adaptive system was presented in IBM's blueprint for autonomic computing \cite{IBM06}.
-Since IBM's autonomic system is widely recognised and respected model, we decided to leverage it as a foundation of our work.
+IBM's autonomic system is widely recognised and respected and hence we decided to leverage it as a foundation of our work.
 A concept of autonomic system is alluring due to its four crucial attributes:
@@
 Application providers can be both end-users or other system interfaces, Cloud-SAP is suitable for both cases.
-Cloud-SAP is entirely cloud providers and resources agnostic as long as they externalise management interfaces.
+Cloud-SAP is entirely cloud provider and resource agnostic as long as they externalise management interfaces.
 
 Successive sections define core layers of a Cloud-SAP and give insight into possible implementation decisions, while the next chapter presents a proof-of-concept implementation.
+While proposed architecture yields a self-adaptive system not an autonomic one, it is based on such model and hence we use terms characteristic to that kind of system as presented in IBM's paper \cite{IBM06}.
 
@@
 \begin{inparaenum}[1)]
-\item provides manageability interface and
-\item implements \emph{sensor} and \emph{effector} behaviour.
+\item provide manageability interface and
+\item implement \emph{sensor} and \emph{effector} behaviour.
 \end{inparaenum}
@@
 More specifically, there may be some random spikes in measurements, caused by a temporary process such as garbage collecting.
-Therefore, Cloud-SAP recommends to transform data using aggregation functions like average, median or filter it by apply low or high-pass filters.
+Therefore, Cloud-SAP recommends to transform data using aggregation functions like average, median or filter by apply low or high-pass filters.
 
-  \item[\textbf{Persistence}] Should collected data be used as a reference point during future control loop iterations, it is persisted, especially along with measures applied by an executor.
+  \item[\textbf{Persistence}] Should collected data be used as a reference point during future control loop iterations, it is persisted along with measures applied by an executor.
 As a consequence, successive assessments can be more effective by taking into account past symptoms and configuration adjustments.
@@
 
-  \item[\textbf{Standard compatibility}] Whilst there is no single, industry accepted standard of monitoring above-mentioned resources ranging from a single application to whole cloud instance, there were initiatives such as OCCI \cite{OCCI} that aims to close that gap and provide unified view of common appliances.
+  \item[\textbf{Standard compatibility}] Whilst there is no single, industry accepted standard of monitoring above-mentioned resources that range from a single application to whole cloud instance, there were initiatives such as OCCI \cite{OCCI} that aims to close that gap and provide unified view of common appliances.
 Having scaling across multiple cloud instances in mind, it is vital to provide compatibility at a monitoring level.
@@
 prediction plans.
-It is possible to incorporate more complex mechanisms, such as machine-learning, to fully use the obtained information for better prediction models.
+It is possible to incorporate more complex mechanisms (e.g.
+machine-learning) to better use the available information.
 For example one or more of the following models can be leveraged \cite{LiWoZh05}:
 \begin{asparaenum}
- \item[\textbf{Queue-based performance models}] This models aims to predict QoS measures using resource states and environmental disturbances such as number of service requests.
+ \item[\textbf{Queue-based performance models}] This models aims to predict QoS measures using resource state and environmental disturbances such as number of service requests.
 Then a configuration satisfying given QoS is found by a search technique using queues and layered queues.
- \item[\textbf{Dynamic models}] Aims to represent QoS as a cost function of the difference between the measured value from sensors and set by effectors.
+ \item[\textbf{Dynamic models}] Aims to represent QoS as a cost function of the difference between the measured value from sensors and value set by effectors.
 Model aspires to minimize that function.
@@
 Self-learning manager should also use a predictive approach \cite{JiPeLiCh11} and track changes made in a container \cite{ZhYaWo05}.
-Cloud-SAP requires to at least threshold-model \cite{tong1978threshold} to be supported to enable simple auto-scaling scenario.
+Cloud-SAP requires at least threshold-model \cite{tong1978threshold} to be supported to enable simple auto-scaling scenario.
 
@@
  \item[\textbf{Priorities}] Multi-tenancy is indispensably correlated with cloud services.
-As a consequence, resource contends again each other to access to lower-level resources.
+As a consequence, resource contends again each other to access lower-level resources.
 This can be mitigated by one of the following: \begin{inparaenum}[1)]
@@
  \item[\textbf{Reservations}] As previous point indicates, different resources may have different priorities.
-Therefore, it is suggested to guarantee that most valuable have their QoS requirements satisfied by resource reservation.
+Therefore, it is suggested to guarantee that most valuable ones have their QoS requirements satisfied by resource reservation.
 Inspired by a computer networks, there are at least to ways to retain a certain resource level:
@@
 \emph{Knowledge source} provides access to knowledge stored, for example, in a registry, dictionary or database.
-It is recommended that knowledge source share knowledge among multiple managers, consequently extending their capabilities by performing additional tasks covering recognising particular symptoms or applying specified policies, for instance.
+It is recommended that knowledge source share knowledge among multiple managers, consequently extending their capabilities by performing additional tasks: recognising particular symptoms or applying specified policies, for instance.
 
@@
 \subsection{Managed resource}
-Application execution environment is a set of tools, APIs and frameworks that provides a generalized approach to creating and running an application.
+Application execution environment is a set of tools, APIs and frameworks that provides a generalized approach to create and run an application.
 It is strictly application specific.
@@
 During monitoring phase, controller collects data from execution environment's touchpoint.
-Depending on application type it can involve different protocols and APIs.
+Depending on application type, it can involve different protocols and APIs.
 Should one monitor application execution environment, issues and concerns specified in overview are taken into consideration.
@@
 
-  \item[\textbf{Active Harmony}] This research project uses a simplex method to find the optimal application server parameter settings.
-Algorithm was validated against a TPC-W benchmark and was proven to optimise application response time by 16\% \cite{chung2004automated}.
+  \item[\textbf{Active Harmony}] This research project uses a simplex method to find the optimal application server settings.
+Algorithm was validated against TPC-W benchmark and was proven to optimise application response time by 16\% \cite{chung2004automated}.
 \end{asparaenum}
@@
 Besides, some actions interfere with each other and their execution should be planned cautiously.
-For example changing cache size and cache invalidation interfere with each other and consequently applying this change at same time may not be a good idea.
+For example changing cache size and cache invalidation time interfere with each other and consequently applying this change at same time may not be a good idea.
 
@@
 \begin{inparaenum}[1)]
- \item supervises lifecycle of a container: provisions, monitors, migrates and destroys a container
- \item modifies container its properties (i.e. increasing CPU, memory) accordingly to a given condition to ensure that service request are served with a sufficient Quality-of-Service
+ \item supervise lifecycle of a container: provisions, monitors, migrates and destroys a container
+ \item modify container's properties (i.e. increasing CPU, memory) accordingly to a given condition to ensure that service request are served with a sufficient Quality-of-Service
 \end{inparaenum}
@@
 
--- lesser overhead possible
+-- least overhead possible
 
@@
 Container's touchpoint externalises hypervisor and operating system APIs.
-Similarly to all touchpoints, sensors for gathering data and effectors for influencing it can be distinguished.
+Similarly to all touchpoints, sensors and effectors can be distinguished.
 It is highly recommended for them to be linked, forming together manageability capabilities.
@@
 \emph{Autonomic container controller} aims to automate container's management function and externalise its configuration through its interfaces.
-In order to do achieve that, it implements a control loop that fully covers container life cycle.
+In order to achieve that, it implements a control loop that fully covers container life cycle.
 It is suggested for a controller to fully support self-management by implementing \emph{self-configuring, self-healing, self-optimising, self-healing} control loops, however, \emph{self-configuring} is the only one required.
-Modular structure of a controller is designated by its main four functions: data monitoring, data analysis, action planning and execution along with a knowledge necessary to perform these operations as shown in figure \ref{img:autonomic-component}.
+Modular structure of a controller is designated by its main four functions: data monitoring, data analysis, action planning and execution along with a knowledge necessary to perform these operations as shown in figure \ref{img:csap-manager}.
 
 \subsubsection{Monitoring}
-During monitoring phase, controller collects data from container's touchpoint. Issues such as data filtering, aggregation should be taken into consideration as it was stated in overview.
+During monitoring phase, controller collects data from a container's touchpoint. Issues such as data filtering, aggregation should be taken into consideration as it was stated in overview.
 
 \subsubsection{Analysis}
-Analysis should incorporate prediction techniques and mathematical models specified in overview, producing change requests optimising container behaviour.
+Analysis should incorporate prediction techniques and mathematical models specified in overview, producing change requests that optimises container behaviour.
 
 \subsubsection{Planning}
-This part schedules change requests submitted by an analysis model producing change procedures.
+This part schedules change requests submitted by an analysis model and produces a change procedure.
 Depending on implementation, it may use reservation and priorities mechanisms to enforce given Quality-of-Service level.
-It vital for a planning phase to be well suited for an underlying hypervisor since not all its types supports dynamic vertical scaling some requires container to be restarted.
-In that case, container should be cloned and then resized, restarted and finally substitute old, non-scaled instance.
+It is vital for a planning phase to be well suited for an underlying hypervisor since one supports dynamic vertical scaling whereas other require container to be restarted.
+In that case, container should be cloned and then resized, restarted and finally substitute an old, non-scaled instance.
 
@@
 \end{itemize}
-Not every hypervisor supports dynamic vertical scaling, some requires a container to be restared. Noticeably, lightweight containers are flexible in th
-
+Not every hypervisor supports dynamic vertical scaling, some requires a container to be restared.
+Noticeably, lightweight containers are flexible in that matter.
 In case of a self-learning system, change plan and its effects should be recorded to serve as a reference point in future.
@@
 \begin{inparaenum}[a)]
- \item monitoring its containers
+ \item monitoring containers
  \item horizontal scaling
@@
 As it was stated in an overview, a stack is a group of containers configured to serve a common purpose (i.e.
-as a server cluster).
+serve as a server cluster).
 In practise, this group consists of a master node and slaves: loadbalancer and java application workers or a master and slave databases, for example.
 Each stack type may leverage different mechanisms to enable that master-slave relation.
-Although specific decisions are left to an implementation, chapter X summarises common load-balancing mechanisms.
 
 \subsubsection{Touchpoint}
-Stack's touchpoint aims to provide information about the state and condition of a whole group of nodes: master as well as its slaves.
-As a consequence, from a structural perspective, it is decentralise and consists of a multiple agents.
+Stack's touchpoint aims to provide information about the state and condition of all nodes.
+As a consequence, from a structural perspective, it is decentralised and consists of a multiple agents.
 Key properties that should be manageable by a touch point include:
@@
  \item average cpu / memory usage in cluster
- \item packet loss from all nodes
+ \item packet loss
 \end{itemize}
@@
 During monitoring phase, controller gathers data from a multiple agents located in stacks' nodes: master and slaves.
-It is important to ensure that received data is synchronised in time, consequently it can be analysed collectively to find a symptoms covering all of the nodes.
+It is important to ensure that received data is synchronised in time, consequently it can be analysed collectively to find a symptoms covering all nodes.
 
@@
  \item changing load-balancing algorithm
- \item changing network-topology to optimize time needed to reach a stack
+ \item changing network topology to optimize time needed to reach a stack
  \item replicating master node and hence eliminating SPOF
@@
 \section{Summary}
-Cloud-SAP fills the gap in existing platforms by providing a standardised approach to scaling applications using a fine-grained actions and operating on different platform levels: starting from execution environment and ending on a cloud federation.
+Cloud-SAP fills the gap in existing platforms by providing a standardised approach to scaling applications using a fine-grained actions and operating on different platform levels: starting from an execution environment and ending on a cloud federation.
 It is based on a concept of an autonomic system, providing methodical but flexible foundation for its implementation.
 What is more, Cloud-SAP leverages core benefits of such a system: promise of a self-configuration, self-optimisation, self-health, self-protection.
-It is driven by a customer-defined policies and consequently ensures that customer is served with demanded Quality-of-Service and with optimised costs.
-
-% TODO Further work?, wady
+It is driven by a customer-defined policies and consequently ensures that customer is served with demanded Quality-of-Service and optimised costs.

### 2014-03-01 thesis/src/chapters/chapter-evaluation.tex
Thesis: summary, eval fixes

@@
 \section{Introduction}
-We carried out a number of tests which aim to prove the solution be better than currently available, especially in terms of:
+We carried out a number of tests which aim to prove the solution to be better than currently available, especially in terms of:
 \begin{enumerate}
@@
 \subsection*{Hardware and virtual machines configuration}
-Many of the carried out tests were run on the same hardware and/or used the same virtual machines so for clarity of presentation their parameters are presented not in a description of every test, but here in table \ref{tab:test-deployment-time-common-hardware-configuration}.
+All tests that were carried out were run on the same hardware and/or used the same virtual machines so for clarity of presentation their parameters are presented not in a description of every test, but here in table \ref{tab:test-deployment-time-common-hardware-configuration}.
 
@@
   \centering
-  \begin{tabular}{ | l | l | l | l | }
-    \hline                        
-    Name & CPU & RAM & HD \\
-    \hline
-    (virtual machine) carina-frontend  & 1 CPU & 1GB & 6GB \\
+  \begin{tabular}{  l  l  l  l  }
+    \specialrule{.1em}{.05em}{.05em}
+
+    \textbf{Name} & \textbf{CPU} & \textbf{RAM} & \textbf{HD} \\
+    \specialrule{.1em}{.05em}{.05em}
+
+    (virtual machine) carina-frontend  & 1 CPU & 1GB & 6GB \\ \hline
     (virtual machine) carina-management-vm  & 1 CPU & 1GB & 9GB \\
@@
 
-\subsubsection{Motivation} This test case aims to prove that scaling across multiple cloud providers is vitally important while ensuring appropriate Quality-of-Service, especially in cases of increased number of service requests.
+\subsubsection{Motivation} This test case aims to prove that scaling across multiple cloud providers is vitally important when ensuring appropriate Quality-of-Service, especially in cases of increased number of service requests.
 Such scaling scenario, known also as cloud-bursting, leverage benefits arising from offloading application load to an external provider.
@@
 While our proof of concept solution features multi-layer scaling and is cloud federation aware, \emph{Carina} adopts only horizontal scaling.
-It is expected that test case prove benefits emerging from enriching application platform provider with an cloud federation awareness.
+It is expected that test case prove benefits emerging from enriching application platform provider with a cloud federation awareness.
 
@@
 \begin{enumerate}
-\item deploy exemplary service
+\item deploy an exemplary service
 \item observe system behaviour under load, simulated by a resource usage mock. It is vital for a resource usage to exceed a single provider capabilities
@@
 However, at some point single cloud provider resources will be surpassed.
-Equation \eqref{eq:cpu-mock-usage-2cp} denotes a resource usage function that is expected to fulfil above-mentioned scenario and is illustrated in figure \ref{fig:auto-scaling-2cp-cpu-usage-function}.
+Equation \eqref{eq:cpu-mock-usage-2cp} denotes a resource usage function that is expected to fulfil above-mentioned conditions and is illustrated in figure \ref{fig:auto-scaling-2cp-cpu-usage-function}.
 \begin{equation}
@@
 
-  \item[\textbf{Deployment diagrams}] Components that took part during testing are show in figures \ref{fig:autoscale-2cp-cloud-sap-deployment-diagram} and \ref{fig:auto-scaling-2cp-carina-deployment-diagram}.
+  \item[\textbf{Deployment diagrams}] Components that took part during test are show in figures \ref{fig:autoscale-2cp-cloud-sap-deployment-diagram} and \ref{fig:auto-scaling-2cp-carina-deployment-diagram}.
 
@@
 \subsubsection*{Test case duration}
-The test case lasts 25 minutes as the cpu usage is constant and equals 25\% after 21 minutes and 44 seconds.
+The test case lasts 25 minutes.
 
@@
   \item[\textbf{Cloud-SAP}] 
-   In total, there were 13 vertical (increasing CPU limit) 2 horizontal (adding a container) scaling events.
+   In total, there were 13 vertical and 2 horizontal scaling events.
 First action took place minute after first violation of scaling policy rule.
@@
   Similarly to a \emph{Cloud-SAP}, first scaling event (slave addition) was noticed in 6th minute of test.
-Scaling jobs were continuously invoked every 2 minutes until capacity of a \emph{CP\-1} hasn been fully exploited.
+Scaling jobs were continuously invoked every 2 minutes until capacity of a \emph{CP\-1} has not been fully exploited.
 Listing \ref{lst:auto-scaling-2cp-results-carina-jobs} summarises jobs performed by \emph{Carina}.

### 2014-03-01 thesis/src/chapters/chapter-summary.tex
Thesis: summary, eval fixes

@@
 
-\chapterintro{This chapter provides the summery of the dissertation and gives information about future work and research directions.}
+\chapterintro{This chapter provides the summary of the dissertation and gives information about future work and research directions.}
 
@@
 In this dissertation we proposed an architecture of a platform that ensures satisfying Quality of Service requirements of users' applications deployed in a cloud computing environment.
-The key attribute of the proposed system is holistic approach to self-adaptability achieved by insightful understanding of platform-as-a-service model.
+The key attribute of the proposed system is holistic approach to a self-adaptability achieved by an insightful understanding of a platform-as-a-service model.
 As a consequence, designed architecture manages variety of resources that users' applications incorporate leading to a system that is capable of vertical, horizontal and most importantly cloud-federation aware scaling.
@@
 All tests of the platform were not simulated on any software, but conducted on the real hardware forming a home-made data center.
-Tests were chosen and design so as to reflect at the highest possible rate real business use cases.
+Tests were chosen so as to reflect at the highest possible rate real business use cases.
 Their results show considerable potential of the proposed solution.

### 2014-03-01 thesis/src/chapters/chapter-implementation.tex
Thesis: impl fixes

@@
 High level overview of system, including its internal and external elements is depicted in figure \ref{fig:hlo-implementation}.
-Diagram \ref{fig:csap-layers-subsystem} illustrates thee relationship between Cloud-SAP managers and above-mentioned subsystems.
+Diagram \ref{fig:csap-layers-subsystem} illustrates the relationship between Cloud-SAP managers and above-mentioned subsystems.
 
@@
   \end{center}
-  \caption{System parts and their relation with autonomic managers}
+  \caption{System parts and their relation with Cloud-SAP's autonomic managers}
   \label{fig:csap-layers-subsystem}
@@
 \begin{asparaenum}
- \item[\textbf{Auto-scaling subsystem}] Auto-scaling subsystem supervises service's life cycle, monitoring it and enacting scaling actions when necessary.
-It is concentrated on all service's aspects: containers, stacks and application instances thus it is in fact a group of autonomic managers composed by container, stack and cloud instance managers.
+ \item[\textbf{Auto-scaling subsystem}] Auto-scaling subsystem supervises service's life cycle, monitors it and enacts scaling actions when necessary.
+It is concentrated on all service's aspects: containers, stacks and application instances thus it is in fact a group of managers composed by container, stack and cloud instance managers.
 
@@
    \item Cloud broker - probes cloud providers and selects best offer according to a given policy.
-In Cloud-SAP model, it plays a role of orchestrating autonomic manager, that is, autonomic cloud federation manager.
+In Cloud-SAP model, it plays a role of orchestrating autonomic manager, that is, a cloud federation manager.
    \item Cloud client - delegates service provisioning request to a cloud broker.
@@
 
- \item[\textbf{Cloud provider}] External system that manages a group of resource such as computing nodes, storage and networking typologies.
+ \item[\textbf{Cloud provider}] External system that manages a group of resources such as computing nodes, storage and network topologies.
 Particularly, it is able to deploy, shutdown, migrate and monitor containers.
@@
 
-As overview states, this subsystem is in fact a loosely coupled group of three autonomic managers, cooperating together to achieve aforementioned subsystem's goals.
+As overview states, this subsystem is in fact a loosely coupled group of three managers, cooperating together to achieve aforementioned subsystem's goals.
 System is structured (figure \ref{fig:auto-scaling-subsystem-deployment}) on top of that observation.
@@
 
- \item[\textbf{Cloud broker}] Is a part of cloud brokerage subsystem responsible for mediating between different auto-scaling subsystems and cloud client. Previous section covers it in more details.
+ \item[\textbf{Cloud broker}] It is a part of cloud brokerage subsystem responsible for mediating between different auto-scaling subsystems and cloud client.
+Previous section covers it in more details.
 
@@
 In case of this implementation it is an OpenNebula frontend and AppFlow server, accessible through a Cloud provider client.
-Apart from that, OpenVZ exposes its manageability capabilities by OpenVZ agent and its REST interfaces.
+Apart from that, OpenVZ exposes its manageability capabilities by an OpenVZ agent and its REST interfaces.
 Next section details cloud provider.
@@
 Noticeably, it may be a facade to different systems, freeing subsystem from accessing cloud provider's internal structure as it is in case of our implementation.
-For example, it delegates service provisioning requests to an Applfow server while monitoring and scaling are passed to an OpenNebula frontend.
+For example, it delegates service provisioning requests to an Appflow server, while monitoring and scaling are passed to an OpenNebula frontend.
 \end{asparaenum}
@@
 Policies (e.g.
-listing ref{lst:service-spec-test-deployment-time}) are based on a threshold model, denoting a valid range for a resource measurement.
-Hence, its evaluation results in \texttt{FITS, LESSER, GREATER}.
+listing \ref{lst:service-spec-test-deployment-time}) are based on a threshold model, denoting a valid range for a resource measurement.
+Hence, depending whether value is in range, is lesser than minimal or greater than maximal it evaluates to: \texttt{FITS, LESSER, GREATER} respectively.
 
@@
 Similarly to a container manager, it attempts to reserve resources needed to implement desired change.
-In case when it is not possible, it problem is further delegated - to a cloud instance manager.
+In case when it is not possible, problem is further delegated to a cloud instance manager.
 
- \item execution - Manager horizontally scales container invoking OpenNebula frontend API through cloud provider client.
+ \item execution - Manager horizontally scales container by invoking OpenNebula frontend API through cloud provider client.
 Noticeably, currently only adding and removing slave instances is possible.
-Hence, attempts to re-scale master instance (i.e.
-load balancer) are not supported.
+Hence, attempt to re-scale master instance (i.e.
+load balancer) is not supported.
 \end{itemize}
@@
   \item stacks's resource are reserved apriori
-  \item cloud provider client is leveraged to create service's AppFlow template and then instantiating it.
+  \item cloud provider client is leveraged to create service's AppFlow template and then instantiate it.
  \end{inparaenum}
@@
 Diagram \ref{fig:case-deployment-seq} depicts sequence of steps involved in service deployment.
-Such request is triggered by a cloud client and involves multiple cloud provider.
-Consequently, cloud broker selects best offer for its client.
+Such request is triggered by a cloud client and involves multiple cloud providers.
+Cloud broker selects best offer for its client.
 
@@
 This is, however, not possible and hence stack container checks if there are enough resources for a horizontal scaling.
-As this is not the case, request is passed to cloud instance manager and then cloud broker to deploy container on another cloud.
-This scenario is also known as a cloud bursting.
+As this is not the case, request is passed to a cloud instance manager and then cloud broker deploys container on another cloud.
+This scenario is known as a cloud bursting.
 
@@
 Although, it supports solely OpenNebula, it remains open for any cloud provider.
-Similarly, platform currently supports vertical scaling of CPU, horizontal scaling and scaling across multiple cloud provider, however, it is expected that next versions support wider range of actions (i.e.
+Similarly, platform currently supports vertical scaling of CPU, horizontal scaling and scaling across multiple cloud provider, however, it is expected that successive versions will support wider range of actions (i.e.
 vertical scaling of memory, application platform tuning).

### 2014-03-02 thesis/src/chapters/chapter-state-of-the-art.tex
Thesis: provider comparision

@@
 \section{Providers comparison}
-In this section we would like to compare all depicted providers as well as other solution in terms of \emph{adaptivity} and \emph{scalability} as none of the examined providers support \emph{cloud federated} deployments.
-Table \ref{tab:cloud-providers-scaling} presents a summary of cloud providers auto-scaling capabilities.
-Interestingly, all of them are focused solely on horizontal scaling, ignoring advantages offered by a fine-grained approach to scaling that leverage scaling up and application tuning.
-Table \ref{tab:cloud-providers-adaptivity} summarises platforms approach to adaptivity: their data analysis mechanisms and policies support.
+In this section we would like to compare all depicted providers in terms of fulfilling functional requirements specified in the overview.
+Table \ref{tab:cloud-providers-adaptivity} takes more detailed approach and summarises how platform tackles adaptivity: their data analysis mechanisms and policies support.
 
 \begin{table}[!htbp]
-\begin{tabularx}{\textwidth}[]{ X  X  X  X}
+\begin{tabularx}{\textwidth}[]{ l  X  X  X  X }
+\specialrule{.1em}{.05em}{.05em} 
+& \textbf{Carina} & \textbf{OneFlow} & \textbf{CloudFoundry} & \textbf{OpenShift} \\
 \specialrule{.1em}{.05em}{.05em}
 
-  & \textbf{Horizontal scaling} & \textbf{Vertical scaling} & \textbf{Application tuning} \\
-\specialrule{.1em}{.05em}{.05em}
+\textbf{Service operations} &  &  &  \\ \specialrule{.1em}{.05em}{.05em} 
+Deploy a service & \checkmark  & \checkmark & \checkmark & \checkmark\\ \hline
+Cancel the service & \checkmark & \checkmark & \checkmark & \checkmark \\ \specialrule{.1em}{.05em}{.05em}
 
-\multicolumn{4}{ l }{\textbf{Infrastructure provider}} \\
-\specialrule{.1em}{.05em}{.05em}
+\textbf{Status check} &  &  &  \\ \specialrule{.1em}{.05em}{.05em} 
+Deployment success & \checkmark & \checkmark &  \checkmark & \checkmark \\ \hline
+Uptime & \checkmark & \checkmark & \checkmark & \checkmark \\ \hline
+Current cost & $\times$ & $\times$ & $\times$ & $\times$ \\ \specialrule{.1em}{.05em}{.05em}
 
-Carina & \checkmark & $\times$ & $\times$ \\ \hline
+\textbf{Client application} &  &  &  \\ \specialrule{.1em}{.05em}{.05em} 
+CLI & \checkmark & \checkmark & \checkmark & \checkmark \\ \hline
+Web service & $\times$ & $\times$ & \checkmark & \checkmark \\ \specialrule{.1em}{.05em}{.05em}
 
-OneFlow & \checkmark & $\times$ & $\times$ \\ \hline
+\textbf{Deployment descriptor} &  &  &  \\ \specialrule{.1em}{.05em}{.05em} 
+Service name & \checkmark & \checkmark & \checkmark & \checkmark \\ \hline
+Technology stack & \checkmark & \checkmark & \checkmark & \checkmark \\ \hline
+Scaling policies & \checkmark & \checkmark & $\times$ & $\times$ \\ \specialrule{.1em}{.05em}{.05em}
 
-AWS EC2 & \checkmark & $\times$ & $\times$ \\ \hline
+\textbf{Deployment cost minimisation} & $\times$ & $\times$ & $\times$ & $\times$ \\ \specialrule{.1em}{.05em}{.05em}
 
-\multicolumn{4}{ l }{\textbf{Platform provider}} \\
-\specialrule{.1em}{.05em}{.05em}
+\textbf{Self-adaptation and scaling} &  &  &  \\ \specialrule{.1em}{.05em}{.05em} 
+Application platform tuning & $\times$ & $\times$ & $\times$ & $\times$ \\ \hline
+Vertical scaling & $\times$ & $\times$ & $\times$ & $\times$ \\ \hline
+Horizontal scaling & \checkmark & \checkmark & \checkmark & \checkmark \\ \hline
+Multiple provider based scaling & $\times$ & $\times$ & $\times$ & $\times$ \\ \specialrule{.1em}{.05em}{.05em}
 
-CloudFoundry & $\times$ & $\times$ & $\times$ \\ \hline
-
-OpenShift & \checkmark & $\times$ & $\times$ \\ \hline
-
-AppEngine & \checkmark & $\times$ & $\times$ \\ \hline
-
-Azure & \checkmark & $\times$ & $\times$ \\ \hline
-
-Heroku & $\times$ & $\times$ & $\times$ \\ \hline
 \end{tabularx}
 
-\caption{Comparison of cloud providers scaling capabilities}
-\label{tab:cloud-providers-scaling}
+\caption{Comparison of cloud providers features}
+\label{tab:cloud-providers-comparision-features}
 
@@
 
-AWS EC2 & 
--- time frame based
-
--- expression based, where expressions corresponds to a AutoScalingGroup
-
--- actions are triggered by a CloudWatch alarms
-
--- customizable adjustments paddings, types, cooldown time
-&
--- threshold model, takes into account minimal and maximal permitted instances of an application as well as application priority
-\\ \hline
-
 \multicolumn{3}{ l }{\textbf{Platform provider}} \\
@@
 
-AppEngine &
-
--- built-in policy based on request queue length
-
--- adjustable minimal, maximal number of application instances, pending latency
-
- &
--- queue-based, new instance is provisioned if queue length got too long 
-  \\ \hline
-
-Azure & 
--- time frame based
-
--- expression based, where expression can involve either CPU usage or Queue length
-
--- customizable adjustments paddings, types, cooldown time
-&
--- threshold model, that takes into account minimal and maximal allowed instances
- \\ \hline
-
-Heroku & $\times$ & $\times$ \\ \hline
 \end{tabularx}

### 2014-03-02 thesis/src/chapters/chapter-adaptivity.tex
Thesis: further adjustments

@@
 
-In the heart of the every self-adaptive systems, according to \cite{brun2009engineering}, lies a feedback loop, concept that originates from a control engineering. As paper's authors advocate:
+In the heart of the every self-adaptive system, according to \cite{brun2009engineering}, lies a feedback loop, concept that originates from a control engineering. As paper's authors advocate:
 \begin{quote}
@@
 
-Currently, service providers employs only simplified models such as threshold model \cite{LiWoZh05}, that defines a valid range for a given metric.
+Currently, service providers employ only simplified models such as threshold model \cite{LiWoZh05}, that defines a valid range for a given metric.
 In cases when it is violated (i.e.
-value is either smaller than minimal or bigger than maximal acceptable) corresponding resource is properly adjusted - figure \ref{fig:threshold-model} illustrates that idea.
+the value is either smaller than minimal or bigger than maximal acceptable) the corresponding resource is properly adjusted - figure \ref{fig:threshold-model} illustrates that idea.
 
@@
 
-Threshold model, while trivial in its form, has been proved useful in a real-world scenarios as it is in case of AWS, OpenShift, Carina or OneCloud.
+Threshold model, while trivial in its form, has been proved useful in real-world scenarios as it is in case of AWS, OpenShift, Carina or OneCloud.
 
 As it was already mentioned, models are often controlled by policies that denote a condition which, when satisfied, triggers an action harnessing environment's disturbance.
-Currently, industry leaders supports \cite{AmazonAutoScaling} two kind of policies:
+Currently, industry leaders support \cite{AmazonAutoScaling} two kind of policies:
+
 \begin{itemize}

### 2014-03-02 thesis/src/chapters/chapter-design.tex
Thesis: cloud-sap
Thesis: further adjustments

@@
 \section{Motivation}
-As one can see, latest advances in providing platform-as-a-service still do not yield a comprehensive solution that embraces a range of fine-grained actions relevant to a given situation in an environment, as it was summarised in previous chapter.
+As one can see, latest advances in providing platform-as-a-service still do not yield a comprehensive solution that embraces a range of fine-grained actions relevant to a given situation in an environment, as it was summarised in the previous chapter.
 All of the presented solutions solely engage horizontal scaling, while other IT processes and corresponding actions such as vertical scaling, cloud bursting or application platform restart remains neglected.
@@
 Successive sections define core layers of a Cloud-SAP and give insight into possible implementation decisions, while the next chapter presents a proof-of-concept implementation.
-While proposed architecture yields a self-adaptive system not an autonomic one, it is based on such model and hence we use terms characteristic to that kind of system as presented in IBM's paper \cite{IBM06}.
+While the proposed architecture yields a self-adaptive system not an autonomic one, it is based on such model and hence we use terms characteristic to that kind of system as presented in IBM's paper \cite{IBM06}.
+This is due to the fact that the proposed architecture does not operate on high level business policies and consequently requires human intervention when changing pricing policies.
 
@@
 
-Before the more detailed diagram of the solution is presented, it is essential to introduce another indispensable function that lies in the heart of any autonomic entity -- its ability to detect an undesirable state and take appropriate actions.
+Before the more detailed diagram of the solution is presented, it is essential to introduce another indispensable function that lies in the heart of any self-adaptive entity -- its ability to detect an undesirable state and take appropriate actions.
 This is achieved by the presence of a \emph{control loop} which collects data about the environment, analyzes it and takes all needed steps to recover the system.
-Not only can the attributes introduced in the previous section be bound to an autonomic system, but they can also be thought of main and broad categories of a control loop \cite{IBM06}.
+Not only can the attributes introduced in the previous section be bound to a self-adaptive system, but they can also be thought of main and broad categories of a control loop \cite{IBM06}.
 

### 2014-03-02 thesis/src/chapters/chapter-introduction.tex
Thesis: add thesis structure
Thesis: further adjustments

@@
 Consequently, it is expected that adaptation at a service provider level is also possible, i.e.
-provider can offload some traffic to a different provider, as long as it satisfies a customer.
+provider can offload some traffic to a different provider, as long as it satisfies customer's needs.
 
-While self-adaptive systems have a long history \cite{Mu04}, it have not been directly applied to a multi-layered problem that exists in a cloud computing environment.
+While self-adaptive systems have a long history \cite{Mu04}, the concept of self-adaption has not been directly applied to a multi-layered problem that exists in a cloud computing environment.
 Especially, the research area at the last layer, which sizes across different service providers, is new.
-Although, architecture known as InterCloud \cite{BuRaCa10} investigates problem of cooperation and negotiation at a cloud level, it neither has been implemented nor presented in a context of self-adaptive system.
+Although, architecture known as InterCloud \cite{BuRaCa10} investigates problem of cooperation and negotiation at a cloud level, it neither has been implemented nor presented in a context of a self-adaptive system.
 
@@
 This data suggests that the subject is attractive for IT industry from the economic point of view.
-However, higher amounts of money spent on cloud services involve higher expectations of theirs quality from customers.
+However, higher amounts of money spent on cloud services involve higher expectations of their quality from customers.
 Although the most significant players in cloud computing have been in the field for quite a long time, it is still possible to outline some deficiencies their products have.
@@
 Amazon EC2, when compared to other companies especially in terms of autoscaling capabilities, really shines.
-OpenShift, RedHat PaaS solution, ensures application scaling but with very limited possibilities of customisation of the process -- the user can only choose if their application should scale and the whole algorithm is solely based on the number of concurrent requests to the application.
+OpenShift, RedHat PaaS solution ensures application scaling, but with very limited possibilities of customisation of the process -- the user can only choose if their application should scale and the whole algorithm is solely based on the number of concurrent requests to the application.
 Users of Heroku, another PaaS solution, have no automation tool that would control the number of instances (\emph{dynos} in Heroku nomenclature) their application is running on -- they can change it manually.

### 2014-03-02 thesis/src/chapters/chapter-state-of-the-art.tex
Thesis: further adjustments

@@
 \section{Requirements}
-One can notice that elements that yields a solution to a problem stated in the first chapter, which is ensuring that users' application provide appropriate Quality-of-Service for its customers in a most-cost effective manner, were gradually introduced in previous chapter:
+One can notice that elements that yield a solution to a problem stated in the first chapter, which is ensuring that users' application provide appropriate Quality-of-Service for its customers in a most-cost effective manner, were gradually introduced in the previous chapter:
 
@@
 Consequently, it is expected that mature and industry-leading solution has all above-mentioned features.
-In the next sections we examine latest advancements used by a platform-as-a-service providers that is: Carina, OneFlow, CloudFoundry and OpenShift.
+In the next sections we examine latest advancements used by a platform-as-a-service providers: Carina, OneFlow, CloudFoundry and OpenShift.
 Apart from getting into detail regarding above-mentioned factors, we scrutinise whether given product satisfies following functional and non-functional requirements.

### 2014-03-02 thesis/src/chapters/chapter-scalability.tex
Thesis: push unmerged change

@@
 Although, all of the aforementioned aspects are vital for any application, our work focuses solely on the first type of scalability.
-The reasoning behind this is that while all of these scalability types lies in direct responsibility of an application developer, the load scalability can be additionally improved by adding additional resources to a system.
-This brings us to a question what kind of resources are used by an application or more appropriately in context of this dissertation: \textit{what kind of resources can we add to improve application performance?} Required resources vary from application to application.
-However, among the most common one we can distinguish:
+The reasoning behind this statement is that, while all of these scalability types lie in direct responsibility of an application developer, the load scalability can be additionally improved by adding additional resources to a system.
+This brings us to a question what kind of resources are used by an application or more appropriately in context of this dissertation: \textit{what kind of resources can we add to improve application performance?} Required resources varies from an application to an application.
+However, among the most common ones we can distinguish:
 \begin{itemize}
@@
 
-Provisioning new server is only a first step in scaling an application, it is required to configure load balancing mechanism to make use of additional node.
-The two important aspects that have to be consider are: load-balancing algorithms and load-balancing scalability.
+Provisioning a new server is only the first step in scaling an application, it is required to configure load balancing mechanism to make use of additional node.
+Two important aspects that have to be consider are: load-balancing algorithms and load-balancing scalability.
 

### 2014-03-06 thesis/src/abstract.tex
Thesis: intro, abstract II

@@
 they support sudden spikes in demand followed by underutilization at other times.
-An architecture of a cloud computing system, that meets these requirements, has a characteristic of a multi-layered self-adaptive system, where different layers corresponds to different levels where cloud operates: starting from an application layer, through the application platform, infrastructure to end on a cloud instance.
+An architecture of a cloud computing system that meets these requirements has a characteristic of a multi-layered self-adaptive system, where different layers corresponds to different levels where cloud operates: starting from an application layer, through the application platform, infrastructure to end on a cloud instance.
 

### 2014-03-06 thesis/src/chapters/chapter-introduction.tex
Thesis: intro, abstract II

@@
 One of the keys factors that has driven transformation of computing industry in the last years is the perception of computing utilities as a commodity\cite{BuYeVeBrBr09}, easily accessible and adjustable to specific needs.
-As a consequence, we observe a profusion of different services, often collectively referred as a cloud computing \cite{MeGr11}.
+As a consequence, we observe a profusion of different services, often collectively referred to as a cloud computing \cite{MeGr11}.
 Similarly to services known from traditional markets, customers expect them to be accessible on demand and in easy manner, while paying only for the consumed goods.
@@
 
-Having customer requirements in mind, it is crucial that a service provider is able to adapt itself to fulfil customer's contract.
+Having customer requirements in mind, it is crucial that a service provider adapt itself to fulfil customer's contract.
 For example, such adaptation can be triggered by a sudden spike in resource demand and may result in provisioning additional application platforms.
-However, due to the problem complexity and number of concerns involved, there are different levels where adaptation is possible:
+However, due to the problem complexity, there are different levels where adaptation is possible:
 \begin{itemize}
@@
 
-While self-adaptive systems have a long history \cite{Mu04}, the concept of self-adaption has not been directly applied to a multi-layered problem that exists in a cloud computing environment.
+While self-adaptive systems have a long history, the concept of self-adaption has not been directly applied to a cloud computing environment, characterised by its multi-layered structure.
 Especially, the research area at the last layer, which sizes across different service providers, is new.
-Although, architecture known as InterCloud \cite{BuRaCa10} investigates problem of cooperation and negotiation at a cloud level, it neither has been implemented nor presented in a context of a self-adaptive system.
+Although architecture known as InterCloud \cite{BuRaCa10} investigates problem of cooperation and negotiation at a cloud level, it neither has been implemented nor presented in a context of a self-adaptive system.
 
@@
 This data suggests that the subject is attractive for IT industry from the economic point of view.
-However, higher amounts of money spent on cloud services involve higher expectations of their quality from customers.
-Although the most significant players in cloud computing have been in the field for quite a long time, it is still possible to outline some deficiencies their products have.
+Higher amounts of money spent on cloud services involve higher expectations of their quality from customers.
+Although the most significant players in cloud computing have been in the field for quite a long time, it is still possible to outline some deficiencies theirs products have.
 Additionally, lack of common standards hinder cooperation among different cloud providers.
-For example it is nearly impossible to create an autoscaling cloud federation with Amazon Web Services (current leader in providing cloud services\cite{GartnerMagicQuadrantSep2013}) and another provider.
-Amazon EC2, when compared to other companies especially in terms of autoscaling capabilities, really shines.
+For example, it is nearly impossible to create an autoscaling cloud federation with Amazon Web Services (current leader in providing cloud services\cite{GartnerMagicQuadrantSep2013}) and another provider.
+Having that said, Amazon EC2, when compared to other companies especially in terms of autoscaling capabilities, really shines.
 OpenShift, RedHat PaaS solution ensures application scaling, but with very limited possibilities of customisation of the process -- the user can only choose if their application should scale and the whole algorithm is solely based on the number of concurrent requests to the application.
@@
 
-The proposed solution in this dissertation tries to deal with the aforementioned providers problems by outlining an example architecture that enables seamless cooperation among cloud providers and provides auto-scaling capabilities.
+The proposed solution in this dissertation tries to deal with the aforementioned providers problems by outlining an exemplary architecture that enables seamless cooperation among cloud providers and provides auto-scaling capabilities.
 
@@
 \section{Impact}
-We hope that the concept of self-adaptation at each level of cloud can be thought-provoking for computing scientists.
+We hope that the concept of self-adaptation in cloud will be thought-provoking for computing scientists.
 What is more, we believe that our successful attempt to implement the proposed architecture will cause a greater gain in interest in models that enable scaling applications across different cloud providers, e.g.
 \emph{InterCloud}.
-Finally, we consider the ideas contained in this work be beneficial to the OpenNebula ecosystem as they provide insights into the ways Quality of Service can be assured:
+Finally, we consider the ideas contained in this work to be beneficial to the OpenNebula ecosystem as they provide insights into the ways how Quality of Service can be assured by:
 
@@
     \setcounter{enumi}{1}
-  \item presents an introduction to the problem domain giving an outline and definition of concepts of \emph{self-adaptive systems}, \emph{application scaling} and \emph{interoperability of clouds}
+  \item presents an introduction to the problem domain giving an outline and definition of concepts of \emph{self-adaptive system}, \emph{application scaling} and \emph{interoperability of clouds}
   \item describes and compares similar solutions in the field with the emphasis on their auto-scaling capabilities
@@
   \item provides information regarding tests of the implementation and their results
-  \item presents conclusions and points further research directions
+  \item presents conclusions and points out further research directions
 \end{enumerate}

### 2014-03-06 thesis/src/chapters/chapter-interoperability.tex
Thesis: domain

@@
 From the perspective of a user of PaaS services it is vital that they are able to deploy seamlessly their applications using libraries, tools and services supported by the cloud provider\cite{MeGr11}.
-Judging by such factors as the popularity of Heroku -- currently one of the most popular PaaS provider which does not offer more advanced features which enable management of the infrastructure underpinning the deployment platform, the fact that Microsoft added auto-scaling to its Azure platform as late as in June 2013, it is perfectly possible most PaaS users are satisfied with the current offers of their providers and do not need another, more sophisticated functionalities.
+Judging by such factors as the popularity of Heroku, which does not offer more advanced features which enable management of the infrastructure underpinning the deployment platform, the fact that Microsoft added auto-scaling to its Azure platform as late as in June 2013, it is perfectly possible most PaaS users are satisfied with the current providers offerings and do not need other, more sophisticated functionalities.
 However, there are more complex applications and systems whose requirements regarding technology stack, availability and scalability are considerably more demanding.
-For such services there ought to be designed slightly specialized features that would require cooperation among different cloud providers.
+For such services there ought to be designed specialised features that would require cooperation among different cloud providers.
 
@@
 One can imagine scenarios in which customers of cloud services know their applications are vulnerable to sudden variations in demand and their responsiveness must be kept at the same level all the time.
-In such cases, they want them to scale dynamically according to current load or other predefined or manually specified metrics.
+In such cases, they want them to scale dynamically according to current load or other specified metrics.
 What is more, in order to ensure high availability of their services, customers do not want to confine themselves to only one provider -- in the best scenario they want their applications (or their logical parts, such as persistence layer) to be spanned across different providers and be able to cooperate with one another at the same time.
@@
 \subsubsection{Deployment models}
-The informal introduction to the concept of a \emph{hybrid cloud} in the previous section requires a strict definition, but it is virtually impossible without defining other deployment models:
+The informal introduction to the concept of a \emph{hybrid cloud} in the previous section requires a strict definition, but creating it is virtually impossible without defining other deployment models:
 \begin{itemize}

### 2014-03-07 thesis/src/chapters/chapter-state-of-the-art.tex
Thesis: state-of-the-art II

@@
 
-\chapterintro{This chapter presents recent achievements in a field of platform-as-a-service model by examining: Carina, OneFlow, CloudFoundry and OpenShift.
+\chapterintro{This chapter presents recent achievements in a field of platform-as-a-service model by examining Carina, OneFlow, CloudFoundry and OpenShift.
 It is especially focused on aspects of adaptivity, scalability and cloud federation awareness.}
@@
 	\item \emph{adaptivity} - ability to adapt (i.e. scale) appropriately to a current usage pattern
-	\item \emph{scalability} - ability to improve application performance by enriching resources
-	\item \emph{cloud-federation awareness} - ability to compose an application deployment using different cloud providers; cooperation with different cloud provider to supply application with extra resources while performing application scaling
+	\item \emph{scalability} - ability to improve application performance by resources enhancement
+	\item \emph{cloud-federation awareness} - ability to compose an application deployment using different cloud providers; cooperation with different cloud providers to supply application with extra resources when performing application scaling
 \end{itemize}
@@
     \end{enumerate}
-  \item One of the elements of the platform is a client application that is used by the user of the platform to communicate with it,
-  \item During the deployment process, the platform takes as an input a description of the service (application) that consists of:
+  \item One of the elements of the platform is a client application that is used by the user of the platform to communicate with it
+  \item During the deployment, the platform takes as an input a description of the service (application) that consists of:
     \begin{itemize}
@@
       \item minimal and maximal number of VMs that are needed for the stack,
-      \item name of the policy (algorithm) which is used for scaling,
+      \item name of the policy (algorithm) used for scaling,
       \item parameters of the policy
@@
     \end{itemize}
-  \item Deployment of a service is done in a way which minimizes the cost from the client's perspective with ensuring Quality-of-Service requirements at the same time,
+  \item Deployment of a service is done in a way which minimises the cost from the client's perspective with ensuring Quality-of-Service requirements at the same time,
   \item The platform monitors the state of the deployed services and based on the results of this process takes appropriate steps in order to meet the auto-scaling requirements. These include
@@
 \subsection{Introduction}
-Carina is an open source project, released under Apache License 2.0, built on top of OpenNebula, which aims to ``(\ldots) standardize the process for automating multi-VM deployments and setting auto-scaling and availability management policies in the cloud.'' \cite{CarinaBlog}.
+Carina is an open source project, released under Apache License 2.0 and built on top of OpenNebula.
+It aims to ``(\ldots) standardize the process for automating multi-VM deployments and setting auto-scaling and availability management policies in the cloud.'' \cite{CarinaBlog}.
 The project is used by the authors at their work at RIM in an OpenNebula-based private cloud.
@@
 \emph{Time-based} policies defines how the system should react in the given window frame(s).
-In each frame we can specify the minimum number of virtual machines that comprises an environment.
+In each frame we can specify the minimal number of virtual machines that comprises an environment.
 On the other hand, \emph{load-based} policies define the way the system reacts to \textbf{average cpu usage} by the environment.
 This is the simplest \emph{threshold-model} as introduced in section 2.1.3.
-The user enters predicates which are evaluated against data gathered by OpenNebula and execute scaling actions  (\emph{scaleup} and \emph{scaledown}) if they are true.
+The user enters predicates which are evaluated against data gathered by OpenNebula and then scaling actions  (\emph{scaleup} and \emph{scaledown}) are executed if predicates are true.
 What is more, the minimal and maximal number of virtual machines must be specified.
@@
 Despite having in its requirements such a position, \emph{Carina} \textbf{does not} support deployment or scaling of services across different providers.
-In the environment configuration we define only \textbf{one} endpoint -- this is can be thought of a reference to a cloud provider.
-As it can be only one per environment, it is not possible to deploy or scale some artifacts across many providers.
+In the environment configuration we define only \textbf{one} endpoint -- it can be thought as a reference to a cloud provider.
+As it can be only one per environment, it is not possible to deploy or scale artifacts across many providers.
 
@@
 The project was initially developed by VMWare and its first release was in 2011, It was then that VMware decided to make it available to the general public and released it under Apache License 2.0.
-For the next consecutive two years, there were maintained two versions of \emph{Cloud Foundry} -- as a hosted solution owned by VMWare and an open-source project.
+For the next consecutive two years, there two versions of \emph{Cloud Foundry} -- as a hosted solution owned by VMWare and an open-source project -- were maintained .
 However, in December 2012 VMware and EMC shared information on the \emph{Pivotal Initiative} \cite{PivotalInitiative} -- a virtual organization of people from both companies with background in \emph{big data} and \emph{cloud application platforms}.
-With the advent of the this new entity, \emph{Pivotal} \cite{GoPivotal}, it was announced the launch of the new product -- Pivotal One, which is powered by \emph{Pivotal CF}, the enterprise version of \emph{Cloud Foundry}.
+With the advent of the this new entity, \emph{Pivotal} \cite{GoPivotal}, it was announced the launch of the new product -- Pivotal One, powered by \emph{Pivotal CF}, the enterprise version of \emph{Cloud Foundry}.
 Now users can choose between the open-source and commercial solutions.
@@
   \item Health management -- continually monitoring application instances and taking according recovery actions if needed
-  \item Standards-based authorization and authentication system which supports various standards such as LDAP, SAML or OAuth 2.0
+  \item Standards-based authorization and authentication system which supports various protocols such as LDAP, SAML or OAuth 2.0
 \end{itemize}
@@
 In terms of monitoring the state of applications, the platform uses \emph{Health Manager} to collect relevant data and then it notifies \emph{Cloud Controller}, which in turn takes appropriate actions.
-This enables, for example, to spawn a new node if an existing one needs a replacement.
+This enables, for example, spawning a new node if an existing one needs a replacement.
 
@@
 In this section we would like to compare all depicted providers in terms of fulfilling functional requirements specified in the overview.
-Table \ref{tab:cloud-providers-adaptivity} takes more detailed approach and summarises how platform tackles adaptivity: their data analysis mechanisms and policies support.
+Table \ref{tab:cloud-providers-comparision-features} presents such comparison.
+Table \ref{tab:cloud-providers-adaptivity} takes more detailed approach and summarises how platform tackles adaptation: their data analysis mechanisms and policies support.
 
@@
 Horizontal scaling & \checkmark & \checkmark & \checkmark & \checkmark \\ \hline
-Multiple provider based scaling & $\times$ & $\times$ & $\times$ & $\times$ \\ \specialrule{.1em}{.05em}{.05em}
+Multiple provider based scaling & $\times$ & $\times$ & \checkmark & $\times$ \\ \specialrule{.1em}{.05em}{.05em}
 

### 2014-03-11 thesis/src/chapters/chapter-design.tex
Thesis: fixes part 2

@@
 As \cite{brun2009engineering} advocates, one of the first architectures of a self-adaptive system was presented in IBM's blueprint for autonomic computing \cite{IBM06}.
-IBM's autonomic system is widely recognised and respected and hence we decided to leverage it as a foundation of our work.
+IBM's autonomic system is widely recognised and respected model, hence, we decided to leverage it as a foundation of our work.
 A concept of autonomic system is alluring due to its four crucial attributes:
@@
 As one can see, Cloud-SAP can be seen as a proxy between the user and cloud instances and as a coordinator of the latter.
-Although the cloud instance itself consists of a cloud manager and additional resources, in general, it is perfectly legitimate to consider it a resource as well.
+Although the cloud instance itself consists of a cloud manager and additional resources, in general, it is perfectly legitimate to consider it as a resource as well.
 
@@
 \end{itemize}
-As resources were identified on various levels of abstraction, starting from the application that is to be deployed and ending on a cloud instance that aggregates different applications and services, we can think of the last three layers (managed resources, touchpoints and touchpoint autonomic managers) as a one logical component which can aggregate and manage other components.
+As resources were identified on various levels of abstraction, starting from the application that is to be deployed and ending on a cloud instance that aggregates different applications and services, the last three layers (managed resources, touchpoints and touchpoint autonomic managers) can be perceived as one logical component which can aggregate and manage other components.
 This recursive definition applies to identified managed resources described in the next section, apart from application as we consider it the most fundamental and indivisible resource.
@@
 \begin{itemize}
-  \item \emph{Application} - application itself can be considered resources which can be managed by an external entity, e.g.
+  \item \emph{Application} - application itself can be considered as a resource which can be managed by an external entity, e.g.
 by restarting/killing it.
@@
   \item \emph{Execution environment of an application} forms an inherent part of an application lifecycle and its proper tuning can greatly influence its performance,
-  \item \emph{Container} -- an isolated and controlled environment in which applications are deployed,
+  \item \emph{Container} -- an isolated and controlled environment in which execution environments are deployed,
   \item \emph{Stack} -- every application has a technology stack associated with it; what is more, if we confine ourselves to map one instance of a container with only one application, we can think of stack as a composition of application (and its execution environment) and a container,
-  \item \emph{Service} -- more complex solutions/applications require interactions with various components that are not necessarily implemented using the same technology stack; service denotes a logical entity that consists of a list of stacks that should form one, indivisible component,
+  \item \emph{Service} -- more complex solutions/applications require interactions with various components that are not necessarily implemented using the same technology stack; service denotes a logical entity that consists of stacks that forms one, indivisible component,
   \item \emph{Cloud instance} -- an entity able to deploy \emph{services}
@@
 The output from one block forms an input to another.
-However, it is perfectly possible for a block to perform an action for obtaining side effects, e.g.
-an action can result in producing relevant information contained in log files, which, in turn, comprises a piece of shared knowledge.
 Each block implements different behaviour based on the abstraction level and resource they operate on (see figure \ref{img:csap-layered-structure}).
-Nevertheless it is possible to define their generic behaviour, which embraces the four phases: monitoring, analysis, planning and execution.
+Nevertheless, it is possible to define their generic behaviour, which embraces the four phases: monitoring, analysis, planning and execution.
 
@@
 \subsection{Orchestrating autonomic manager}
-While an autonomic manager which manages a single resource may be satisfactory enough in many business cases, there are situations when additional coordination among managers is essential to attain the certain goal.
+While an autonomic manager which manages a single resource may be satisfactory in many business cases, there are situations when additional coordination among managers is essential to attain the certain goal.
 This can result in incorporating autonomic behaviour within an entity in a system-wide scope.
@@
 
-As the exemplary actions can be regarded complex ones, they cannot be associated with actions taken by autonomic managers of only one type.
+As the exemplary actions can be regarded as complex, they cannot be associated with actions taken by autonomic managers of only one type.
 Thus, an orchestrating autonomic manager should coordinate other managers of any type -- an arbitrary mixture of self-healing, self-configuring, self-optimizing and self-protecting managers.
@@
 \section{Autonomic cloud instance manager}
-The main responsibility of this manager is the management of a life cycle of resources of a given cloud provider.
+The main responsibility of this manager is the management of resources' life cycle of a given cloud provider.
 It could be thought of an autonomic entity which automates the process of managing resources of data centers that forms the services of a cloud instance.
@@
 \subsection{Managed resource}
-The resources of a cloud provider can be associated with a data center comprising physical nodes, i.e. clusters, hosts, etc.
+The resources of a cloud provider associated with a data center comprising physical nodes, i.e. clusters, hosts, etc.
 
@@
 
-Constraint Programming can be another method that can be implemented \cite{nguyen2009autonomic} to tackle this problem.
+Constraint Programming is another method that can be implemented \cite{nguyen2009autonomic} to tackle this problem.
 This is done by trying to optimize the global utility function of resource provisioning costs and application-specific requirements and constraints.
@@
 \emph{Autonomic cloud federation manager} is a building block of the proposed architecture that forms the highest layer of its model (see figure \ref{img:csap-layered-structure}).
-Additionally, it is an element which directly cooperates with a client by capturing and handling their requests.
+Additionally, it is an element which directly cooperates with a client by capturing and handling his requests.
 
-This manager has two main responsibilities, which include
+This manager has two main responsibilities:
 \begin{inparaenum}[1)]
@@
 \subsection{Managed resource}
-As it was stated in the previous sections and can be seen in figure \ref{img:orchestrating-autonomic-manager}, \emph{cloud instance autonomic managers} are managed resources of this manager.
+As it was stated in the previous sections and can be seen in figure \ref{img:orchestrating-autonomic-manager}, \emph{cloud instance autonomic managers} are managed resources.
 Each cloud instance is a deployment environment for a service or a stack, i.e.
@@
 \subsubsection*{Touchpoints}
-As the cloud federation manager has to gain information about the state of a given cloud instance and be able to order the deployment of a stack or a service, the cloud instance exposes its manageability interface and its two types:
+As the cloud federation manager has to gain information about the state of a given cloud instance and be able to order the deployment of a stack or a service, the cloud instance exposes its manageability interface that consists of:
 \begin{asparaenum}
@@
 The first one is the characteristic of resources available in a cloud -- not only are they geographically distributed, but they can be under different administrative domains and can vary in accessibility.
-This places a heavy burden on cooperation possibilities across cloud providers, both in a technical and non-technical way as providers need to maintain the high level of trust in order to effectively  deliver its products.
+This places a heavy burden on cooperation across cloud providers, both in a technical and non-technical way as providers need to maintain the high level of trust in order to effectively  deliver its products.
 Additionally, different entities forms the resources -- starting from individual devices, through virtual machines and whole workstations, ending with clusters and data centers.
@@
 This should include resource capacity and utilization rates.
-  \item[\textbf{Service scaling}] When an application needs to be scaled, there is available additional knowledge that this component should use of -- detailed information about service's performance up to that point, e.g.
+  \item[\textbf{Service scaling}] When an application needs to be scaled, there is  additional knowledge available that this component should use -- detailed information about service's performance up to that point, e.g.
 service resource utilization level (CPU, I/O), response time, its users profile including their geographical location and so on.
@@
 In such circumstances the cloud provider should be able to quickly adapt to new conditions to prevent the service from being unable to work properly, e.g.
-take into account the geographical location of the application's users and deploy a service in close proximity to them so as to minimize the latency.
+take into account the geographical location of the application's users and deploy a service in proximity to them so as to minimise the latency.
 
@@
 One way to divide prediction models into categories is to classify each one as a \emph{reactive} or a \emph{proactive} one.
-Since in the introductory section there were given reasons why the reactive approach is inefficient, here is a short overview of proactive approaches to the issue.
+Since in the introductory section there were given reasons why the reactive approach is inefficient, here is a short overview of proactive approaches.
 
@@
 
-The knowledge that the component obtains is about current \emph{topology} (which cloud instances make the cloud federation, what is their geographical location, how are they related, etc.) and \emph{policies} (data required to be compared against policies of deployed serviced).
+The knowledge that the component obtains is about current \emph{topology} (which cloud instances make the cloud federation, what is their geographical location, how are they related, etc.) and \emph{policies} (data required to be compared against policies of deployed service).
 What is more, for the system to be able to perform auto-scaling, the gathered information can be classified as \emph{problem determination knowledge} \cite{IBM06}.
@@
 In the introductory section a brief overview of some possible techniques and algorithms that can be used in the analyse component was presented.
-We do not impose a particular method on this component -- it is up to the system complexity and design, which solution would be more suitable to the given needs.
+We do not impose a particular method on this component -- it is up to the system design, which solution would be more suitable to the given needs.
 
 \section{Summary}
-Cloud-SAP fills the gap in existing platforms by providing a standardised approach to scaling applications using a fine-grained actions and operating on different platform levels: starting from an execution environment and ending on a cloud federation.
+Cloud-SAP fills the gap in existing platforms by providing a standardised approach to scaling applications using a fine-grained actions achieved by operating on different platform levels: starting from an execution environment and ending on a cloud federation.
 It is based on a concept of an autonomic system, providing methodical but flexible foundation for its implementation.
-What is more, Cloud-SAP leverages core benefits of such a system: promise of a self-configuration, self-optimisation, self-health, self-protection.
+What is more, Cloud-SAP leverages core benefits of such system: promise of a self-configuration, self-optimisation, self-health, self-protection.
 It is driven by a customer-defined policies and consequently ensures that customer is served with demanded Quality-of-Service and optimised costs.

### 2014-03-11 thesis/src/chapters/chapter-evaluation.tex
Thesis: fixes part 2

@@
 \subsection*{Hardware and virtual machines configuration}
-All tests that were carried out were run on the same hardware and/or used the same virtual machines so for clarity of presentation their parameters are presented not in a description of every test, but here in table \ref{tab:test-deployment-time-common-hardware-configuration}.
+All tests that were carried out were run on the same hardware and/or used the same virtual machines as presented in table \ref{tab:test-deployment-time-common-hardware-configuration}.
 
@@
 \subsection*{Description}
-The aim of this test case is to test the primary use case of the proposed solution -- deployment of a service with the emphasis of client's \textbf{cost}.
+The aim of this test case is to test the primary use case of the proposed solution -- deployment of a service with the emphasis on client's \textbf{cost}.
 It should show that the platform chooses the best mapping between the stacks and cloud providers so that the client's pays the \textbf{lowest} possible price.
@@
 To answer this question we want to look into the description of the virtual machines comprising the deployed environment.
-It states that each VM can use up to 30\% power of the CPU of a host.
+It states that each VM can use up to 30\% power of a single host's CPU.
 Thus, we can be fairly sure that adding another virtual machine that uses 0.3 CPU capacity of its host is equivalent with changing the CPU usage settings of already deployed VM from 30 to 60\%.
@@
 Listing \ref{lst:auto-scaling-1cp-one-config} shows the most important OpenNebula excerpt from settings file used in this test case -- polling interval specification, which was set to 30 seconds.
-To ensure that both products can actually use up-to-date data, they should figuate their policies every 30 seconds or longer.
+To ensure that both products can actually use up-to-date data, they should figure their policies every 30 seconds or longer.
 In Carina the user cannot specify this value, because it is hard-coded in source code to 60 seconds.
@@
 \subsubsection*{Test case duration}
-The test case lasts 25 minutes as the cpu usage is constant and equals 25\% after 21 minutes and 44 seconds.
+The test case lasts 25 minutes as the CPU usage is constant and equals 25\% after 21 minutes and 44 seconds.
 
@@
 Then it gradually goes down, which simulates that all taken actions have successfully completed.
-Then, for another several minutes, the cycle recurs, but this time we expect the proposed solution to scale horizontally.
+Then, after another several minutes, the cycle recurs, but this time we expect the proposed solution to scale horizontally.
 \emph{Carina} is expected to behave as in the previous cycle.
@@
 \subsubsection*{Cost}
-In our scenario cost is roughly equivalent with the CPU usage, so the plot of a cost virtually presents the cpu usage.
-Figure \ref{fig:auto-scaling-1cp-cost-comparison} shows cpu usage by environment when deployed and configured by two competing products.
+In our scenario cost is roughly equivalent with the CPU usage, so the plot of a cost virtually presents the CPU usage.
+Figure \ref{fig:auto-scaling-1cp-cost-comparison} shows CPU usage by environment when deployed and configured by two competing products.
 \begin{figure}[!ht]
@@
 \subsection*{Conclusion}
-The obtained results shows that the proposed solution has enormous potential in terms of better utilization of available resources of the cloud and reducing cost paid by cloud consumers.
+The obtained results shows that the proposed solution has enormous potential in terms of better utilisation of available resources of the cloud and reducing cost paid by cloud consumers.
 
@@
 \subsection*{Conclusion}
-As one can can notice, deployment of a java stack takes using \emph{Cloud-SAP} lasts longer than when using \emph{Carina} for the same purpose.
+As one can can notice, deployment of a java stack using \emph{Cloud-SAP} lasts longer than when using \emph{Carina} for the same purpose.
 Moreover, the more instances are in such stack, the greater difference is.
@@
 
-Noticeably, deployment time never was a crucial issue for \emph{Cloud-SAP}, hence, this issue is of a little importance.
+Noticeably, deployment time never was a crucial issue for \emph{Cloud-SAP}, hence, this matter is of a little importance.
 What is actually important for \emph{Cloud-SAP} is deployment cost and service performance.

### 2014-03-11 thesis/src/chapters/chapter-implementation.tex
Thesis: fixes part 2

@@
 
-      The reason we chose this technology is because of its simplicity, maturity and great support from Ruby platform.
+      The reason why we chose this technology is because of its simplicity, maturity and great support from Ruby platform.
 The other solutions that could be used in place of this one include some message-oriented middleware standards such as AMQP or JMS, other standards that facilitate communications among systems/components such as CORBA, and other such as RMI or RPC.
@@
 OpenNebula is an open source respond to this need -- first released in 2008, with good support from the community, seemed a good candidate for a tool.
-Our familiarity with it was an additional factor which influenced our final choice of OpenNebula as the tool for the job.
+Our familiarity with it was an additional factor which influenced our final choice.
     \item[\textbf{Chef}] As manual virtual machines provisioning is a tedious and error-prone task, this is a perfect match for a tool that would automate the process.
@@
     \item[\textbf{OpenVZ (drivers)}] The significance of choosing linux containers as a virtualisation technology was fully described in the preceding chapters.
-However, the drivers for the OpenVZ are not supported by the OpenNebula maintainers, so we had to use our own-developed tool while taking a course at the university.
-This tools forms an official add-on for OpenNebula 3.8.
+However, the drivers for the OpenVZ are not supported by the OpenNebula maintainers, so we had to use our own-developed drivers while taking a course at the university.
+They form an official add-on for OpenNebula 3.8.
     \end{asparaenum}
@@
   \end{center}
-  \caption{Deployment diagram of a Cloud brokerage subsystem}
+   \caption{Deployment diagram of cloud brokerage subsystem}
   \label{fig:csap-cloud-brokerage-deployment-diagram}
@@
 \subsection{Cloud client}
-The client is a command line application that allows a user to communicate with a cloud broker which will further process the message and act on behalf the user.
+The client is a command line application that allows a user to communicate with a cloud broker which will further process the message and act on the user behalf.
 By means of this application, the user is able to deploy a service by preparing its specification in a JSON format and request its deployment to the cloud broker.
@@
 
-The broker uses Advanced Message Queuing Protocol (AMQP) for communication between itself and cloud providers.
-It was considered the best choice for this purpose as it ensure asynchronous message processing and sending messages according to fan-out paradigm which is ideal for this use case.
+The broker uses Advanced Message Queuing Protocol (AMQP) for communication with cloud providers.
+It was considered the best choice for this purpose as it ensures asynchronous message processing and sending according to fan-out paradigm which is ideal for this use case.
 
@@
 This happens when a service is to be deployed or horizontally scaled.
-As this implementation merely wants to be a minimal viable product, we took only the cost attribute into consideration when performing this function.
+As this implementation merely wants to be a minimal viable product, we took only the cost into consideration when performing this function.
 
@@
 The offer is sent back to the cloud broker for further processing.
-\item[\textbf{Horizontal-scaling message}] This message is send when it is not possible to continue having deployed a stack on the given cloud instance by the lack of sufficient resources or violation of a policy.
+\item[\textbf{Horizontal-scaling message}] This message is send when it is not possible to continue scaling of deployed a stack on the given cloud instance due to the lack of sufficient resources or violation of a policy.
 It contains the stack specification as the platform needs this information to redeploy the stack to another cloud instance.

### 2014-03-11 thesis/src/chapters/chapter-summary.tex
Thesis: fixes part 2

@@
 It is expected that such cooperation broaden pool of supported cloud providers and thus truly enable diversified federation of clouds.
-However, for cloud providers to cooperate, there is a need for discussion and consideration regarding aspects such as authentication, authorization, resource
- access, resource discovery, to name a few.
+However, for cloud providers to cooperate, there is a need for discussion and consideration regarding aspects such as authentication, authorization, resource access and resource discovery, to name a few.

### 2014-03-11 thesis/src/chapters/chapter-state-of-the-art.tex
Thesis: degradation of ch3

@@
-\chapter{Related work}
+\section{Related work}
 
@@
 
-\section{Requirements}
+\subsection{Requirements}
 One can notice that elements that yield a solution to a problem stated in the first chapter, which is ensuring that users' application provide appropriate Quality-of-Service for its customers in a most-cost effective manner, were gradually introduced in the previous chapter:
@@
 
-\subsection{Functional requirements}
+\subsubsection{Functional requirements}
 \begin{enumerate}
@@
 
-\subsection{Non-functional}
+\subsubsection{Non-functional}
 \begin{itemize}
@@
 % radek
-\section{Carina}
+\subsection{Carina}
 
-\subsection{Introduction}
+\subsubsection{Introduction}
 Carina is an open source project, released under Apache License 2.0 and built on top of OpenNebula.
@@
 
-\subsection{Features}
+\subsubsection{Features}
 As it is stated in the requirements of the solution, \emph{Carina} should support variety of features which can be considered worth scrutinizing carefully as they are closely related to notions of adaptivity and scalability.
@@
 
-\subsection{Key technical concepts}
+\subsubsection{Key technical concepts}
 
-\subsubsection{Adaptivity}
+\begin{asparaenum}
+\item[\textbf{Adaptivity}]
 In \emph{Carina} there are mechanisms which can perform application scaling, both in manual and automatic fashion.
@@
 
-\subsubsection{Scaling}
+\item[\textbf{Scaling}]
 Section 2.2 introduced and discussed various types of application scaling.
@@
 
-\subsubsection{Cloud federation awareness}
+\item[\textbf{Cloud federation awareness}]
 In this paper we strongly advocate spanning cloud resources across multiple vendors/providers.
@@
 
-\subsection{Design}
+\end{asparaenum}
+
+\subsubsection{Design}
 To complete the discussion about \emph{Carina} it is essential to include its architecture overview. This is shown in figure \ref{sota:carina-design}.
@@
 
-\section{OneFlow}
+\subsection{OneFlow}
 
-\subsection{Introduction}
+\subsubsection{Introduction}
 \emph{OneFlow} \cite{OneFlow} is a management system integrated into \emph{OpenNebula} that aspires to enhance it with multi-tiered applications (service) deployment and auto-scaling capabilities.
@@
 
-\subsection{Features}
+\subsubsection{Features}
 In a context of this thesis, \emph{OneFlow} enriches \emph{OpenNebula} with only one, yet powerful feature: application scaling based on policies.
@@
 
-\subsection{Key technical concepts}
-\subsubsection{Adaptivity}
+\subsubsection{Key technical concepts}
+
+\begin{asparaenum}
+
+\item[\textbf{Adaptivity}]
 Adaptivity is achieved by specifying policies, that violated, trigger events impacting service environment. There are two types of policies available:
@@
 
-\subsubsection{Scaling}
+\item[\textbf{Scaling}]
 \emph{OneFlow} is capable of \textbf{only} horizontal scaling: it adds or removes virtual machine using \emph{OpenNebula API}.
@@
 
-\subsubsection{Cloud federation awareness}
+\item[\textbf{Cloud federation awareness}]
 \emph{OneFlow} itself \textbf{does not} cooperate with different cloud instance.
@@
 Consequently, it is not possible to supervise provisioned instances.
+\end{asparaenum}
 
@@
 
-\section{OpenShift}
+\subsection{OpenShift}
 
-\subsection{Introduction}
+\subsubsection{Introduction}
 \emph{OpenShift} is a developed by \emph{RedHat} Platform-as-a-Service solution. It is available in two different flavours:
@@
 
-\subsection{Features}
+\subsubsection{Features}
 High level features that distinguishes \emph{OpenShift} are as follows:
@@
 
-\subsection{Key technical concepts}
+\subsubsection{Key technical concepts}
+\begin{asparaenum}
 
-\subsubsection{Adaptivity}
+\item[\textbf{Adaptivity}]
 \emph{OpenShift}'s adaptivity capabilities are constrained to a simple case: while creating a new application, it is possible to specify whether application should scale.
@@
 
-\subsubsection{Scalability}
+\item[\textbf{Scalability}]
 As previous point implies, OpenShift \textbf{solely} uses horizontal scaling: it adds additional gears if number of requests increases beyond a certain level.
 
-\subsubsection{Cloud federation awareness}
+\item[\textbf{Cloud federation awareness}]
 OpenShift does not use any notion of a cloud provider or datacenter.
@@
 
-\subsubsection{Multi-tenancy}
+\item[\textbf{Multi-tenancy}]
 In OpenShift, each application is represented as a set of gears.
@@
 
-\subsection{Summary}
+\end{asparaenum}
+
+\subsubsection{Summary}
 Summing up, key advantage of the OpenShift seems to be it simplicity: creating and deploying applications using OpenShift is as simple as it gets.
@@
 % radek
-\section{CloudFoundry}
-\subsection{Overview}
+\subsection{CloudFoundry}
+\subsubsection{Overview}
 \emph{Cloud Foundry} is a platform-as-a-service solution which can be installed on local or off-premises infrastructure, such as Amazon Web Services, OpenStack or vSphere.
@@
 
-\subsection{History}
+\subsubsection{History}
 The project was initially developed by VMWare and its first release was in 2011, It was then that VMware decided to make it available to the general public and released it under Apache License 2.0.
@@
 
-\subsection{Features}
+\subsubsection{Features}
 Below there are listed some of the main functionalities offered by \emph{Cloud Foundry}:
@@
 
-\subsection{Key technical concepts}
+\subsubsection{Key technical concepts}
 
-\subsubsection{Cloud Foundry-specific}
 \begin{asparaenum}
-\item[\textbf{Warden}] The main goal of \emph{Warden} is to manage isolated and resource-controller environments called containers.
+
+\item[\textbf{Cloud Foundry-specific}] Technical solutions that makes CF interesting are as follows:
+\begin{itemize}
+\item Warden - The main goal of \emph{Warden} is to manage isolated and resource-controller environments called containers.
 It is responsible for management of the whole lifecycle of an environment and provides API for actually performing any operation on a container.
 At first, \emph{Warden} used LXC technology under the hood, but now containers are an own product of \emph{Cloud Foundry} community.
-\item[\textbf{Droplet Execution Agent}] \emph{DEA} manages \emph{Warden} containers, stages user's application and runs droplets -- wrappers around already deployed applications.
-\item[\textbf{BOSH}] \emph{BOSH} is a tool chain for release, deployment and lifecycle management of distributed services. Its functionalities can be compared to those offered by \emph{Chef}.
-\end{asparaenum}
+\item Droplet Execution Agent - \emph{DEA} manages \emph{Warden} containers, stages user's application and runs droplets -- wrappers around already deployed applications.
+\item BOSH - \emph{BOSH} is a tool chain for release, deployment and lifecycle management of distributed services. Its functionalities can be compared to those offered by \emph{Chef}.
+\end{itemize}
 
-\subsubsection{Adaptivity}
+\item[\textbf{Adaptivity}]
 \emph{Cloud foundry} does not support auto-scaling -- it is only possible for the cloud consumer to scale the application manually.
@@
 
-\subsubsection{Scalability}
+\item[\textbf{Scalability}]
 \emph{Cloud foundry} supports horizontal scaling not only within one provider, but also across many. This feature is called \emph{Multi-Cloud} in Cloud Foundry nomenclature.
 
-\subsubsection{Cloud-federation awareness}
+\item[\textbf{Cloud-federation awareness}]
 As it was stated in previous sections, Cloud Foundry is \emph{cloud-federation aware}, i.e.
 it allows users to span their applications across many providers, which results in a hybrid solution that comes in many flavours: \emph{public-public}, \emph{public-private} or \emph{private-private}.
+\end{asparaenum}
 
-\subsection{Summary}
+\subsubsection{Summary}
 Owing to the variety of useful features, great support from community and significant IT enterprises such as IBM, VMware to name a few, \emph{Cloud Foundry} can be considered a mature product which can compete with current leaders in the fields, such as Amazon AWS.
@@
 % radek
-\section{Providers comparison}
+\subsection{Providers comparison}
 In this section we would like to compare all depicted providers in terms of fulfilling functional requirements specified in the overview.
